### 2017-10-26 docs/manuscript.tex
manuscript: wip tex

\documentclass[oneside, abstracton, titlepage]{scrartcl}

\usepackage[left=2.5cm,right=2.6cm,top=3cm,bottom=3cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[intlimits]{amsmath}
\usepackage{amssymb}
\usepackage{color}

% \usepackage{graphicx}

\begin{document}
	\title{Discover governing reactions from concentration data}
	\maketitle

	\section*{Introduction}
	When presented with a time series of possibly noisy concentrations as output of, e.g.,  measurements or simulations that were parameterized by microscopic rates \textcolor{red}{(cite ReaDDy?)}, one can ask for the corresponding macroscopic rates and generating reaction network. 
	In this paper we present an application of the shallow learning method SINDy \cite{Brunton2015} which is able to identify generating parsimonious nonlinear dynamics in data that stems from dynamical systems.
	In our application we, opposed to the original method, do not only look for macroscopic rates of net species change but investigate the specific reactions that might have lead to the observations.
	We demonstrate the algorithm on a toy problem demonstrating the identification abilities.

	\section*{The method}
	The underlying model that we want to fit the data to is a law of mass action type dynamical system. To this end, let $S$ be number of species and $T$ be the number observations. The concentration data at a time $t\in [1, T]$ can be represented by a vector
	\[
	\mathbf{x}(t)=\begin{pmatrix}
	x_1(t)\\ \vdots \\ x_S(t)
	\end{pmatrix}\in \mathbb{R}^S.
	\]
	Further, one can choose $R$ possible ansatz reactions with their respective reaction function
	\[
	\theta_r(\textbf{x}(t))=\begin{pmatrix}
	\theta_{r,1}(\textbf{x}(t)) \\ \vdots \\ \theta_{r,S}(\textbf{x}(t))
	\end{pmatrix}
	\]
	so that the change of concentration for species $i$ at time $t$ is represented by the dynamical system
	\[
	\dot{\textbf{x}}_i(t) = \sum_{r=1}^{R}\theta_{r,i}(\textbf{x}(t))\xi_r,
	\]
	where $\xi_r$ are the to-be estimated macroscopic reaction rates.

	\section*{Examples}

	\subsection*{Regression without regularization}

	\subsection*{Regression with regularization}

	\newpage
	\bibliographystyle{alpha}
	\bibliography{bibliography.bib}

\end{document}

### 2017-10-26 readdy_learn/analyze/derivative.py

@@ -185,7 +185,7 @@ def test_finite_differences():
     testf = np.array([np.sin(x) for x in x0])
     true_deriv = [np.cos(x) for x in x0]
 
-    wwidth = 3
+    wwidth = 5
     deriv = np.empty_like(x0)
     for ix, (wx, wy) in enumerate(zip(window_evensteven(x0, width=wwidth), window_evensteven(testf, width=wwidth))):
         x = x0[ix]
@@ -216,4 +216,4 @@ def test_ld_derivative():
 
 if __name__ == '__main__':
     test_finite_differences()
-    # test_ld_derivative()
+    #test_ld_derivative()

### 2017-10-27 docs/manuscript.tex
manuscript: wip tex

@@
 % \usepackage{graphicx}
+\DeclareMathOperator*{\argmin}{arg\,min}
 
@@
 	The underlying model that we want to fit the data to is a law of mass action type dynamical system.
-To this end, let $S$ be number of species and $T$ be the number observations.
-The concentration data at a time $t\in [1, T]$ can be represented by a vector
+To this end, let $S$ be the number of species, then the concentration data at a time $t$ can be represented by a vector
 	\[
@@
 	\[
-	\theta_r(\textbf{x}(t))=\begin{pmatrix}
-	\theta_{r,1}(\textbf{x}(t)) \\ \vdots \\ \theta_{r,S}(\textbf{x}(t))
+	\textbf{y}_r(\textbf{x}(t))=\begin{pmatrix}
+	y_{r,1}(\textbf{x}(t)) \\ \vdots \\ y_{r,S}(\textbf{x}(t))
 	\end{pmatrix}
 	\]
-	so that the change of concentration for species $i$ at time $t$ is represented by the dynamical system
+	so that the change of concentration for species $i$ at time $t_i$, is represented by the dynamical system
 	\[
-	\dot{\textbf{x}}_i(t) = \sum_{r=1}^{R}\theta_{r,i}(\textbf{x}(t))\xi_r,
+	\dot{\textbf{x}}_i(t) = \sum_{r=1}^{R}y_{r,i}(\textbf{x}(t))\xi_r,\quad i=1,\ldots, S,
 	\]
-	where $\xi_r$ are the to-be estimated macroscopic reaction rates.
+	where $\xi_r$ are the to-be estimated macroscopic rates.
+
+	When presented with a time series with $T$ observations, the data can be represented as a matrix
+	\[
+	\textbf{X} = \begin{pmatrix}
+		x_1(t_1) & x_2(t_1) & \cdots & x_S(t_1) \\
+		x_1(t_2) & x_2(t_2) & \cdots & x_S(t_2) \\
+		\vdots   & \vdots   & \ddots & \vdots   \\
+		x_1(t_T) & x_2(t_T) & \cdots & x_S(t_T)
+	\end{pmatrix} \in \mathbb{R}^{T\times S}.
+	\]
+	Given this matrix, one can propose a library $\Theta(\textbf{X}) = \begin{pmatrix} \theta_1(\textbf{X}) & \theta_2(\textbf{X}) & \cdots & \theta_R(\textbf{X}) \end{pmatrix}$ of $R$ ansatz reactions with corresponding reaction functions
+	\[
+	\theta_r(\textbf{X}) = \begin{pmatrix}
+	\textbf{y}_r(\textbf{X}_1)^T \\ \vdots \\ \textbf{y}_r(\textbf{X}_T)^T
+	\end{pmatrix}\in \mathbb{R}^{T\times S},\quad r=1,\ldots,R,
+	\]
+	where $\textbf{X}_i$ denotes the $i$-th row in $X$.
+Applying the concentration trajectory to the library yields a data tensor $\Theta(\textbf{X})\in\mathbb{R}^{T\times S\times R}$.
+Following the approach of SINDy, the goal is now to find coefficients $\Xi = \begin{pmatrix} \xi_1 & \xi_2 & \cdots & \xi_R
+	\end{pmatrix}^T$, so that
+	\[
+	\dot{\textbf{X}} = \Theta(\textbf{X})\Xi = \sum_{r=1}^{R}\theta_r(\textbf{X})\xi_r.
+	\]
+	In particular, the system is linear in the coefficients $\Xi$, which makes sparse regression tools such as LASSO \cite{Tibshirani1996, Hastie2009} applicable.
+To this end, one can consider the minimization problem to find $\hat{\Xi}$ such that
+	\begin{align}
+		\hat{\Xi} = \argmin_{\Xi}\left( \frac{1}{2T}\left\| \dot{\textbf{X}} - \Theta(\textbf{X})\Xi \right\|_F^2 + \alpha\|\Xi\|_1 \right) \quad \text{subject to }\Xi \geq 0,
+	\label{method:minimizationproblem}\end{align}
+	where $\alpha\geq 0$ is a sparsity inducing hyperparameter.
+For $\alpha=0$ this problem reduces to constrained least-squares.
+For solving (\ref{method:minimizationproblem}) we applied the sequential least squares minimizer SLSQP, originally described in \cite{Kraft1988}, contained in the software package SciPy \cite{SciPy}.
 
@@
 	\newpage
-	\bibliographystyle{alpha}
+% 	\bibliographystyle{alpha}
+	\bibliographystyle{abbrv}
 	\bibliography{bibliography.bib}

### 2017-10-27 readdy_learn/generate/generate_tools/kinetic_monte_carlo.py
method to calculate smallest time step

@@ -418,6 +418,10 @@ class ReactionDiffusionSystem:
     def sequence(self):
         return self._event_list, self._time_list, self._state_list
 
+    @property
+    def smallest_dt(self):
+        return np.min(np.diff(self._time_list))
+
     # @todo move this into a result(event sequence)-object
     def convert_events_to_time_series(self, time_step=None, n_frames=None):
         """

### 2017-10-27 docs/manuscript.tex
update manuscript

@@
 
-% \usepackage{graphicx}
-\DeclareMathOperator*{\argmin}{arg\,min}
+\usepackage{graphicx}
 
@@
 	\begin{align}
-		\hat{\Xi} = \argmin_{\Xi}\left( \frac{1}{2T}\left\| \dot{\textbf{X}} - \Theta(\textbf{X})\Xi \right\|_F^2 + \alpha\|\Xi\|_1 \right) \quad \text{subject to }\Xi \geq 0,
+		\hat{\Xi} = \underset{\Xi}{\arg\min}\left( \frac{1}{2T}\left\| \dot{\textbf{X}} - \Theta(\textbf{X})\Xi \right\|_F^2 + \alpha\|\Xi\|_1 \right) \quad \text{subject to }\Xi \geq 0,
 	\label{method:minimizationproblem}\end{align}
@@
 For $\alpha=0$ this problem reduces to constrained least-squares.
-For solving (\ref{method:minimizationproblem}) we applied the sequential least squares minimizer SLSQP, originally described in \cite{Kraft1988}, contained in the software package SciPy \cite{SciPy}.
+For solving (\ref{method:minimizationproblem}) we apply the sequential least squares minimizer SLSQP, originally described in \cite{Kraft1988}, contained in, e.g., the software package SciPy \cite{SciPy}.
 
@@
 
+	\begin{figure}
+		\begin{center}
+			\includegraphics[width=.5\textwidth]{./figures_tex/aux_linear_approximation.eps}
+		\end{center}
+		\caption{This figure depicts a possible output yielded by an application of the Gillespie method and the performed piecewise linear approximation in order to obtain a better behaving derivative.}
+		\label{examples:linear-approximation}
+	\end{figure}
+
+	For generating time series data of concentrations we use the Gillespie method \cite{Gillespie1976}.
+For a given set of initial conditions, we produce several realizations which are then converted to a trajectory with fixed time step and averaged.
+Since these trajectories are piecewise constant, one has to take special care when approximating the temporal derivative.
+Simply taking finite differences has the effect that, with decreasing time step, the derivative is most of the time constant zero and approaches infinity at the jump discontinuities.
+To counter this effect, we perform a linear approximation between each two adjacent discontinuities, as depicted in Figure \ref{examples:linear-approximation}.
 	\subsection*{Regression without regularization}
@@
 	\subsection*{Regression with regularization}
+
+	\section*{Conclusion}
+	dfdf
 

### 2017-10-27 readdy_learn/analyze/generate.py

@@ -8,7 +8,7 @@ def generate_continuous_counts(rates, initial_condition, bfc, timestep, n_steps)
         theta = _np.array([f(data) for f in bfc.functions])
         return _np.matmul(rates, theta)
 
-    xs = _np.linspace(0, n_steps*timestep, num=timestep, endpoint=False)
+    xs = _np.linspace(0, n_steps*timestep, num=n_steps, endpoint=False)
     initial_condition = _np.array(initial_condition).squeeze()
     return xs, _odeint(fun_reference, initial_condition, xs)
 

### 2017-10-27 readdy_learn/test/test_deriv.py

@@ -17,7 +17,7 @@ class TestDeriv(unittest.TestCase):
         """
         test the jac implementation against an approximated one!
         """
-        plot = False
+        plot = True
         desired_rates = np.array([.02, .07])
         init_condition = [[70, 50]]
         n_steps = 5000
@@ -41,7 +41,6 @@ class TestDeriv(unittest.TestCase):
         for i in range(len(rates_ab)):
             ys[i] = jac(np.array([rates_ab[i], rates_ba[i]]))
             ys_approx[i] = jac_approx(np.array([rates_ab[i], rates_ba[i]]))
-        print(ys.shape)
         if plot:
             plt.plot(rates_ab, ys[:, 0], label='analytical jac d(a->b)')
             plt.plot(rates_ab, ys_approx[:, 0], label='approximated jac d(a->b)')

### 2017-10-27 docs/manuscript.tex
wip

@@
 Simply taking finite differences has the effect that, with decreasing time step, the derivative is most of the time constant zero and approaches infinity at the jump discontinuities.
-To counter this effect, we perform a linear approximation between each two adjacent discontinuities, as depicted in Figure \ref{examples:linear-approximation}.
+To counter this effect, we first perform a linear approximation between each two adjacent discontinuities, as depicted in Figure \ref{examples:linear-approximation}.
+(\textcolor{red}{Use regularized derivative?})
+
 	\subsection*{Regression without regularization}

### 2017-10-27 moritz/convergence_simple.py

@@ -4,7 +4,7 @@ import os
 import numpy as np
 
 import readdy_learn.generate.generate_tools.kinetic_monte_carlo as kmc
-from readdy_learn.analyze.sklearn import BasisFunctionConfiguration
+from readdy_learn.analyze.basis import BasisFunctionConfiguration
 from readdy_learn.sample_tools import Suite
 
 
@@ -13,16 +13,13 @@ def set_up_system():
                                       species_names=["A", "B"])
     sys.add_conversion("A", "B", np.array([4.]))
     sys.add_conversion("B", "A", np.array([0.5]))
-
-    bfc = BasisFunctionConfiguration(n_species=sys.n_species)
-    bfc.add_conversion(0, 1)  # A -> B
-    bfc.add_conversion(1, 0)  # B -> A
-    return sys, bfc
-
+    return sys
 
 if __name__ == '__main__':
 
-    suite = Suite(set_up_system, interp_degree='pw_linear')
+    bfc = BasisFunctionConfiguration(n_species=2)
+    bfc.add_conversion(0, 1)  # A -> B
+    bfc.add_conversion(1, 0)  # B -> A
 
     parser = argparse.ArgumentParser()
     parser.add_argument("-o", "--outfile", help="the outfile", type=str)
@@ -47,6 +44,7 @@ if __name__ == '__main__':
 
     if args.plot:
         print("---> got plot argument, just plotting the outfile if it exists")
+        suite = Suite(set_up_system, bfc, interp_degree='pw_linear')
         suite.plot(outfile)
     else:
         if not args.n_steps:
@@ -55,11 +53,13 @@ if __name__ == '__main__':
             n_steps = args.n_steps
 
         if not args.n_realizations:
-            n_realizations = 20
+            n_realizations = 8
         else:
             n_realizations = args.n_realizations
 
         print("---> running analysis for n_steps={} with n_realizations={}".format(n_steps, n_realizations))
         timesteps = [.00001, .0001] + [x for x in np.arange(.001, .1, step=.001)]
+
+        suite = Suite(set_up_system, bfc, interp_degree='pw_linear')
         suite.calculate(outfile, timesteps=timesteps, n_steps=n_steps, n_realizations=n_realizations,
-                        verbose=args.verbose, write_concentrations_for_time_step=.001)
+                        verbose=args.verbose, write_concentrations_for_time_step=.001, n_gillespie_realizations=8)

### 2017-10-27 readdy_learn/sample_tools.py

@@ -8,6 +8,8 @@ from pathos.multiprocessing import Pool
 import readdy_learn.analyze.tools as pat
 from readdy_learn.analyze.sklearn import ReaDDyElasticNetEstimator
 
+import readdy_learn.analyze.generate as generate
+
 
 class Suite(object):
     def __init__(self, set_up_system, bfc, alpha=0., l1_ratio=1., maxiter=30000, tol=1e-12, interp_degree=10,
@@ -54,10 +56,10 @@ class Suite(object):
                                         options={'ftol': self._tol}, rescale=False, init_xi=self._init_xi)
         return est
 
-    def run(self, sys, bfc, verbose=True, n_frames=None, timestep=None):
+    def run(self, times, counts, bfc, verbose=True, n_frames=None, timestep=None, realization=None):
         if verbose:
             print("---- begin suite run")
-        counts, times, config = sys.get_counts_config(n_frames=n_frames, timestep=timestep)
+        # counts, times, config = sys.get_counts_config(n_frames=n_frames, timestep=timestep)
 
         traj = pat.Trajectory.from_counts(counts, times[1] - times[0], verbose=verbose,
                                           interp_degree=self._interp_degree)
@@ -148,7 +150,7 @@ class Suite(object):
         plt.show()
 
     def calculate(self, file, timesteps, n_steps, n_realizations=1, write_concentrations_for_time_step=None,
-                  verbose=True, save=True, njobs=8):
+                  verbose=True, save=True, njobs=8, n_gillespie_realizations=1):
         if os.path.exists(file):
             raise ValueError("File already existed: {}".format(file))
 
@@ -169,14 +171,16 @@ class Suite(object):
                 return self.run(**args)
 
             params = []
-            for n in range(n_realizations):
-                system = self._get_system()
-                system.simulate(n_steps)
-                for dt in timesteps:
-                    params.append({'sys': system, 'bfc': self._bfc, 'timestep': dt, 'verbose': verbose})
+            for dt in timesteps:
+                for n in range(n_realizations):
+                    times, counts = generate.generate_averaged_kmc_counts(self._set_up_system, n_steps, timestep=dt,
+                                                                          n_realizations=n_gillespie_realizations)
+
+                    params.append({'times': times, 'counts': counts, 'bfc': self._bfc, 'timestep': dt,
+                                   'verbose': verbose, 'realization': n})
                     if dt == write_concentrations_for_time_step:
[…]

### 2017-10-30 docs/manuscript.tex
wip

@@
 	\section*{Introduction}
-	When presented with a time series of possibly noisy concentrations as output of, e.g.,  measurements or simulations that were parameterized by microscopic rates \textcolor{red}{(cite ReaDDy?)}, one can ask for the corresponding macroscopic rates and generating reaction network.
-	In this paper we present an application of the shallow learning method SINDy \cite{Brunton2015} which is able to identify generating parsimonious nonlinear dynamics in data that stems from dynamical systems.
+	When presented with a time series of possibly noisy concentration values as output of, e.g., measurements or simulations that were parameterized by microscopic rates \textcolor{red}{(cite ReaDDy?)}, one can ask for the corresponding macroscopic rates and generating reaction network.
+	In this paper we present an application of the shallow learning method SINDy \cite{Brunton2015} which is able to identify the generating parsimonious nonlinear dynamics in data that stems from dynamical systems, thus providing an interpretable result.
 	In our application we, opposed to the original method, do not only look for macroscopic rates of net species change but investigate the specific reactions that might have lead to the observations.

### 2017-10-30 readdy_learn/analyze/derivative.py

@@ -62,6 +62,13 @@ def ld_derivative2(data, xs, alpha, maxit=1000, verbose=False):
 
     u = np.gradient(data, xs)
 
+    wwidth = 2
+    deriv = np.empty_like(xs)
+    for ix, (wx, wy) in enumerate(zip(window_evensteven(xs, width=wwidth), window_evensteven(data, width=wwidth))):
+        x = xs[ix]
+        coeff = fd_coeff(x, wx, k=1)
+        deriv[ix] = coeff.dot(wy)
+
 
 def ld_derivative(data, timestep, alpha, maxit=1000, verbose=False):
     assert isinstance(data, np.ndarray)
@@ -199,21 +206,29 @@ def test_finite_differences():
 
 def test_ld_derivative():
     x0 = np.arange(0, 2.0 * np.pi, 0.05)
+    xx = []
+    for x in x0:
+        if np.random.random() < .2:
+            xx.append(x)
+    x0 = np.array(xx)
+
     testf = np.array([np.sin(x) for x in x0])
     testf = testf + np.random.normal(0.0, 0.04, x0.shape)
     true_deriv = [np.cos(x) for x in x0]
 
-    deriv_sm = ld_derivative(testf, timestep=0.05, alpha=5e-4, verbose=False)
-    deriv_lrg = ld_derivative(testf, timestep=0.05, alpha=1e-1)
+    ld_derivative2(testf, x0, alpha=1e-1)
 
-    plt.plot(testf, label='fun')
-    plt.plot(deriv_sm, label='alpha=5e-4')
-    plt.plot(deriv_lrg, label='alpha=1e-1')
-    plt.plot(true_deriv, label='derivative')
-    plt.legend()
-    plt.show()
+    deriv_sm = ld_derivative(testf, timestep=0.05, alpha=5e-4, verbose=False)
+    #deriv_lrg = ld_derivative(testf, timestep=0.05, alpha=1e-1)
+
+    #plt.plot(testf, label='fun')
+    #plt.plot(deriv_sm, label='alpha=5e-4')
+    #plt.plot(deriv_lrg, label='alpha=1e-1')
+    #plt.plot(true_deriv, label='derivative')
+    #plt.legend()
+    #plt.show()
 
 
 if __name__ == '__main__':
-    test_finite_differences()
-    #test_ld_derivative()
+    #test_finite_differences()
+    test_ld_derivative()

### 2017-10-30 docs/manuscript.tex
manuscript wip

@@
 
-	\section*{Introduction}
+	\section{Introduction}
 	When presented with a time series of possibly noisy concentration values as output of, e.g., measurements or simulations that were parameterized by microscopic rates \textcolor{red}{(cite ReaDDy?)}, one can ask for the corresponding macroscopic rates and generating reaction network.
@@
 
-	\section*{The method}
+	\section{The method}
 	The underlying model that we want to fit the data to is a law of mass action type dynamical system.
@@
 
-	\section*{Examples}
+	\section{Examples}
 
@@
 	For generating time series data of concentrations we use the Gillespie method \cite{Gillespie1976}.
-For a given set of initial conditions, we produce several realizations which are then converted to a trajectory with fixed time step and averaged.
+For a given set of initial conditions, we produce several realizations which then are converted to a trajectory with fixed time step and averaged.
 Since these trajectories are piecewise constant, one has to take special care when approximating the temporal derivative.
 Simply taking finite differences has the effect that, with decreasing time step, the derivative is most of the time constant zero and approaches infinity at the jump discontinuities.
-To counter this effect, we first perform a linear approximation between each two adjacent discontinuities, as depicted in Figure \ref{examples:linear-approximation}.
+To counter this effect, we beforehand perform a linear approximation between each two adjacent discontinuities, as depicted in Figure \ref{examples:linear-approximation}.
 (\textcolor{red}{Use regularized derivative?})
 
-	\subsection*{Regression without regularization}
+	\subsection{Regression without regularization}
+	\label{sec:ex1}
 
-	\subsection*{Regression with regularization}
+	Here we demonstrate that, with decreasing time step $\Delta t$ and $\alpha=0$, the estimated rates converge to the true rates if there is no ambiguity in the system.
+The systems contains two species $A$ and $B$ which can convert into each other with macroscopic rates of $r_1=4.0$ and $r_2=0.5$, respectively.
+The initial amount of particles is set to $70$ of type $A$ and $0$ of type $B$.
+A possibly resulting change of concentrations over time is depicted in Figure \ref{ex1:concentration-curves}.
 
-	\section*{Conclusion}
+	\begin{figure}
+		\begin{center}
+			\includegraphics[width=.5\textwidth]{./figures_tex/aux_ex1_concentration_curves}
+		\end{center}
+		\caption{A realization of the example given in Section \ref{sec:ex1}.}
+		\label{ex1:concentration-curves}
+	\end{figure}
+
+	For each $\Delta t$, we averaged eight different realizations with identical starting conditions into one single trajectory and estimated the rates for reactions $A\to B$ and $B\to A$.
+This procedure, we repeated $30$.
+In Figure \ref{ex1:rates-convergence}, the average estimates out of these realizations alongside with their standard deviations are shown.
+One can see that, with decreasing $\Delta t$, also the estimate for the rates becomes better and they can be recovered.
+
+	\begin{figure}
+		\begin{center}
+			\includegraphics[width=.5\textwidth]{./figures_tex/aux_rates_convergence}
+		\end{center}
+		\caption{The \textcolor{red}{todo}}
+		\label{ex1:rates-convergence}
+	\end{figure}
+
+	\subsection{Regression with regularization}
+
+	\section{Conclusion}
 	dfdf

### 2017-10-31 docs/manuscript.tex
wip

@@
 	\section{Introduction}
-	When presented with a time series of possibly noisy concentration values as output of, e.g., measurements or simulations that were parameterized by microscopic rates \textcolor{red}{(cite ReaDDy?)}, one can ask for the corresponding macroscopic rates and generating reaction network.
+	\textcolor{red}{non-equilibrium.}
+	When presented with a time series of possibly noisy concentration fluctuations of some species as output of, e.g., measurements from experiments or simulations that were parameterized by microscopic rates \textcolor{red}{(cite ReaDDy?)}, one can ask for the corresponding macroscopic rates and generating reaction network.
 	In this paper we present an application of the shallow learning method SINDy \cite{Brunton2015} which is able to identify the generating parsimonious nonlinear dynamics in data that stems from dynamical systems, thus providing an interpretable result.
 	In our application we, opposed to the original method, do not only look for macroscopic rates of net species change but investigate the specific reactions that might have lead to the observations.
-	We demonstrate the algorithm on a toy problem demonstrating the identification abilities.
+	We demonstrate the algorithm on two toy problems - one problem showing that when there is no ambiguity in the system, one converges to the correct rates with increasing resolution of concentration fluctuations and one in which we compute a sparse reaction network for given data.
 
@@
 		\end{center}
-		\caption{The \textcolor{red}{todo}}
+		\caption{Convergence of estimated rates to the ground truth with decreasing $\Delta t$.}
 		\label{ex1:rates-convergence}

### 2017-11-01 docs/manuscript.tex
wip

@@
 	\section{Introduction}
-	\textcolor{red}{non-equilibrium.}
-	When presented with a time series of possibly noisy concentration fluctuations of some species as output of, e.g., measurements from experiments or simulations that were parameterized by microscopic rates \textcolor{red}{(cite ReaDDy?)}, one can ask for the corresponding macroscopic rates and generating reaction network.
-	In this paper we present an application of the shallow learning method SINDy \cite{Brunton2015} which is able to identify the generating parsimonious nonlinear dynamics in data that stems from dynamical systems, thus providing an interpretable result.
+	When presented with a non-equilibrium time series of possibly noisy concentration fluctuations of some species as output of, e.g., measurements from experiments or simulations that were parameterized by microscopic rates \textcolor{red}{(cite ReaDDy?)}, one can ask for the corresponding macroscopic rates and generating reaction network.
+	In this paper we present an application of the shallow learning method SINDy \cite{Brunton2015}.
+It is able to identify the generating parsimonious nonlinear dynamics in data that stems from dynamical systems, thus providing an interpretable result.
 	In our application we, opposed to the original method, do not only look for macroscopic rates of net species change but investigate the specific reactions that might have lead to the observations.
@@
 For solving (\ref{method:minimizationproblem}) we apply the sequential least squares minimizer SLSQP, originally described in \cite{Kraft1988}, contained in, e.g., the software package SciPy \cite{SciPy}.
+As initial guess for the regulated problem we use the solution of (\ref{method:minimizationproblem}) with $\alpha=0$.
 
@@
 	For each $\Delta t$, we averaged eight different realizations with identical starting conditions into one single trajectory and estimated the rates for reactions $A\to B$ and $B\to A$.
-This procedure, we repeated $30$.
-In Figure \ref{ex1:rates-convergence}, the average estimates out of these realizations alongside with their standard deviations are shown.
+This procedure, we repeated $30$ times.
+In Figure \ref{ex1:rates-convergence}, the average estimates alongside with their standard deviations are shown.
 One can see that, with decreasing $\Delta t$, also the estimate for the rates becomes better and they can be recovered.
@@
 	\subsection{Regression with regularization}
+	In this example we have four species $A$, $B$, $C$, and $D$ with reactions $A\xrightarrow{k_1}D$, $D\xrightarrow{k_2}A$, $D\xrightarrow{k_3}B$, $A+B\xrightarrow{k_4}C$, and $C+D\xrightarrow{k_5}B$, where $k_1=k_2=2\cdot 10^{-2}$, $k_3 = 1\cdot 10^{-2}$, and $k_4=k_5=1\cdot 10^{-4}$.
 
 	\section{Conclusion}
-	dfdf
+	todo.
 

### 2017-11-01 docs/manuscript.tex
wip

@@
 \usepackage{color}
+\usepackage{subcaption}
 
@@
 	\subsection{Regression with regularization}
-	In this example we have four species $A$, $B$, $C$, and $D$ with reactions $A\xrightarrow{k_1}D$, $D\xrightarrow{k_2}A$, $D\xrightarrow{k_3}B$, $A+B\xrightarrow{k_4}C$, and $C+D\xrightarrow{k_5}B$, where $k_1=k_2=2\cdot 10^{-2}$, $k_3 = 1\cdot 10^{-2}$, and $k_4=k_5=1\cdot 10^{-4}$.
+	\label{sec:ex2}
+	In this example there are four species $A$, $B$, $C$, and $D$ with reactions $A\xrightarrow{k_1}D$, $D\xrightarrow{k_2}A$, $D\xrightarrow{k_3}B$, $A+B\xrightarrow{k_4}C$, and $C+D\xrightarrow{k_5}B$, where $k_1=k_2=2\cdot 10^{-2}$, $k_3 = 1\cdot 10^{-2}$, and $k_4=k_5=1\cdot 10^{-4}$.
+These reactions are inserted into the ansatz library $\Theta (\cdot)$.
+Additionally the reactions $A\leftrightarrow B$, $A\leftrightarrow C$, and $A+C\leftrightarrow D$ are inserted into the library.
+
+	\begin{figure}
+		\centering
+		\begin{subfigure}[b]{.45\textwidth}
+			\includegraphics[width=\textwidth]{./figures_tex/cv_concentrations_train}
+			\caption{Train data set.}
+		\end{subfigure}
+		~
+		\begin{subfigure}[b]{.45\textwidth}
+			\includegraphics[width=\textwidth]{./figures_tex/cv_concentrations_test}
+			\caption{Test data set.}
+		\end{subfigure}
+		\caption{Train and test data sets for the example given in Section \ref{sec:ex2}.}
+		\label{ex2:test-train-concentration-curves}
+	\end{figure}
+
+	The problem now is to find a suitable $\alpha$ such that a parsimonious solution is picked.
+To this end, cross-validation is applied.
+The trajectories that were used as training and test data set are depicted in Figure \ref{ex2:test-train-concentration-curves}.
 

### 2017-11-01 docs/manuscript.tex

@@
 To this end, let $S$ be the number of species, then the concentration data at a time $t$ can be represented by a vector
-	\[
+	\begin{align}
 	\mathbf{x}(t)=\begin{pmatrix}
@@
 	\end{pmatrix}\in \mathbb{R}^S.
-	\]
+	\end{align}
 	Further, one can choose $R$ possible ansatz reactions with their respective reaction function
-	\[
+	\begin{align}
 	\textbf{y}_r(\textbf{x}(t))=\begin{pmatrix}
@@
 	\end{pmatrix}
-	\]
+	\end{align}
 	so that the change of concentration for species $i$ at time $t_i$, is represented by the dynamical system
-	\[
+	\begin{align}
 	\dot{\textbf{x}}_i(t) = \sum_{r=1}^{R}y_{r,i}(\textbf{x}(t))\xi_r,\quad i=1,\ldots, S,
-	\]
+	\end{align}
 	where $\xi_r$ are the to-be estimated macroscopic rates.
@@
 	When presented with a time series with $T$ observations, the data can be represented as a matrix
-	\[
+	\begin{align}
 	\textbf{X} = \begin{pmatrix}
@@
 	\end{pmatrix} \in \mathbb{R}^{T\times S}.
-	\]
+	\end{align}
 	Given this matrix, one can propose a library $\Theta(\textbf{X}) = \begin{pmatrix} \theta_1(\textbf{X}) & \theta_2(\textbf{X}) & \cdots & \theta_R(\textbf{X}) \end{pmatrix}$ of $R$ ansatz reactions with corresponding reaction functions
-	\[
+	\begin{align}
 	\theta_r(\textbf{X}) = \begin{pmatrix}
@@
 	\end{pmatrix}\in \mathbb{R}^{T\times S},\quad r=1,\ldots,R,
-	\]
+	\end{align}
 	where $\textbf{X}_i$ denotes the $i$-th row in $X$.
@@
 	\end{pmatrix}^T$, so that
-	\[
+	\begin{align}
 	\dot{\textbf{X}} = \Theta(\textbf{X})\Xi = \sum_{r=1}^{R}\theta_r(\textbf{X})\xi_r.
-	\]
+	\end{align}
 	In particular, the system is linear in the coefficients $\Xi$, which makes sparse regression tools such as LASSO \cite{Tibshirani1996, Hastie2009} applicable.
@@
 	\label{method:minimizationproblem}\end{align}
-	where $\alpha\geq 0$ is a sparsity inducing hyperparameter.
+	where $\alpha\geq 0$ is a sparsity inducing hyperparameter and $\|\cdot\|_F$ denotes the Frobenius norm.
 For $\alpha=0$ this problem reduces to constrained least-squares.
@@
 The trajectories that were used as training and test data set are depicted in Figure \ref{ex2:test-train-concentration-curves}.
+\textcolor{red}{Perhaps use several test trajectories with different initial conditions to get something like a confidence interval?} As an indicator to the estimation error we use the scoring function
+	\begin{align}
+		f_\alpha = \frac{1}{T}\|\mathbf{\dot{X}}_\text{test} - \Theta (\textbf{X}_\text{test}) \Xi_\text{train}\|_F^2.
+		\label{ex2:score-function}
+	\end{align}
+	For a selection of values $\alpha\in [0, 100]$, the model parameters $\Xi$ were estimated and the scoring function (\ref{ex2:score-function}) evaluated, yielding a graph as in Figure \ref{ex2:score-graph}.
+
+	\begin{figure}
+		\begin{center}
+			\includegraphics[width=.5\textwidth]{./figures_tex/cv_score}
+		\end{center}
+		\caption{Score function (\ref{ex2:score-function}) evaluated for a range of $\alpha$ values.}
+		\label{ex2:score-graph}
+	\end{figure}
+
+	One can see that the minimum is at about
+	\begin{align}
+		\hat{\alpha} = \underset{\alpha\in[0,100]}{\arg\min}f_\alpha\approx 10.
+	\end{align}
+	When evaluating the minimization problem (\ref{method:minimizationproblem}) with this hyperparameter and applying a cut-off of $10^{-8}$ to the yielded parameters, one obtains
 

### 2017-11-02 docs/manuscript.tex
wip: manuscript

@@
 \begin{document}
-	\title{Discover governing reactions from concentration data}
+	\title{Discovering governing reactions from concentration data}
 	\maketitle
@@
 	\dot{\textbf{x}}_i(t) = \sum_{r=1}^{R}y_{r,i}(\textbf{x}(t))\xi_r,\quad i=1,\ldots, S,
-	\end{align}
+	\label{method:the-system}\end{align}
 	where $\xi_r$ are the to-be estimated macroscopic rates.
@@
 		\end{pmatrix}\in \mathbb{R}^{T\times S},\quad r=1,\ldots,R,
-	\end{align}
+	\label{method:the-reactions}\end{align}
 	where $\textbf{X}_i$ denotes the $i$-th row in $X$.
@@
 	\label{sec:ex2}
-	In this example there are four species $A$, $B$, $C$, and $D$ with reactions $A\xrightarrow{k_1}D$, $D\xrightarrow{k_2}A$, $D\xrightarrow{k_3}B$, $A+B\xrightarrow{k_4}C$, and $C+D\xrightarrow{k_5}B$, where $k_1=k_2=2\cdot 10^{-2}$, $k_3 = 1\cdot 10^{-2}$, and $k_4=k_5=1\cdot 10^{-4}$.
+	In this example there are four species $A$, $B$, $C$, and $D$ with reactions $A\xrightarrow{k_1}D$, $D\xrightarrow{k_2}A$, $D\xrightarrow{k_3}B$, $A+B\xrightarrow{k_4}C$, and $C\xrightarrow{k_5}D+B$, where $k_1=k_2=2\cdot 10^{-2}$, $k_3 = 1\cdot 10^{-2}$, and $k_4=k_5=1\cdot 10^{-4}$.
 These reactions are inserted into the ansatz library $\Theta (\cdot)$.
@@
 	\end{align}
-	For a selection of values $\alpha\in [0, 100]$, the model parameters $\Xi$ were estimated and the scoring function (\ref{ex2:score-function}) evaluated, yielding a graph as in Figure \ref{ex2:score-graph}.
+	For a selection of values $\alpha\in [0, 100]$, the model parameters $\Xi$ were estimated and the scoring function (\ref{ex2:score-function}) evaluated, yielding the graph that is depicted in Figure \ref{ex2:score-graph}.
 
@@
 	\end{align}
-	When evaluating the minimization problem (\ref{method:minimizationproblem}) with this hyperparameter and applying a cut-off of $10^{-8}$ to the yielded parameters, one obtains
+	When evaluating the minimization problem (\ref{method:minimizationproblem}) with $\hat{\alpha}$ as hyperparameter for the regularizer and applying a cut-off of $10^{-10}$ to the yielded parameters, one obtains the simpler set of reactions $A\xrightarrow{\hat{k}_1}D$, $D\xrightarrow{\hat{k}_3} B$, and $A+B\xrightarrow{\hat{k}_4}C$ with rates $\hat{k}_1 \approx 3.5\cdot 10^{-3}$, $\hat{k}_3\approx 9.1\cdot 10^{-3}$, and $\hat{k}_4\approx 8.7\cdot 10^{-5}$, respectively.
+
+	One can see in Figure \ref{ex2:result} that with this simpler set of reactions the observed concentration curves can be explained as well.
+It can be argued that the second reaction of the generating model $D\xrightarrow{k_2}A$ can be compensated by a smaller rate $\hat{k}_1 < k_1$ and that the fifth reaction of the generating model can be neglected altogether in the observed timescale as its rate is two orders of magnitude smaller than the rates of the other unimolecular reactions.
+	The rates $k_3$ and $k_4$ are approximately recovered by $\hat{k}_3$ and $\hat{k}_4$, respectively.
+
+	\begin{figure}
+		\begin{center}
+			\includegraphics[width=.5\textwidth]{./figures_tex/cv_expected_estimated_concentrations}
+		\end{center}
+		\caption{Estimated and generating concentration curves for the example in Section \ref{sec:ex2}.}
+		\label{ex2:result}
+	\end{figure}
 
 	\section{Conclusion}
-	todo.
+	In this work we have successfully applied and extended the SINDy method to not only parsimoniously detect potentially nonlinear terms in a dynamical system from possibly noisy data, but also pick up on, in this case, a sparse set of generating reactions (\ref{method:the-reactions}) and their respective rates.
+
+	In two examples it was demonstrated that, despite noisy data and unavailability of derivative measurements, a parsimonious generating reaction network that is able to explain the observed data can be recovered.
+	In particular it was shown in the example given in Section \ref{sec:ex1} that if there is no ambiguity in the underlying model and ansatz reaction library, the actual rates can be recovered.
+In the example of Section \ref{sec:ex2} an even simpler model could be obtained by making use of sparse regression and cross-validation.
 

### 2017-11-02 docs/manuscript.tex
wip: manuscript

@@
 	\section{Introduction}
-	When presented with a non-equilibrium time series of possibly noisy concentration fluctuations of some species as output of, e.g., measurements from experiments or simulations that were parameterized by microscopic rates \textcolor{red}{(cite ReaDDy?)}, one can ask for the corresponding macroscopic rates and generating reaction network.
+	When presented with a time series of possibly noisy non-equilibrium concentration fluctuations of some species as output of, e.g., measurements from experiments or simulations that were parameterized by microscopic rates \textcolor{red}{(cite ReaDDy?)}, one can ask for the corresponding macroscopic rates and a generating reaction network.
 	In this paper we present an application of the shallow learning method SINDy \cite{Brunton2015}.
-It is able to identify the generating parsimonious nonlinear dynamics in data that stems from dynamical systems, thus providing an interpretable result.
-	In our application we, opposed to the original method, do not only look for macroscopic rates of net species change but investigate the specific reactions that might have lead to the observations.
+By sparse regression, it is able to identify generating nonlinear dynamics in data that stems from dynamical systems.
+The parsimonious nature of the results avoids overfitting and provides interpretability.
+	In our application we, as opposed to the original method, do not only look for macroscopic rates of net species change but investigate the specific reactions that might have lead to the observations.
 	We demonstrate the algorithm on two toy problems - one problem showing that when there is no ambiguity in the system, one converges to the correct rates with increasing resolution of concentration fluctuations and one in which we compute a sparse reaction network for given data.
@@
 	\section{The method}
-	The underlying model that we want to fit the data to is a law of mass action type dynamical system.
-To this end, let $S$ be the number of species, then the concentration data at a time $t$ can be represented by a vector
+	The underlying model is a law of mass action type dynamical system. To this end, let $S$ be the number of species, then the observed concentration at a time $t$ can be represented by a vector
 	\begin{align}
@@
 	\end{align}
-	so that the change of concentration for species $i$ at time $t_i$, is represented by the dynamical system
+	so that the change of concentration for species $i$ at time $t$, is represented by the dynamical system
 	\begin{align}
@@
 
-	When presented with a time series with $T$ observations, the data can be represented as a matrix
+	When presented with a time series consisting of $T$ observations, the data can be represented as a matrix
 	\begin{align}
@@
 	where $\textbf{X}_i$ denotes the $i$-th row in $X$.
-Applying the concentration trajectory to the library yields a data tensor $\Theta(\textbf{X})\in\mathbb{R}^{T\times S\times R}$.
+Applying the concentration trajectory to the library yields $\Theta(\textbf{X})\in\mathbb{R}^{T\times S\times R}$.
 Following the approach of SINDy, the goal is now to find coefficients $\Xi = \begin{pmatrix} \xi_1 & \xi_2 & \cdots & \xi_R
@@
 	\label{method:minimizationproblem}\end{align}
-	where $\alpha\geq 0$ is a sparsity inducing hyperparameter and $\|\cdot\|_F$ denotes the Frobenius norm.
-For $\alpha=0$ this problem reduces to constrained least-squares.
-For solving (\ref{method:minimizationproblem}) we apply the sequential least squares minimizer SLSQP, originally described in \cite{Kraft1988}, contained in, e.g., the software package SciPy \cite{SciPy}.
-As initial guess for the regulated problem we use the solution of (\ref{method:minimizationproblem}) with $\alpha=0$.
+	where $\|\cdot\|_F$ denotes the Frobenius norm and $\alpha\geq 0$ is a sparsity inducing hyperparameter.
+For $\alpha=0$ the minimization problem reduces to constrained least-squares.
+In order to solve (\ref{method:minimizationproblem}) we apply the sequential least-squares minimizer SLSQP, originally described in \cite{Kraft1988}, by the software package SciPy \cite{SciPy}.
+As initial guess for the regularized problem we use the solution of (\ref{method:minimizationproblem}) with $\alpha=0$.
 
@@
 	For generating time series data of concentrations we use the Gillespie method \cite{Gillespie1976}.
-For a given set of initial conditions, we produce several realizations which then are converted to a trajectory with fixed time step and averaged.
-Since these trajectories are piecewise constant, one has to take special care when approximating the temporal derivative.
-Simply taking finite differences has the effect that, with decreasing time step, the derivative is most of the time constant zero and approaches infinity at the jump discontinuities.
-To counter this effect, we beforehand perform a linear approximation between each two adjacent discontinuities, as depicted in Figure \ref{examples:linear-approximation}.
+For given initial conditions, we produce several realizations which then are converted to a trajectory with fixed time step and averaged.
+Since these trajectories are piecewise constant, one has to take special care when approximating the data's temporal derivative that is needed in (\ref{method:minimizationproblem}).
+Applying finite differences has the effect that, with decreasing time step, the derivative mostly constant zero and approaches infinity at the jump discontinuities.
+To counter this effect, we beforehand perform a linear approximation between each two adjacent discontinuities, as depicted in Figure \ref{examples:linear-approximation} and thus obtain a piecewise constant derivative.
 (\textcolor{red}{Use regularized derivative?})
@@
 	Here we demonstrate that, with decreasing time step $\Delta t$ and $\alpha=0$, the estimated rates converge to the true rates if there is no ambiguity in the system.
-The systems contains two species $A$ and $B$ which can convert into each other with macroscopic rates of $r_1=4.0$ and $r_2=0.5$, respectively.
+Let there be two species $A$ and $B$ which can unimolecularly convert into each other with macroscopic rates of $k_1=4.0$ and $k_2=0.5$, respectively.
 The initial amount of particles is set to $70$ of type $A$ and $0$ of type $B$.
-A possibly resulting change of concentrations over time is depicted in Figure \ref{ex1:concentration-curves}.
+The to this system corresponding law of mass action solution and a Gillespie realization are  depicted in Figure \ref{ex1:concentration-curves}.
 
@@
 		\end{center}
-		\caption{A realization of the example given in Section \ref{sec:ex1}.}
+		\caption{The law of mass action solution and a Gillespie realization of the example given in Section \ref{sec:ex1}.}
 		\label{ex1:concentration-curves}
@@
 
-	For each $\Delta t$, we averaged eight different realizations with identical starting conditions into one single trajectory and estimated the rates for reactions $A\to B$ and $B\to A$.
-This procedure, we repeated $30$ times.
-In Figure \ref{ex1:rates-convergence}, the average estimates alongside with their standard deviations are shown.
-One can see that, with decreasing $\Delta t$, also the estimate for the rates becomes better and they can be recovered.
+	For each $\Delta t$, eight different realizations with identical starting conditions are averaged into a single trajectory, of which then the rates for reactions $A\to B$ and $B\to A$ are estimated.
+This procedure was repeated $30$ times.
+In Figure \ref{ex1:rates-convergence}, the average estimates with their standard deviations are shown.
+One can see that, with decreasing $\Delta t$, the estimates become better and the rates can be recovered.
+Further it can be observed that in this case a time step of $\Delta t\approx 10^{-2}$ suffices.
 
@@
 		\end{center}
-		\caption{Convergence of estimated rates to the ground truth with decreasing $\Delta t$.}
+		\caption{Convergence of estimated rates with decreasing $\Delta t$.}
 		\label{ex1:rates-convergence}
@@
 
-	The problem now is to find a suitable $\alpha$ such that a parsimonious solution is picked.
+	The problem now is to find a suitable $\alpha$ such that a parsimonious solution is yielded.
 To this end, cross-validation is applied.
-The trajectories that were used as training and test data set are depicted in Figure \ref{ex2:test-train-concentration-curves}.
-\textcolor{red}{Perhaps use several test trajectories with different initial conditions to get something like a confidence interval?} As an indicator to the estimation error we use the scoring function
+The trajectories that are used as training and test data set are depicted in Figure \ref{ex2:test-train-concentration-curves}.
+\textcolor{red}{Perhaps use several test trajectories with different initial conditions to get something like a confidence interval?} As an indicator to the estimation error one can use the scoring function
 	\begin{align}
-		f_\alpha = \frac{1}{T}\|\mathbf{\dot{X}}_\text{test} - \Theta (\textbf{X}_\text{test}) \Xi_\text{train}\|_F^2.
+		f_\alpha = \frac{1}{T}\|\mathbf{\dot{X}}_\text{test} - \Theta (\textbf{X}_\text{test}) \Xi_\text{train}\|_F^2,
 		\label{ex2:score-function}
 	\end{align}
-	For a selection of values $\alpha\in [0, 100]$, the model parameters $\Xi$ were estimated and the scoring function (\ref{ex2:score-function}) evaluated, yielding the graph that is depicted in Figure \ref{ex2:score-graph}.
+	where the best possible score for some $f_\alpha = 0$ and larger values indicate worse models.
+	For a selection of hyperparameters $\alpha\in [0, 100]$, the model parameters $\Xi$ are estimated and the scoring function (\ref{ex2:score-function}) evaluated, yielding a dependency of the score on $\alpha$ as depicted in Figure \ref{ex2:score-graph}.
 
@@
 
-	One can see that the minimum is at about
+	One can see that the minimum score is achieved at
 	\begin{align}

### 2017-11-02 docs/manuscript.tex
wip: manuscript

@@
 	One can see in Figure \ref{ex2:result} that with this simpler set of reactions the observed concentration curves can be explained as well.
-It can be argued that the second reaction of the generating model $D\xrightarrow{k_2}A$ can be compensated by a smaller rate $\hat{k}_1 < k_1$ and that the fifth reaction of the generating model can be neglected altogether in the observed timescale as its rate is two orders of magnitude smaller than the rates of the other unimolecular reactions.
-	The rates $k_3$ and $k_4$ are approximately recovered by $\hat{k}_3$ and $\hat{k}_4$, respectively.
+Further, it can be argued that the second reaction of the original model $D\xrightarrow{k_2}A$ is compensated by a smaller rate $\hat{k}_1 < k_1$ and that the fifth reaction of the generating model can be neglected altogether in the observed timescale as its rate is two orders of magnitude smaller than the other unimolecular reactions' rates.
+	The parameters $k_3$ and $k_4$ are approximately recovered by $\hat{k}_3$ and $\hat{k}_4$, respectively.
 
@@
 		\end{center}
-		\caption{Estimated and generating concentration curves for the example in Section \ref{sec:ex2}.}
+		\caption{Estimated and generating concentration curves as well as the corresponding Gillespie realization for the example in Section \ref{sec:ex2}.}
 		\label{ex2:result}
@@
 	\section{Conclusion}
-	In this work we have successfully applied and extended the SINDy method to not only parsimoniously detect potentially nonlinear terms in a dynamical system from possibly noisy data, but also pick up on, in this case, a sparse set of generating reactions (\ref{method:the-reactions}) and their respective rates.
+	In this work we have successfully applied and extended the SINDy method to not only parsimoniously detect potentially nonlinear terms in a dynamical system from possibly noisy data, but also yield, in this case, a sparse set of rates with respect to generating reactions (\ref{method:the-reactions}).
 
-	In two examples it was demonstrated that, despite noisy data and unavailability of derivative measurements, a parsimonious generating reaction network that is able to explain the observed data can be recovered.
-	In particular it was shown in the example given in Section \ref{sec:ex1} that if there is no ambiguity in the underlying model and ansatz reaction library, the actual rates can be recovered.
-In the example of Section \ref{sec:ex2} an even simpler model could be obtained by making use of sparse regression and cross-validation.
+	In two examples it was demonstrated that, despite noisy data and unavailability of derivative measurements, a parsimonious generating reaction network that is qualitatively able to explain the observed data can be estimated.
+	In particular it was shown in the first example that if there is no ambiguity in the underlying model and ansatz reaction library, the actual rates can be recovered with decreasing time step, i.e., increasing resolution of the jump process.
+	In the second example we could obtain an even simpler model than what was used to generate data by making use of sparse regression and cross-validation.
 

### 2017-11-02 docs/manuscript.tex
wip: manuscript

@@
 The initial amount of particles is set to $70$ of type $A$ and $0$ of type $B$.
-The to this system corresponding law of mass action solution and a Gillespie realization are  depicted in Figure \ref{ex1:concentration-curves}.
+The law of mass action solution and Gillespie realization corresponding to this system are depicted in Figure \ref{ex1:concentration-curves}.
 
@@
 	\label{sec:ex2}
-	In this example there are four species $A$, $B$, $C$, and $D$ with reactions $A\xrightarrow{k_1}D$, $D\xrightarrow{k_2}A$, $D\xrightarrow{k_3}B$, $A+B\xrightarrow{k_4}C$, and $C\xrightarrow{k_5}D+B$, where $k_1=k_2=2\cdot 10^{-2}$, $k_3 = 1\cdot 10^{-2}$, and $k_4=k_5=1\cdot 10^{-4}$.
+	In this example there are four species $A$, $B$, $C$, and $D$.
+The system was generated with reactions $A\xrightarrow{k_1}D$, $D\xrightarrow{k_2}A$, $D\xrightarrow{k_3}B$, $A+B\xrightarrow{k_4}C$, and $C\xrightarrow{k_5}D+B$, where the rates are given in Table \ref{tab:estimated-rates}.
 These reactions are inserted into the ansatz library $\Theta (\cdot)$.
@@
 	\end{align}
-	where the best possible score for some $f_\alpha = 0$ and larger values indicate worse models.
+	where the best possible score is $f_\alpha = 0$ and larger values indicate worse models.
 	For a selection of hyperparameters $\alpha\in [0, 100]$, the model parameters $\Xi$ are estimated and the scoring function (\ref{ex2:score-function}) evaluated, yielding a dependency of the score on $\alpha$ as depicted in Figure \ref{ex2:score-graph}.
@@
 	\end{align}
-	When evaluating the minimization problem (\ref{method:minimizationproblem}) with $\hat{\alpha}$ as hyperparameter for the regularizer and applying a cut-off of $10^{-10}$ to the yielded parameters, one obtains the simpler set of reactions $A\xrightarrow{\hat{k}_1}D$, $D\xrightarrow{\hat{k}_3} B$, and $A+B\xrightarrow{\hat{k}_4}C$ with rates $\hat{k}_1 \approx 3.5\cdot 10^{-3}$, $\hat{k}_3\approx 9.1\cdot 10^{-3}$, and $\hat{k}_4\approx 8.7\cdot 10^{-5}$, respectively.
+	When evaluating the minimization problem (\ref{method:minimizationproblem}) with $\hat{\alpha}$ as hyperparameter for the regularizer and applying a cut-off of $10^{-10}$ to the yielded parameters, one obtains the simpler set of reactions $A\xrightarrow{\hat{k}_1}D$, $D\xrightarrow{\hat{k}_3} B$, and $A+B\xrightarrow{\hat{k}_4}C$ with rates as in Table \ref{tab:estimated-rates}.
+
+	\begin{table}
+		\centering
+		\begin{tabular}{ccc}
+			Reaction & original rate & estimated rate \\ \hline\noalign{\vskip .1cm}
+			$A \to D$ & $k_1 = 2\cdot 10^{-2}$ & $\hat{k}_1 \approx 3.5\cdot 10^{-3}$\\
+			$D \to A$ & $k_2 = 2\cdot 10^{-2}$ & $\hat{k}_2 = 0$ \\
+			$D\to B$ & $k_3 = 1\cdot 10^{-2}$ & $\hat{k}_3\approx 9.1 \cdot 10^{-3}$ \\
+			$A+B\to C$ & $k_4 = 1\cdot 10^{-4}$ & $\hat{k}_4\approx 8.7 \cdot 10^{-5}$ \\
+			$C \to D+B$ & $k_5 = 1\cdot 10^{-4}$ & $\hat{k}_5 = 0$ 
+		\end{tabular}
+		\caption{Original and estimated reaction rates for the example in Section \ref{sec:ex2}.}
+		\label{tab:estimated-rates}
+	\end{table}
 
@@
 		\end{center}
-		\caption{Estimated and generating concentration curves as well as the corresponding Gillespie realization for the example in Section \ref{sec:ex2}.}
+		\caption{Integrated law of mass action equations (\ref{method:the-system}) for the sparse estimated parameters and the corresponding Gillespie realization for the example in Section \ref{sec:ex2}.}
 		\label{ex2:result}

### 2017-11-02 docs/manuscript.tex
wip

@@
 	\section{Conclusion}
-	In this work we have successfully applied and extended the SINDy method to not only parsimoniously detect potentially nonlinear terms in a dynamical system from possibly noisy data, but also yield, in this case, a sparse set of rates with respect to generating reactions (\ref{method:the-reactions}).
+	In this work we have successfully applied and extended the SINDy method to not only parsimoniously detect potentially nonlinear terms in a dynamical system from noisy data, but also yield, in this case, a sparse set of rates with respect to generating reactions (\ref{method:the-reactions}).
 
-	In two examples it was demonstrated that, despite noisy data and unavailability of derivative measurements, a parsimonious generating reaction network that is qualitatively able to explain the observed data can be estimated.
+	In two examples it was demonstrated that despite noisy data and unavailable derivative measurements, a parsimonious generating reaction network that is qualitatively able to explain the observed data can be estimated.
 	In particular it was shown in the first example that if there is no ambiguity in the underlying model and ansatz reaction library, the actual rates can be recovered with decreasing time step, i.e., increasing resolution of the jump process.

### 2018-01-22 cpp/include/lasso_minimizer_objective_fun.h
reg network

@@ -48,14 +48,15 @@ void least_squares_function(input_array &result, const input_array &propensities
                             const input_array &dX);
 
 input_array elastic_net_objective_function_jac(const input_array &propensities,
-                                               const double alpha, const double l1_ratio, const input_array &theta,
-                                               const input_array &dX);
+                                               const double alpha, const double l1_ratio,
+                                               const std::vector<input_array> &theta,
+                                               const std::vector<input_array> &dX);
 
 double elastic_net_objective_function(const input_array &propensities, const double alpha, const double l1_ratio,
-                                      const input_array &theta, const input_array &dX);
+                                      const std::vector<input_array> &theta, const std::vector<input_array> &dX);
 
 double lasso_cost_fun(const input_array &propensities, const double alpha,
-                      const input_array &theta, const input_array &dX);
+                      const std::vector<input_array> &theta, const std::vector<input_array> &dX);
 
 
 }

### 2018-01-22 cpp/src/lasso_minimizer_objective_fun.cpp

@@ -104,43 +104,66 @@ void least_squares_function(input_array &result, const input_array &propensities
 }
 
 input_array elastic_net_objective_function_jac(const input_array &propensities,
-                                               const double alpha, const double l1_ratio, const input_array &theta,
-                                               const input_array &dX) {
-    input_array result;
-    const auto n_timesteps = static_cast<std::size_t>(theta.shape()[0]);
-    const auto n_reactions = static_cast<std::size_t>(theta.shape()[1]);
-    const auto n_species = static_cast<std::size_t>(theta.shape()[2]);
+                                               const double alpha, const double l1_ratio,
+                                               const std::vector<input_array> &thetas,
+                                               const std::vector<input_array> &dXs) {
 
-    {
-        std::vector<std::size_t> shape;
-        shape.push_back(static_cast<std::size_t>(n_reactions));
-        result.resize(shape);
+    if(thetas.size() != dXs.size()) {
+        throw std::invalid_argument("n thetas and n dXs must match.");
     }
-    // auto data = result.mutable_data();
+    if(thetas.empty()) {
+        throw std::invalid_argument("At least one set of theta<->dX required.");
+    }
 
-    // for each element in the 1xr jacobi matrix...
-    for(std::size_t i = 0; i < n_reactions; ++i) {
-        // first set it to 0
-        result.mutable_at(i) = 0;
-        // then calculate the frobenius part
-        for (std::size_t t = 0; t < n_timesteps; ++t) {
-            for (std::size_t s = 0; s < n_species; ++s) {
-                auto x = dX.at(t, s);
-                auto theta_t_i_s = theta.at(t, i, s);
-                for(std::size_t r = 0; r < n_reactions; ++r) {
-                    x -= propensities.at(r) * theta.at(t, r, s);
+    input_array result;
+    const auto n_reactions = static_cast<std::size_t>(thetas.at(0).shape()[1]);
+    const auto n_species = static_cast<std::size_t>(thetas.at(0).shape()[2]);
+
+    std::vector<std::size_t> shape;
+    shape.push_back(static_cast<std::size_t>(n_reactions));
+    result.resize(shape);
+
+    input_array update;
+    update.resize(shape);
+
+    auto itThetas = thetas.begin();
+    auto itDXs = dXs.begin();
+    for(; itThetas != thetas.end(); ++itThetas, ++itDXs) {
+        const auto &theta = *itThetas;
[…]

### 2018-01-22 moritz/multiple_trajs.py

@@ -1,0 +1,54 @@
+import os
+import numpy as np
+import matplotlib as mpl
+
+import readdy_learn.analyze.analyze as ana
+import readdy_learn.analyze.basis as basis
+
+import pynumtools.kmc as kmc
+
+#mpl.rcParams['figure.figsize'] = (13, 13)
+import matplotlib.pyplot as plt
+import scipy.signal as ss
+from readdy_learn.example.regulation_network import RegulationNetwork
+
+regulation_network = RegulationNetwork()
+
+def plot_overview(plot_functions, n_rows=2, n_cols=2, size_factor=1.):
+    plt.figure(1, figsize=(5*n_cols*size_factor,3.5*n_rows*size_factor))
+    n_plots = n_rows * n_cols
+    idx = 1
+    for pf in plot_functions:
+        if idx > n_plots:
+            break
+        plt.subplot(n_rows, n_cols, idx)
+        pf()
+        plt.legend(loc="best")
+        idx += 1
+
+    #plt.subplots_adjust(top=0.88, bottom=0.08, left=0.10, right=0.95, hspace=0.4, wspace=0.35)
+    plt.tight_layout(pad=0.6, w_pad=2.0, h_pad=2.0)
+
+def plot_and_persist_lma_traj(t):
+    plt.plot(t.counts[:, 0], label=regulation_network.species_names[0])
+    plt.plot(t.counts[:, 1], label=regulation_network.species_names[1])
+    plt.plot(t.counts[:, 2], label=regulation_network.species_names[2])
+    plt.plot(t.counts[:, 3], label=regulation_network.species_names[3])
+    plt.plot(t.counts[:, 4], label=regulation_network.species_names[4])
+    plt.plot(t.counts[:, 5], label=regulation_network.species_names[5])
+    plt.plot(t.counts[:, 6], label=regulation_network.species_names[6])
+    plt.plot(t.counts[:, 7], label=regulation_network.species_names[7])
+    plt.plot(t.counts[:, 8], label=regulation_network.species_names[8])
+    plt.legend(loc="best")
+    plt.show()
+    t.persist()
+
+
+if __name__ == '__main__':
+    analysis = regulation_network.generate_analysis_object()
+    for i in range(4):
+        print("traj {}".format(i))
+        t = analysis.generate_or_load_traj_lma(i, regulation_network.target_time,
+                                               noise_variance=regulation_network.noise_variance,
+                                               realizations=regulation_network.realisations)
+        plot_and_persist_lma_traj(t)

### 2018-01-22 readdy_learn/analyze/analyze.py

@@ -29,7 +29,7 @@ def split(a, n):
     return (a[i * k + min(i, m):(i + 1) * k + min(i + 1, m)] for i in range(n))
 
 
-def obtain_derivative(traj, desired_n_counts=6000, alpha=1000, atol=1e-10, tol=1e-10, maxit=1000, alpha_search_depth=5,
+def obtain_derivative(traj, alpha=1000, atol=1e-10, tol=1e-10, maxit=1000, alpha_search_depth=5,
                       interp_degree='regularized_derivative', variance=None, verbose=False, x0=None,
                       species=None, override=False, subdivisions=None, reuse_deriv=True, solver='spsolve'):
     if species is None:
@@ -41,12 +41,8 @@ def obtain_derivative(traj, desired_n_counts=6000, alpha=1000, atol=1e-10, tol=1
             interp_degree = traj.interpolation_degree
             traj.interpolation_degree = None
             traj.update()
-            stride = max(traj.counts.shape[0] // desired_n_counts, 1)
-            strided_times = traj.times[::stride]
-            strided_counts = traj.counts[::stride, :]
-            strided_dt = strided_times[1] - strided_times[0]
-            print("got {} counts (and {} corresp. time steps), dt=".format(strided_counts.shape[0], len(strided_times)),
-                  strided_dt)
+            print("got {} counts (and {} corresp. time steps), dt=".format(traj.counts.shape[0], len(traj.times)),
+                  traj.time_step)
 
             x0 = np.asarray(x0).squeeze()
             used_alphas = []
@@ -54,29 +50,29 @@ def obtain_derivative(traj, desired_n_counts=6000, alpha=1000, atol=1e-10, tol=1
                 if int(s) in traj.separate_derivs.keys() and not override:
                     print("skipping species {} as it already has derivatives".format(s))
                     continue
-                ys = strided_counts[:, s]
+                ys = traj.counts[:, s]
                 kw = {'maxit': maxit, 'linalg_solver_maxit': 500000, 'tol': tol, 'atol': atol, 'rtol': None,
                       'solver': solver, 'verbose': False, 'show_progress': verbose}
                 if isinstance(alpha, np.ndarray):
                     if len(alpha) > 1:
                         if subdivisions is None:
-                            best_alpha, ld, scores = deriv.best_tv_derivative(ys, strided_times, alpha,
+                            best_alpha, ld, scores = deriv.best_tv_derivative(ys, traj.times, alpha,
                                                                               n_iters=alpha_search_depth,
[…]

### 2018-01-22 readdy_learn/analyze/basis.py

@@ -101,16 +101,20 @@ class BasisFunctionConfiguration(object):
 
     def add_conversion(self, type1, type2):
         self._basis_functions.append(ConversionReaction(type1, type2, self._n_species))
+        return self
 
     def add_fusion(self, type_from1, type_from2, type_to):
         self._basis_functions.append(FusionReaction(type_from1, type_from2, type_to, self._n_species))
+        return self
 
     def add_fission(self, type_from, type_to1, type_to2):
         self._basis_functions.append(FissionReaction(type_from, type_to1, type_to2, self._n_species))
+        return self
 
     def add_intercept(self, type):
         self._basis_functions.append(Intercept(type, self._n_species))
-
+        return self
 
     def add_decay(self, type_from):
         self._basis_functions.append(DecayReaction(type_from, self._n_species))
+        return self

### 2018-01-22 readdy_learn/analyze/estimator.py

@@ -143,12 +143,6 @@ class ReaDDyElasticNetEstimator(BaseEstimator):
 
         data, expected = self._get_slice(traj_range)
 
-        if self.rescale:
-            # confine values in data to [0, 100]
-            xmax = np.max(data)
-            data /= xmax / 100.
-            expected /= xmax / 100.
-
         large_theta = self.get_theta(data)
 
         if self.constrained:
@@ -167,7 +161,6 @@ class ReaDDyElasticNetEstimator(BaseEstimator):
 
         def objective(x):
             obj = opt.elastic_net_objective_fun(x, self.alpha, self.l1_ratio, large_theta, expected)
-            # print("got {}".format(obj))
             return obj
 
         result = so.minimize(

### 2018-01-22 readdy_learn/analyze/generate.py

@@ -4,14 +4,49 @@ from pathos.multiprocessing import Pool as _Pool
 import readdy_learn.analyze.progress as _pr
 
 
-def generate_continuous_counts(rates, initial_condition, bfc, timestep, n_steps):
-    def fun_reference(data, _):
-        theta = _np.array([f(data) for f in bfc.functions])
-        return _np.matmul(rates, theta)
+def generate_continuous_counts(rates, initial_condition, bfc, timestep, n_steps, noise_variance=0., n_realizations=1,
+                               njobs=8):
+    if n_realizations == 1 or noise_variance == 0.:
+        def fun_reference(data, _):
+            theta = _np.array([f(data) for f in bfc.functions])
+            return _np.matmul(rates, theta)
 
-    xs = _np.linspace(0, n_steps * timestep, num=n_steps, endpoint=False)
-    initial_condition = _np.array(initial_condition).squeeze()
-    return xs, _odeint(fun_reference, initial_condition, xs)
+        xs = _np.linspace(0, n_steps * timestep, num=n_steps, endpoint=False)
+        initial_condition = _np.array(initial_condition).squeeze()
+        ys = _np.array(_odeint(fun_reference, initial_condition, xs)).squeeze()
+        if noise_variance > 0.:
+            ys += _np.random.normal(0.0, _np.sqrt(noise_variance), size=ys.shape)
+        return xs, ys
+    else:
+        def fun_reference(data, _):
+            theta = _np.array([f(data) for f in bfc.functions])
+            return _np.matmul(rates, theta)
+
+        xs = _np.linspace(0, n_steps * timestep, num=n_steps, endpoint=False)
+        initial_condition = _np.array(initial_condition).squeeze()
+
+        def generate_wrapper(args):
+            ys = _np.array(_odeint(fun_reference, initial_condition, xs)).squeeze()
+            if noise_variance > 0.:
+                _np.random.seed()
+                ys += _np.random.normal(0.0, _np.sqrt(noise_variance), size=ys.shape)
+            return ys
+
+        avgcounts = None
+        progress = _pr.Progress(n=n_realizations, label="generate averaged lma counts with additive noise")
+        N = 0.
+
+        params = [() for _ in range(n_realizations)]
+        with _Pool(processes=njobs) as p:
+            for counts in p.imap(generate_wrapper, params, 1):
+                progress.increase(1)
+                N += 1.
+                if avgcounts is None:
+                    avgcounts = counts
[…]

### 2018-01-22 readdy_learn/example/regulation_network.py

@@ -1,0 +1,215 @@
+
+import numpy as _np
+import pynumtools.kmc as kmc
+import readdy_learn.analyze.basis as basis
+
+class RegulationNetwork(object):
+
+    def __init__(self):
+        # species DA  MA  A  DB  MB  B  DC  MC  C
+        # ids     0   1   2  3   4   5  6   7   8
+        self.n_species = 9
+        self.species_names = ["DA", "MA", "A", "DB", "MB", "B", "DC", "MC", "C"]
+        self.desired_rates = _np.array([
+            1.8,  # DA -> DA + MA, transcription A
+            2.1,  # MA -> MA + A, translation A
+            1.3,  # MA -> 0, decay
+            1.5,  # A -> 0, decay
+            2.2,  # DB -> DB + MB, transcription B
+            2.0,  # MB -> MB + B, translation B
+            2.0,  # MB -> 0, decay
+            2.5,  # B -> 0, decay
+            3.2,  # DC -> DC + MC, transcription C
+            3.0,  # MC -> MC + C, translation C
+            2.3,  # MC -> 0, decay
+            2.5,  # C -> 0, decay
+            # self regulation
+            0.,  # MA + A -> A, A regulates A
+            0.,  # MB + B -> B, B regulates B
+            0.,  # MC + C -> C, C regulates C
+            # cyclic forward
+            0.,  # MB + A -> A, A regulates B
+            0.,  # MC + B -> B, B regulates C
+            0.,  # MA + C -> C, C regulates A
+            # cyclic backward
+            6.,  # MC + A -> A, A regulates C
+            4.,  # MB + C -> C, C regulates B
+            3.,  # MA + B -> B, B regulates A
+            # nonsense reactions, mRNA eats protein self
+            0., 0., 0.,
+            # nonsense reactions, mRNA eats protein cyclic forward
+            0., 0., 0.,
+            # nonsense reactions, mRNA eats protein  cyclic backward
+            0., 0., 0.,
+            # nonsense reactions, protein eats protein self
+            0., 0., 0.,
+            # nonsense reactions, protein eats protein cyclic forward
+            0., 0., 0.,
+            # nonsense reactions, protein eats protein cyclic backward
[…]

### 2018-01-22 setup.py

@@ -67,5 +67,5 @@ setup(
     ext_modules=ext_modules,
     packages=find_packages(),
     zip_safe=False,
-    install_requires=['numpy', 'scipy', 'pathos', 'scikit-learn', 'h5py', 'matplotlib']
+    install_requires=['numpy', 'scipy', 'pathos', 'scikit-learn', 'h5py', 'matplotlib', 'progress_reporter']
 )

### 2018-01-22 cpp/include/lasso_minimizer_objective_fun.h
reg network

@@ -34,6 +34,8 @@
 
 #include <pybind11/pybind11.h>
 #include <pybind11/numpy.h>
+#include <pybind11/stl.h>
+#include <pybind11/functional.h>
 
 namespace analyze_tools {
 namespace opt {

### 2018-01-22 cpp/src/lasso_minimizer_objective_fun.cpp

@@ -1,5 +1,8 @@
 #include "lasso_minimizer_objective_fun.h"
 
+#include <pybind11/stl.h>
+#include <pybind11/functional.h>
+
 #include <iostream>
 #include <vector>
 

### 2018-01-22 moritz/multiple_trajs.py

@@ -14,36 +14,6 @@ from readdy_learn.example.regulation_network import RegulationNetwork
 
 regulation_network = RegulationNetwork()
 
-def plot_overview(plot_functions, n_rows=2, n_cols=2, size_factor=1.):
-    plt.figure(1, figsize=(5*n_cols*size_factor,3.5*n_rows*size_factor))
-    n_plots = n_rows * n_cols
-    idx = 1
-    for pf in plot_functions:
-        if idx > n_plots:
-            break
-        plt.subplot(n_rows, n_cols, idx)
-        pf()
-        plt.legend(loc="best")
-        idx += 1
-
-    #plt.subplots_adjust(top=0.88, bottom=0.08, left=0.10, right=0.95, hspace=0.4, wspace=0.35)
-    plt.tight_layout(pad=0.6, w_pad=2.0, h_pad=2.0)
-
-def plot_and_persist_lma_traj(t):
-    plt.plot(t.counts[:, 0], label=regulation_network.species_names[0])
-    plt.plot(t.counts[:, 1], label=regulation_network.species_names[1])
-    plt.plot(t.counts[:, 2], label=regulation_network.species_names[2])
-    plt.plot(t.counts[:, 3], label=regulation_network.species_names[3])
-    plt.plot(t.counts[:, 4], label=regulation_network.species_names[4])
-    plt.plot(t.counts[:, 5], label=regulation_network.species_names[5])
-    plt.plot(t.counts[:, 6], label=regulation_network.species_names[6])
-    plt.plot(t.counts[:, 7], label=regulation_network.species_names[7])
-    plt.plot(t.counts[:, 8], label=regulation_network.species_names[8])
-    plt.legend(loc="best")
-    plt.show()
-    t.persist()
-
-
 if __name__ == '__main__':
     analysis = regulation_network.generate_analysis_object()
     for i in range(4):
@@ -51,4 +21,11 @@ if __name__ == '__main__':
         t = analysis.generate_or_load_traj_lma(i, regulation_network.target_time,
                                                noise_variance=regulation_network.noise_variance,
                                                realizations=regulation_network.realisations)
-        plot_and_persist_lma_traj(t)
+        t.persist()
+        # analysis.plot_and_persist_lma_traj(t)
+        regulation_network.compute_gradient_derivatives(analysis)
+        # regulation_network.plot_concentrations(analysis, i)
[…]

### 2018-01-22 readdy_learn/analyze/analyze.py

@@ -127,7 +127,7 @@ def obtain_derivative(traj, alpha=1000, atol=1e-10, tol=1e-10, maxit=1000, alpha
 class ReactionAnalysis(object):
     def __init__(self, bfc, desired_rates, initial_states, set_up_system, recompute=False, recompute_traj=False,
                  fname_prefix="", fname_postfix="", n_species=4, timestep=5e-4,
-                 interp_degree='regularized_derivative', ld_derivative_config=None):
+                 interp_degree='regularized_derivative', ld_derivative_config=None, species_names=None):
         if initial_states is None or len(initial_states) <= 1:
             raise ValueError("needs at least two initial states!")
         if ld_derivative_config is None:
@@ -157,6 +157,11 @@ class ReactionAnalysis(object):
         self._trajs = []
         self._initial_states = initial_states
         self._best_alphas = {}
+        self._species_names = species_names
+
+    @property
+    def species_names(self):
+        return self._species_names
 
     @property
     def ld_derivative_config(self):
@@ -276,7 +281,7 @@ class ReactionAnalysis(object):
             self._trajs.append(None)
         self._trajs[n] = traj
 
-    def plot_cv_results(cv, mainscore=0):
+    def plot_cv_results(self, cv, mainscore=0):
         xs = {}
         ys = {}
         allys = {}
@@ -329,6 +334,35 @@ class ReactionAnalysis(object):
         ax.set_xticklabels(["{}".format(i) for i in ind])
         plt.show()
 
+    def plot_overview(self, plot_functions, n_rows=2, n_cols=2, size_factor=1.):
+        plt.figure(1, figsize=(5 * n_cols * size_factor, 3.5 * n_rows * size_factor))
+        n_plots = n_rows * n_cols
+        idx = 1
+        for pf in plot_functions:
+            if idx > n_plots:
+                break
+            plt.subplot(n_rows, n_cols, idx)
+            pf()
+            plt.legend(loc="best")
+            idx += 1
+
+        # plt.subplots_adjust(top=0.88, bottom=0.08, left=0.10, right=0.95, hspace=0.4, wspace=0.35)
+        plt.tight_layout(pad=0.6, w_pad=2.0, h_pad=2.0)
+
+    def plot_and_persist_lma_traj(self, t):
+        plt.plot(t.counts[:, 0], label=self.species_names[0])
+        plt.plot(t.counts[:, 1], label=self.species_names[1])
[…]

### 2018-01-22 readdy_learn/analyze/estimator.py

@@ -85,10 +85,8 @@ class ReaDDyElasticNetEstimator(BaseEstimator):
                 data = self.trajs[0].counts[traj_range]
                 expected = self.trajs[0].dcounts_dt[traj_range]
         else:
-            data = self.trajs[0].counts
-            expected = self.trajs[0].dcounts_dt
-        data = np.ascontiguousarray(data)
-        expected = np.ascontiguousarray(expected)
+            data = [np.ascontiguousarray(t.counts) for t in self.trajs]
+            expected = [np.ascontiguousarray(t.dcounts_dt) for t in self.trajs]
         return data, expected
 
     def fit(self, X, y=None):
@@ -101,9 +99,17 @@ class ReaDDyElasticNetEstimator(BaseEstimator):
         return self.fit_trajs(X)
 
     def get_theta(self, data):
-        large_theta = np.array([f(data) for f in self.basis_function_configuration.functions])
-        large_theta = np.ascontiguousarray(np.transpose(large_theta, axes=(1, 0, 2)))
-        return large_theta
+        if isinstance(data, list):
+            thetas = []
+            for d in data:
+                large_theta = np.array([f(d) for f in self.basis_function_configuration.functions])
+                large_theta = np.ascontiguousarray(np.transpose(large_theta, axes=(1, 0, 2)))
+                thetas.append(large_theta)
+            return thetas
+        else:
+            large_theta = np.array([f(data) for f in self.basis_function_configuration.functions])
+            large_theta = np.ascontiguousarray(np.transpose(large_theta, axes=(1, 0, 2)))
+            return large_theta
 
     def get_analytical_jac(self):
         data, expected = self._get_slice(None)
@@ -199,13 +205,15 @@ class CrossTrajSplit(object):
 
 
 class CV(object):
-    def __init__(self, traj, bfc, alphas, l1_ratios, n_splits, init_xi, n_jobs=8, show_progress=True,
-                 mode='k_fold', verbose=False, method='SLSQP', test_traj=None, maxiter=300000, rescale=True, tol=1e-16):
+    def __init__(self, trajs, bfc, alphas, l1_ratios, n_splits, init_xi, n_jobs=8, show_progress=True,
+                 mode='k_fold', verbose=False, method='SLSQP', maxiter=300000, rescale=True, tol=1e-16):
         self.alphas = alphas
         self.l1_ratios = l1_ratios
         self.n_jobs = n_jobs
         self.n_splits = n_splits
[…]

### 2018-01-22 readdy_learn/example/regulation_network.py

@@ -2,6 +2,9 @@
 import numpy as _np
 import pynumtools.kmc as kmc
 import readdy_learn.analyze.basis as basis
+import matplotlib as mpl
+#mpl.rcParams['figure.figsize'] = (13, 13)
+import matplotlib.pyplot as plt
 
 class RegulationNetwork(object):
 
@@ -69,12 +72,12 @@ class RegulationNetwork(object):
                 'ld_derivative_use_preconditioner': False
             }
 
-        self.noise_variance = 0.001
+        self.noise_variance = 0.0001
         self.target_time = 3.
         self.train_n = 1
         self.test_n = 2
         self.traj_n_s = [self.train_n, self.test_n]
-        self.realisations = 1000
+        self.realisations = 60
         self.timestep = 1e-3
 
     def set_up_system(self, init_state):
@@ -206,10 +209,45 @@ class RegulationNetwork(object):
             "got {} basis functions but only {} desired rates".format(bfc.n_basis_functions, len(self.desired_rates))
         return bfc
 
+    def plot_concentrations(self, analysis, traj_n):
+        traj_number = traj_n
+        pfs = [
+            lambda t=traj_number: analysis.plot_derivatives(t, species=[0]),
+            lambda t=traj_number: analysis.plot_concentration_curves(t, species=[0]),
+            lambda t=traj_number: analysis.plot_derivatives(t, species=[3]),
+            lambda t=traj_number: analysis.plot_concentration_curves(t, species=[3]),
+            lambda t=traj_number: analysis.plot_derivatives(t, species=[6]),
+            lambda t=traj_number: analysis.plot_concentration_curves(t, species=[6]),
+            lambda t=traj_number: analysis.plot_derivatives(t, species=[1, 2]),
+            lambda t=traj_number: analysis.plot_concentration_curves(t, species=[1, 2]),
+            lambda t=traj_number: analysis.plot_derivatives(t, species=[4, 5]),
+            lambda t=traj_number: analysis.plot_concentration_curves(t, species=[4, 5]),
+            lambda t=traj_number: analysis.plot_derivatives(t, species=[7, 8]),
+            lambda t=traj_number: analysis.plot_concentration_curves(t, species=[7, 8])
+        ]
+        analysis.plot_overview(pfs, n_cols=2, n_rows=6, size_factor=1.3)
+        plt.show()
+
     def generate_analysis_object(self, fname_prefix="regulation_network", fname_postfix=""):
         import readdy_learn.analyze.analyze as ana
[…]

### 2018-01-22 readdy_learn/analyze/estimator.py
reg network

@@ -258,22 +258,6 @@ class CV(object):
                 print("status %s: %s" % (estimator.result_.status, estimator.result_.message))
                 print("%s / %s iterations" % (estimator.result_.nit, self.maxiter))
 
-        # trajs = [self.traj] + list(self.test_traj)
-        # for train_idx, test_idx in splitter.split(trajs):
-        #     estimator = ReaDDyElasticNetEstimator([trajs[ix] for ix in train_idx], self.bfc, -1, alpha=alpha,
-        #                                           l1_ratio=l1_ratio, init_xi=self.init_xi, verbose=self.verbose,
-        #                                           method=self.method)
-        #     # fit the whole thing
-        #     estimator.fit(None)
-        #     if estimator.success_:
-        #         for idx in test_idx:
-        #             testimator = ReaDDyElasticNetEstimator(trajs[idx], self.bfc, -1, alpha=alpha,
-        #                                                    l1_ratio=l1_ratio, init_xi=self.init_xi, verbose=self.verbose,
-        #                                                    method=self.method)
-        #             testimator.coefficients_ = estimator.coefficients_
-        #             scores.append(
-        #                 testimator.score(range(0, trajs[idx].n_time_steps), trajs[idx].dcounts_dt)
-        #             )
         return {'scores': scores, 'alpha': alpha, 'l1_ratio': l1_ratio}
 
     def fit_cross_trajs(self):

### 2018-01-22 readdy_learn/analyze/analyze.py
f

@@ -715,7 +715,7 @@ class ReactionAnalysis(object):
             stride = max(len(traj.times) // n_points, 1)
         estimator = suite.get_estimator(verbose=True, interp_degree=self._interp_degree)
         data, expected = estimator._get_slice(None)
-        theta = estimator.get_theta(data)
+        theta = estimator.get_theta(data[traj_n])
         theta = np.transpose(theta, axes=(0, 2, 1))
         dx = theta.dot(self._desired_rates)
 

### 2018-01-22 readdy_learn/analyze/estimator.py

@@ -99,7 +99,7 @@ class ReaDDyElasticNetEstimator(BaseEstimator):
         return self.fit_trajs(X)
 
     def get_theta(self, data):
-        if isinstance(data, list):
+        if not isinstance(data, np.ndarray):
             thetas = []
             for d in data:
                 large_theta = np.array([f(d) for f in self.basis_function_configuration.functions])

### 2018-01-22 readdy_learn/analyze/generate.py
f

@@ -28,7 +28,7 @@ def generate_continuous_counts(rates, initial_condition, bfc, timestep, n_steps,
         def generate_wrapper(args):
             ys = _np.array(_odeint(fun_reference, initial_condition, xs)).squeeze()
             if noise_variance > 0.:
-                _np.random.seed()
+                _np.random.seed(*args)
                 ys += _np.random.normal(0.0, _np.sqrt(noise_variance), size=ys.shape)
             return ys
 
@@ -36,7 +36,7 @@ def generate_continuous_counts(rates, initial_condition, bfc, timestep, n_steps,
         progress = _pr.Progress(n=n_realizations, label="generate averaged lma counts with additive noise")
         N = 0.
 
-        params = [() for _ in range(n_realizations)]
+        params = [(i,) for i in range(n_realizations)]
         with _Pool(processes=njobs) as p:
             for counts in p.imap(generate_wrapper, params, 1):
                 progress.increase(1)

### 2018-01-22 readdy_learn/analyze/analyze.py
f

@@ -704,18 +704,17 @@ class ReactionAnalysis(object):
     def plot_derivatives(self, traj_n, n_points=None, species=None):
         if species is None:
             species = [i for i in range(self.n_species)]
-        traj = self._trajs[traj_n]
-        if isinstance(traj, str):
-            traj = tools.Trajectory(traj, self.timestep, interpolation_degree=self.interp_degree, verbose=False)
+        traj = self.get_traj(traj_n)
         init = self._initial_states[traj_n]
         system = self._set_up_system(init)
+
         suite = sample_tools.Suite.from_trajectory(traj, system, self._bfc, interp_degree=self._interp_degree,
                                                    tol=1e-12, init_xi=np.ones_like(self._desired_rates) * .0)
         if n_points is not None:
             stride = max(len(traj.times) // n_points, 1)
         estimator = suite.get_estimator(verbose=True, interp_degree=self._interp_degree)
         data, expected = estimator._get_slice(None)
-        theta = estimator.get_theta(data[traj_n])
+        theta = estimator.get_theta(data)
         theta = np.transpose(theta, axes=(0, 2, 1))
         dx = theta.dot(self._desired_rates)
 

### 2018-01-22 readdy_learn/analyze/analyze.py
f

@@ -714,7 +714,7 @@ class ReactionAnalysis(object):
             stride = max(len(traj.times) // n_points, 1)
         estimator = suite.get_estimator(verbose=True, interp_degree=self._interp_degree)
         data, expected = estimator._get_slice(None)
-        theta = estimator.get_theta(data)
+        theta = estimator.get_theta(data[0])
         theta = np.transpose(theta, axes=(0, 2, 1))
         dx = theta.dot(self._desired_rates)
 

### 2018-01-22 moritz/multiple_trajs.py
f

@@ -7,7 +7,7 @@ import readdy_learn.analyze.basis as basis
 
 import pynumtools.kmc as kmc
 
-#mpl.rcParams['figure.figsize'] = (13, 13)
+# mpl.rcParams['figure.figsize'] = (13, 13)
 import matplotlib.pyplot as plt
 import scipy.signal as ss
 from readdy_learn.example.regulation_network import RegulationNetwork
@@ -15,17 +15,25 @@ from readdy_learn.example.regulation_network import RegulationNetwork
 regulation_network = RegulationNetwork()
 
 if __name__ == '__main__':
-    analysis = regulation_network.generate_analysis_object()
-    for i in range(4):
-        print("traj {}".format(i))
-        t = analysis.generate_or_load_traj_lma(i, regulation_network.target_time,
-                                               noise_variance=regulation_network.noise_variance,
-                                               realizations=regulation_network.realisations)
-        t.persist()
-        # analysis.plot_and_persist_lma_traj(t)
-        regulation_network.compute_gradient_derivatives(analysis)
-        # regulation_network.plot_concentrations(analysis, i)
-    cv_res = analysis.elastic_net_cv([i for i in range(4)], alphas=np.concatenate((
-        np.linspace(0.000001,0.0001,num=16),
-        np.linspace(0.0002,0.012,num=16),
-        np.linspace(0.012,0.1,num=8))), l1_ratios=np.linspace(0., 1., num=3))
+    from sklearn.model_selection import LeaveOneOut
+    loo = LeaveOneOut()
+    for train, test in loo.split([1,2,3,4]):
+        print("train: {}, test: {}".format(train, test))
+
+    if False:
+        analysis = regulation_network.generate_analysis_object()
+        for i in range(4):
+            print("traj {}".format(i))
+            t = analysis.generate_or_load_traj_lma(i, regulation_network.target_time,
+                                                   noise_variance=regulation_network.noise_variance,
+                                                   realizations=regulation_network.realisations)
+            t.persist()
+            # analysis.plot_and_persist_lma_traj(t)
+            regulation_network.compute_gradient_derivatives(analysis)
+            # regulation_network.plot_concentrations(analysis, i)
+        cv_res = analysis.elastic_net_cv([i for i in range(4)],
+                                         alphas=np.concatenate(
+                                             (np.linspace(0.000001, 0.0001, num=16),
+                                              np.linspace(0.0002, 0.012, num=16),
[…]

### 2018-01-22 readdy_learn/analyze/analyze.py

@@ -649,8 +649,6 @@ class ReactionAnalysis(object):
             cv.result = np.load(persist_fname)
         return cv
 
-
-
     def elastic_net(self, train_n, alphas, l1_ratios, test_n=None, initial_guess=None, tol=1e-16, njobs=8):
         if test_n is None:
             test_n = [self._trajs[i] for i in range(len(self._trajs)) if i != train_n]
@@ -678,18 +676,17 @@ class ReactionAnalysis(object):
         return self.fname_prefix + "_solution_{}_".format(n) + self.fname_postfix + ".npy"
 
     def solve(self, n, alpha, l1_ratio, tol=1e-12, constrained=True, recompute=False):
+        if not isinstance(n, (list, tuple)):
+            n = [n]
 
         if not recompute:
             if not self.recompute and os.path.exists(self.get_solve_fname(n)):
                 return np.load(self.get_solve_fname(n))
 
         system = self._set_up_system(self.initial_states[n])
-        traj = self._trajs[n]
-        if isinstance(traj, str):
-            traj = tools.Trajectory(traj, self.timestep, interpolation_degree=self.interp_degree, verbose=False)
-            traj.update()
+        trajs = [self.get_traj(x) for x in n]
 
-        optsuite = sample_tools.Suite.from_trajectory(traj, system, self._bfc, interp_degree=self.interp_degree,
+        optsuite = sample_tools.Suite.from_trajectory(trajs, system, self._bfc, interp_degree=self.interp_degree,
                                                       tol=tol, alpha=alpha, l1_ratio=l1_ratio,
                                                       init_xi=np.zeros_like(self.desired_rates))
         estimator = optsuite.get_estimator(verbose=True, interp_degree=self.interp_degree, constrained=constrained)
@@ -743,6 +740,43 @@ def plot_cv_results(cv, mainscore=0, best_params_ix_l1=1.):
             else:
                 xs[l1_ratio] = [r['alpha']]
                 ys[l1_ratio] = [r['scores'][mainscore]]
+                allys[l1_ratio] = [r['scores']]
+    for l1_ratio in xs.keys():
+        l1xs = np.array(xs[l1_ratio])
+        l1ys = np.array(ys[l1_ratio])
+        l1allys = np.array([np.array(arr) for arr in allys[l1_ratio]]).T
+        sorts = np.argsort(l1xs)
[…]

### 2018-01-22 readdy_learn/analyze/estimator.py

@@ -226,54 +226,50 @@ class CV(object):
         self.tol = tol
 
     def compute_cv_result_cross_trajs(self, params):
+        alpha, l1_ratio, (train_ix, test_ix) = params
+        scores = []
+
+        train_trajs = [self.trajs[ix] for ix in train_ix]
+        test_trajs = [self.trajs[ix] for ix in test_ix]
+        estimator = ReaDDyElasticNetEstimator(train_trajs, self.bfc, alpha=alpha, maxiter=self.maxiter,
+                                              l1_ratio=l1_ratio, init_xi=self.init_xi, verbose=self.verbose,
+                                              method=self.method, rescale=self.rescale, tol=self.tol)
+        # fit the whole thing
+        estimator.fit(None)
+        if estimator.success_:
+            testimator = ReaDDyElasticNetEstimator(test_trajs, self.bfc, alpha=alpha,
+                                                   l1_ratio=l1_ratio, init_xi=self.init_xi,
+                                                   verbose=self.verbose,
+                                                   method=self.method, rescale=self.rescale, tol=self.tol)
+            testimator.coefficients_ = estimator.coefficients_
+            ttraj = testimator.trajs[0]
+            score = testimator.score(range(0, ttraj.n_time_steps), ttraj.dcounts_dt)
+            scores.append(score)
+
+        else:
+            print("no success for alpha={}, l1_ratio={}".format(alpha, l1_ratio))
+            print("status %s: %s" % (estimator.result_.status, estimator.result_.message))
+            print("%s / %s iterations" % (estimator.result_.nit, self.maxiter))
+
+        return {'scores': scores, 'alpha': alpha, 'l1_ratio': l1_ratio, 'train_trajs': train_ix, 'test_trajs': test_ix}
+
+    def fit_cross_trajs(self):
+        from readdy_learn.analyze.progress import Progress
         from sklearn.model_selection import LeaveOneOut
 
         loo = LeaveOneOut()
-
-        alpha, l1_ratio = params
-        scores = []
-
-        N = np.array(range(len(self.trajs)), dtype=int)
-
-        for train_ix, test_ix in loo.split(N):
-            train_trajs = [self.trajs[ix] for ix in train_ix]
-            test_trajs = [self.trajs[ix] for ix in test_ix]
-            estimator = ReaDDyElasticNetEstimator(train_trajs, self.bfc, alpha=alpha, maxiter=self.maxiter,
-                                                  l1_ratio=l1_ratio, init_xi=self.init_xi, verbose=self.verbose,
[…]

### 2018-01-22 readdy_learn/sample_tools.py

@@ -40,7 +40,11 @@ class Suite(object):
     def get_estimator(self, timestep=-1, interp_degree=10, verbose=False, constrained=True):
 
         if self._trajectory is not None:
-            self._trajectory.update()
+            if isinstance(self._trajectory, (list, tuple)):
+                for traj in self._trajectory:
+                    traj.update()
+            else:
+                self._trajectory.update()
             est = ReaDDyElasticNetEstimator(self._trajectory, self._bfc, alpha=self._alpha, l1_ratio=self._l1_ratio,
                                             maxiter=self._maxiter, method='SLSQP', verbose=verbose, approx_jac=False,
                                             options={'ftol': self._tol}, rescale=False, init_xi=self._init_xi,

### 2018-01-22 readdy_learn/analyze/analyze.py
f

@@ -675,7 +675,7 @@ class ReactionAnalysis(object):
     def get_solve_fname(self, n):
         return self.fname_prefix + "_solution_{}_".format(n) + self.fname_postfix + ".npy"
 
-    def solve(self, n, alpha, l1_ratio, tol=1e-12, constrained=True, recompute=False):
+    def solve(self, n, alpha, l1_ratio, tol=1e-12, constrained=True, recompute=False, verbose=True):
         if not isinstance(n, (list, tuple)):
             n = [n]
 
@@ -683,13 +683,13 @@ class ReactionAnalysis(object):
             if not self.recompute and os.path.exists(self.get_solve_fname(n)):
                 return np.load(self.get_solve_fname(n))
 
-        system = self._set_up_system(self.initial_states[n])
         trajs = [self.get_traj(x) for x in n]
 
-        optsuite = sample_tools.Suite.from_trajectory(trajs, system, self._bfc, interp_degree=self.interp_degree,
-                                                      tol=tol, alpha=alpha, l1_ratio=l1_ratio,
-                                                      init_xi=np.zeros_like(self.desired_rates))
-        estimator = optsuite.get_estimator(verbose=True, interp_degree=self.interp_degree, constrained=constrained)
+        estimator = rlas.ReaDDyElasticNetEstimator(trajs, self._bfc, alpha=alpha, l1_ratio=l1_ratio,
+                                                   maxiter=30000, method='SLSQP', verbose=verbose, approx_jac=False,
+                                                   options={'ftol': tol}, rescale=False, init_xi=np.zeros_like(self.desired_rates),
+                                                   constrained=constrained)
+
         estimator.fit(None)
         if estimator.success_:
             rates = estimator.coefficients_

### 2018-01-22 readdy_learn/analyze/analyze.py
f

@@ -590,30 +590,31 @@ class ReactionAnalysis(object):
             plt.savefig(fname)
 
     def get_lsq_fname(self, n):
-        return self._fname_prefix + "_lsq_{}_".format(n) + self._fname_postfix + ".npz"
+        nstr = "_".join(str(ns) for ns in n)
+        return self._fname_prefix + "_lsq_{}_".format(nstr) + self._fname_postfix + ".npz"
 
-    def least_squares(self, n, tol=1e-12, initial_guess=None):
-        fname = self.get_lsq_fname(n)
-        if self._recompute or not os.path.exists(fname):
-            system = self._set_up_system(self._initial_states[n])
-            traj = self._trajs[n]
-            if isinstance(traj, str):
-                traj = tools.Trajectory(traj, self.timestep, interpolation_degree=self.interp_degree, verbose=False)
-            if initial_guess is None:
-                initial_guess = np.zeros_like(self._desired_rates)
-            suite = sample_tools.Suite.from_trajectory(traj, system, self._bfc, interp_degree=self._interp_degree,
-                                                       tol=tol,
-                                                       init_xi=initial_guess)
-            estimator = suite.get_estimator(verbose=True, interp_degree=self._interp_degree)
-            estimator.fit(None)
-            if estimator.success_:
-                rates = estimator.coefficients_
-                np.save(fname, rates)
-            else:
-                raise ValueError("Didn't converge :-/")
+    def least_squares(self, n, tol=1e-12, initial_guess=None, recompute=True):
+        if not isinstance(n, (list, tuple)):
+            n = [n]
+
+        if not recompute:
+            if not self.recompute and os.path.exists(self.get_lsq_fname(n)):
+                return np.load(self.get_lsq_fname(n))
+
+        trajs = [self.get_traj(x) for x in n]
+
+        estimator = rlas.ReaDDyElasticNetEstimator(trajs, self._bfc, alpha=0., l1_ratio=1.,
+                                                   maxiter=30000, method='SLSQP', verbose=True, approx_jac=False,
+                                                   options={'ftol': tol}, rescale=False, init_xi=np.zeros_like(self.desired_rates),
+                                                   constrained=True)
+
+        estimator.fit(None)
+        if estimator.success_:
+            rates = estimator.coefficients_
+            np.save(self.get_lsq_fname(n), rates)
+            return rates
         else:
-            rates = np.load(fname)
-        return rates
[…]

### 2018-01-22 readdy_learn/analyze/analyze.py
removed unnecessary argument

@@ -593,7 +593,7 @@ class ReactionAnalysis(object):
         nstr = "_".join(str(ns) for ns in n)
         return self._fname_prefix + "_lsq_{}_".format(nstr) + self._fname_postfix + ".npz"
 
-    def least_squares(self, n, tol=1e-12, initial_guess=None, recompute=True):
+    def least_squares(self, n, tol=1e-12, recompute=True):
         if not isinstance(n, (list, tuple)):
             n = [n]
 

### 2018-01-22 readdy_learn/analyze/analyze.py
removed unnecessary argument

@@ -674,7 +674,8 @@ class ReactionAnalysis(object):
         return cv
 
     def get_solve_fname(self, n):
-        return self.fname_prefix + "_solution_{}_".format(n) + self.fname_postfix + ".npy"
+        nstr = "_".join(str(ns) for ns in n)
+        return self.fname_prefix + "_solution_{}_".format(nstr) + self.fname_postfix + ".npy"
 
     def solve(self, n, alpha, l1_ratio, tol=1e-12, constrained=True, recompute=False, verbose=True):
         if not isinstance(n, (list, tuple)):

### 2018-01-22 readdy_learn/analyze/analyze.py
removed unnecessary argument

@@ -591,7 +591,7 @@ class ReactionAnalysis(object):
 
     def get_lsq_fname(self, n):
         nstr = "_".join(str(ns) for ns in n)
-        return self._fname_prefix + "_lsq_{}_".format(nstr) + self._fname_postfix + ".npz"
+        return self._fname_prefix + "_lsq_{}_".format(nstr) + self._fname_postfix + ".npy"
 
     def least_squares(self, n, tol=1e-12, recompute=True):
         if not isinstance(n, (list, tuple)):

### 2018-01-22 readdy_learn/analyze/analyze.py
removed unnecessary argument

@@ -501,7 +501,7 @@ class ReactionAnalysis(object):
             traj.update()
             traj.persist()
 
-        while len(self.trajs) < n+1:
+        while len(self.trajs) < n + 1:
             self.trajs.append(None)
         self.trajs[n] = traj
         return traj
@@ -529,6 +529,7 @@ class ReactionAnalysis(object):
         def fun_reference(data, _):
             theta = np.array([f(data) for f in bfc.functions])
             return np.matmul(self._desired_rates, theta)
+
         traj = n
         if isinstance(traj, int):
             traj = self._trajs[traj]
@@ -555,7 +556,7 @@ class ReactionAnalysis(object):
 
         if outfile is not None:
             f.savefig(outfile)
-        #plt.show()
+        # plt.show()
         return traj
 
     def plot_concentration_curves(self, n, fname=None, species=None, plot_estimated=True):
@@ -569,9 +570,9 @@ class ReactionAnalysis(object):
         estimated = sample_tools.Suite.estimated_behavior(self._desired_rates, self._bfc,
                                                           self.initial_states[n].squeeze(), traj.times)
 
-        #fig, ax1 = plt.subplots(nrows=1, ncols=1)
+        # fig, ax1 = plt.subplots(nrows=1, ncols=1)
 
-        #fig.suptitle('Training trajectory')
+        # fig.suptitle('Training trajectory')
         plt.xlabel('time')
         plt.ylabel('concentration')
 
@@ -585,7 +586,7 @@ class ReactionAnalysis(object):
                     integrated_ld = deriv.integrate.cumtrapz(traj.separate_derivs[type_id], x=traj.times, initial=0) \
                                     + self.initial_states[n].squeeze()[type_id]
                     plt.plot(traj.times, integrated_ld, "r--", label="integrated derivative")
-        #plt.legend(loc="upper right")
+        # plt.legend(loc="upper right")
         if fname is not None:
             plt.savefig(fname)
 
@@ -605,7 +606,8 @@ class ReactionAnalysis(object):
 
         estimator = rlas.ReaDDyElasticNetEstimator(trajs, self._bfc, alpha=0., l1_ratio=1.,
                                                    maxiter=30000, method='SLSQP', verbose=True, approx_jac=False,
-                                                   options={'ftol': tol}, rescale=False, init_xi=np.zeros_like(self.desired_rates),
[…]

### 2018-01-22 readdy_learn/analyze/analyze.py
removed unnecessary argument

@@ -857,7 +857,8 @@ def best_params(cv):
                 scores.append(d['scores'][0])
             scores = np.array(scores)
             meanscore = np.mean(scores)
-            if meanscore > current_best_score:
+            print(meanscore)
+            if current_best_score == -1 or meanscore > current_best_score:
                 best_alpha = alpha
                 best_l1_ratio = l1_ratio
                 current_best_score = meanscore

### 2018-01-22 docs/manuscript.tex
manuscript update

@@
 Applying the concentration trajectory to the library yields $\Theta(\textbf{X})\in\mathbb{R}^{T\times S\times R}$.
-Following the approach of SINDy, the goal is now to find coefficients $\Xi = \begin{pmatrix} \xi_1 & \xi_2 & \cdots & \xi_R
+Following the approach of SINDy, the goal is to find coefficients $\Xi = \begin{pmatrix} \xi_1 & \xi_2 & \cdots & \xi_R
 	\end{pmatrix}^T$, so that
@@
 	\end{align}
-	In particular, the system is linear in the coefficients $\Xi$, which makes sparse regression tools such as LASSO \cite{Tibshirani1996, Hastie2009} applicable.
+	In particular, the system is linear in the coefficients $\Xi$, which makes potentially sparse regression tools such as elastic net regularization \cite{Zou2005} applicable.
 To this end, one can consider the minimization problem to find $\hat{\Xi}$ such that
 	\begin{align}
-		\hat{\Xi} = \underset{\Xi}{\arg\min}\left( \frac{1}{2T}\left\| \dot{\textbf{X}} - \Theta(\textbf{X})\Xi \right\|_F^2 + \alpha\|\Xi\|_1 \right) \quad \text{subject to }\Xi \geq 0,
+		\hat{\Xi} = \underset{\Xi}{\arg\min}\left( \frac{1}{2T}\left\| \dot{\textbf{X}} - \Theta(\textbf{X})\Xi \right\|_F^2 + \alpha\lambda\|\Xi\|_1 + \alpha(1-\lambda)\|\Xi\|_2^2 \right) \quad \text{subject to }\Xi \geq 0,
 	\label{method:minimizationproblem}\end{align}
-	where $\|\cdot\|_F$ denotes the Frobenius norm and $\alpha\geq 0$ is a sparsity inducing hyperparameter.
+	where $\|\cdot\|_F$ denotes the Frobenius norm, $\lambda\in[0,1]$ a hyperparameter that interpolates linearly between LASSO \cite{Tibshirani1996, Hastie2009} and Ridge \cite{Hoerl1} methods, and $\alpha\geq 0$ is a hyperparameter that, depending on $\lambda$, can induce sparsity and give preference to smaller solutions in the $L_1$ or $L_2$ sense.
+
 	For $\alpha=0$ the minimization problem reduces to constrained least-squares.
@@
 
-	\begin{figure}
-		\begin{center}
-			\includegraphics[width=.5\textwidth]{./figures_tex/aux_linear_approximation.eps}
-		\end{center}
-		\caption{This figure depicts a possible output yielded by an application of the Gillespie method and the performed piecewise linear approximation in order to obtain a better behaving derivative.}
-		\label{examples:linear-approximation}
-	\end{figure}
+	%\begin{figure}
+	%	\begin{center}
+	%		\includegraphics[width=.5\textwidth]{./figures_tex/aux_linear_approximation.eps}
+	%	\end{center}
+	%	\caption{This figure depicts a possible output yielded by an application of the Gillespie method and the performed piecewise linear approximation in order to obtain a better behaving derivative.}
+	%	\label{examples:linear-approximation}
+	%\end{figure}
+
+	(\textcolor{red}{this has changed, rewrite!})
 

### 2018-01-22 readdy_learn/example/regulation_network.py
slightly altered initial condition 3, use gradient for all species

@@ -58,7 +58,7 @@ class RegulationNetwork(object):
 
         initial_states = [
             [1, 0, 0, 1, 0, 0, 1, 0, 0], [1, 2, 0, 1, 0, 3, 1, 0, 0], [1, 1, 2, 1, 0, 2.5, 1, 0, 2],
-            [1, 1, 2, 1, 0, 0, 1, 4, 0]
+            [1, 1, 2, 1, 0, 0, 1, 3, 0]
         ]
         self.initial_states = [_np.array([arr]) for arr in initial_states]
 
@@ -240,11 +240,11 @@ class RegulationNetwork(object):
     def compute_gradient_derivatives(self, analysis):
         for t in range(len(self.initial_states)):
             traj = analysis.get_traj(t)
-            for sp in [0, 3, 6]:
-                dx = _np.zeros_like(traj.counts[:, sp])
-                print("species {} dx.shape {}".format(sp, dx.shape))
-                traj.separate_derivs[sp] = dx
-            for sp in [1, 2, 4, 5, 7, 8]:
+            #for sp in [0, 3, 6]:
+            #    dx = _np.zeros_like(traj.counts[:, sp])
+            #    print("species {} dx.shape {}".format(sp, dx.shape))
+            #    traj.separate_derivs[sp] = dx
+            for sp in [1, 2, 4, 5, 7, 8, 0, 3, 6]:
                 x = traj.counts[:, sp]
                 dx = _np.zeros_like(x)
                 dt = traj.time_step

### 2018-01-23 docs/manuscript.tex
manuscript update

@@
 
+	When presented with multiple trajectories one can extend the objective function by a sum over these and obtain a more general minimization problem that reads: Find $\hat{\Xi}$ such that
+	\begin{align}
+		\hat{\Xi} = \underset{\Xi}{\arg\min}\left( \sum_{\text{traj}}\frac{1}{2T_{\text{traj}}}\left\| \dot{\textbf{X}_\text{traj}} - \Theta(\textbf{X}_\text{traj})\Xi \right\|_F^2 + \alpha\lambda\|\Xi\|_1 + \alpha(1-\lambda)\|\Xi\|_2^2 \right)
+	\end{align}
+	subject to $\Xi\geq 0$. (\textcolor{red}{this probably should be motivated better})
+
 	For $\alpha=0$ the minimization problem reduces to constrained least-squares.
@@
 
-	%\begin{figure}
-	%	\begin{center}
-	%		\includegraphics[width=.5\textwidth]{./figures_tex/aux_linear_approximation.eps}
-	%	\end{center}
-	%	\caption{This figure depicts a possible output yielded by an application of the Gillespie method and the performed piecewise linear approximation in order to obtain a better behaving derivative.}
-	%	\label{examples:linear-approximation}
-	%\end{figure}
-
-	(\textcolor{red}{this has changed, rewrite!})
-
-	For generating time series data of concentrations we use the Gillespie method \cite{Gillespie1976}.
-For given initial conditions, we produce several realizations which then are converted to a trajectory with fixed time step and averaged.
-Since these trajectories are piecewise constant, one has to take special care when approximating the data's temporal derivative that is needed in (\ref{method:minimizationproblem}).
-Applying finite differences has the effect that, with decreasing time step, the derivative mostly constant zero and approaches infinity at the jump discontinuities.
-To counter this effect, we beforehand perform a linear approximation between each two adjacent discontinuities, as depicted in Figure \ref{examples:linear-approximation} and thus obtain a piecewise constant derivative.
-(\textcolor{red}{Use regularized derivative?})
+	For generating time series data of concentrations we integrate the law of mass action equations and add Gaussian noise.
+In that manner we produce several realizations for a couple of initial conditions.
+The derivatives are evaluated by finite differences of first order.
 

### 2018-01-23 readdy_learn/analyze/analyze.py
f

@@ -128,8 +128,8 @@ class ReactionAnalysis(object):
     def __init__(self, bfc, desired_rates, initial_states, set_up_system, recompute=False, recompute_traj=False,
                  fname_prefix="", fname_postfix="", n_species=4, timestep=5e-4,
                  interp_degree='regularized_derivative', ld_derivative_config=None, species_names=None):
-        if initial_states is None or len(initial_states) <= 1:
-            raise ValueError("needs at least two initial states!")
+        #if initial_states is None or len(initial_states) <= 1:
+        #    raise ValueError("needs at least two initial states!")
         if ld_derivative_config is None:
             ld_derivative_config = {
                 'ld_derivative_atol': 1e-8,

### 2018-01-23 readdy_learn/analyze/analyze.py
f

@@ -857,7 +857,6 @@ def best_params(cv):
                 scores.append(d['scores'][0])
             scores = np.array(scores)
             meanscore = np.mean(scores)
-            print(meanscore)
             if current_best_score == -1 or meanscore > current_best_score:
                 best_alpha = alpha
                 best_l1_ratio = l1_ratio

### 2018-01-23 readdy_learn/example/regulation_network.py
f

@@ -75,8 +75,6 @@ class RegulationNetwork(object):
         self.noise_variance = 0.0001
         self.target_time = 3.
         self.train_n = 1
-        self.test_n = 2
-        self.traj_n_s = [self.train_n, self.test_n]
         self.realisations = 60
         self.timestep = 1e-3
 

### 2018-01-23 readdy_learn/example/regulation_network.py
f

@@ -1,10 +1,8 @@
-
 import numpy as _np
 import pynumtools.kmc as kmc
 import readdy_learn.analyze.basis as basis
-import matplotlib as mpl
-#mpl.rcParams['figure.figsize'] = (13, 13)
 import matplotlib.pyplot as plt
+
 
 class RegulationNetwork(object):
 
@@ -63,18 +61,17 @@ class RegulationNetwork(object):
         self.initial_states = [_np.array([arr]) for arr in initial_states]
 
         self.ld_derivative_config = {
-                'ld_derivative_atol': 1e-4,
-                'ld_derivative_rtol': None,
-                'ld_derivative_linalg_solver_maxit': 100000,
-                'ld_derivative_alpha': 1e-1,
-                'ld_derivative_solver': 'spsolve',
-                'ld_derivative_linalg_solver_tol': 1e-10,
-                'ld_derivative_use_preconditioner': False
-            }
+            'ld_derivative_atol': 1e-4,
+            'ld_derivative_rtol': None,
+            'ld_derivative_linalg_solver_maxit': 100000,
+            'ld_derivative_alpha': 1e-1,
+            'ld_derivative_solver': 'spsolve',
+            'ld_derivative_linalg_solver_tol': 1e-10,
+            'ld_derivative_use_preconditioner': False
+        }
 
         self.noise_variance = 0.0001
         self.target_time = 3.
-        self.train_n = 1
         self.realisations = 60
         self.timestep = 1e-3
 
@@ -238,7 +235,7 @@ class RegulationNetwork(object):
     def compute_gradient_derivatives(self, analysis):
         for t in range(len(self.initial_states)):
             traj = analysis.get_traj(t)
-            #for sp in [0, 3, 6]:
+            # for sp in [0, 3, 6]:
             #    dx = _np.zeros_like(traj.counts[:, sp])
             #    print("species {} dx.shape {}".format(sp, dx.shape))
             #    traj.separate_derivs[sp] = dx

### 2018-01-23 readdy_learn/analyze/analyze.py
f

@@ -518,6 +518,26 @@ class ReactionAnalysis(object):
 
         return result
 
+    def compute_L2_error(self, n, rates):
+        from scipy.integrate import odeint, trapz
+        bfc = self._bfc
+
+        def fun(data, _):
+            theta = np.array([f(data) for f in bfc.functions])
+            return np.matmul(rates, theta)
+
+        def fun_reference(data, _):
+            theta = np.array([f(data) for f in bfc.functions])
+            return np.matmul(self._desired_rates, theta)
+
+        traj = self.get_traj(n)
+        xs = traj.times
+        num_solution = np.array(odeint(fun, self.initial_states[n].squeeze(), xs))
+        reference_soln = np.array(odeint(fun_reference, self.initial_states[n].squeeze(), xs))
+
+        return trapz(np.abs(num_solution - reference_soln), x=xs)
+
+
     def plot_results(self, n, rates, title=None, outfile=None):
         from scipy.integrate import odeint
         bfc = self._bfc

### 2018-01-23 docs/gene_regulation.tex
description of gene regulatory network

\documentclass[oneside, abstracton, titlepage]{scrartcl}

\usepackage[left=2.5cm,right=2.6cm,top=3cm,bottom=3cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[intlimits]{amsmath}
\usepackage{amssymb}
\usepackage{color}
\usepackage{subcaption}
\usepackage[version=4]{mhchem}

\usepackage{graphicx}

\begin{document}

\section{Example: Regulation of gene expression}\label{sec:generegulation}

To demonstrate our method we estimate a gene-regulatory network from data. The data is time-series of molecule-concentrations obtained from law of mass action (LMA) equations with additive gaussian noise.

Let A, B and C be the three species of proteins. Each of them is translated from its according mRNA molecule. Each mRNA molecule is transcribed from its according DNA. The proteins and mRNA molecules decay over time. For each species one can formulate the following basic reactions
\begin{align*}
    \ce{ DNA_i } & \ce{ -> DNA_i + mRNA_i } &\text{(transcription)} \\
    \ce{ mRNA_i } & \ce{ -> mRNA_i + Protein_i } &\text{(translation)} \\
    \ce{ mRNA_i } & \ce{ -> } \emptyset &\text{(decay of mRNA)} \\
    \ce{ Protein_i } & \ce{ -> } \emptyset &\text{(decay of Protein)}
\end{align*}

A, B and C regulate each other in a negative way, by hindering the transcription process. In the LMA model we account for this by a reaction. A repression of species $j$ by species $i$ is modeled as
\begin{equation*}
    \ce{ Protein_i + mRNA_j -> Protein_i } \quad \text{(regulation by repression)}
\end{equation*}

In our example A proteins regulate the mRNA$_\mathrm{B}$ molecules, B proteins regulate the mRNA$_\mathrm{A}$ molecules and C proteins regulate the mRNA$_\mathrm{A}$ molecules. The network of all species and reactions is depicted in figure ???????. This will serve as our reference model to generate the time-series of concentrations.
\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{./figures_tex/gene_regulation_full.pdf}
    \caption{The regulation network example described in Section \ref{sec:generegulation}. Each circle is a species, each arrow corresponds to one reaction. The red arrows describe the reference regulatory network which generates our data.}
    \label{fig:network}
\end{figure}

\end{document}

### 2018-01-25 docs/manuscript.tex
inserted description of gene regulatory network

@@
 \usepackage{subcaption}
+\usepackage[version=4]{mhchem}
 
@@
 As initial guess for the regularized problem we use the solution of (\ref{method:minimizationproblem}) with $\alpha=0$.
+
+    \section{Example system: Regulation of gene expression}\label{sec:generegulation}
+
+    To demonstrate our method we estimate a gene-regulatory network from data.
+The data is time-series of molecule-concentrations.
+We generate the data by integrating the law of mass action (LMA) equations of the true model and add gaussian noise.
+
+    We describe the gene-regulatory network as follows.
+Let A, B and C be the three species of proteins.
+Each of them is translated from its according mRNA molecule.
+Each mRNA molecule is transcribed from its according DNA.
+The proteins and mRNA molecules decay over time.
+For each species one can formulate the following basic reactions
+    \begin{align*}
+        \ce{ DNA_i } & \ce{ -> DNA_i + mRNA_i } &\text{(transcription)} \\
+        \ce{ mRNA_i } & \ce{ -> mRNA_i + Protein_i } &\text{(translation)} \\
+        \ce{ mRNA_i } & \ce{ -> } \emptyset &\text{(decay of mRNA)} \\
+        \ce{ Protein_i } & \ce{ -> } \emptyset &\text{(decay of Protein)}
+    \end{align*}
+
+    A, B and C regulate each other in a negative way, by hindering the transcription process.
+In the LMA model we account for this by a reaction.
+A repression of species $j$ by species $i$ is modeled as
+    \begin{equation*}
+        \ce{ Protein_i + mRNA_j -> Protein_i } \quad \text{(regulation by repression)}
+    \end{equation*}
+
+    In our example A proteins regulate the mRNA$_\mathrm{B}$ molecules, B proteins regulate the mRNA$_\mathrm{A}$ molecules and C proteins regulate the mRNA$_\mathrm{A}$ molecules.
+The network of all species and reactions is depicted in figure \ref{fig:network}.
+This will serve as our reference model to generate the time-series of concentrations.
+    \begin{figure}
+        \centering
+        \includegraphics[width=.5\textwidth]{./figures_tex/gene_regulation_full.pdf}
+        \caption{The regulation network example described in Section \ref{sec:generegulation}.
+Each circle is a species, each arrow corresponds to one reaction.
+The red arrows describe the regulatory network which generates our data.}
+        \label{fig:network}
+    \end{figure}
 

### 2018-01-25 readdy_learn/example/regulation_network.py
f

@@ -250,13 +250,13 @@ class RegulationNetwork(object):
                 traj.persist()
 
 
-def sample_lsq_rates(realizations, base_variance=.1, samples_per_variance=8, njobs=8):
+def sample_lsq_rates(realizations, base_variance=.1, samples_per_variance=8, njobs=8, timestep=6e-3):
     if not isinstance(realizations, (list, tuple, _np.ndarray)):
         realizations = [realizations]
 
     def get_regulation_network(n):
         regulation_network = RegulationNetwork()
-        regulation_network.timestep = 6e-3
+        regulation_network.timestep = timestep
         regulation_network.realisations = 1.
         regulation_network.noise_variance = base_variance / n
         regulation_network.initial_states = [regulation_network.initial_states[1]]

### 2018-01-25 docs/manuscript.tex
f
described different scenarios of measurement

@@
     \end{figure}
+
+    \section{Scenarios of measurement}
+
+    The estimation method we propose shall be fit to find reaction networks and rates from experimental data.
+Therefore we define the following measurement scenarios, that differentiate how many timeseries are available and how much noise is contained in the data:
+    \begin{enumerate}
+        \item The initial state of the measurement apparatus can be reproduced exactly.
+There is only one set of generating reaction rates.
+Hence we can observe the same process $N$ times.
+Each observation has a certain level of noise in each datapoint.
+One can average the $N$ observations.
+Thus the outcome of this scenarios is one time-series of concentrations with a noise-level, that decreases with $N$.
+        \item The initial state of the measurement apparatus differs in each observation. The outcome of this scenario is $N$ time-series each with a given noise level and different initial states.
+        \item The initial state of the measurement apparatus differs in each observation.
+Additionally the generating reaction rates slightly differ in each observation.
+The outcome of this scenario is $N$ time-series each with a given noise level, different initial states and a slighty different underlying process.
+    \end{enumerate}
 

### 2018-01-25 docs/manuscript.tex
different wording

--- 2018-01-25 docs/manuscript.tex
@@
 	\begin{align}
-		\hat{\Xi} = \underset{\Xi}{\arg\min}\left( \sum_{\text{traj}}\frac{1}{2T_{\text{traj}}}\left\| \dot{\textbf{X}_\text{traj}} - \Theta(\textbf{X}_\text{traj})\Xi \right\|_F^2 + \alpha\lambda\|\Xi\|_1 + \alpha(1-\lambda)\|\Xi\|_2^2 \right)
+		\hat{\Xi} = \underset{\Xi}{\arg\min}\left( \sum_{\text{traj}}\frac{1}{2T_{\text{traj}}}\left\| \dot{\textbf{X}}_\text{traj} - \Theta(\textbf{X}_\text{traj})\Xi \right\|_F^2 + \alpha\lambda\|\Xi\|_1 + \alpha(1-\lambda)\|\Xi\|_2^2 \right)
 	\end{align}
-	subject to $\Xi\geq 0$. (\textcolor{red}{this probably should be motivated better})
-
+	subject to $\Xi\geq 0$.
 	For $\alpha=0$ the minimization problem reduces to constrained least-squares.
@@
     \end{figure}
+
+    \section{Scenarios of measurement}
+
+    The proposed estimation method is designed to find reaction networks and rates from experimental data.
+Therefore one can define the following measurement scenarios, that differentiate how many timeseries are available and how much noise is contained in the data:
+    \begin{enumerate}
+        \item The initial state of the measurement apparatus can be reproduced exactly.
+There is only one set of generating reaction rates.
+Each datapoint is subject to noise.
+The outcome of this scenarios is one time-series of concentrations with a certain noise-level.
+        \item The process is observed multiple times.
+The initial state of the measurement apparatus differs in each observation.
+The outcome of this scenario is multiple time-series each with a given noise level and different initial states.
+        \item The process is observed multiple times.
+The initial state of the measurement apparatus differs in each observation.
+Additionally the generating reaction rates slightly differ in each observation.
+The outcome of this scenario is multiple time-series each with a given noise level, different initial states and slighty perturbed underlying reaction rates.
+    \end{enumerate}
 

### 2018-01-25 readdy_learn/example/regulation_network.py
f

@@ -286,7 +286,7 @@ def sample_lsq_rates(realizations, base_variance=.1, samples_per_variance=8, njo
             print("initial states:")
             for init in regulation_network.initial_states:
                 print("\t{}".format(init))
-        lsq_rates = analysis.least_squares([0], tol=1e-14, recompute=True, persist=False, verbose=False)
+        lsq_rates = analysis.least_squares([0], tol=1e-16, recompute=True, persist=False, verbose=False)
         l2_err = analysis.compute_L2_error(0, lsq_rates)
         if verbose:
             print("|LMA-LMA_est|_2 = {}".format(l2_err))

### 2018-01-25 readdy_learn/example/regulation_network.py
f

@@ -286,7 +286,19 @@ def sample_lsq_rates(realizations, base_variance=.1, samples_per_variance=8, njo
             print("initial states:")
             for init in regulation_network.initial_states:
                 print("\t{}".format(init))
-        lsq_rates = analysis.least_squares([0], tol=1e-16, recompute=True, persist=False, verbose=False)
+        try:
+            lsq_rates = analysis.least_squares([0], tol=1e-16, recompute=True, persist=False, verbose=False)
+        except ValueError:
+            try:
+                lsq_rates = analysis.least_squares([0], tol=1e-15, recompute=True, persist=False, verbose=False)
+            except ValueError:
+                try:
+                    lsq_rates = analysis.least_squares([0], tol=1e-14, recompute=True, persist=False, verbose=False)
+                except ValueError:
+                    try:
+                        lsq_rates = analysis.least_squares([0], tol=1e-13, recompute=True, persist=False, verbose=False)
+                    except ValueError:
+                        raise ValueError("This takes the cake")
         l2_err = analysis.compute_L2_error(0, lsq_rates)
         if verbose:
             print("|LMA-LMA_est|_2 = {}".format(l2_err))

### 2018-01-25 docs/manuscript.tex
f

@@
 \usepackage{color}
+\usepackage{xcolor}
 \usepackage{subcaption}
@@
 \usepackage{graphicx}
+
+\newsavebox{\selvestebox}
+\newenvironment{colbox}[1]
+{\newcommand\colboxcolor{#1}%
+	\begin{lrbox}{\selvestebox}%
+		\begin{minipage}{\dimexpr\columnwidth-2\fboxsep\relax}}
+		{\end{minipage}\end{lrbox}%
+	\begin{center}
+		\colorbox[HTML]{\colboxcolor}{\usebox{\selvestebox}}
+\end{center}}
+\newcommand\restr[2]{{% we make the whole thing an ordinary symbol
+		\left.\kern-\nulldelimiterspace % automatically resize the bar with \right
+		#1 % the function
+		\vphantom{\big|} % pretend it's a little taller at normal size
+		\right|_{#2} % this is the delimiter
+}}
 
@@
 	\section{Introduction}
-	When presented with a time series of possibly noisy non-equilibrium concentration fluctuations of some species as output of, e.g., measurements from experiments or simulations that were parameterized by microscopic rates \textcolor{red}{(cite ReaDDy?)}, one can ask for the corresponding macroscopic rates and a generating reaction network.
+	When presented with a time series of possibly noisy non-equilibrium concentration fluctuations of some species as output of, e.g., measurements from experiments or simulations that were parameterized by microscopic rates, one can ask for the corresponding macroscopic rates and a generating reaction network.
 	In this paper we present an application of the shallow learning method SINDy \cite{Brunton2015}.
@@
 	For $\alpha=0$ the minimization problem reduces to constrained least-squares.
-In order to solve (\ref{method:minimizationproblem}) we apply the sequential least-squares minimizer SLSQP, originally described in \cite{Kraft1988}, by the software package SciPy \cite{SciPy}.
-As initial guess for the regularized problem we use the solution of (\ref{method:minimizationproblem}) with $\alpha=0$.
+In order to solve (\ref{method:minimizationproblem}) the sequential least-squares minimizer SLSQP \cite{Kraft1988} is applied via the software package SciPy \cite{SciPy}.
+Since only the concentration data $\mathbf{X}$ is available but not its temporal derivative $\dot{\mathbf{X}}$, it is approximated numerically by second order finite differences with the exception of boundary data.
 
@@
 
-    To demonstrate our method we estimate a gene-regulatory network from data.
-The data is time-series of molecule-concentrations.
-We generate the data by integrating the law of mass action (LMA) equations of the true model and add gaussian noise.
+    To demonstrate the method a gene-regulatory network is estimated from time series of molecule-concentrations.
+The data of is generated by integrating the law of mass action equations of the underlying model and adding Gaussian noise.
 
-    We describe the gene-regulatory network as follows.
-Let A, B and C be the three species of proteins.
-Each of them is translated from its according mRNA molecule.
-Each mRNA molecule is transcribed from its according DNA.
-The proteins and mRNA molecules decay over time.
-For each species one can formulate the following basic reactions
+    To this end, let $A$, $B$, and $C$ be three species of proteins which are being translated from a corresponding mRNA molecule.
+Each mRNA in turn has a corresponding DNA which it is transcribed from.
+The proteins and mRNA molecules decay over time whereas the DNA concentration remains constant.
+
+    For each of the three species one can formulate the following basic reactions
 	\begin{align*}
-        \ce{ DNA_i } & \ce{ -> DNA_i + mRNA_i } &\text{(transcription)} \\
-        \ce{ mRNA_i } & \ce{ -> mRNA_i + Protein_i } &\text{(translation)} \\
-        \ce{ mRNA_i } & \ce{ -> } \emptyset &\text{(decay of mRNA)} \\
-        \ce{ Protein_i } & \ce{ -> } \emptyset &\text{(decay of Protein)}
+        \ce{ DNA_i } & \ce{ -> DNA_i + mRNA_i } &\text{(transcription)}, \\
+        \ce{ mRNA_i } & \ce{ -> mRNA_i + Protein_i } &\text{(translation)}, \\
+        \ce{ mRNA_i } & \ce{ -> } \emptyset &\text{(decay of mRNA)}, \\
+        \ce{ Protein_i } & \ce{ -> } \emptyset &\text{(decay of Protein)},
     \end{align*}
+    for $i=A,B,C$.
 
-    A, B and C regulate each other in a negative way, by hindering the transcription process.
-In the LMA model we account for this by a reaction.
-A repression of species $j$ by species $i$ is modeled as
+    The protein concentrations of $A$, $B$, and $C$ regulate each other in a negative way by hindering the transcription process.
+In the law of mass action model we account for this by a reaction.
+A repression of species $j$ by species $i$ can be modeled as
     \begin{equation*}
-        \ce{ Protein_i + mRNA_j -> Protein_i } \quad \text{(regulation by repression)}
+        \ce{ Protein_i + mRNA_j -> Protein_i } \quad \text{(regulation by repression)}.
     \end{equation*}
 
-    In our example A proteins regulate the mRNA$_\mathrm{B}$ molecules, B proteins regulate the mRNA$_\mathrm{A}$ molecules and C proteins regulate the mRNA$_\mathrm{A}$ molecules.
-The network of all species and reactions is depicted in figure \ref{fig:network}.
-This will serve as our reference model to generate the time-series of concentrations.
+    In our example proteins of type $A$ regulate the mRNA$_\mathrm{B}$ molecules, proteins of type $B$ regulate the mRNA$_\mathrm{A}$ molecules and proteins of type $C$ regulate the mRNA$_\mathrm{A}$ molecules.
+The network of all species and reactions is depicted in Figure \ref{fig:network}.
+This serves as reference model to generate the time series of concentrations.
     \begin{figure}
@@
         \caption{The regulation network example described in Section \ref{sec:generegulation}.
-Each circle is a species, each arrow corresponds to one reaction.
-The red arrows describe the regulatory network which generates our data.}
+Each circle depicts a species, each arrow corresponds to one reaction.
+The red arrows describe the regulatory network which generates the example data.}
         \label{fig:network}
@@
 
-    \section{Scenarios of measurement}
+    \section{Scenarios of measurement}\label{sec:scenarios-of-measurement}
 
     The proposed estimation method is designed to find reaction networks and rates from experimental data.
-Therefore one can define the following measurement scenarios, that differentiate how many timeseries are available and how much noise is contained in the data:
+In the following measurement scenarios are considered that differ in how many time series are available and the level of noise that the data is subject to:
     \begin{enumerate}
@@
 There is only one set of generating reaction rates.
-Each datapoint is subject to noise.
-The outcome of this scenarios is one time-series of concentrations with a certain noise-level.
+Each frame of a time series is subject to noise.
+The outcome of this scenario is one time series of concentrations with a certain noise-level.
         \item The process is observed multiple times.
 The initial state of the measurement apparatus differs in each observation.
-The outcome of this scenario is multiple time-series each with a given noise level and different initial states.
+The outcome of this scenario is multiple time series each with a certain noise level and different initial states.
         \item The process is observed multiple times.
 The initial state of the measurement apparatus differs in each observation.
-Additionally the generating reaction rates slightly differ in each observation.
-The outcome of this scenario is multiple time-series each with a given noise level, different initial states and slighty perturbed underlying reaction rates.
+Additionally the generating reaction rates differ slightly in each observation.
+The outcome of this scenario is multiple time series each with a given noise level, different initial states and slightly perturbed underlying reaction rates.
     \end{enumerate}
@@
 
-	For generating time series data of concentrations we integrate the law of mass action equations and add Gaussian noise.
-In that manner we produce several realizations for a couple of initial conditions.
-The derivatives are evaluated by finite differences of first order.
+	In the following two examples are presented. The first example considers data according to the first scenario of measurement as presented in Section \ref{sec:scenarios-of-measurement}.
+	The minimization problem (\ref{method:minimizationproblem}) with $\alpha=0$ is solved for different magnitudes of noise variance.
 
@@
 	\label{sec:ex1}
+
+	In this example it is demonstrated that given enough data, the original model can be recovered in the limit of the noise variance $\sigma^2\to 0$.
+
+	\begin{colbox}{F8E0E0}
+		todo
+	\end{colbox}
 
@@
 	\begin{align}
-		f_\alpha = \frac{1}{T}\|\mathbf{\dot{X}}_\text{test} - \Theta (\textbf{X}_\text{test}) \Xi_\text{train}\|_F^2,
+		f_\alpha = \frac{1}{T}\|\dot{\mathbf{X}}_\text{test} - \Theta (\textbf{X}_\text{test}) \Xi_\text{train}\|_F^2,
 		\label{ex2:score-function}

### 2018-01-26 readdy_learn/analyze/analyze.py
f

@@ -7,6 +7,7 @@ import readdy_learn.analyze.derivative as deriv
 import readdy_learn.analyze.estimator as rlas
 import readdy_learn.analyze.generate as generate
 import readdy_learn.analyze.tools as tools
+import readdy_learn.analyze.basis as basis
 import readdy_learn.sample_tools as sample_tools
 
 
@@ -128,7 +129,7 @@ class ReactionAnalysis(object):
     def __init__(self, bfc, desired_rates, initial_states, set_up_system, recompute=False, recompute_traj=False,
                  fname_prefix="", fname_postfix="", n_species=4, timestep=5e-4,
                  interp_degree='regularized_derivative', ld_derivative_config=None, species_names=None):
-        #if initial_states is None or len(initial_states) <= 1:
+        # if initial_states is None or len(initial_states) <= 1:
         #    raise ValueError("needs at least two initial states!")
         if ld_derivative_config is None:
             ld_derivative_config = {
@@ -540,7 +541,6 @@ class ReactionAnalysis(object):
             err += trapz(np.abs(num_solution[:, i] - reference_soln[:, i]), x=xs)
         return err
 
-
     def plot_results(self, n, rates, title=None, outfile=None):
         from scipy.integrate import odeint
         bfc = self._bfc
@@ -617,7 +617,33 @@ class ReactionAnalysis(object):
         nstr = "_".join(str(ns) for ns in n)
         return self._fname_prefix + "_lsq_{}_".format(nstr) + self._fname_postfix + ".npy"
 
-    def least_squares(self, n, tol=1e-12, recompute=True, persist=True, verbose=True):
+    def least_squares_iter(self, n, cutoff, tol=1e-12, verbose=True):
+        N = np.arange(self.bfc.n_basis_functions, dtype=int)
+        bfcs = basis.BasisFunctionConfiguration(self.n_species)
+        bfcs.functions = self.bfc.functions
+        myrates = np.zeros((len(N),), dtype=np.float64)
+        while True:
+            rates = self.least_squares(n, tol=tol, recompute=True, persist=False, verbose=verbose, bfc=bfcs)
+
+            indices = np.argwhere(rates >= cutoff)
+            not_indices = np.argwhere(rates < cutoff)
+            newbfcs = basis.BasisFunctionConfiguration(self.n_species)
+            newbfcs.functions = [f for idx, f in enumerate(bfcs.functions) if idx in indices]
+            bfcs = newbfcs
+            if verbose:
[…]

### 2018-01-26 readdy_learn/analyze/basis.py

@@ -95,6 +95,10 @@ class BasisFunctionConfiguration(object):
     def functions(self):
         return self._basis_functions
 
+    @functions.setter
+    def functions(self, value):
+        self._basis_functions = value
+
     @property
     def n_basis_functions(self):
         return len(self._basis_functions)

### 2018-01-26 readdy_learn/example/regulation_network.py

@@ -315,7 +315,9 @@ def sample_lsq_rates(realizations, base_variance=.1, samples_per_variance=8, njo
 
     def worker(args):
         n, r = args
-        _np.random.seed(int(time.time()) + n*samples_per_variance + r)
+        seed = int(time.time()) + n*samples_per_variance + r
+        seed = seed % (2**32 - 1)
+        _np.random.seed(seed)
         return do_for_n_realizations(n)
 
     from collections import defaultdict

### 2018-01-26 readdy_learn/example/regulation_network.py
smooth

@@ -251,6 +251,61 @@ class RegulationNetwork(object):
                 traj.persist()
 
 
+def smooth(x, window_len=11, window='hanning'):
+    """smooth the data using a window with requested size.
+
+    This method is based on the convolution of a scaled window with the signal.
+    The signal is prepared by introducing reflected copies of the signal
+    (with the window size) in both ends so that transient parts are minimized
+    in the begining and end part of the output signal.
+
+    input:
+        x: the input signal
+        window_len: the dimension of the smoothing window; should be an odd integer
+        window: the type of window from 'flat', 'hanning', 'hamming', 'bartlett', 'blackman'
+            flat window will produce a moving average smoothing.
+
+    output:
+        the smoothed signal
+
+    example:
+
+    t=linspace(-2,2,0.1)
+    x=sin(t)+randn(len(t))*0.1
+    y=smooth(x)
+
+    see also:
+
+    numpy.hanning, numpy.hamming, numpy.bartlett, numpy.blackman, numpy.convolve
+    scipy.signal.lfilter
+
+    TODO: the window parameter could be the window itself if an array instead of a string
+    NOTE: length(output) != length(input), to correct this: return y[(window_len/2-1):-(window_len/2)] instead of just y.
+    """
+
+    if x.ndim != 1:
+        raise ValueError("smooth only accepts 1 dimension arrays.")
+
+    if x.size < window_len:
+        raise ValueError("Input vector needs to be bigger than window size.")
+
+    if window_len < 3:
+        return x
+
+    if not window in ['flat', 'hanning', 'hamming', 'bartlett', 'blackman']:
+        raise ValueError("Window is on of 'flat', 'hanning', 'hamming', 'bartlett', 'blackman'")
+
+    s = _np.r_[x[window_len - 1:0:-1], x, x[-2:-window_len - 1:-1]]
+    # print(len(s))
+    if window == 'flat':  # moving average
+        w = _np.ones(window_len, 'd')
+    else:
+        w = eval('numpy.' + window + '(window_len)')
+
+    y = _np.convolve(w / w.sum(), s, mode='valid')
+    return y
+
+
 def sample_lsq_rates(realizations, base_variance=.1, samples_per_variance=8, njobs=8, timestep=6e-3):
     if not isinstance(realizations, (list, tuple, _np.ndarray)):
         realizations = [realizations]

### 2018-01-26 docs/manuscript.tex
f
wip

@@
 
+	The second example deals with the second scenario of measurement and solves the minimization problem (\ref{method:minimizationproblem}) with regularization parameters being determined by cross validation.
+
+	In each of the considered examples the library of ansatz reactions $\Theta$ contains
+	\begin{colbox}{F8E0E0}
+		introduce ansatz functions
+	\end{colbox}
+
 	\subsection{Regression without regularization}
@@
 
-	In this example it is demonstrated that given enough data, the original model can be recovered in the limit of the noise variance $\sigma^2\to 0$.
+	In this example it is demonstrated that given enough data, the original model can be recovered in the limit of the noise variance $\sigma^2\to 0$. To this end, we first solve
 

### 2018-01-29 docs/manuscript.tex

@@
 
-	\section{Examples}
+	\section{Examples}\label{sec:examples}
 
@@
 
-	In each of the considered examples the library of ansatz reactions $\Theta$ contains
-	\begin{colbox}{F8E0E0}
-		introduce ansatz functions
-	\end{colbox}
+	In each of the considered examples the library $\Theta$ includes, among others, the reactions of the example system described in Section \ref{sec:generegulation}.
+The full set of reactions can be found in the Appendix, Table \ref{tab:reaction-library}.
 
@@
 
+	\section{Appendix}
+	\begin{table}
+		\centering
+		\begin{tabular}{rclcc}
+			\multicolumn{3}{c}{Reaction} & rate & description \\ \hline\noalign{\vskip .1cm}
+			$\ce{DNA_A}$ &$\ce{->}$& $\ce{DNA_A + mRNA_A }$ & $k_1 = 1.8$ & transcription of $\ce{mRNA_A}$\\
+			$\ce{mRNA_A}$ &$\ce{->}$& $\ce{mRNA_A + A}$ & $k_2 = 2.1$ & translation of $\ce{A}$ proteins\\
+			$\ce{mRNA_A}$ &$\ce{->}$& $\ce{\emptyset}$ & $k_3 = 1.3$ & $\ce{mRNA_A}$ decay\\
+			$\ce{A}$ &$\ce{->}$& $\ce{\emptyset }$ & $k_4 = 1.5$ & decay of $\ce{A}$ proteins\\
+			$\ce{DNA_B}$ &$\ce{->}$& $\ce{DNA_B + mRNA_B }$ & $k_5 = 2.2$ & transcription of $\ce{mRNA_B}$\\
+			$\ce{mRNA_B}$ &$\ce{->}$& $\ce{mRNA_B + B}$ & $k_6 = 2.0$ & translation of $\ce{B}$ proteins\\
+			$\ce{mRNA_B}$ &$\ce{->}$& $\ce{\emptyset }$ & $k_7 = 2.0$ & $\ce{mRNA_B}$ decay\\
+			$\ce{B}$ &$\ce{->}$& $\ce{\emptyset }$ & $k_8 = 2.5$ & decay of $\ce{B}$ proteins\\
+			$\ce{DNA_C}$ &$\ce{->}$& $\ce{DNA_C + mRNA_C }$ & $k_9 = 3.2$ & transcription of $\ce{mRNA_C}$\\
+			$\ce{mRNA_C}$ &$\ce{->}$& $\ce{mRNA_C + C}$ & $k_{10} = 3.0$ & translation of $\ce{C}$ proteins\\
+			$\ce{mRNA_C}$ &$\ce{->}$& $\ce{\emptyset }$ & $k_{11} = 2.3$ & $\ce{mRNA_C}$ decay\\
+			$\ce{C}$ &$\ce{->}$& $\ce{\emptyset }$ & $k_{12} = 2.5$ & decay of $\ce{C}$ proteins\\
+			$\ce{mRNA_A + A}$ &$\ce{->}$& $A$ & $k_{13} = 0$ & self regulation of $\ce{A}$ proteins\\
+			$\ce{mRNA_B + B}$ &$\ce{->}$& $B$ & $k_{14} = 0$ & self regulation of $\ce{B}$ proteins\\
+			$\ce{mRNA_C + C}$ &$\ce{->}$& $C$ & $k_{15} = 0$ & self regulation of $\ce{C}$ proteins\\
+			$\ce{mRNA_B + A}$ &$\ce{->}$& $A$ & $k_{16} = 0$ & cyclic regulation of $\ce{A}$ proteins\\
+			$\ce{mRNA_C + B}$ &$\ce{->}$& $B$ & $k_{17} = 0$ & cyclic regulation of $\ce{B}$ proteins\\
+			$\ce{mRNA_A + C}$ &$\ce{->}$& $C$ & $k_{18} = 0$ & cyclic regulation of $\ce{C}$ proteins\\
+			$\ce{mRNA_C + A}$ &$\ce{->}$& $A$ & $k_{16} = 6.0$ & cyclic regulation of $\ce{A}$ proteins\\
+			$\ce{mRNA_B + C}$ &$\ce{->}$& $C$ & $k_{17} = 4.0$ & cyclic regulation of $\ce{C}$ proteins\\
+			$\ce{mRNA_A + B}$ &$\ce{->}$& $B$ & $k_{18} = 3.0$ & cyclic regulation of $\ce{B}$ proteins\\
+			$\ce{mRNA_A + A}$ &$\ce{->}$& $\ce{mRNA_A}$ & $k_{19} = 0$ & artificial fusion\\
+			$\ce{mRNA_B + B}$ &$\ce{->}$& $\ce{mRNA_B}$ & $k_{20} = 0$ & artificial fusion\\
+			$\ce{mRNA_C + C}$ &$\ce{->}$& $\ce{mRNA_C}$ & $k_{21} = 0$ & artificial fusion\\
+			$\ce{mRNA_A + B}$ &$\ce{->}$& $\ce{mRNA_A}$ & $k_{22} = 0$ & artificial fusion\\
+			$\ce{mRNA_B + C}$ &$\ce{->}$& $\ce{mRNA_B}$ & $k_{23} = 0$ & artificial fusion\\
+			$\ce{mRNA_C + A}$ &$\ce{->}$& $\ce{mRNA_C}$ & $k_{24} = 0$ & artificial fusion\\
+			$\ce{mRNA_A + C}$ &$\ce{->}$& $\ce{mRNA_A}$ & $k_{25} = 0$ & artificial fusion\\
+			$\ce{mRNA_B + A}$ &$\ce{->}$& $\ce{mRNA_B}$ & $k_{26} = 0$ & artificial fusion\\
+			$\ce{mRNA_C + B}$ &$\ce{->}$& $\ce{mRNA_C}$ & $k_{27} = 0$ & artificial fusion\\
+			$\ce{A + A}$ &$\ce{->}$& $\ce{A}$ & $k_{28} = 0$ & $\ce{A}$ regulates $\ce{A}$\\
+			$\ce{B + B}$ &$\ce{->}$& $\ce{B}$ & $k_{29} = 0$ & $\ce{B}$ regulates $\ce{B}$\\
+			$\ce{C + C}$ &$\ce{->}$& $\ce{C}$ & $k_{30} = 0$ & $\ce{C}$ regulates $\ce{C}$\\
+			$\ce{B + A}$ &$\ce{->}$& $\ce{A}$ & $k_{31} = 0$ & artificial fusion between proteins\\
+			$\ce{C + B}$ &$\ce{->}$& $\ce{B}$ & $k_{32} = 0$ & artificial fusion between proteins\\
+			$\ce{A + C}$ &$\ce{->}$& $\ce{C}$ & $k_{33} = 0$ & artificial fusion between proteins\\
+			$\ce{C + A}$ &$\ce{->}$& $\ce{A}$ & $k_{34} = 0$ & artificial fusion between proteins\\
+			$\ce{B + C}$ &$\ce{->}$& $\ce{C}$ & $k_{35} = 0$ & artificial fusion between proteins\\
+			$\ce{A + B}$ &$\ce{->}$& $\ce{B}$ & $k_{36} = 0$ & artificial fusion between proteins\\
+			$\ce{A}$ &$\ce{->}$& $\ce{B}$ & $k_{37} = 0$ & artificial conversion between proteins\\
+			$\ce{B}$ &$\ce{->}$& $\ce{C}$ & $k_{38} = 0$ & artificial conversion between proteins\\
+			$\ce{C}$ &$\ce{->}$& $\ce{A}$ & $k_{39} = 0$ & artificial conversion between proteins\\
+			$\ce{A}$ &$\ce{->}$& $\ce{C}$ & $k_{40} = 0$ & artificial conversion between proteins\\
+			$\ce{C}$ &$\ce{->}$& $\ce{B}$ & $k_{41} = 0$ & artificial conversion between proteins\\
+			$\ce{B}$ &$\ce{->}$& $\ce{A}$ & $k_{42} = 0$ & artificial conversion between proteins
+		\end{tabular}
+		\caption{Full set of ansatz reactions $\Theta$ used in Section \ref{sec:examples}.}
+		\label{tab:reaction-library}
+	\end{table}
 \end{document}

### 2018-02-01 docs/manuscript.tex
manuscript update case 1

@@
 	\label{method:the-reactions}\end{align}
-	where $\textbf{X}_i$ denotes the $i$-th row in $X$.
+	where $\textbf{X}_i$ denotes the $i$-th row in $\textbf{X}$.
 Applying the concentration trajectory to the library yields $\Theta(\textbf{X})\in\mathbb{R}^{T\times S\times R}$.
@@
 
-	When presented with multiple trajectories one can extend the objective function by a sum over these and obtain a more general minimization problem that reads: Find $\hat{\Xi}$ such that
-	\begin{align}
-		\hat{\Xi} = \underset{\Xi}{\arg\min}\left( \sum_{\text{traj}}\frac{1}{2T_{\text{traj}}}\left\| \dot{\textbf{X}}_\text{traj} - \Theta(\textbf{X}_\text{traj})\Xi \right\|_F^2 + \alpha\lambda\|\Xi\|_1 + \alpha(1-\lambda)\|\Xi\|_2^2 \right)
-	\end{align}
-	subject to $\Xi\geq 0$.
+	When presented with multiple trajectories the data matrices $\mathbf{X}$ and $\dot{\mathbf{X}}$ can be stacked and inserted into the minimization problem.
+
 	For $\alpha=0$ the minimization problem reduces to constrained least-squares.
@@
 
-    \section{Scenarios of measurement}\label{sec:scenarios-of-measurement}
+	\section{Proof of concept}\label{sec:proof-of-concept}
 
 	The proposed estimation method is designed to find reaction networks and rates from experimental data.
-In the following measurement scenarios are considered that differ in how many time series are available and the level of noise that the data is subject to:
-    \begin{enumerate}
-        \item The initial state of the measurement apparatus can be reproduced exactly.
+In the following the method is applied to data generated from the example described in Section \ref{sec:generegulation}.
+
+	First it is assumed that the concentration data can be obtained without noise and it is shown that while the minimization problem (\ref{method:minimizationproblem}) without regularization, i.e., $\alpha=0$, yields rates that fit well to the input data, the original rates cannot be recovered while adding $L_1$ regularization improves the situation and the original model \textcolor{red}{almost} can be recovered.
+
+	In Section \ref{sec:case-2} the example data contains additive Gaussian noise.
+It is assumed that the initial state can be reproduced exactly and one has several measurements.
+It is shown that in the limit of small noise one can obtain a good fit and regularization helps in the reliability of the estimated reaction rates.
+
+	The last example is described in Section \ref{sec:case-3} and can be seen as an extension of the situation in Section \ref{sec:case-3} to multiple initial conditions.
+
+	%measurement scenarios are considered that differ in how many time series are available and the level of noise that the data is subject to:
+	%\begin{enumerate}
+	%	\item The initial state of the measurement apparatus can be reproduced exactly.
 There is only one set of generating reaction rates.
@@
 The outcome of this scenario is one time series of concentrations with a certain noise-level.
-        \item The process is observed multiple times.
+	%	\item The process is observed multiple times.
 The initial state of the measurement apparatus differs in each observation.
 The outcome of this scenario is multiple time series each with a certain noise level and different initial states.
-        \item The process is observed multiple times.
+	%\item The process is observed multiple times.
 The initial state of the measurement apparatus differs in each observation.
@@
 The outcome of this scenario is multiple time series each with a given noise level, different initial states and slightly perturbed underlying reaction rates.
-    \end{enumerate}
+	%\end{enumerate}
 
-	\section{Examples}\label{sec:examples}
+	\subsection{The noiseless case}\label{sec:case-1}
 
-	In the following two examples are presented. The first example considers data according to the first scenario of measurement as presented in Section \ref{sec:scenarios-of-measurement}.
-	The minimization problem (\ref{method:minimizationproblem}) with $\alpha=0$ is solved for different magnitudes of noise variance.
+	In this section noiseless data generated from the law of mass action system described in Section \ref{sec:generegulation} is applied to the minimization problem (\ref{method:minimizationproblem}).
+When setting the hyperparameter $\alpha=0$ one obtains least squares regression with a constraint on the non-negativity of the estimated reaction rates.
 
-	The second example deals with the second scenario of measurement and solves the minimization problem (\ref{method:minimizationproblem}) with regularization parameters being determined by cross validation.
+	\begin{figure}
+		\centering
+		\includegraphics[width=.7\textwidth]{./figures_tex/bubbles}
+		\caption{Generating and estimated reaction rates in the situation of Section \ref{sec:case-1}.
+The y and x axes contain reaction educts and products, respectively.
+The black squares denote the reactions with which the system was generated and their size indicates the respective reaction rate.
+The blue circles are least squares estimated reaction rates and the red rings depict rates which were obtained by some $\alpha>0$ in the minimization problem (\ref{method:minimizationproblem}).}
+		\label{fig:case-1-sparsity-pattern}
+	\end{figure}
 
-	In each of the considered examples the library $\Theta$ includes, among others, the reactions of the example system described in Section \ref{sec:generegulation}.
-The full set of reactions can be found in the Appendix, Table \ref{tab:reaction-library}.
+	When applying no regularization, one can observe in Figure \ref{fig:case-1-sparsity-pattern} that the sparsity pattern does not match the generating reaction network and the reaction responsible for the decay of $\mathrm{A}$ particles is completely ignored.
+For a suitable choice of the hyperparameter $\alpha > 0$ one can obtain a sparser solution and also recover the decay reaction as indicated by the red rings in the figure.
 
-	\subsection{Regression without regularization}
-	\label{sec:ex1}
+	\subsection{Multiple measurements at the same noise level}\label{sec:case-2}
 
-	In this example it is demonstrated that given enough data, the original model can be recovered in the limit of the noise variance $\sigma^2\to 0$. To this end, we first solve
+	\subsection{Multiple initial conditions}\label{sec:case-3}
 
+	\section{Conclusion}
 	\begin{colbox}{F8E0E0}
@@
 	\end{colbox}
-
-	Here we demonstrate that, with decreasing time step $\Delta t$ and $\alpha=0$, the estimated rates converge to the true rates if there is no ambiguity in the system.
-Let there be two species $A$ and $B$ which can unimolecularly convert into each other with macroscopic rates of $k_1=4.0$ and $k_2=0.5$, respectively.
-The initial amount of particles is set to $70$ of type $A$ and $0$ of type $B$.
-The law of mass action solution and Gillespie realization corresponding to this system are depicted in Figure \ref{ex1:concentration-curves}.
-
-	\begin{figure}
-		\begin{center}
-			\includegraphics[width=.5\textwidth]{./figures_tex/aux_ex1_concentration_curves}
-		\end{center}
-		\caption{The law of mass action solution and a Gillespie realization of the example given in Section \ref{sec:ex1}.}
-		\label{ex1:concentration-curves}
-	\end{figure}
-
-	For each $\Delta t$, eight different realizations with identical starting conditions are averaged into a single trajectory, of which then the rates for reactions $A\to B$ and $B\to A$ are estimated.
-This procedure was repeated $30$ times.
-In Figure \ref{ex1:rates-convergence}, the average estimates with their standard deviations are shown.
-One can see that, with decreasing $\Delta t$, the estimates become better and the rates can be recovered.
-Further it can be observed that in this case a time step of $\Delta t\approx 10^{-2}$ suffices.
-
-	\begin{figure}
-		\begin{center}
-			\includegraphics[width=.5\textwidth]{./figures_tex/aux_rates_convergence}
-		\end{center}
-		\caption{Convergence of estimated rates with decreasing $\Delta t$.}
-		\label{ex1:rates-convergence}
-	\end{figure}
-
-	\subsection{Regression with regularization}
-	\label{sec:ex2}
-	In this example there are four species $A$, $B$, $C$, and $D$.
-The system was generated with reactions $A\xrightarrow{k_1}D$, $D\xrightarrow{k_2}A$, $D\xrightarrow{k_3}B$, $A+B\xrightarrow{k_4}C$, and $C\xrightarrow{k_5}D+B$, where the rates are given in Table \ref{tab:estimated-rates}.
-These reactions are inserted into the ansatz library $\Theta (\cdot)$.
-Additionally the reactions $A\leftrightarrow B$, $A\leftrightarrow C$, and $A+C\leftrightarrow D$ are inserted into the library.
-
-	\begin{figure}
-		\centering
-		\begin{subfigure}[b]{.45\textwidth}
-			\includegraphics[width=\textwidth]{./figures_tex/cv_concentrations_train}
-			\caption{Train data set.}
-		\end{subfigure}
-		~
-		\begin{subfigure}[b]{.45\textwidth}
-			\includegraphics[width=\textwidth]{./figures_tex/cv_concentrations_test}
-			\caption{Test data set.}
-		\end{subfigure}
-		\caption{Train and test data sets for the example given in Section \ref{sec:ex2}.}
-		\label{ex2:test-train-concentration-curves}
-	\end{figure}
-
-	The problem now is to find a suitable $\alpha$ such that a parsimonious solution is yielded.
-To this end, cross-validation is applied.
-The trajectories that are used as training and test data set are depicted in Figure \ref{ex2:test-train-concentration-curves}.
-\textcolor{red}{Perhaps use several test trajectories with different initial conditions to get something like a confidence interval?} As an indicator to the estimation error one can use the scoring function
-	\begin{align}
-		f_\alpha = \frac{1}{T}\|\dot{\mathbf{X}}_\text{test} - \Theta (\textbf{X}_\text{test}) \Xi_\text{train}\|_F^2,
-		\label{ex2:score-function}
-	\end{align}
-	where the best possible score is $f_\alpha = 0$ and larger values indicate worse models.
-	For a selection of hyperparameters $\alpha\in [0, 100]$, the model parameters $\Xi$ are estimated and the scoring function (\ref{ex2:score-function}) evaluated, yielding a dependency of the score on $\alpha$ as depicted in Figure \ref{ex2:score-graph}.
-
-	\begin{figure}
-		\begin{center}
-			\includegraphics[width=.5\textwidth]{./figures_tex/cv_score}
-		\end{center}
-		\caption{Score function (\ref{ex2:score-function}) evaluated for a range of $\alpha$ values.}
-		\label{ex2:score-graph}
-	\end{figure}
-
-	One can see that the minimum score is achieved at
-	\begin{align}
-		\hat{\alpha} = \underset{\alpha\in[0,100]}{\arg\min}f_\alpha\approx 10.
-	\end{align}
-	When evaluating the minimization problem (\ref{method:minimizationproblem}) with $\hat{\alpha}$ as hyperparameter for the regularizer and applying a cut-off of $10^{-10}$ to the yielded parameters, one obtains the simpler set of reactions $A\xrightarrow{\hat{k}_1}D$, $D\xrightarrow{\hat{k}_3} B$, and $A+B\xrightarrow{\hat{k}_4}C$ with rates as in Table \ref{tab:estimated-rates}.
-
-	\begin{table}
-		\centering
-		\begin{tabular}{ccc}
-			Reaction & original rate & estimated rate \\ \hline\noalign{\vskip .1cm}
-			$A \to D$ & $k_1 = 2\cdot 10^{-2}$ & $\hat{k}_1 \approx 3.5\cdot 10^{-3}$\\
-			$D \to A$ & $k_2 = 2\cdot 10^{-2}$ & $\hat{k}_2 = 0$ \\
-			$D\to B$ & $k_3 = 1\cdot 10^{-2}$ & $\hat{k}_3\approx 9.1 \cdot 10^{-3}$ \\
-			$A+B\to C$ & $k_4 = 1\cdot 10^{-4}$ & $\hat{k}_4\approx 8.7 \cdot 10^{-5}$ \\
-			$C \to D+B$ & $k_5 = 1\cdot 10^{-4}$ & $\hat{k}_5 = 0$
-		\end{tabular}
-		\caption{Original and estimated reaction rates for the example in Section \ref{sec:ex2}.}
-		\label{tab:estimated-rates}
-	\end{table}
-
-	One can see in Figure \ref{ex2:result} that with this simpler set of reactions the observed concentration curves can be explained as well.
-Further, it can be argued that the second reaction of the original model $D\xrightarrow{k_2}A$ is compensated by a smaller rate $\hat{k}_1 < k_1$ and that the fifth reaction of the generating model can be neglected altogether in the observed timescale as its rate is two orders of magnitude smaller than the other unimolecular reactions' rates.
-	The parameters $k_3$ and $k_4$ are approximately recovered by $\hat{k}_3$ and $\hat{k}_4$, respectively.
-
-	\begin{figure}
-		\begin{center}
-			\includegraphics[width=.5\textwidth]{./figures_tex/cv_expected_estimated_concentrations}
-		\end{center}
-		\caption{Integrated law of mass action equations (\ref{method:the-system}) for the sparse estimated parameters and the corresponding Gillespie realization for the example in Section \ref{sec:ex2}.}
-		\label{ex2:result}
-	\end{figure}
-
-	\section{Conclusion}
 	In this work we have successfully applied and extended the SINDy method to not only parsimoniously detect potentially nonlinear terms in a dynamical system from noisy data, but also yield, in this case, a sparse set of rates with respect to generating reactions (\ref{method:the-reactions}).
@@
 	\bibliography{bibliography.bib}
-
-	\section{Appendix}
-	\begin{table}
+	\newpage	
+	\section{Appendix}\label{sec:appendix}
+	\begin{table}[h]
 		\centering
+		\scalebox{0.7}{
 			\begin{tabular}{rclcc}
@@
 			\end{tabular}
+		}
 		\caption{Full set of ansatz reactions $\Theta$ used in Section \ref{sec:examples}.}

### 2018-02-08 readdy_learn/analyze/validation.py
seed

@@ -141,7 +141,7 @@ def get_scores(result, alpha, l1_ratio):
     return [r['score'] for r in result if r['alpha'] == alpha and r['l1_ratio'] == l1_ratio]
 
 
-def plot_validation_result(result):
+def plot_validation_result(result, ax=None):
     import matplotlib.pyplot as plt
     l1_ratios = get_distinct_l1_ratios(result)
     alphas = get_distinct_alphas(result)
@@ -152,4 +152,7 @@ def plot_validation_result(result):
         for ix, alpha in enumerate(alphas):
             ys[ix] = _np.mean(get_scores(result, alpha, l1_ratio))
             yerr[ix] = _np.std(get_scores(result, alpha, l1_ratio))
-        plt.errorbar(xs, ys, yerr=yerr, label=r'$\lambda={:.3f}$'.format(l1_ratio))
+        if ax is None:
+            plt.errorbar(xs, ys, yerr=yerr, label=r'$\lambda={:.3f}$'.format(l1_ratio))
+        else:
+            ax.errorbar(xs, ys, yerr=yerr, label=r'$\lambda={:.3f}$'.format(l1_ratio))

### 2018-02-08 readdy_learn/analyze/validation.py
seed

@@ -152,7 +152,10 @@ def plot_validation_result(result, ax=None):
         for ix, alpha in enumerate(alphas):
             ys[ix] = _np.mean(get_scores(result, alpha, l1_ratio))
             yerr[ix] = _np.std(get_scores(result, alpha, l1_ratio))
+        ymin_ix = _np.argmin(ys)
         if ax is None:
             plt.errorbar(xs, ys, yerr=yerr, label=r'$\lambda={:.3f}$'.format(l1_ratio))
+            plt.plot(xs[ymin_ix], ys[ymin_ix], 'o')
         else:
             ax.errorbar(xs, ys, yerr=yerr, label=r'$\lambda={:.3f}$'.format(l1_ratio))
+            ax.plot(xs[ymin_ix], ys[ymin_ix], 'o')

### 2018-02-08 readdy_learn/analyze/analyze.py
wip

@@ -9,6 +9,7 @@ import readdy_learn.analyze.generate as generate
 import readdy_learn.analyze.tools as tools
 import readdy_learn.analyze.basis as basis
 import readdy_learn.sample_tools as sample_tools
+import readdy_learn.analyze.interface as _interface
 
 
 def estimate_noise_variance(xs, ys):
@@ -125,7 +126,7 @@ def obtain_derivative(traj, alpha=1000, atol=1e-10, tol=1e-10, maxit=1000, alpha
         return [], traj
 
 
-class ReactionAnalysis(object):
+class ReactionAnalysis(_interface.ReactionAnalysisObject):
     def __init__(self, bfc, desired_rates, initial_states, set_up_system, recompute=False, recompute_traj=False,
                  fname_prefix="", fname_postfix="", n_species=4, timestep=5e-4,
                  interp_degree='regularized_derivative', ld_derivative_config=None, species_names=None):

### 2018-02-08 readdy_learn/analyze/cross_validation.py

@@ -1,0 +1,95 @@
+import time as _time
+
+import numpy as _np
+import itertools as _itertools
+
+import readdy_learn.analyze.interface as _interface
+import readdy_learn.analyze.progress as _progress
+
+import pathos.multiprocessing as _multiprocessing
+
+class CrossValidation(object):
+
+    def __init__(self, regulation_network: _interface.AnalysisObjectGenerator, show_progress: bool = True,
+                 njobs: int = 8):
+        self._regulation_network = regulation_network
+        self._show_progress = show_progress
+        self._njobs = njobs
+        self.result_ = None
+
+    def _cross_validate(self, args):
+        alpha, l1_ratio, cutoff, i = args
+
+        hash = (i + int(_time.time())) % (2 ** 32 - 1)
+        _np.random.seed(hash)
+
+        regulation_network = self.regulation_network
+        analysis_train = self.regulation_network.generate_analysis_object(None, None)
+        for i in range(len(regulation_network.initial_states)):
+            analysis_train.generate_or_load_traj_lma(i, regulation_network.target_time,
+                                                     noise_variance=regulation_network.noise_variance,
+                                                     realizations=1)
+        regulation_network.compute_gradient_derivatives(analysis_train, persist=False)
+
+
+
+    def cross_validate(self, alphas, lambdas, cutoffs=None, realizations: int = 10):
+        if cutoffs is None:
+            cutoffs = _np.zeros((1,), dtype=float)
+        else:
+            cutoffs = _np.array(cutoffs).squeeze()
+        alphas = _np.array(alphas).squeeze()
+        lambdas = _np.array(lambdas).squeeze()
+
+        assert _np.alltrue(lambdas <= 1), "some lambdas were greater than 1"
+        assert _np.alltrue(lambdas >= 0), "some lambdas were smaller than 0"
+
+        print("validating across grid with {} alphas, {} lambdas, {} cutoffs with {} realizations"
+              .format(len(alphas), len(lambdas), len(cutoffs), realizations))
+
+        params = _itertools.product(alphas, lambdas, cutoffs)
+        params = [(p[0], p[1], p[2], j + i * realizations) for i, p in enumerate(params)
+                  for j, _ in enumerate(range(realizations))]
+
+        result = []
+
+        progress = None
+        if self.show_progress:
+            progress = _progress.Progress(len(params),
+                                          label="validation", nstages=1)
+
+        with _multiprocessing.Pool(processes=self.njobs) as p:
[…]

### 2018-02-08 readdy_learn/analyze/estimator.py

@@ -31,7 +31,6 @@ import numpy as np
 import scipy.optimize as so
 from pathos.multiprocessing import Pool
 from readdy_learn.analyze_tools import opt
-import readdy_learn.analyze.tools as tools
 import readdy_learn.analyze.derivative as deriv
 from sklearn.linear_model.base import BaseEstimator
 from sklearn.model_selection import KFold, LeaveOneOut

### 2018-02-08 readdy_learn/analyze/interface.py

@@ -1,9 +1,144 @@
-import readdy_learn.analyze.analyze as _ana
+import typing as _typing
+import numpy as _np
+
+
+class ReactionAnalysisObject(object):
+
+    @property
+    def bfc(self):
+        raise NotImplementedError("please implement me")
+
+    @bfc.setter
+    def bfc(self, value):
+        raise NotImplementedError("please implement me")
+
+    @property
+    def recompute(self):
+        raise NotImplementedError("please implement me")
+
+    @recompute.setter
+    def recompute(self, value):
+        raise NotImplementedError("please implement me")
+
+    @property
+    def set_up_system(self):
+        raise NotImplementedError("please implement me")
+
+    @set_up_system.setter
+    def set_up_system(self, value):
+        raise NotImplementedError("please implement me")
+
+    @property
+    def recompute_traj(self):
+        raise NotImplementedError("please implement me")
+
+    @recompute_traj.setter
+    def recompute_traj(self, value):
+        raise NotImplementedError("please implement me")
+
+    @property
+    def fname_prefix(self):
+        raise NotImplementedError("please implement me")
+
+    @fname_prefix.setter
+    def fname_prefix(self, value):
+        raise NotImplementedError("please implement me")
+
+    @property
+    def fname_postfix(self):
+        raise NotImplementedError("please implement me")
+
+    @fname_postfix.setter
+    def fname_postfix(self, value):
+        raise NotImplementedError("please implement me")
+
+    @property
+    def n_species(self):
+        raise NotImplementedError("please implement me")
+
+    @n_species.setter
+    def n_species(self, value):
+        raise NotImplementedError("please implement me")
+
+    @property
+    def species_names(self):
+        raise NotImplementedError("please implement me")
+
+    @property
+    def desired_rates(self):
+        raise NotImplementedError("please implement me")
+
+    @desired_rates.setter
+    def desired_rates(self, value):
+        raise NotImplementedError("please implement me")
+
+    @property
+    def best_alphas(self):
+        raise NotImplementedError("please implement me")
+
+    @property
+    def initial_states(self):
+        raise NotImplementedError("please implement me")
+
+    @initial_states.setter
+    def initial_states(self, value):
+        raise NotImplementedError("please implement me")
+
+    @property
+    def trajs(self):
+        raise NotImplementedError("please implement me")
+
[…]

### 2018-02-08 readdy_learn/analyze/tools.py

@@ -37,6 +37,7 @@ from pathos.multiprocessing import Pool
 
 import readdy_learn.analyze.derivative as deriv
 from readdy_learn import analyze_tools as opt
+import readdy_learn.analyze.interface as _interface
 
 def get_count_trajectory(fname, cache_fname=None):
     if cache_fname and os.path.exists(cache_fname):
@@ -83,16 +84,25 @@ class TrajectoryConfig(object):
         self.reactions = ioutils.get_reactions(file_path)
 
 
-class Trajectory(object):
-    def __init__(self, counts, time_step, interpolation_degree=10, verbose=True,
-                 ld_derivative_atol=1e-7, ld_derivative_rtol=1e-10, ld_derivative_alpha=3e-3,
-                 ld_derivative_solver='lgmres', ld_derivative_maxit=10000, ld_derivative_linalg_solver_maxit=100,
-                 ld_derivative_linalg_solver_tol=1e-4, ld_derivative_use_preconditioner=True,
-                 fname=None):
+class Trajectory(_interface.ReactionLearnDataContainer):
+    def __init__(self, counts, time_step, interpolation_degree=10, verbose=True, ld_derivative_atol=1e-7,
+                 ld_derivative_rtol=1e-10, ld_derivative_alpha=3e-3, ld_derivative_solver='lgmres',
+                 ld_derivative_maxit=10000, ld_derivative_linalg_solver_maxit=100, ld_derivative_linalg_solver_tol=1e-4,
+                 ld_derivative_use_preconditioner=True, fname=None):
         if isinstance(counts, str):
             fname = counts
             counts = None
-        self._counts = counts
+        if fname is not None and os.path.exists(fname):
+            archive = np.load(fname)
+            counts = archive['counts']
+            super().__init__(counts)
+            for s in range(self.n_species):
+                if 'dcounts_dt_{}'.format(s) in archive.keys():
+                    self._separate_derivs[s] = archive['dcounts_dt_{}'.format(s)]
+            if 'dt' in archive.keys():
+                self._time_step = archive['dt']
+        else:
+            super().__init__(counts)
         self._box_size = [15., 15., 15.]
         self._time_step = time_step
         self._thetas = []
@@ -110,14 +120,7 @@ class Trajectory(object):
                                      'alpha': ld_derivative_alpha, 'solver': ld_derivative_solver,
                                      'maxit': ld_derivative_maxit, 'linalg_solver_maxit': ld_derivative_linalg_solver_maxit,
[…]

### 2018-02-08 readdy_learn/analyze/cross_validation.py
CV

@@ -1,21 +1,64 @@
 import time as _time
+import typing as _typing
 
 import numpy as _np
 import itertools as _itertools
 
 import readdy_learn.analyze.interface as _interface
 import readdy_learn.analyze.progress as _progress
+import readdy_learn.analyze.estimator as _estimator
 
 import pathos.multiprocessing as _multiprocessing
 
+from sklearn.model_selection import KFold as _KFold
+
+_TrajList = _typing.List[_interface.ReactionLearnDataContainer]
+
+
 class CrossValidation(object):
 
-    def __init__(self, regulation_network: _interface.AnalysisObjectGenerator, show_progress: bool = True,
-                 njobs: int = 8):
-        self._regulation_network = regulation_network
+    def __init__(self, trajs: _TrajList, bfc, _n_splits: int = 10, show_progress: bool = True, njobs: int = 8):
         self._show_progress = show_progress
         self._njobs = njobs
         self.result_ = None
+
+        self._counts = _np.stack([t.counts for t in trajs], axis=0)
+        self._dcounts_dt = _np.stack([t.dcounts_dt for t in trajs], axis=0)
+        self._n_splits = _n_splits
+        self._bfc = bfc
+
+    def _obtain_trajs_subset(self, indices: _np.ndarray):
+        counts = self._counts[indices]
+        dcounts_dt = self._dcounts_dt[indices]
+
+        container = _interface.ReactionLearnDataContainer(counts)
+        container.dcounts_dt = dcounts_dt
+        return container
+
+    def _solve(self, traj, alpha, l1_ratio, tol=1e-16):
+        initial_value = _np.zeros((self._bfc.n_basis_functions,), dtype=_np.float64)
+        estimator = _estimator.ReaDDyElasticNetEstimator(traj, self._bfc, alpha=alpha, l1_ratio=l1_ratio,
+                                                         maxiter=30000, method='SLSQP', verbose=False, approx_jac=False,
+                                                         options={'ftol': tol}, rescale=False,
+                                                         init_xi=initial_value,
+                                                         constrained=True)
+
+        estimator.fit(None)
+        if estimator.success_:
+            rates = estimator.coefficients_
+            return rates
+        else:
+            raise ValueError('*_*: {}, {}'.format(estimator.result_.status, estimator.result_.message))
+
+    def _score(self, traj, rates):
[…]

### 2018-02-08 readdy_learn/analyze/estimator.py

@@ -31,10 +31,11 @@ import numpy as np
 import scipy.optimize as so
 from pathos.multiprocessing import Pool
 from readdy_learn.analyze_tools import opt
-import readdy_learn.analyze.derivative as deriv
 from sklearn.linear_model.base import BaseEstimator
-from sklearn.model_selection import KFold, LeaveOneOut
+from sklearn.model_selection import KFold
 from sklearn.model_selection import TimeSeriesSplit
+
+import readdy_learn.analyze.derivative as deriv
 
 _epsilon = np.sqrt(np.finfo(float).eps)
 

### 2018-02-08 readdy_learn/analyze/interface.py

@@ -135,6 +135,10 @@ class ReactionLearnDataContainer(object):
     def n_species(self) -> int:
         return self.counts.shape[1]
 
+    @property
+    def n_time_steps(self):
+        return self.counts.shape[0]
+
 
 class AnalysisObjectGenerator(object):
 

### 2018-02-08 readdy_learn/analyze/tools.py

@@ -338,10 +338,6 @@ class Trajectory(_interface.ReactionLearnDataContainer):
     def counts(self):
         return self._counts
 
-    @property
-    def n_time_steps(self):
-        return self.counts.shape[0]
-
     @counts.setter
     def counts(self, value):
         self._counts = value

### 2018-02-08 readdy_learn/analyze/validation.py

@@ -155,7 +155,7 @@ def plot_validation_result(result, ax=None):
         ymin_ix = _np.argmin(ys)
         if ax is None:
             plt.errorbar(xs, ys, yerr=yerr, label=r'$\lambda={:.3f}$'.format(l1_ratio))
-            plt.plot(xs[ymin_ix], ys[ymin_ix], 'o')
+            plt.plot(xs[ymin_ix], ys[ymin_ix], 'o', color='k')
         else:
             ax.errorbar(xs, ys, yerr=yerr, label=r'$\lambda={:.3f}$'.format(l1_ratio))
-            ax.plot(xs[ymin_ix], ys[ymin_ix], 'o')
+            ax.plot(xs[ymin_ix], ys[ymin_ix], 'o', color='k')

### 2018-02-08 readdy_learn/analyze/cross_validation.py
CV

@@ -71,7 +71,7 @@ class CrossValidation(object):
 
         splitter = _KFold(n_splits=self.n_splits)
         scores = []
-        for train, test in splitter.split(indices):
+        for train, test in splitter.split(indices[:, _np.newaxis]):
             train_traj = self._obtain_trajs_subset(train)
             test_traj = self._obtain_trajs_subset(test)
             tolerances_to_try = [1e-16, 1e-15, 1e-14, 1e-13, 1e-12, 1e-11, 1e-10, 1e-9, 1e-8]

### 2018-02-08 readdy_learn/analyze/cross_validation.py
CV

@@ -12,12 +12,14 @@ import pathos.multiprocessing as _multiprocessing
 
 from sklearn.model_selection import KFold as _KFold
 
-_TrajList = _typing.List[_interface.ReactionLearnDataContainer]
+_TrajList = _typing.Union[_typing.List[_interface.ReactionLearnDataContainer], _interface.ReactionLearnDataContainer]
 
 
 class CrossValidation(object):
 
     def __init__(self, trajs: _TrajList, bfc, _n_splits: int = 10, show_progress: bool = True, njobs: int = 8):
+        if not isinstance(trajs, (list, tuple)):
+            trajs = [trajs]
         self._show_progress = show_progress
         self._njobs = njobs
         self.result_ = None
@@ -71,7 +73,7 @@ class CrossValidation(object):
 
         splitter = _KFold(n_splits=self.n_splits)
         scores = []
-        for train, test in splitter.split(indices[:, _np.newaxis]):
+        for train, test in splitter.split(range(0, n_steps_total)):
             train_traj = self._obtain_trajs_subset(train)
             test_traj = self._obtain_trajs_subset(test)
             tolerances_to_try = [1e-16, 1e-15, 1e-14, 1e-13, 1e-12, 1e-11, 1e-10, 1e-9, 1e-8]

### 2018-02-08 readdy_learn/analyze/cross_validation.py
CV

@@ -73,7 +73,7 @@ class CrossValidation(object):
 
         splitter = _KFold(n_splits=self.n_splits)
         scores = []
-        for train, test in splitter.split(range(0, n_steps_total)):
+        for train, test in splitter.split(indices):
             train_traj = self._obtain_trajs_subset(train)
             test_traj = self._obtain_trajs_subset(test)
             tolerances_to_try = [1e-16, 1e-15, 1e-14, 1e-13, 1e-12, 1e-11, 1e-10, 1e-9, 1e-8]

### 2018-02-08 readdy_learn/analyze/cross_validation.py
CV

@@ -69,11 +69,11 @@ class CrossValidation(object):
         _np.random.seed(hash)
 
         n_steps_total = self._counts.shape[0]
-        indices = _np.arange(n_steps_total)
+        print("n_steps_total = {}".format(n_steps_total))
 
         splitter = _KFold(n_splits=self.n_splits)
         scores = []
-        for train, test in splitter.split(indices):
+        for train, test in splitter.split(_np.arange(n_steps_total)):
             train_traj = self._obtain_trajs_subset(train)
             test_traj = self._obtain_trajs_subset(test)
             tolerances_to_try = [1e-16, 1e-15, 1e-14, 1e-13, 1e-12, 1e-11, 1e-10, 1e-9, 1e-8]

### 2018-02-08 readdy_learn/analyze/cross_validation.py
CV

@@ -69,7 +69,7 @@ class CrossValidation(object):
         _np.random.seed(hash)
 
         n_steps_total = self._counts.shape[0]
-        print("n_steps_total = {}".format(n_steps_total))
+        print("n_steps_total = {}, counts shape: {}".format(n_steps_total, self._counts.shape))
 
         splitter = _KFold(n_splits=self.n_splits)
         scores = []

### 2018-02-08 readdy_learn/analyze/cross_validation.py
CV

@@ -24,8 +24,8 @@ class CrossValidation(object):
         self._njobs = njobs
         self.result_ = None
 
-        self._counts = _np.stack([t.counts for t in trajs], axis=0)
-        self._dcounts_dt = _np.stack([t.dcounts_dt for t in trajs], axis=0)
+        self._counts = _np.stack([t.counts for t in trajs], axis=0).squeeze()
+        self._dcounts_dt = _np.stack([t.dcounts_dt for t in trajs], axis=0).squeeze()
         self._n_splits = _n_splits
         self._bfc = bfc
 
@@ -69,7 +69,6 @@ class CrossValidation(object):
         _np.random.seed(hash)
 
         n_steps_total = self._counts.shape[0]
-        print("n_steps_total = {}, counts shape: {}".format(n_steps_total, self._counts.shape))
 
         splitter = _KFold(n_splits=self.n_splits)
         scores = []

### 2018-02-08 readdy_learn/analyze/cross_validation.py
CV

@@ -15,6 +15,17 @@ from sklearn.model_selection import KFold as _KFold
 _TrajList = _typing.Union[_typing.List[_interface.ReactionLearnDataContainer], _interface.ReactionLearnDataContainer]
 
 
+def get_cross_validation_object(regulation_network: _interface.AnalysisObjectGenerator):
+    analysis = regulation_network.generate_analysis_object(None, None)
+    for i in range(len(regulation_network.initial_states)):
+        analysis.generate_or_load_traj_lma(i, regulation_network.target_time,
+                                           noise_variance=regulation_network.noise_variance,
+                                           realizations=1)
+    regulation_network.compute_gradient_derivatives(analysis, persist=False)
+    traj = analysis.get_traj(0)
+    cv = CrossValidation([traj], regulation_network.get_bfc())
+    return cv
+
 class CrossValidation(object):
 
     def __init__(self, trajs: _TrajList, bfc, _n_splits: int = 10, show_progress: bool = True, njobs: int = 8):
@@ -75,6 +86,10 @@ class CrossValidation(object):
         for train, test in splitter.split(_np.arange(n_steps_total)):
             train_traj = self._obtain_trajs_subset(train)
             test_traj = self._obtain_trajs_subset(test)
+            assert train_traj.n_time_steps == len(train)
+            assert test_traj.n_time_steps == len(test)
+            assert train_traj.dcounts_dt.shape[0] == len(train)
+            assert test_traj.dcounts_dt.shape[0] == len(test)
             tolerances_to_try = [1e-16, 1e-15, 1e-14, 1e-13, 1e-12, 1e-11, 1e-10, 1e-9, 1e-8]
 
             rates = None

### 2018-02-08 readdy_learn/analyze/interface.py

@@ -97,6 +97,9 @@ class ReactionAnalysisObject(object):
     def score(self, test_traj_n, rates):
         raise NotImplementedError("please implement me")
 
+    def get_traj(self, n:int):
+        raise NotImplementedError("please implement me")
+
 
 class ReactionLearnDataContainer(object):
 
@@ -178,4 +181,7 @@ class AnalysisObjectGenerator(object):
 
     @desired_rates.setter
     def desired_rates(self, value):
+        raise NotImplementedError("desired rates please")
+
+    def get_bfc(self):
         raise NotImplementedError("desired rates please")

### 2018-02-08 readdy_learn/analyze/cross_validation.py
CV

@@ -10,7 +10,7 @@ import readdy_learn.analyze.estimator as _estimator
 
 import pathos.multiprocessing as _multiprocessing
 
-from sklearn.model_selection import KFold as _KFold
+from sklearn.model_selection import ShuffleSplit as _ShuffleSplit
 
 _TrajList = _typing.Union[_typing.List[_interface.ReactionLearnDataContainer], _interface.ReactionLearnDataContainer]
 
@@ -81,7 +81,7 @@ class CrossValidation(object):
 
         n_steps_total = self._counts.shape[0]
 
-        splitter = _KFold(n_splits=self.n_splits)
+        splitter = _ShuffleSplit(n_splits=self.n_splits)
         scores = []
         for train, test in splitter.split(_np.arange(n_steps_total)):
             train_traj = self._obtain_trajs_subset(train)

### 2018-02-08 readdy_learn/analyze/cross_validation.py
Derp

@@ -26,6 +26,7 @@ def get_cross_validation_object(regulation_network: _interface.AnalysisObjectGen
     cv = CrossValidation([traj], regulation_network.get_bfc())
     return cv
 
+
 class CrossValidation(object):
 
     def __init__(self, trajs: _TrajList, bfc, _n_splits: int = 10, show_progress: bool = True, njobs: int = 8):
@@ -76,8 +77,8 @@ class CrossValidation(object):
     def _cross_validate(self, args):
         alpha, l1_ratio, cutoff, i = args
 
-        hash = (i + int(_time.time())) % (2 ** 32 - 1)
-        _np.random.seed(hash)
+        seed = (i + int(_time.time())) % (2 ** 32 - 1)
+        _np.random.seed(seed)
 
         n_steps_total = self._counts.shape[0]
 

### 2018-02-08 docs/manuscript.tex
Updated

@@
 
-	First it is assumed that the concentration data can be obtained without noise and it is shown that while the minimization problem (\ref{method:minimizationproblem}) without regularization, i.e., $\alpha=0$, yields rates that fit well to the input data, the original rates cannot be recovered while adding $L_1$ regularization improves the situation and the original model \textcolor{red}{almost} can be recovered.
+	First it is assumed that the concentration data can be obtained without noise and it is shown that while the minimization problem (\ref{method:minimizationproblem}) without regularization, i.e., $\alpha=0$, yields rates that fit well to the input data, the original rates cannot be recovered while adding $L_1$ regularization and a certain cutoff improves the situation and the original model can be recovered.
 
@@
 		\centering
-		\includegraphics[width=.7\textwidth]{./figures_tex/bubbles}
+		\includegraphics[width=.9\textwidth]{./figures_tex/bubbles}
 		\caption{Generating and estimated reaction rates in the situation of Section \ref{sec:case-1}.
 The y and x axes contain reaction educts and products, respectively.
-The black squares denote the reactions with which the system was generated and their size indicates the respective reaction rate.
-The blue circles are least squares estimated reaction rates and the red rings depict rates which were obtained by some $\alpha>0$ in the minimization problem (\ref{method:minimizationproblem}).}
+The black outlines denote the reactions with which the system was generated and the area of the circles scales linearly with the reaction rate.
+The blue circles are least squares estimated reaction rates and the orange circles depict rates which were obtained by some $\alpha>0$ in the minimization problem (\ref{method:minimizationproblem}).
+If a certain rate was estimated in both cases, two wedges instead of once circle as displayed.}
 		\label{fig:case-1-sparsity-pattern}
@@
 	When applying no regularization, one can observe in Figure \ref{fig:case-1-sparsity-pattern} that the sparsity pattern does not match the generating reaction network and the reaction responsible for the decay of $\mathrm{A}$ particles is completely ignored.
-For a suitable choice of the hyperparameter $\alpha > 0$ one can obtain a sparser solution and also recover the decay reaction as indicated by the red rings in the figure.
+For a suitable choice of the hyperparameter $\alpha > 0$ and a cutoff one can obtain a sparser solution and also recover the decay reaction as indicated by the red rings in the figure.
 
@@
 
+	\begin{itemize}
+		\item Same setup as in Section \ref{sec:case-1}
+		\item Cross validation with a shuffle split and 3 folds
+		\item comparison to LSQ: LSQ unreliable
+		\item in the limit of small noise one gets a good fit
+		\item obtain cutoff in noise free case, compute failure rate (number of false positives and negatives)
+	\end{itemize}
+
 	\subsection{Multiple initial conditions}\label{sec:case-3}
+
+	\begin{itemize}
+		\item extend situation of Section \ref{sec:case-3} to multiple initial conditions
+		\item compute failure rate
+	\end{itemize}
 
@@
 	\section{Appendix}\label{sec:appendix}
+	\begin{colbox}{F8E0E0}
+		the reaction table needs an update
+	\end{colbox}
 	\begin{table}[h]

### 2018-02-08 readdy_learn/analyze/cross_validation.py
Updated

@@ -82,7 +82,7 @@ class CrossValidation(object):
 
         n_steps_total = self._counts.shape[0]
 
-        splitter = _ShuffleSplit(n_splits=self.n_splits)
+        splitter = _ShuffleSplit(n_splits=self.n_splits, test_size=.3)
         scores = []
         for train, test in splitter.split(_np.arange(n_steps_total)):
             train_traj = self._obtain_trajs_subset(train)

### 2018-08-13 docs/manuscript.tex
wip

@@
 	In our application we, as opposed to the original method, do not only look for macroscopic rates of net species change but investigate the specific reactions that might have lead to the observations.
-	We demonstrate the algorithm on two toy problems - one problem showing that when there is no ambiguity in the system, one converges to the correct rates with increasing resolution of concentration fluctuations and one in which we compute a sparse reaction network for given data.
+	We demonstrate the algorithm on three toy problems: When there is no noise in the data we can find all relevant processes of the ground truth.
+If there is noise in the data we converge to the correct reaction network and rates with decreasing noise.
+The last toy problem deals with the case that there are two realizations with different initial conditions, for which can also show convergence to the correct model with decreasing levels of noise.
 
@@
 	\end{align}
-	Further, one can choose $R$ possible ansatz reactions with their respective reaction function
+	Further, one can choose $R$ possible ansatz reactions with their respective reaction functions
 	\begin{align}
@@
 	y_{r,1}(\textbf{x}(t)) \\ \vdots \\ y_{r,S}(\textbf{x}(t))
-	\end{pmatrix}
+	\end{pmatrix},\quad r=1,\ldots,R,
 	\end{align}
-	so that the change of concentration for species $i$ at time $t$, is represented by the dynamical system
+	so that the change of concentration for species $i$ at time $t$ is represented by the dynamical system
 	\begin{align}
@@
 
-	When presented with a time series consisting of $T$ observations, the data can be represented as a matrix
+	When presented with a time series consisting of $T$ observations at points in time $t_1<\ldots < t_T$, the data can be represented as a matrix
 	\begin{align}
@@
 	\end{align}
-	Given this matrix, one can propose a library $\Theta(\textbf{X}) = \begin{pmatrix} \theta_1(\textbf{X}) & \theta_2(\textbf{X}) & \cdots & \theta_R(\textbf{X}) \end{pmatrix}$ of $R$ ansatz reactions with corresponding reaction functions
+	Given this matrix, a library $\Theta(\textbf{X}) = \begin{pmatrix} \theta_1(\textbf{X}) & \theta_2(\textbf{X}) & \cdots & \theta_R(\textbf{X}) \end{pmatrix}$ of $R$ ansatz reactions can be proposed with corresponding reaction functions
 	\begin{align}
@@
 	\end{align}
-	In particular, the system is linear in the coefficients $\Xi$, which makes potentially sparse regression tools such as elastic net regularization \cite{Zou2005} applicable.
+	In particular, the system is linear in the coefficients $\Xi$, which makes regression tools such as elastic net regularization \cite{Zou2005} applicable.
 To this end, one can consider the minimization problem to find $\hat{\Xi}$ such that

### 2018-08-13 docs/makefigures/case_config.py

@@ -1,0 +1,270 @@
+import numpy as np
+import readdy_learn.analyze.basis as basis
+from readdy_learn.example.regulation_network import RegulationNetwork
+from tabulate import tabulate
+
+SPECIES = ["DNA_A", "mRNA_A", "A", "DNA_B", "mRNA_B", "B", "DNA_C", "mRNA_C", "C"]
+
+
+def get_bfc_custom():
+    # species DA  MA  A  DB  MB  B  DC  MC  C
+    # ids     0   1   2  3   4   5  6   7   8
+    bfc = basis.BasisFunctionConfiguration(9)
+    # usual stuff A
+    bfc.add_fission(0, 0, 1)  # 0   DA -> DA + MA, transcription A
+    bfc.add_fission(1, 1, 2)  # 1   MA -> MA + A, translation A
+    bfc.add_decay(1)  # 2   MA -> 0, decay
+    bfc.add_decay(2)  # 3   A -> 0, decay
+    # usual stuff B
+    bfc.add_fission(3, 3, 4)  # 4   DB -> DB + MB, transcription B
+    bfc.add_fission(4, 4, 5)  # 5   MB -> MB + B, translation B
+    bfc.add_decay(4)  # 6   MB -> 0, decay
+    bfc.add_decay(5)  # 7   B -> 0, decay
+    # usual stuff C
+    bfc.add_fission(6, 6, 7)  # 8   DC -> DC + MC, transcription C
+    bfc.add_fission(7, 7, 8)  # 9   MC -> MC + C, translation C
+    bfc.add_decay(7)  # 10  MC -> 0, decay
+    bfc.add_decay(8)  # 11  C -> 0, decay
+
+    # all possible regulations
+    # self regulation
+    bfc.add_fusion(1, 2, 2)  # 12  MA + A -> A, A regulates A
+    bfc.add_fusion(4, 5, 5)  # 13  MB + B -> B, B regulates B
+    bfc.add_fusion(7, 8, 8)  # 14  MC + C -> C, C regulates C
+    # cyclic forward
+    bfc.add_fusion(4, 2, 2)  # 15  MB + A -> A, A regulates B
+    bfc.add_fusion(7, 5, 5)  # 16  MC + B -> B, B regulates C
+    bfc.add_fusion(1, 8, 8)  # 17  MA + C -> C, C regulates A
+    # cyclic backward
[…]

### 2018-08-21 docs/makefigures/case_config.py
upd

@@ -54,6 +54,16 @@ DESIRED_RATES = np.array([
 ])
 DESIRED_RATES = np.append(DESIRED_RATES, np.zeros((N_ADDITIONAL_FUNS,)))
 
+def failure_rate(estimated_rates):
+    mask = np.zeros_like(DESIRED_RATES)
+    active_processes = np.where(DESIRED_RATES > 0)
+    mask[active_processes] = 1.
+    
+    estimated_mask = np.zeros_like(DESIRED_RATES)
+    estimated_mask[np.where(estimated_rates >= CASE1_CUTOFF)] = 1.
+    
+    diff = np.sum(np.abs(mask - estimated_mask))
+    return diff
 
 def get_bfc_custom():
     # species DA  MA  A  DB  MB  B  DC  MC  C

### 2018-08-21 docs/makefigures/case_config.py
This is my code. My code is amazing.

@@ -1,14 +1,29 @@
 import numpy as np
+import h5py as h5
 import readdy_learn.analyze.basis as basis
 from readdy_learn.example.regulation_network import RegulationNetwork
 from tabulate import tabulate
 
 SPECIES = ["DNA_A", "mRNA_A", "A", "DNA_B", "mRNA_B", "B", "DNA_C", "mRNA_C", "C"]
-CASE1_CUTOFF = 0.38
+SPECIES_TEX = {
+    -1: '\emptyset',
+    0: '\mathrm{DNA}_\mathrm{A}',
+    1: '\mathrm{mRNA}_\mathrm{A}',
+    2: '\mathrm{A}',
+    3: '\mathrm{DNA}_\mathrm{B}',
+    4: '\mathrm{mRNA}_\mathrm{B}',
+    5: '\mathrm{B}',
+    6: '\mathrm{DNA}_\mathrm{C}',
+    7: '\mathrm{mRNA}_\mathrm{C}',
+    8: '\mathrm{C}'
+}
+CASE1_CUTOFF = 0.22
+CASE1_MIN_L1_ERR = 2.4077475813223574
 N_ADDITIONAL_FUNS = 2
 TARGET_TIME = 2.
 SCALE = 500.
 TIMESTEP = 3e-3
+INITIAL_CONDITIONS = [[1, 2, 0, 1, 0, 3, 1, 0, 0], [1, 1.5, 0, 1, 0, 2., 1, 0, 1.]]
 
 DESIRED_RATES = np.array([
     1.8,  # DA -> DA + MA, transcription A
@@ -54,16 +69,18 @@ DESIRED_RATES = np.array([
 ])
 DESIRED_RATES = np.append(DESIRED_RATES, np.zeros((N_ADDITIONAL_FUNS,)))
 
+
 def failure_rate(estimated_rates):
     mask = np.zeros_like(DESIRED_RATES)
     active_processes = np.where(DESIRED_RATES > 0)
     mask[active_processes] = 1.
-    
+
     estimated_mask = np.zeros_like(DESIRED_RATES)
     estimated_mask[np.where(estimated_rates >= CASE1_CUTOFF)] = 1.
-    
+
     diff = np.sum(np.abs(mask - estimated_mask))
     return diff
+
 
 def get_bfc_custom():
     # species DA  MA  A  DB  MB  B  DC  MC  C
@@ -274,3 +291,24 @@ def get_regulation_network_gillespie(gillespie_realisations=1):
[…]

### 2018-08-24 docs/makefigures/illustrative_traj.py
wip

@@ -5,42 +5,53 @@ import case_config
 
 plt.style.use("lm.mplstyle")
 
-#DATA_ROOT = "/home/chris/Dropbox/readdy_learn/reaction_learn_data"
-DATA_ROOT = "/srv/public/chrisfr/workspace/data/reaction_learn_data"
+DATA_ROOT = "/home/chris/Dropbox/readdy_learn/reaction_learn_data"
+# DATA_ROOT = "/srv/public/chrisfr/workspace/data/reaction_learn_data"
 
 traj = case_config.get_traj_from_file(os.path.join(DATA_ROOT, "gillespie_trajs_init_1.h5"), 20000, 0)
 
-#which = "mrna"
-#which = "protein"
-which = "all"
 
-if which=="mrna":
-    species = np.array([1, 4, 7])
-    ylim = (-0.05, 0.85)
-elif which=="protein":
-    species = np.array([2, 5, 8])
-    ylim = (-0.05, 0.85)
-elif which=="all":
-    species = np.array([1, 2, 4, 5, 7, 8])
-    ylim = (-0.05, 0.85)
-else:
-    raise ValueError("blub")
+def plot_species(species):
+    plot_format = ["-", "--"] * 6
+    for ix, s in enumerate(species):
+        name = case_config.SPECIES_TEX[s]
+        plt.plot(t[::stride], traj[0][:, s][::stride], plot_format[ix], label=r"${}$".format(name),
+                 #color='C{}'.format(ix)
+                 color="black"
+                 )
+    plt.legend()
 
-filename = "illustrative_traj_init_1_"+which+".pdf"
 
-plot_format = ["-"] * 6
-#plot_format = ["-", "--", ":","-", "--", ":"]
-
-t = np.arange(len(traj[0]))*case_config.TIMESTEP
+t = np.arange(len(traj[0])) * case_config.TIMESTEP
 stride = 1
-for ix, s in enumerate(species):
-    name = case_config.SPECIES_TEX[s]
-    plt.plot(t[::stride], traj[0][:,s][::stride], plot_format[ix], label=r"${}$".format(name), color='C{}'.format(ix))
-plt.legend(loc="lower right")
-plt.ylim(ylim)
-plt.xlabel('Time in a.u.')
-plt.ylabel('Concentration in a.u.')
-plt.gcf().tight_layout()
 
-plt.savefig(filename)
[…]

### 2018-08-24 docs/makefigures/illustrative_traj.py
update scheme fig

@@ -3,7 +3,7 @@ import os
 import numpy as np
 import case_config
 
-plt.style.use("lm.mplstyle")
+plt.style.use("rlearn.mplstyle")
 
 DATA_ROOT = "/home/chris/Dropbox/readdy_learn/reaction_learn_data"
 # DATA_ROOT = "/srv/public/chrisfr/workspace/data/reaction_learn_data"
@@ -28,7 +28,7 @@ stride = 1
 # width = 7.1 # double columns
 width = 7.1 / 2.  # single column
 n_rows = 3
-height = n_rows * 1.5
+height = n_rows * 1.3
 n_cols = 1
 
 fig, axarr = plt.subplots(n_rows, n_cols, figsize=(width, height), sharex=True)
@@ -40,7 +40,7 @@ plot_species(np.array([2, 1]))
 
 plt.sca(axlist[1])
 plot_species(np.array([5, 4]))
-#plt.ylabel('Concentration in a.u.')
+plt.ylabel('Concentration in a.u.', labelpad=20)
 
 plt.sca(axlist[2])
 plot_species(np.array([8, 7]))

### 2018-08-24 docs/makefigures/illustrative_traj.py
update scheme fig

@@ -19,7 +19,6 @@ def plot_species(species):
                  #color='C{}'.format(ix)
                  color="black"
                  )
-    plt.legend()
 
 
 t = np.arange(len(traj[0])) * case_config.TIMESTEP
@@ -27,6 +26,7 @@ stride = 1
 
 # width = 7.1 # double columns
 width = 7.1 / 2.  # single column
+width = 4.5
 n_rows = 3
 height = n_rows * 1.3
 n_cols = 1
@@ -37,15 +37,18 @@ axlist = axarr.flatten()
 plt.sca(axlist[0])
 plot_species(np.array([2, 1]))
 #plt.ylabel('Concentration in a.u.')
+plt.legend(loc="center left", bbox_to_anchor=(1,0.5))
 
 plt.sca(axlist[1])
 plot_species(np.array([5, 4]))
 plt.ylabel('Concentration in a.u.', labelpad=20)
+plt.legend(loc="center left", bbox_to_anchor=(1,0.5))
 
 plt.sca(axlist[2])
 plot_species(np.array([8, 7]))
 plt.xlabel('Time in a.u.')
 #plt.ylabel('Concentration in a.u.')
+plt.legend(loc="center left", bbox_to_anchor=(1,0.5))
 
 fig.tight_layout()
 fig.subplots_adjust(hspace=0)

### 2018-08-24 docs/manuscript.tex
upd
wip
captions etc

@@
         \centering
-        \includegraphics[width=.5\textwidth]{./figures_tex/gene_regulation_full.pdf}
-        \caption{The regulation network example described in Section \ref{sec:generegulation}.
-Each circle depicts a species, each arrow corresponds to one reaction.
-The red arrows describe the regulatory network which generates the example data.}
+        \includegraphics[width=.5\textwidth]{./figures_tex/scheme.pdf}
+        \caption{The regulation network example described in Sec.~\ref{sec:generegulation}.
+\textbf{(a)}: Each circle depicts a species, each arrow corresponds to one reaction.
+Blue arrows denote transcription from DNA to mRNA, green arrows denote translation from mRNA to protein, and red arrows denote the regulatory network.
+\textbf{(b)}: Concentration timeseries generated from the reaction network shown in~(a).
+The initial condition prescribes positive concentration values only for B protein and $\mathrm{mRNA}_\mathrm{A}$ species.
+This initial condition is used in the subsequent sections for further analysis.}
         \label{fig:network}
@@
 
-	\section{Proof of concept}\label{sec:proof-of-concept}
+	\section{Results}\label{sec:results}
 
@@
 		\centering
-		\includegraphics[width=.9\textwidth]{./figures_tex/bubbles}
-		\caption{Generating and estimated reaction rates in the situation of Section \ref{sec:case-1}.
+		\includegraphics[width=.9\textwidth]{./figures_tex/sparsity_pattern}
+		\caption{Estimated reaction rates in the system described in Sec.~\ref{sec:case-1}.
 The y and x axes contain reaction educts and products, respectively.
-The black outlines denote the reactions with which the system was generated and the area of the circles scales linearly with the reaction rate.
-The blue circles are least squares estimated reaction rates and the orange circles depict rates which were obtained by some $\alpha>0$ in the minimization problem (\ref{method:minimizationproblem}).
-If a certain rate was estimated in both cases, two wedges instead of once circle as displayed.}
+A circle at position $(i,j)$ represents a reaction $i\rightharpoonup j$ whose rate has a linear relation with the area of the circle.
+The black outlines denote the reactions with which the system was generated and contain the respective rate value.
+Red crosses denote reactions that were used as additional ansatz reactions.
+Blue circles are estimated by linear least-squares and orange circles depict rates which were obtained by solving the minimization problem (\ref{method:minimizationproblem}).
+The latter rates are subject to a cutoff $c=0.22$ corresponding to the green circle's area under which a sparse solution with the correct processes can be recovered.
+If a certain rate was estimated in both cases, two wedges instead of one circle are displayed.}
 		\label{fig:case-1-sparsity-pattern}
@@
 
+	\begin{figure}
+		\centering
+		\includegraphics[width=.5\columnwidth]{./figures_tex/case2}
+		\caption{lalala}
+	\end{figure}
+
 	\subsection{Multiple initial conditions}\label{sec:case-3}
@@
 	\end{itemize}
+
+	\begin{figure}
+		\centering
+		\includegraphics[width=.5\columnwidth]{./figures_tex/case3}
+	\end{figure}
 

### 2018-08-24 docs/makefigures/illustrative_traj.py
update fig case3

@@ -8,7 +8,7 @@ plt.style.use("rlearn.mplstyle")
 DATA_ROOT = "/home/chris/Dropbox/readdy_learn/reaction_learn_data"
 # DATA_ROOT = "/srv/public/chrisfr/workspace/data/reaction_learn_data"
 
-traj = case_config.get_traj_from_file(os.path.join(DATA_ROOT, "gillespie_trajs_init_1.h5"), 20000, 0)
+traj = case_config.get_traj_from_file(os.path.join(DATA_ROOT, "gillespie_trajs_init_3.h5"), 1, 0)
 
 
 def plot_species(species):
@@ -54,7 +54,7 @@ fig.tight_layout()
 fig.subplots_adjust(hspace=0)
 plt.setp([a.get_xticklabels() for a in fig.axes[:-1]], visible=False)
 
-plt.savefig("./illustrative-traj.pdf", transparent=True)
+plt.savefig("./illustrative-traj-case3-init3.pdf", transparent=True)
 plt.show()
 plt.clf()
 plt.close(fig)

### 2018-08-24 docs/manuscript.tex
A

@@
 		\includegraphics[width=.5\columnwidth]{./figures_tex/case2}
-		\caption{lalala}
+		\caption{Convergence of the estimation error when estimating the system described in Sec.~\ref{sec:case-1} with decreasing levels of noise by application of the minimization problem (\ref{method:minimizationproblem}).
+The level of noise is regulated by the number of measurements being averaged.
+The hyperparameters for each level of noise were estimated using 10-fold cross validation.
+This was repeated 10 times giving rise to the mean and standard deviation depicted by dark blue graph and shaded blue area, respectively.
+The failure rate is the number of incorrectly identified processes.
+The estimation error is given by the $L_1$ difference of the generating reaction rates $\hat{\xi}$ and the estimated reaction rates $\xi_\mathrm{CV}$.}
+		\label{fig:case2-convergence}
 	\end{figure}
@@
 		\centering
-		\includegraphics[width=.5\columnwidth]{./figures_tex/case3}
+		\includegraphics[width=\columnwidth]{./figures_tex/case3}
+		\caption{Convergence of the estimation error when using two concatenated concentration timeseries with different initial conditions.
+\textbf{(a)}: A realization of one of the initial conditions that were used.
+It is identical to the initial conditions used in Sec.~\ref{sec:case-1} and Sec.~\ref{sec:case-2}.
+\textbf{(b)}: A realization of the other initial condition that was used.
+It prescribes positive initial concentrations for $\mathrm{mRNA}_\mathrm{A}$, $\mathrm{B}$, and $\mathrm{C}$ species.
+\textbf{(c)}: Analogously to Fig.~\ref{fig:case2-convergence} with the difference that 20-fold cross validation was used for hyperparameter estimation.}
 	\end{figure}

### 2018-08-27 docs/manuscript.tex
wip

@@
 \usepackage{subcaption}
-\usepackage[version=4]{mhchem}
 
@@
 	For $\alpha=0$ the minimization problem reduces to constrained least-squares.
-In order to solve (\ref{method:minimizationproblem}) the sequential least-squares minimizer SLSQP \cite{Kraft1988} is applied via the software package SciPy \cite{SciPy}.
+In order to solve (\ref{method:minimizationproblem}) the numerical sequential least-squares minimizer SLSQP \cite{Kraft1988} is applied via the software package SciPy \cite{SciPy}.
 Since only the concentration data $\mathbf{X}$ is available but not its temporal derivative $\dot{\mathbf{X}}$, it is approximated numerically by second order finite differences with the exception of boundary data.
@@
 
-    To this end, let $A$, $B$, and $C$ be three species of proteins which are being translated from a corresponding mRNA molecule.
-Each mRNA in turn has a corresponding DNA which it is transcribed from.
-The proteins and mRNA molecules decay over time whereas the DNA concentration remains constant.
+    To this end, let $\mathrm{A}$, $\mathrm{B}$, and $\mathrm{C}$ be three species of proteins which are being translated from a corresponding $\mathrm{mRNA}$ molecule.
+Each $\mathrm{mRNA}$ in turn has a corresponding $\mathrm{DNA}$ which it is transcribed from.
+The proteins and $\mathrm{mRNA}$ molecules decay over time whereas the $\mathrm{DNA}$ concentration remains constant.
 
-    For each of the three species one can formulate the following basic reactions
+    The following basic reactions can be formulated
 	\begin{align*}
-        \ce{ DNA_i } & \ce{ -> DNA_i + mRNA_i } &\text{(transcription)}, \\
-        \ce{ mRNA_i } & \ce{ -> mRNA_i + Protein_i } &\text{(translation)}, \\
-        \ce{ mRNA_i } & \ce{ -> } \emptyset &\text{(decay of mRNA)}, \\
-        \ce{ Protein_i } & \ce{ -> } \emptyset &\text{(decay of Protein)},
+		\mathrm{DNA}_i &\rightharpoonup \mathrm{DNA}_i + \mathrm{mRNA}_i &&\text{(transcription)},\\
+		\mathrm{mRNA}_i &\rightharpoonup \mathrm{mRNA}_i + \mathrm{Protein}_i &&\text{(translation)},\\
+		\mathrm{mRNA}_i &\rightharpoonup \emptyset &&\text{(decay of mRNA)},\\
+		\mathrm{Protein}_i &\rightharpoonup \emptyset &&\text{(decay of Protein)},
     \end{align*}
-    for $i=A,B,C$.
+    for each of the species $i\in S:=\{\mathrm{A},\mathrm{B},\mathrm{C}\}$.
 
-    The protein concentrations of $A$, $B$, and $C$ regulate each other in a negative way by hindering the transcription process.
+    The protein concentrations of $\mathrm{A}$, $\mathrm{B}$, and $\mathrm{C}$ regulate each other in a negative way by hindering the transcription process.
 In the law of mass action model we account for this by a reaction.
-A repression of species $j$ by species $i$ can be modeled as
-    \begin{equation*}
-        \ce{ Protein_i + mRNA_j -> Protein_i } \quad \text{(regulation by repression)}.
-    \end{equation*}
+A repression of species $j\in S$ by species $i\in S$ can be modeled as
+    \begin{align*}
+    i + \mathrm{mRNA}_j \rightharpoonup i &&&\text{(regulation by repression).}
+    \end{align*}
 
-    In our example proteins of type $A$ regulate the mRNA$_\mathrm{B}$ molecules, proteins of type $B$ regulate the mRNA$_\mathrm{A}$ molecules and proteins of type $C$ regulate the mRNA$_\mathrm{A}$ molecules.
+    In our example proteins of type $\mathrm{A}$ regulate the $\mathrm{mRNA}_\mathrm{B}$ molecules, proteins of type $\mathrm{B}$ regulate the $\mathrm{mRNA}_\mathrm{A}$ molecules and proteins of type $\mathrm{C}$ regulate the $\mathrm{mRNA}_\mathrm{A}$ molecules.
 The network of all species and reactions is depicted in Figure \ref{fig:network}.
@@
 				\multicolumn{3}{c}{Reaction} & rate & description \\ \hline\noalign{\vskip .1cm}
-				$\ce{DNA_A}$ &$\ce{->}$& $\ce{DNA_A + mRNA_A }$ & $k_1 = 1.8$ & transcription of $\ce{mRNA_A}$\\
-				$\ce{mRNA_A}$ &$\ce{->}$& $\ce{mRNA_A + A}$ & $k_2 = 2.1$ & translation of $\ce{A}$ proteins\\
-				$\ce{mRNA_A}$ &$\ce{->}$& $\ce{\emptyset}$ & $k_3 = 1.3$ & $\ce{mRNA_A}$ decay\\
-				$\ce{A}$ &$\ce{->}$& $\ce{\emptyset }$ & $k_4 = 1.5$ & decay of $\ce{A}$ proteins\\
-				$\ce{DNA_B}$ &$\ce{->}$& $\ce{DNA_B + mRNA_B }$ & $k_5 = 2.2$ & transcription of $\ce{mRNA_B}$\\
-				$\ce{mRNA_B}$ &$\ce{->}$& $\ce{mRNA_B + B}$ & $k_6 = 2.0$ & translation of $\ce{B}$ proteins\\
-				$\ce{mRNA_B}$ &$\ce{->}$& $\ce{\emptyset }$ & $k_7 = 2.0$ & $\ce{mRNA_B}$ decay\\
-				$\ce{B}$ &$\ce{->}$& $\ce{\emptyset }$ & $k_8 = 2.5$ & decay of $\ce{B}$ proteins\\
-				$\ce{DNA_C}$ &$\ce{->}$& $\ce{DNA_C + mRNA_C }$ & $k_9 = 3.2$ & transcription of $\ce{mRNA_C}$\\
-				$\ce{mRNA_C}$ &$\ce{->}$& $\ce{mRNA_C + C}$ & $k_{10} = 3.0$ & translation of $\ce{C}$ proteins\\
-				$\ce{mRNA_C}$ &$\ce{->}$& $\ce{\emptyset }$ & $k_{11} = 2.3$ & $\ce{mRNA_C}$ decay\\
-				$\ce{C}$ &$\ce{->}$& $\ce{\emptyset }$ & $k_{12} = 2.5$ & decay of $\ce{C}$ proteins\\
-				$\ce{mRNA_A + A}$ &$\ce{->}$& $A$ & $k_{13} = 0$ & self regulation of $\ce{A}$ proteins\\
-				$\ce{mRNA_B + B}$ &$\ce{->}$& $B$ & $k_{14} = 0$ & self regulation of $\ce{B}$ proteins\\
-				$\ce{mRNA_C + C}$ &$\ce{->}$& $C$ & $k_{15} = 0$ & self regulation of $\ce{C}$ proteins\\
-				$\ce{mRNA_B + A}$ &$\ce{->}$& $A$ & $k_{16} = 0$ & cyclic regulation of $\ce{A}$ proteins\\
-				$\ce{mRNA_C + B}$ &$\ce{->}$& $B$ & $k_{17} = 0$ & cyclic regulation of $\ce{B}$ proteins\\
-				$\ce{mRNA_A + C}$ &$\ce{->}$& $C$ & $k_{18} = 0$ & cyclic regulation of $\ce{C}$ proteins\\
-				$\ce{mRNA_C + A}$ &$\ce{->}$& $A$ & $k_{16} = 6.0$ & cyclic regulation of $\ce{A}$ proteins\\
-				$\ce{mRNA_B + C}$ &$\ce{->}$& $C$ & $k_{17} = 4.0$ & cyclic regulation of $\ce{C}$ proteins\\
-				$\ce{mRNA_A + B}$ &$\ce{->}$& $B$ & $k_{18} = 3.0$ & cyclic regulation of $\ce{B}$ proteins\\
-				$\ce{mRNA_A + A}$ &$\ce{->}$& $\ce{mRNA_A}$ & $k_{19} = 0$ & artificial fusion\\
-				$\ce{mRNA_B + B}$ &$\ce{->}$& $\ce{mRNA_B}$ & $k_{20} = 0$ & artificial fusion\\
-				$\ce{mRNA_C + C}$ &$\ce{->}$& $\ce{mRNA_C}$ & $k_{21} = 0$ & artificial fusion\\
-				$\ce{mRNA_A + B}$ &$\ce{->}$& $\ce{mRNA_A}$ & $k_{22} = 0$ & artificial fusion\\
-				$\ce{mRNA_B + C}$ &$\ce{->}$& $\ce{mRNA_B}$ & $k_{23} = 0$ & artificial fusion\\
-				$\ce{mRNA_C + A}$ &$\ce{->}$& $\ce{mRNA_C}$ & $k_{24} = 0$ & artificial fusion\\
-				$\ce{mRNA_A + C}$ &$\ce{->}$& $\ce{mRNA_A}$ & $k_{25} = 0$ & artificial fusion\\
-				$\ce{mRNA_B + A}$ &$\ce{->}$& $\ce{mRNA_B}$ & $k_{26} = 0$ & artificial fusion\\
-				$\ce{mRNA_C + B}$ &$\ce{->}$& $\ce{mRNA_C}$ & $k_{27} = 0$ & artificial fusion\\
-				$\ce{A + A}$ &$\ce{->}$& $\ce{A}$ & $k_{28} = 0$ & $\ce{A}$ regulates $\ce{A}$\\
-				$\ce{B + B}$ &$\ce{->}$& $\ce{B}$ & $k_{29} = 0$ & $\ce{B}$ regulates $\ce{B}$\\
-				$\ce{C + C}$ &$\ce{->}$& $\ce{C}$ & $k_{30} = 0$ & $\ce{C}$ regulates $\ce{C}$\\
-				$\ce{B + A}$ &$\ce{->}$& $\ce{A}$ & $k_{31} = 0$ & artificial fusion between proteins\\
-				$\ce{C + B}$ &$\ce{->}$& $\ce{B}$ & $k_{32} = 0$ & artificial fusion between proteins\\
-				$\ce{A + C}$ &$\ce{->}$& $\ce{C}$ & $k_{33} = 0$ & artificial fusion between proteins\\
-				$\ce{C + A}$ &$\ce{->}$& $\ce{A}$ & $k_{34} = 0$ & artificial fusion between proteins\\
-				$\ce{B + C}$ &$\ce{->}$& $\ce{C}$ & $k_{35} = 0$ & artificial fusion between proteins\\
-				$\ce{A + B}$ &$\ce{->}$& $\ce{B}$ & $k_{36} = 0$ & artificial fusion between proteins\\
-				$\ce{A}$ &$\ce{->}$& $\ce{B}$ & $k_{37} = 0$ & artificial conversion between proteins\\
-				$\ce{B}$ &$\ce{->}$& $\ce{C}$ & $k_{38} = 0$ & artificial conversion between proteins\\
-				$\ce{C}$ &$\ce{->}$& $\ce{A}$ & $k_{39} = 0$ & artificial conversion between proteins\\
-				$\ce{A}$ &$\ce{->}$& $\ce{C}$ & $k_{40} = 0$ & artificial conversion between proteins\\
-				$\ce{C}$ &$\ce{->}$& $\ce{B}$ & $k_{41} = 0$ & artificial conversion between proteins\\
-				$\ce{B}$ &$\ce{->}$& $\ce{A}$ & $k_{42} = 0$ & artificial conversion between proteins
+				$\mathrm{DNA}_\mathrm{A}$ &$\rightharpoonup$& $\mathrm{DNA}_\mathrm{A} + \mathrm{mRNA}_\mathrm{A}$ & $k_1 = 1.8$ & transcription of $\mathrm{mRNA}_\mathrm{A}$\\
+				$\mathrm{mRNA}_\mathrm{A}$ &$\rightharpoonup$& $\mathrm{mRNA}_\mathrm{A} + \mathrm{A}$ & $k_2 = 2.1$ & translation of $\mathrm{A}$ proteins\\
+				$\mathrm{mRNA}_\mathrm{A}$ &$\rightharpoonup$& $\emptyset$ & $k_3 = 1.3$ & $\mathrm{mRNA}_\mathrm{A}$ decay\\
+				$\mathrm{A}$ &$\rightharpoonup$& $\emptyset$ & $k_4 = 1.5$ & decay of $\mathrm{A}$ proteins\\
+				$\mathrm{DNA}_\mathrm{B}$ &$\rightharpoonup$& $\mathrm{DNA}_\mathrm{B} + \mathrm{mRNA}_\mathrm{B}$ & $k_5 = 2.2$ & transcription of $\mathrm{mRNA}_\mathrm{B}$\\
+				$\mathrm{mRNA}_\mathrm{B}$ &$\rightharpoonup$& $\mathrm{mRNA}_\mathrm{B} + \mathrm{B}$ & $k_6 = 2.0$ & translation of $\mathrm{B}$ proteins\\
+				$\mathrm{mRNA}_\mathrm{B}$ &$\rightharpoonup$& $\emptyset$ & $k_7 = 2.0$ & $\mathrm{mRNA}_\mathrm{B}$ decay\\
+				$\mathrm{B}$ &$\rightharpoonup$& $\emptyset$ & $k_8 = 2.5$ & decay of $\mathrm{B}$ proteins\\
+				$\mathrm{DNA}_\mathrm{C}$ &$\rightharpoonup$& $\mathrm{DNA}_\mathrm{C} + \mathrm{mRNA}_\mathrm{C}$ & $k_9 = 3.2$ & transcription of $\mathrm{mRNA}_\mathrm{C}$\\
+				$\mathrm{mRNA}_\mathrm{C}$ &$\rightharpoonup$& $\mathrm{mRNA}_\mathrm{C} + \mathrm{C}$ & $k_{10} = 3.0$ & translation of $\mathrm{C}$ proteins\\
+				$\mathrm{mRNA}_\mathrm{C}$ &$\rightharpoonup$& $\emptyset$ & $k_{11} = 2.3$ & $\mathrm{mRNA}_\mathrm{C}$ decay\\
+				$\mathrm{C}$ &$\rightharpoonup$& $\emptyset$ & $k_{12} = 2.5$ & decay of $\mathrm{C}$ proteins\\
+				$\mathrm{mRNA}_\mathrm{A} + \mathrm{A}$ &$\rightharpoonup$& $\mathrm{A}$ & $k_{13} = 0$ & self regulation of $\mathrm{A}$ proteins\\
+				$\mathrm{mRNA}_\mathrm{B} + \mathrm{B}$ &$\rightharpoonup$& $\mathrm{B}$ & $k_{14} = 0$ & self regulation of $\mathrm{B}$ proteins\\
+				$\mathrm{mRNA}_\mathrm{C} + \mathrm{C}$ &$\rightharpoonup$& $\mathrm{C}$ & $k_{15} = 0$ & self regulation of $\mathrm{C}$ proteins\\
+				$\mathrm{mRNA}_\mathrm{B} + \mathrm{A}$ &$\rightharpoonup$& $\mathrm{A}$ & $k_{16} = 0$ & cyclic regulation of $\mathrm{A}$ proteins\\
+				$\mathrm{mRNA}_\mathrm{C} + \mathrm{B}$ &$\rightharpoonup$& $\mathrm{B}$ & $k_{17} = 0$ & cyclic regulation of $\mathrm{B}$ proteins\\
+				$\mathrm{mRNA}_\mathrm{A} + \mathrm{C}$ &$\rightharpoonup$& $\mathrm{C}$ & $k_{18} = 0$ & cyclic regulation of $\mathrm{C}$ proteins\\
+				$\mathrm{mRNA}_\mathrm{C} + \mathrm{A}$ &$\rightharpoonup$& $\mathrm{A}$ & $k_{16} = 6.0$ & cyclic regulation of $\mathrm{A}$ proteins\\
+				$\mathrm{mRNA}_\mathrm{B} + \mathrm{C}$ &$\rightharpoonup$& $\mathrm{C}$ & $k_{17} = 4.0$ & cyclic regulation of $\mathrm{C}$ proteins\\
+				$\mathrm{mRNA}_\mathrm{A} + \mathrm{B}$ &$\rightharpoonup$& $\mathrm{B}$ & $k_{18} = 3.0$ & cyclic regulation of $\mathrm{B}$ proteins\\
+				$\mathrm{mRNA}_\mathrm{A} + \mathrm{A}$ &$\rightharpoonup$& $\mathrm{mRNA}_\mathrm{A}$ & $k_{19} = 0$ & artificial fusion\\
+				$\mathrm{mRNA}_\mathrm{B} + \mathrm{B}$ &$\rightharpoonup$& $\mathrm{mRNA}_\mathrm{B}$ & $k_{20} = 0$ & artificial fusion\\
+				$\mathrm{mRNA}_\mathrm{A} + \mathrm{B}$ &$\rightharpoonup$& $\mathrm{mRNA}_\mathrm{A}$ & $k_{21} = 0$ & artificial fusion\\
+				$\mathrm{mRNA}_\mathrm{B} + \mathrm{C}$ &$\rightharpoonup$& $\mathrm{mRNA}_\mathrm{B}$ & $k_{22} = 0$ & artificial fusion\\
+				$\mathrm{mRNA}_\mathrm{C} + \mathrm{A}$ &$\rightharpoonup$& $\mathrm{mRNA}_\mathrm{C}$ & $k_{23} = 0$ & artificial fusion\\
+				$\mathrm{mRNA}_\mathrm{A} + \mathrm{C}$ &$\rightharpoonup$& $\mathrm{mRNA}_\mathrm{A}$ & $k_{24} = 0$ & artificial fusion\\
+				$\mathrm{mRNA}_\mathrm{B} + \mathrm{A}$ &$\rightharpoonup$& $\mathrm{mRNA}_\mathrm{B}$ & $k_{25} = 0$ & artificial fusion\\
+				$\mathrm{A} + \mathrm{A}$ &$\rightharpoonup$& $\mathrm{A}$ & $k_{26} = 0$ & $\mathrm{A}$ regulates $\mathrm{A}$\\
+				$\mathrm{B} + \mathrm{B}$ &$\rightharpoonup$& $\mathrm{B}$ & $k_{27} = 0$ & $\mathrm{B}$ regulates $\mathrm{B}$\\
+				$\mathrm{C} + \mathrm{C}$ &$\rightharpoonup$& $\mathrm{C}$ & $k_{28} = 0$ & $\mathrm{C}$ regulates $\mathrm{C}$\\
+				$\mathrm{B} + \mathrm{A}$ &$\rightharpoonup$& $\mathrm{A}$ & $k_{29} = 0$ & artificial fusion between proteins\\
+				$\mathrm{C} + \mathrm{B}$ &$\rightharpoonup$& $\mathrm{B}$ & $k_{30} = 0$ & artificial fusion between proteins\\
+				$\mathrm{A} + \mathrm{C}$ &$\rightharpoonup$& $\mathrm{C}$ & $k_{31} = 0$ & artificial fusion between proteins\\
+				$\mathrm{C} + \mathrm{A}$ &$\rightharpoonup$& $\mathrm{A}$ & $k_{32} = 0$ & artificial fusion between proteins\\
+				$\mathrm{B} + \mathrm{C}$ &$\rightharpoonup$& $\mathrm{C}$ & $k_{33} = 0$ & artificial fusion between proteins\\
+				$\mathrm{A} + \mathrm{B}$ &$\rightharpoonup$& $\mathrm{B}$ & $k_{34} = 0$ & artificial fusion between proteins\\
+				$\mathrm{A}$ &$\rightharpoonup$& $\mathrm{B}$ & $k_{35} = 0$ & artificial conversion between proteins\\
+				$\mathrm{B}$ &$\rightharpoonup$& $\mathrm{C}$ & $k_{36} = 0$ & artificial conversion between proteins\\
+				$\mathrm{C}$ &$\rightharpoonup$& $\mathrm{A}$ & $k_{37} = 0$ & artificial conversion between proteins\\
+				$\mathrm{A}$ &$\rightharpoonup$& $\mathrm{C}$ & $k_{38} = 0$ & artificial conversion between proteins\\
+				$\mathrm{C}$ &$\rightharpoonup$& $\mathrm{B}$ & $k_{39} = 0$ & artificial conversion between proteins\\
+				$\mathrm{B}$ &$\rightharpoonup$& $\mathrm{A}$ & $k_{40} = 0$ & artificial conversion between proteins\\
+				$\mathrm{mRNA}_\mathrm{B} + \mathrm{mRNA}_\mathrm{C}$ &$\rightharpoonup$& $\mathrm{mRNA}_\mathrm{A}$ & $k_{41} = 0$ & artificial fusion\\
+				$\mathrm{mRNA}_\mathrm{C} + \mathrm{mRNA}_\mathrm{B}$ &$\rightharpoonup$& $\mathrm{mRNA}_\mathrm{C}$ & $k_{42} = 0$ & artificial fusion\\
+				$\mathrm{mRNA}_\mathrm{C} + \mathrm{A}$ &$\rightharpoonup$& $\mathrm{C}$ & $k_{43} = 0$ & artificial fusion\\
+				$\mathrm{mRNA}_\mathrm{C} + \mathrm{C}$ &$\rightharpoonup$& $\mathrm{C}$ & $k_{43} = 0$ & artificial fusion
 			\end{tabular}

### 2018-08-27 docs/manuscript.tex
wip

@@
 \textbf{(a)}: Each circle depicts a species, each arrow corresponds to one reaction.
-Blue arrows denote transcription from DNA to mRNA, green arrows denote translation from mRNA to protein, and red arrows denote the regulatory network.
+Blue arrows denote transcription from DNA to $\mathrm{mRNA}$, green arrows denote translation from $\mathrm{mRNA}$ to protein, and red arrows denote the regulatory network.
 \textbf{(b)}: Concentration timeseries generated from the reaction network shown in~(a).
-The initial condition prescribes positive concentration values only for B protein and $\mathrm{mRNA}_\mathrm{A}$ species.
+The initial condition prescribes positive concentration values only for $\mathrm{B}$ protein and $\mathrm{mRNA}_\mathrm{A}$ species.
 This initial condition is used in the subsequent sections for further analysis.}
@@
 
-	First it is assumed that the concentration data can be obtained without noise and it is shown that while the minimization problem (\ref{method:minimizationproblem}) without regularization, i.e., $\alpha=0$, yields rates that fit well to the input data, the original rates cannot be recovered while adding $L_1$ regularization and a certain cutoff improves the situation and the original model can be recovered.
+	First it is assumed that the concentration data can be obtained without noise and it is shown that while the minimization problem (\ref{method:minimizationproblem}) without regularization, i.e., $\alpha=0$, yields rates that fit well to the input data, the original rates cannot be recovered while adding $L_1$ regularization and applying a cutoff recovers the underlying reaction network \textcolor{red}{(and rates?)}.
 
-	In Section \ref{sec:case-2} the example data contains additive Gaussian noise.
-It is assumed that the initial state can be reproduced exactly and one has several measurements.
-It is shown that in the limit of small noise one can obtain a good fit and regularization helps in the reliability of the estimated reaction rates.
+	In Section \ref{sec:case-2} the example data was generated using the Gillespie SSA \textbf{[cite]}.
+While the concentrations now contain stochastic noise, it is assumed that the initial state can be reproduced exactly and multiple measurements with that exact initial condition can be performed.
+It is shown that in the limit of many measurements a good fit can reliably be obtained.
 
-	The last example is described in Section \ref{sec:case-3} and can be seen as an extension of the situation in Section \ref{sec:case-3} to multiple initial conditions.
+	The last example is described in Section \ref{sec:case-3} and is an extension of the situation in Section \ref{sec:case-2} to multiple initial conditions.
+Due to more varied data and more statistics a good fit can be obtained with fewer measurements than in the case of just one initial condition.
 

### 2018-08-27 docs/makefigures/illustrative_traj.py
illustrative traj concatenated

@@ -8,17 +8,28 @@ plt.style.use("rlearn.mplstyle")
 DATA_ROOT = "/home/chris/Dropbox/readdy_learn/reaction_learn_data"
 # DATA_ROOT = "/srv/public/chrisfr/workspace/data/reaction_learn_data"
 
-traj = case_config.get_traj_from_file(os.path.join(DATA_ROOT, "gillespie_trajs_init_3.h5"), 1, 0)
+traj = case_config.get_traj_from_file(os.path.join(DATA_ROOT, "gillespie_trajs_conced_1_3_normal.h5"), 1, 0)
 
 
 def plot_species(species):
     plot_format = ["-", "--"] * 6
     for ix, s in enumerate(species):
         name = case_config.SPECIES_TEX[s]
-        plt.plot(t[::stride], traj[0][:, s][::stride], plot_format[ix], label=r"${}$".format(name),
-                 #color='C{}'.format(ix)
-                 color="black"
-                 )
+        if False:
+            plt.plot(t[::stride], traj[0][:, s][::stride], plot_format[ix], label=r"${}$".format(name),
+                     #color='C{}'.format(ix)
+                     color="black"
+                     )
+        else:
+            half = len(t)//2
+            plt.plot(t[:half:stride], traj[0][:, s][:half:stride], plot_format[ix], label=r"${}$".format(name),
+                     # color='C{}'.format(ix)
+                     color="black"
+                     )
+            plt.plot(t[half::stride], traj[0][:, s][half::stride], plot_format[ix],
+                     # color='C{}'.format(ix)
+                     color="black"
+                     )
 
 
 t = np.arange(len(traj[0])) * case_config.TIMESTEP
@@ -51,10 +62,10 @@ plt.xlabel('Time in a.u.')
 plt.legend(loc="center left", bbox_to_anchor=(1,0.5))
 
 fig.tight_layout()
-fig.subplots_adjust(hspace=0)
+fig.subplots_adjust(hspace=0.2)
 plt.setp([a.get_xticklabels() for a in fig.axes[:-1]], visible=False)
 
-plt.savefig("./illustrative-traj-case3-init3.pdf", transparent=True)
+plt.savefig("./illustrative-traj-conced.pdf", transparent=True)
 plt.show()
 plt.clf()
 plt.close(fig)

### 2018-08-27 docs/manuscript.tex
wip

@@
 
+	When applying no regularization, one can observe in Figure \ref{fig:case-1-sparsity-pattern} that the sparsity pattern does not match the generating reaction network and the reaction responsible for the decay of $\mathrm{A}$ particles is completely ignored.
+For a suitable choice of hyperparameters $\alpha \approx 1.91\cdot 10^{-7}$, $\lambda=1$, and a cutoff $\kappa = 0.22$ a sparser solution can be obtained and also the decay reaction as indicated by the red rings in the figure can be recovered.
+The cutoff $\kappa$ is applied so that for all reactions with an estimated rate $\hat{\xi} < \kappa$ a rate of $\hat{\xi} = 0$ is assigned.
+A bubble whose size corresponds to a rate of $\kappa$ is depicted in Fig.~\ref{fig:case-1-sparsity-pattern} in green.
+
 	\begin{figure}
@@
 Blue circles are estimated by linear least-squares and orange circles depict rates which were obtained by solving the minimization problem (\ref{method:minimizationproblem}).
-The latter rates are subject to a cutoff $c=0.22$ corresponding to the green circle's area under which a sparse solution with the correct processes can be recovered.
+The latter rates are subject to a cutoff $\kappa=0.22$ corresponding to the green circle's area under which a sparse solution with the correct processes can be recovered.
 If a certain rate was estimated in both cases, two wedges instead of one circle are displayed.}
@@
 	\end{figure}
-
-	When applying no regularization, one can observe in Figure \ref{fig:case-1-sparsity-pattern} that the sparsity pattern does not match the generating reaction network and the reaction responsible for the decay of $\mathrm{A}$ particles is completely ignored.
-For a suitable choice of the hyperparameter $\alpha > 0$ and a cutoff one can obtain a sparser solution and also recover the decay reaction as indicated by the red rings in the figure.
 
@@
 	\section{Appendix}\label{sec:appendix}
-	\begin{colbox}{F8E0E0}
-		the reaction table needs an update
-	\end{colbox}
 	\begin{table}[h]
 		\centering
-		\scalebox{0.7}{
+		\scalebox{.75}{
 			\begin{tabular}{rclcc}
@@
 		}
-		\caption{Full set of ansatz reactions $\Theta$ used in Section \ref{sec:examples}.}
+		\caption{Full set of ansatz reactions $\Theta$ used in Sec.~\ref{sec:results}.}
 		\label{tab:reaction-library}

### 2018-08-27 docs/manuscript.tex
this should fix it
betterer code

@@
 	\section{Introduction}
+	\textcolor{red}{Cite and mention previous works.}
+
 	When presented with a time series of possibly noisy non-equilibrium concentration fluctuations of some species as output of, e.g., measurements from experiments or simulations that were parameterized by microscopic rates, one can ask for the corresponding macroscopic rates and a generating reaction network.
@@
 The parsimonious nature of the results avoids overfitting and provides interpretability.
-	In our application we, as opposed to the original method, do not only look for macroscopic rates of net species change but investigate the specific reactions that might have lead to the observations.
-	We demonstrate the algorithm on three toy problems: When there is no noise in the data we can find all relevant processes of the ground truth.
+	In our application we, as opposed to the SINDy method, estimate parameters which are coupled across the equations of the arising ODE system, i.e., we look for specific reactions and their rate constants that might have lead to the observations instead of net flux across species.
+	We demonstrate the algorithm on a biologically motivated reaction kinetics system in three different scenarios of measurement: When there is no noise in the data we can find all relevant processes of the ground truth.
 If there is noise in the data we converge to the correct reaction network and rates with decreasing noise.
-The last toy problem deals with the case that there are two realizations with different initial conditions, for which can also show convergence to the correct model with decreasing levels of noise.
+The last scenario deals with the case that there are two realizations with different initial conditions, for which one can also show convergence to the correct model with decreasing levels of noise.
 
 	\section{The method}
-	The underlying model is a law of mass action type dynamical system. To this end, let $S$ be the number of species, then the observed concentration at a time $t$ can be represented by a vector
+	The underlying model are reaction-rate equations subject to the law of mass action.
+To this end, let $S$ be the number of species, then the observed concentration at a time $t$ can be represented by a vector
 	\begin{align}
@@
 	\label{method:the-system}\end{align}
-	where $\xi_r$ are the to-be estimated macroscopic rates.
+	where $\xi_r$ are the to-be estimated macroscopic rate constants.
 
@@
 	\end{align}
-	Given this matrix, a library $\Theta(\textbf{X}) = \begin{pmatrix} \theta_1(\textbf{X}) & \theta_2(\textbf{X}) & \cdots & \theta_R(\textbf{X}) \end{pmatrix}$ of $R$ ansatz reactions can be proposed with corresponding reaction functions
+	Given this matrix, a library $\Theta: \mathbb{R}^{T\times S}\to \mathbb{R}^{T \times S\times R},\;\mathbf{A}\mapsto \begin{pmatrix} \theta_1(\mathbf{A}) & \theta_2(\mathbf{A}) & \cdots & \theta_R(\mathbf{A}) \end{pmatrix}$ of $R$ ansatz reactions can be proposed with corresponding reaction functions
 	\begin{align}
-		\theta_r(\textbf{X}) = \begin{pmatrix}
-		\textbf{y}_r(\textbf{X}_1)^T \\ \vdots \\ \textbf{y}_r(\textbf{X}_T)^T
+		\theta_r(\mathbf{A}) = \begin{pmatrix}
+		\textbf{y}_r(\mathbf{A}_{1*})^T \\ \vdots \\ \textbf{y}_r(\mathbf{A}_{T*})^T
 		\end{pmatrix}\in \mathbb{R}^{T\times S},\quad r=1,\ldots,R,
 	\label{method:the-reactions}\end{align}
-	where $\textbf{X}_i$ denotes the $i$-th row in $\textbf{X}$.
+	where $\textbf{A}_{i*}$ denotes the $i$-th row in $\textbf{A}$.
 Applying the concentration trajectory to the library yields $\Theta(\textbf{X})\in\mathbb{R}^{T\times S\times R}$.
@@
 
-	When presented with multiple trajectories the data matrices $\mathbf{X}$ and $\dot{\mathbf{X}}$ can be stacked and inserted into the minimization problem.
-
 	For $\alpha=0$ the minimization problem reduces to constrained least-squares.
@@
 Since only the concentration data $\mathbf{X}$ is available but not its temporal derivative $\dot{\mathbf{X}}$, it is approximated numerically by second order finite differences with the exception of boundary data.
+Once the pair $(\mathbf{X}, \dot{\mathbf{X}})$ is obtained, the problem becomes invariant under temporal reordering.
+Hence, when presented with multiple trajectories the data matrices $\mathbf{X}_i$ and $\dot{\mathbf{X}}_i$ can simply be concatenated.
 
@@
     To demonstrate the method a gene-regulatory network is estimated from time series of molecule-concentrations.
-The data of is generated by integrating the law of mass action equations of the underlying model and adding Gaussian noise.
 
@@
 		\mathrm{DNA}_i &\rightharpoonup \mathrm{DNA}_i + \mathrm{mRNA}_i &&\text{(transcription)},\\
-		\mathrm{mRNA}_i &\rightharpoonup \mathrm{mRNA}_i + \mathrm{Protein}_i &&\text{(translation)},\\
+		\mathrm{mRNA}_i &\rightharpoonup \mathrm{mRNA}_i + i &&\text{(translation)},\\
 		\mathrm{mRNA}_i &\rightharpoonup \emptyset &&\text{(decay of mRNA)},\\
-		\mathrm{Protein}_i &\rightharpoonup \emptyset &&\text{(decay of Protein)},
+		i &\rightharpoonup \emptyset &&\text{(decay of protein)},
     \end{align*}
@@
     \begin{align*}
-    i + \mathrm{mRNA}_j \rightharpoonup i &&&\text{(regulation by repression).}
+    i + \mathrm{mRNA}_j \rightharpoonup i.
     \end{align*}
 
-    In our example proteins of type $\mathrm{A}$ regulate the $\mathrm{mRNA}_\mathrm{B}$ molecules, proteins of type $\mathrm{B}$ regulate the $\mathrm{mRNA}_\mathrm{A}$ molecules and proteins of type $\mathrm{C}$ regulate the $\mathrm{mRNA}_\mathrm{A}$ molecules.
-The network of all species and reactions is depicted in Figure \ref{fig:network}.
+    In our example proteins of type $\mathrm{A}$ regulate the $\mathrm{mRNA}_\mathrm{B}$ molecules, proteins of type $\mathrm{B}$ regulate the $\mathrm{mRNA}_\mathrm{C}$ molecules and proteins of type $\mathrm{C}$ regulate the $\mathrm{mRNA}_\mathrm{A}$ molecules.
+The network of all species and reactions is depicted in Fig.~\ref{fig:network}.
 This serves as reference model to generate the time series of concentrations.
@@
 
-	\section{Results}\label{sec:results}
+	\section{Results} \label{sec:results}
 
-	The proposed estimation method is designed to find reaction networks and rates from experimental data.
+	The proposed estimation method is designed to find reaction networks and rate constants from experimental data.
 In the following the method is applied to data generated from the example described in Section \ref{sec:generegulation}.
 
-	First it is assumed that the concentration data can be obtained without noise and it is shown that while the minimization problem (\ref{method:minimizationproblem}) without regularization, i.e., $\alpha=0$, yields rates that fit well to the input data, the original rates cannot be recovered while adding $L_1$ regularization and applying a cutoff recovers the underlying reaction network \textcolor{red}{(and rates?)}.
+	In Sec.~\ref{sec:case-1} is assumed that the concentration data can be obtained without noise and it is shown that while the minimization problem (\ref{method:minimizationproblem}) without regularization, i.e., $\alpha=0$, yields rates that fit well to the input data, the original rates cannot be recovered while adding $L_1$ regularization and applying a cutoff recovers the underlying reaction network.
 
-	In Section \ref{sec:case-2} the example data was generated using the Gillespie SSA \textbf{[cite]}.
+	In Sec.~\ref{sec:case-2} the example data was generated using the Gillespie SSA \textbf{[cite]}.
 While the concentrations now contain stochastic noise, it is assumed that the initial state can be reproduced exactly and multiple measurements with that exact initial condition can be performed.
@@
 
-	The last example is described in Section \ref{sec:case-3} and is an extension of the situation in Section \ref{sec:case-2} to multiple initial conditions.
+	The last example is described in Sec.~\ref{sec:case-3} and is an extension of the situation in Sec.~\ref{sec:case-2} to multiple initial conditions.
 Due to more varied data and more statistics a good fit can be obtained with fewer measurements than in the case of just one initial condition.
@@
 
-	In this section noiseless data generated from the law of mass action system described in Section \ref{sec:generegulation} is applied to the minimization problem (\ref{method:minimizationproblem}).
-When setting the hyperparameter $\alpha=0$ one obtains least squares regression with a constraint on the non-negativity of the estimated reaction rates.
+	In this section noiseless data is generated from integrating the reaction-rate equation (\ref{method:the-system}) of the system described in Sec.~\ref{sec:generegulation}.
+The minimization problem (\ref{method:minimizationproblem}) is solved with respect to this data.
+When setting the hyperparameter $\alpha=0$ one obtains linear least-squares regression with a constraint on the non-negativity of the estimated reaction rates.
 
-	When applying no regularization, one can observe in Figure \ref{fig:case-1-sparsity-pattern} that the sparsity pattern does not match the generating reaction network and the reaction responsible for the decay of $\mathrm{A}$ particles is completely ignored.
-For a suitable choice of hyperparameters $\alpha \approx 1.91\cdot 10^{-7}$, $\lambda=1$, and a cutoff $\kappa = 0.22$ a sparser solution can be obtained and also the decay reaction as indicated by the red rings in the figure can be recovered.
+	When applying constrained linear least-squares regression, one can observe that the sparsity pattern does not match the generating reaction network and the reaction responsible for the decay of $\mathrm{A}$ particles is completely ignored (Fig.~\ref{fig:case-1-sparsity-pattern}).
+For a suitable choice of hyperparameters $\alpha \approx 1.91\cdot 10^{-7}$, $\lambda=1$, and a cutoff $\kappa = 0.22$ a sparse solution can be obtained and also the decay reaction can be recovered.
 The cutoff $\kappa$ is applied so that for all reactions with an estimated rate $\hat{\xi} < \kappa$ a rate of $\hat{\xi} = 0$ is assigned.
-A bubble whose size corresponds to a rate of $\kappa$ is depicted in Fig.~\ref{fig:case-1-sparsity-pattern} in green.
+A circle with size corresponding to a rate of $\kappa$ is depicted in Fig.~\ref{fig:case-1-sparsity-pattern} in green.
 
@@
 
-	\subsection{Multiple measurements at the same noise level}\label{sec:case-2}
+	\textbf{Todo}:
+	\begin{enumerate}
+		\item mention finite statistics
+		\item how does LSQ compensate for $A\to\emptyset$
+		\item ansatzfunctions used in $\Theta$
+	\end{enumerate}
+
+	\subsection{Data with stochastic noise}\label{sec:case-2}
+	For Gillespie, the initial condition contains hundreds of particles such that the average over many realizations converges against the reaction-rate equation.
 
@@
 		\item Same setup as in Section \ref{sec:case-1}
-		\item Cross validation with a shuffle split and 3 folds
-		\item comparison to LSQ: LSQ unreliable
+		\item def. of failure rate
+		\item $10$-fold cross validation discarding fold with test set at the (temporal) beginning (reason: more robust hyperparameter estimation as statisticial dependence of our data is higher at the initial equilibration phase, shuffle didn't really work well)
+		\item CV was independently carried out 10 times to obtain avg/std
+		\item comparison to LSQ: LSQ unreliable \textbf{[todo]}
 		\item in the limit of small noise one gets a good fit
-		\item obtain cutoff in noise free case, compute failure rate (number of false positives and negatives)
+		\item obtain cutoff in noise free case -> compute failure rate (number of false positives and negatives)
 	\end{itemize}
@@
 	\begin{itemize}
-		\item extend situation of Section \ref{sec:case-3} to multiple initial conditions
+		\item extend situation of Section \ref{sec:case-3} to multiple initial conditions (zipped / reorder)
+		\item $20$-fold CV discarding 1st fold (see sec 2)
 		\item compute failure rate
+		\item model can reliably be recovered in the limit of small noise
 	\end{itemize}
@@
 		\caption{Convergence of the estimation error when using two concatenated concentration timeseries with different initial conditions.
-\textbf{(a)}: A realization of one of the initial conditions that were used.
-It is identical to the initial conditions used in Sec.~\ref{sec:case-1} and Sec.~\ref{sec:case-2}.
-\textbf{(b)}: A realization of the other initial condition that was used.
-It prescribes positive initial concentrations for $\mathrm{mRNA}_\mathrm{A}$, $\mathrm{B}$, and $\mathrm{C}$ species.
-\textbf{(c)}: Analogously to Fig.~\ref{fig:case2-convergence} with the difference that 20-fold cross validation was used for hyperparameter estimation.}
+\textbf{(a)}: A realization of the concatenated timeseries.
+The first initial condition is identical to what was used in Sec.~\ref{sec:case-1} and Sec.~\ref{sec:case-2}.
+The second initial condition prescribes positive initial concentrations for $\mathrm{mRNA}_\mathrm{A}$, $\mathrm{B}$, and $\mathrm{C}$ species.
+\textbf{(b)}: Analogously to Fig.~\ref{fig:case2-convergence} with the difference that 20-fold cross validation was used for hyperparameter estimation.}
 	\end{figure}
@@
 
-	In two examples it was demonstrated that despite noisy data and unavailable derivative measurements, a parsimonious generating reaction network that is qualitatively able to explain the observed data can be estimated.
-	In particular it was shown in the first example that if there is no ambiguity in the underlying model and ansatz reaction library, the actual rates can be recovered with decreasing time step, i.e., increasing resolution of the jump process.
-	In the second example we could obtain an even simpler model than what was used to generate data by making use of sparse regression and cross-validation.
+	%In two examples it was demonstrated that despite noisy data and unavailable derivative measurements, a parsimonious generating reaction network that is qualitatively able to explain the observed data can be estimated.
+	%In particular it was shown in the first example that if there is no ambiguity in the underlying model and ansatz reaction library, the actual rates can be recovered with decreasing time step, i.e., increasing resolution of the jump process.
+	%In the second example we could obtain an even simpler model than what was used to generate data by making use of sparse regression and cross-validation.
 

### 2018-08-28 docs/manuscript.tex
lsq rates
wip

@@
 
-	For $\alpha=0$ the minimization problem reduces to constrained least-squares.
+	For $\alpha=0$ the minimization problem reduces to constrained least-squares (LSQ).
 In order to solve (\ref{method:minimizationproblem}) the numerical sequential least-squares minimizer SLSQP \cite{Kraft1988} is applied via the software package SciPy \cite{SciPy}.
@@
 		\centering
-		\includegraphics[width=.5\columnwidth]{./figures_tex/case2}
-		\caption{Convergence of the estimation error when estimating the system described in Sec.~\ref{sec:case-1} with decreasing levels of noise by application of the minimization problem (\ref{method:minimizationproblem}).
+		\includegraphics[width=.7\columnwidth]{./figures_tex/case2}
+		\caption{Convergence of the estimation error when estimating the system described in Sec.~\ref{sec:case-1} with decreasing levels of noise by application of the minimization problem (\ref{method:minimizationproblem}) for $\alpha=0$ (LSQ) and a $(\alpha, \lambda)$ combination found through $10$-fold cross-validation (regularized).
 The level of noise is regulated by the number of measurements being averaged.
-The hyperparameters for each level of noise were estimated using 10-fold cross validation.
-This was repeated 10 times giving rise to the mean and standard deviation depicted by dark blue graph and shaded blue area, respectively.
-The failure rate is the number of incorrectly identified processes.
-The estimation error is given by the $L_1$ difference of the generating reaction rates $\hat{\xi}$ and the estimated reaction rates $\xi_\mathrm{CV}$.}
+The procedure was repeated 10 times with different realizations giving rise to the mean and standard deviation depicted by dark blue graph and shaded blue area, respectively.
+The failure rate is the number of incorrectly identified processes up to the cutoff value introduced in Sec.~\ref{sec:case-1}.
+The estimation error is given by the $L_1$ difference of the generating reaction rates $\xi$ and the estimated reaction rates $\hat{\xi}$.}
 		\label{fig:case2-convergence}

### 2018-09-03 docs/manuscript.tex
fixed some stuff in the document

@@
-\documentclass[oneside, abstracton, titlepage]{scrartcl}
-
-\usepackage[left=2.5cm,right=2.6cm,top=3cm,bottom=3cm]{geometry}
-\usepackage[utf8]{inputenc}
-\usepackage[intlimits]{amsmath}
+%% LyX 2.3.0 created this file.  For more info, see http://www.lyx.org/.
+%% Do not edit unless you really know what you are doing.
+\documentclass[abstracton,titlepage]{scrartcl}
+\usepackage[utf8]{luainputenc}
+\usepackage{geometry}
+\geometry{verbose}
+\usepackage{color}
+\usepackage{booktabs}
+\usepackage{amsmath}
 \usepackage{amssymb}
-\usepackage{color}
-\usepackage{xcolor}
-\usepackage{subcaption}
-
 \usepackage{graphicx}
 
-\newsavebox{\selvestebox}
-\newenvironment{colbox}[1]
-{\newcommand\colboxcolor{#1}%
-	\begin{lrbox}{\selvestebox}%
-		\begin{minipage}{\dimexpr\columnwidth-2\fboxsep\relax}}
-		{\end{minipage}\end{lrbox}%
-	\begin{center}
-		\colorbox[HTML]{\colboxcolor}{\usebox{\selvestebox}}
-\end{center}}
-\newcommand\restr[2]{{% we make the whole thing an ordinary symbol
-		\left.\kern-\nulldelimiterspace % automatically resize the bar with \right
-		#1 % the function
-		\vphantom{\big|} % pretend it's a little taller at normal size
-		\right|_{#2} % this is the delimiter
-}}
+\makeatletter
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
+%% Because html converters don't know tabularnewline
+\providecommand{\tabularnewline}{\\}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
 
 \begin{document}
+
 \title{Discovering governing reactions from concentration data}
@@
 \section{Introduction}
+
 \textcolor{red}{Cite and mention previous works.}
 
-	When presented with a time series of possibly noisy non-equilibrium concentration fluctuations of some species as output of, e.g., measurements from experiments or simulations that were parameterized by microscopic rates, one can ask for the corresponding macroscopic rates and a generating reaction network.
-	In this paper we present an application of the shallow learning method SINDy \cite{Brunton2015}.
-By sparse regression, it is able to identify generating nonlinear dynamics in data that stems from dynamical systems.
-The parsimonious nature of the results avoids overfitting and provides interpretability.
-	In our application we, as opposed to the SINDy method, estimate parameters which are coupled across the equations of the arising ODE system, i.e., we look for specific reactions and their rate constants that might have lead to the observations instead of net flux across species.
-	We demonstrate the algorithm on a biologically motivated reaction kinetics system in three different scenarios of measurement: When there is no noise in the data we can find all relevant processes of the ground truth.
-If there is noise in the data we converge to the correct reaction network and rates with decreasing noise.
-The last scenario deals with the case that there are two realizations with different initial conditions, for which one can also show convergence to the correct model with decreasing levels of noise.
+When presented with a time series of possibly noisy non-equilibrium
+concentration fluctuations of some species as output of, e.g., measurements
+from experiments or simulations that were parameterized by microscopic
+rates, one can ask for the corresponding macroscopic rates and a generating
+reaction network. In this paper we present an application of the shallow
+learning method SINDy~\cite{Brunton2015}. By sparse regression,
+it is able to identify generating nonlinear dynamics in data that
+stems from dynamical systems. The parsimonious nature of the results
+avoids overfitting and provides interpretability. In our application
+we, as opposed to the SINDy method, estimate parameters which are
+coupled across the equations of the arising ODE system, i.e., we look
+for specific reactions and their rate constants that might have lead
+to the observations instead of net flux across species. We demonstrate
+the algorithm on a biologically motivated reaction kinetics system
+in three different scenarios of measurement: When there is no noise
+in the data we can find all relevant processes of the ground truth.
+If there is noise in the data we converge to the correct reaction
+network and rates with decreasing noise. The last scenario deals with
+the case that there are two realizations with different initial conditions,
+for which one can also show convergence to the correct model with
+decreasing levels of noise.
 
-	\section{The method}
-	The underlying model are reaction-rate equations subject to the law of mass action.
-To this end, let $S$ be the number of species, then the observed concentration at a time $t$ can be represented by a vector
+\section{Sparse learning of reaction kinetics}
+
+The underlying model are reaction-rate equations subject to the law
+of mass action. To this end, let $S$ be the number of species, then
+the observed concentration at a time $t$ can be represented by a
+vector 
 \begin{align}
-	\mathbf{x}(t)=\begin{pmatrix}
-	x_1(t)\\ \vdots \\ x_S(t)
-	\end{pmatrix}\in \mathbb{R}^S.
+\mathbf{x}(t)=\begin{pmatrix}x_{1}(t)\\
+\vdots\\
+x_{S}(t)
+\end{pmatrix}\in\mathbb{R}^{S}.
 \end{align}
-	Further, one can choose $R$ possible ansatz reactions with their respective reaction functions
+Further, one can choose $R$ possible ansatz reactions with their
+respective reaction functions 
 \begin{align}
-	\textbf{y}_r(\textbf{x}(t))=\begin{pmatrix}
-	y_{r,1}(\textbf{x}(t)) \\ \vdots \\ y_{r,S}(\textbf{x}(t))
+\textbf{y}_{r}(\textbf{x}(t))=\begin{pmatrix}y_{r,1}(\textbf{x}(t))\\
+\vdots\\
+y_{r,S}(\textbf{x}(t))
 \end{pmatrix},\quad r=1,\ldots,R,
 \end{align}
-	so that the change of concentration for species $i$ at time $t$ is represented by the dynamical system
+so that the change of concentration for species $i$ at time $t$
+is represented by the dynamical system 
 \begin{align}
-	\dot{\textbf{x}}_i(t) = \sum_{r=1}^{R}y_{r,i}(\textbf{x}(t))\xi_r,\quad i=1,\ldots, S,
-	\label{method:the-system}\end{align}
-	where $\xi_r$ are the to-be estimated macroscopic rate constants.
+\dot{\textbf{x}}_{i}(t)=\sum_{r=1}^{R}y_{r,i}(\textbf{x}(t))\xi_{r},\quad i=1,\ldots,S,\label{method:the-system}
+\end{align}
+where $\xi_{r}$ are the to-be estimated macroscopic rate constants.
 
-	When presented with a time series consisting of $T$ observations at points in time $t_1<\ldots < t_T$, the data can be represented as a matrix
+When presented with a time series consisting of $T$ observations
+at points in time $t_{1}<\cdots<t_{T}$, the data can be represented
+as a matrix 
 \begin{align}
-	\textbf{X} = \begin{pmatrix}
-		x_1(t_1) & x_2(t_1) & \cdots & x_S(t_1) \\
-		x_1(t_2) & x_2(t_2) & \cdots & x_S(t_2) \\
-		\vdots   & \vdots   & \ddots & \vdots   \\
-		x_1(t_T) & x_2(t_T) & \cdots & x_S(t_T)
-	\end{pmatrix} \in \mathbb{R}^{T\times S}.
+\textbf{X}=\begin{pmatrix}x_{1}(t_{1}) & x_{2}(t_{1}) & \cdots & x_{S}(t_{1})\\
+x_{1}(t_{2}) & x_{2}(t_{2}) & \cdots & x_{S}(t_{2})\\
+\vdots & \vdots & \ddots & \vdots\\
+x_{1}(t_{T}) & x_{2}(t_{T}) & \cdots & x_{S}(t_{T})
+\end{pmatrix}\in\mathbb{R}^{T\times S}.
 \end{align}
-	Given this matrix, a library $\Theta: \mathbb{R}^{T\times S}\to \mathbb{R}^{T \times S\times R},\;\mathbf{A}\mapsto \begin{pmatrix} \theta_1(\mathbf{A}) & \theta_2(\mathbf{A}) & \cdots & \theta_R(\mathbf{A}) \end{pmatrix}$ of $R$ ansatz reactions can be proposed with corresponding reaction functions
+Given this matrix, a library $\Theta:\mathbb{R}^{T\times S}\to\mathbb{R}^{T\times S\times R},\;\mathbf{A}\mapsto\begin{pmatrix}\theta_{1}(\mathbf{A}) & \theta_{2}(\mathbf{A}) & \cdots & \theta_{R}(\mathbf{A})\end{pmatrix}$
+of $R$ ansatz reactions can be proposed with corresponding reaction
+functions 
 \begin{align}
-		\theta_r(\mathbf{A}) = \begin{pmatrix}
-		\textbf{y}_r(\mathbf{A}_{1*})^T \\ \vdots \\ \textbf{y}_r(\mathbf{A}_{T*})^T
-		\end{pmatrix}\in \mathbb{R}^{T\times S},\quad r=1,\ldots,R,
-	\label{method:the-reactions}\end{align}
+\theta_{r}(\mathbf{A})=\begin{pmatrix}\textbf{y}_{r}(\mathbf{A}_{1*})^{T}\\
+\vdots\\
+\textbf{y}_{r}(\mathbf{A}_{T*})^{T}
+\end{pmatrix}\in\mathbb{R}^{T\times S},\quad r=1,\ldots,R,\label{method:the-reactions}
+\end{align}
 where $\textbf{A}_{i*}$ denotes the $i$-th row in $\textbf{A}$.
 Applying the concentration trajectory to the library yields $\Theta(\textbf{X})\in\mathbb{R}^{T\times S\times R}$.
-Following the approach of SINDy, the goal is to find coefficients $\Xi = \begin{pmatrix} \xi_1 & \xi_2 & \cdots & \xi_R
-	\end{pmatrix}^T$, so that
+Following the approach of SINDy, the goal is to find coefficients
+$\Xi=\begin{pmatrix}\xi_{1} & \xi_{2} & \cdots & \xi_{R}\end{pmatrix}^{T}$,
+so that 
 \begin{align}
-	\dot{\textbf{X}} = \Theta(\textbf{X})\Xi = \sum_{r=1}^{R}\theta_r(\textbf{X})\xi_r.
+\dot{\textbf{X}}=\Theta(\textbf{X})\Xi=\sum_{r=1}^{R}\theta_{r}(\textbf{X})\xi_{r}.
 \end{align}
-	In particular, the system is linear in the coefficients $\Xi$, which makes regression tools such as elastic net regularization \cite{Zou2005} applicable.
-To this end, one can consider the minimization problem to find $\hat{\Xi}$ such that
+In particular, the system is linear in the coefficients $\Xi$, which
+makes regression tools such as elastic net regularization~\cite{Zou2005}
+applicable. To this end, one can consider the minimization problem
+to find $\hat{\Xi}$ such that 
 \begin{align}
-		\hat{\Xi} = \underset{\Xi}{\arg\min}\left( \frac{1}{2T}\left\| \dot{\textbf{X}} - \Theta(\textbf{X})\Xi \right\|_F^2 + \alpha\lambda\|\Xi\|_1 + \alpha(1-\lambda)\|\Xi\|_2^2 \right) \quad \text{subject to }\Xi \geq 0,
-	\label{method:minimizationproblem}\end{align}
-	where $\|\cdot\|_F$ denotes the Frobenius norm, $\lambda\in[0,1]$ a hyperparameter that interpolates linearly between LASSO \cite{Tibshirani1996, Hastie2009} and Ridge \cite{Hoerl1} methods, and $\alpha\geq 0$ is a hyperparameter that, depending on $\lambda$, can induce sparsity and give preference to smaller solutions in the $L_1$ or $L_2$ sense.
+\hat{\Xi}=\underset{\Xi}{\arg\min}\left(\frac{1}{2T}\left\Vert \dot{\textbf{X}}-\Theta(\textbf{X})\Xi\right\Vert _{F}^{2}+\alpha\lambda\|\Xi\|_{1}+\alpha(1-\lambda)\|\Xi\|_{2}^{2}\right)\quad\text{subject to }\Xi\geq0,\label{method:minimizationproblem}
+\end{align}
+where $\|\cdot\|_{F}$ denotes the Frobenius norm, $\lambda\in[0,1]$
+a hyperparameter that interpolates linearly between LASSO~\cite{Tibshirani1996,Hastie2009}
+and Ridge~\cite{Hoerl1} methods, and $\alpha\geq0$ is a hyperparameter
+that, depending on $\lambda$, can induce sparsity and give preference
+to smaller solutions in the $L_{1}$ or $L_{2}$ sense.
 
-	For $\alpha=0$ the minimization problem reduces to constrained least-squares (LSQ).
-In order to solve (\ref{method:minimizationproblem}) the numerical sequential least-squares minimizer SLSQP \cite{Kraft1988} is applied via the software package SciPy \cite{SciPy}.
-Since only the concentration data $\mathbf{X}$ is available but not its temporal derivative $\dot{\mathbf{X}}$, it is approximated numerically by second order finite differences with the exception of boundary data.
-Once the pair $(\mathbf{X}, \dot{\mathbf{X}})$ is obtained, the problem becomes invariant under temporal reordering.
-Hence, when presented with multiple trajectories the data matrices $\mathbf{X}_i$ and $\dot{\mathbf{X}}_i$ can simply be concatenated.
+For $\alpha=0$ the minimization problem reduces to constrained least-squares
+(LSQ). In order to solve (\ref{method:minimizationproblem}) the numerical
+sequential least-squares minimizer SLSQP~\cite{Kraft1988} is applied
+via the software package SciPy~\cite{SciPy}. Since only the concentration
+data $\mathbf{X}$ is available but not its temporal derivative $\dot{\mathbf{X}}$,
+it is approximated numerically by second order finite differences
+with the exception of boundary data. Once the pair $(\mathbf{X},\dot{\mathbf{X}})$
+is obtained, the problem becomes invariant under temporal reordering.
+Hence, when presented with multiple trajectories the data matrices
+$\mathbf{X}_{i}$ and $\dot{\mathbf{X}}_{i}$ can simply be concatenated.
 
-    \section{Example system: Regulation of gene expression}\label{sec:generegulation}
+\section{Results\label{sec:results}}
 
-    To demonstrate the method a gene-regulatory network is estimated from time series of molecule-concentrations.
-
-    To this end, let $\mathrm{A}$, $\mathrm{B}$, and $\mathrm{C}$ be three species of proteins which are being translated from a corresponding $\mathrm{mRNA}$ molecule.
-Each $\mathrm{mRNA}$ in turn has a corresponding $\mathrm{DNA}$ which it is transcribed from.
-The proteins and $\mathrm{mRNA}$ molecules decay over time whereas the $\mathrm{DNA}$ concentration remains constant.
+To demonstrate the method a gene-regulatory network is estimated from
+time series of molecule-concentrations. To this end, let $S:=\{A,B,C\}$
+be a set of three species of proteins which are being translated from
+a corresponding $\mathrm{mRNA}$ molecule. Each $\mathrm{mRNA}$ in
+turn has a corresponding $\mathrm{DNA}$ which it is transcribed from.
+The proteins and $\mathrm{mRNA}$ molecules decay over time whereas
+the $\mathrm{DNA}$ concentration remains constant.
 
@@
 \begin{align*}
-		\mathrm{DNA}_i &\rightharpoonup \mathrm{DNA}_i + \mathrm{mRNA}_i &&\text{(transcription)},\\
-		\mathrm{mRNA}_i &\rightharpoonup \mathrm{mRNA}_i + i &&\text{(translation)},\\
-		\mathrm{mRNA}_i &\rightharpoonup \emptyset &&\text{(decay of mRNA)},\\
-		i &\rightharpoonup \emptyset &&\text{(decay of protein)},
+\mathrm{DNA}_{i} & \rightharpoonup\mathrm{DNA}_{i}+\mathrm{mRNA}_{i} &  & \text{(transcription)},\\
+\mathrm{mRNA}_{i} & \rightharpoonup\mathrm{mRNA}_{i}+i &  & \text{(translation)},\\
+\mathrm{mRNA}_{i} & \rightharpoonup\emptyset &  & \text{(decay of mRNA)},\\
+i & \rightharpoonup\emptyset &  & \text{(decay of protein)},\\
+i+\mathrm{mRNA}_{j} & \rightharpoonup i &  & \text{(regulation of protein }j\in S\text{),}
 \end{align*}
-    for each of the species $i\in S:=\{\mathrm{A},\mathrm{B},\mathrm{C}\}$.
-
-    The protein concentrations of $\mathrm{A}$, $\mathrm{B}$, and $\mathrm{C}$ regulate each other in a negative way by hindering the transcription process.
-In the law of mass action model we account for this by a reaction.
-A repression of species $j\in S$ by species $i\in S$ can be modeled as
-    \begin{align*}
-    i + \mathrm{mRNA}_j \rightharpoonup i.
-    \end{align*}
-
-    In our example proteins of type $\mathrm{A}$ regulate the $\mathrm{mRNA}_\mathrm{B}$ molecules, proteins of type $\mathrm{B}$ regulate the $\mathrm{mRNA}_\mathrm{C}$ molecules and proteins of type $\mathrm{C}$ regulate the $\mathrm{mRNA}_\mathrm{A}$ molecules.
-The network of all species and reactions is depicted in Fig.~\ref{fig:network}.
-This serves as reference model to generate the time series of concentrations.
+for each of the species $i\in S$. The regulation reaction models
+a repression of species $j$ by hindering its transcription process.
+In our example proteins of type $\mathrm{A}$ regulate the $\mathrm{mRNA}_{\mathrm{B}}$
+molecules, proteins of type $\mathrm{B}$ regulate the $\mathrm{mRNA}_{\mathrm{C}}$
+molecules and proteins of type $\mathrm{C}$ regulate the $\mathrm{mRNA}_{\mathrm{A}}$
+molecules. The network of all species and reactions is depicted in
+Fig.~\ref{fig:network}. This serves as reference model to generate
+time series of concentrations. 
 \begin{figure}
-        \centering
-        \includegraphics[width=.5\textwidth]{./figures_tex/scheme.pdf}
-        \caption{The regulation network example described in Sec.~\ref{sec:generegulation}.
-\textbf{(a)}: Each circle depicts a species, each arrow corresponds to one reaction.
-Blue arrows denote transcription from DNA to $\mathrm{mRNA}$, green arrows denote translation from $\mathrm{mRNA}$ to protein, and red arrows denote the regulatory network.
-\textbf{(b)}: Concentration timeseries generated from the reaction network shown in~(a).
-The initial condition prescribes positive concentration values only for $\mathrm{B}$ protein and $\mathrm{mRNA}_\mathrm{A}$ species.
-This initial condition is used in the subsequent sections for further analysis.}
-        \label{fig:network}
+\centering \includegraphics[width=1\columnwidth]{figures_tex/scheme}
+\caption{\label{fig:network}The regulation network example described in Sec.~\ref{sec:results}.
+\textbf{(a)}: Each circle depicts a species, each arrow corresponds
+to one reaction. Blue arrows denote transcription from DNA to $\mathrm{mRNA}$,
+green arrows denote translation from $\mathrm{mRNA}$ to protein,
+and red arrows denote the regulatory network. \textbf{(b)}: Concentration
+time series generated from the reaction network shown in~(a). The
+initial condition prescribes positive concentration values only for
+$\mathrm{B}$ protein and $\mathrm{mRNA}_{\mathrm{A}}$ species. This
+initial condition is used in the subsequent sections for further analysis.}
 \end{figure}
 
-	\section{Results} \label{sec:results}
+These time series of concentrations are then applied to the proposed
+estimation method which is designed to find reaction networks and
+rate constants from experimental data.
 
-	The proposed estimation method is designed to find reaction networks and rate constants from experimental data.
-In the following the method is applied to data generated from the example described in Section \ref{sec:generegulation}.
+\subsection{The noiseless case\label{sec:case-1}}
 
-	In Sec.~\ref{sec:case-1} is assumed that the concentration data can be obtained without noise and it is shown that while the minimization problem (\ref{method:minimizationproblem}) without regularization, i.e., $\alpha=0$, yields rates that fit well to the input data, the original rates cannot be recovered while adding $L_1$ regularization and applying a cutoff recovers the underlying reaction network.
+In this section noiseless data is generated from integrating the reaction-rate
+equation (\ref{method:the-system}) of the system described in Sec.~\ref{sec:results}.
+The minimization problem (\ref{method:minimizationproblem}) is solved
+with respect to this data. When setting the hyperparameter $\alpha=0$
+one obtains linear least-squares regression with a constraint on the
+non-negativity of the estimated reaction rates.
 
-	In Sec.~\ref{sec:case-2} the example data was generated using the Gillespie SSA \textbf{[cite]}.
-While the concentrations now contain stochastic noise, it is assumed that the initial state can be reproduced exactly and multiple measurements with that exact initial condition can be performed.
-It is shown that in the limit of many measurements a good fit can reliably be obtained.
-
-	The last example is described in Sec.~\ref{sec:case-3} and is an extension of the situation in Sec.~\ref{sec:case-2} to multiple initial conditions.
-Due to more varied data and more statistics a good fit can be obtained with fewer measurements than in the case of just one initial condition.
-
-	%measurement scenarios are considered that differ in how many time series are available and the level of noise that the data is subject to:
-	%\begin{enumerate}
-	%	\item The initial state of the measurement apparatus can be reproduced exactly.
-There is only one set of generating reaction rates.
-Each frame of a time series is subject to noise.
-The outcome of this scenario is one time series of concentrations with a certain noise-level.
-	%	\item The process is observed multiple times.
-The initial state of the measurement apparatus differs in each observation.
-The outcome of this scenario is multiple time series each with a certain noise level and different initial states.
-	%\item The process is observed multiple times.
-The initial state of the measurement apparatus differs in each observation.
-Additionally the generating reaction rates differ slightly in each observation.
-The outcome of this scenario is multiple time series each with a given noise level, different initial states and slightly perturbed underlying reaction rates.
-	%\end{enumerate}
-
-	\subsection{The noiseless case}\label{sec:case-1}
-
-	In this section noiseless data is generated from integrating the reaction-rate equation (\ref{method:the-system}) of the system described in Sec.~\ref{sec:generegulation}.
-The minimization problem (\ref{method:minimizationproblem}) is solved with respect to this data.
-When setting the hyperparameter $\alpha=0$ one obtains linear least-squares regression with a constraint on the non-negativity of the estimated reaction rates.
-
-	When applying constrained linear least-squares regression, one can observe that the sparsity pattern does not match the generating reaction network and the reaction responsible for the decay of $\mathrm{A}$ particles is completely ignored (Fig.~\ref{fig:case-1-sparsity-pattern}).
-For a suitable choice of hyperparameters $\alpha \approx 1.91\cdot 10^{-7}$, $\lambda=1$, and a cutoff $\kappa = 0.22$ a sparse solution can be obtained and also the decay reaction can be recovered.
-The cutoff $\kappa$ is applied so that for all reactions with an estimated rate $\hat{\xi} < \kappa$ a rate of $\hat{\xi} = 0$ is assigned.
-A circle with size corresponding to a rate of $\kappa$ is depicted in Fig.~\ref{fig:case-1-sparsity-pattern} in green.
+When applying constrained linear least-squares regression, one can
+observe that the sparsity pattern does not match the generating reaction
+network and the reaction responsible for the decay of $\mathrm{A}$
+particles is completely ignored (Fig.~\ref{fig:case-1-sparsity-pattern}).
+For a suitable choice of hyperparameters $\alpha\approx1.91\cdot10^{-7}$,
+$\lambda=1$, and a cutoff $\kappa=0.22$ a sparse solution can be
+obtained and also the decay reaction can be recovered. The cutoff
+$\kappa$ is applied so that for all reactions with an estimated rate
+$\hat{\xi}<\kappa$ a rate of $\hat{\xi}=0$ is assigned. A circle
+with size corresponding to a rate of $\kappa$ is depicted in Fig.~\ref{fig:case-1-sparsity-pattern}
+in green.
 
 \begin{figure}
-		\centering
-		\includegraphics[width=.9\textwidth]{./figures_tex/sparsity_pattern}
-		\caption{Estimated reaction rates in the system described in Sec.~\ref{sec:case-1}.
-The y and x axes contain reaction educts and products, respectively.
-A circle at position $(i,j)$ represents a reaction $i\rightharpoonup j$ whose rate has a linear relation with the area of the circle.
-The black outlines denote the reactions with which the system was generated and contain the respective rate value.
-Red crosses denote reactions that were used as additional ansatz reactions.
-Blue circles are estimated by linear least-squares and orange circles depict rates which were obtained by solving the minimization problem (\ref{method:minimizationproblem}).
-The latter rates are subject to a cutoff $\kappa=0.22$ corresponding to the green circle's area under which a sparse solution with the correct processes can be recovered.
-If a certain rate was estimated in both cases, two wedges instead of one circle are displayed.}
-		\label{fig:case-1-sparsity-pattern}
+\centering \includegraphics[width=0.9\linewidth]{figures_tex/sparsity_pattern}
+\caption{\label{fig:case-1-sparsity-pattern}Estimated reaction rates in the
+system described in Sec.~\ref{sec:case-1}. The y and x axes contain
+reaction educts and products, respectively. A circle at position $(i,j)$
+represents a reaction $i\rightharpoonup j$ whose rate has a linear
+relation with the area of the circle. The black outlines denote the
+reactions with which the system was generated and contain the respective
+rate value. Red crosses denote reactions that were used as additional
+ansatz reactions. Blue circles are estimated by linear least-squares
+and orange circles depict rates which were obtained by solving the
+minimization problem (\ref{method:minimizationproblem}). The latter
+rates are subject to a cutoff $\kappa=0.22$ corresponding to the
+green circle's area under which a sparse solution with the correct
+processes can be recovered. If a certain rate was estimated in both
+cases, two wedges instead of one circle are displayed.}
 \end{figure}
@@
 
-	\subsection{Data with stochastic noise}\label{sec:case-2}
-	For Gillespie, the initial condition contains hundreds of particles such that the average over many realizations converges against the reaction-rate equation.
+\subsection{Data with stochastic noise\label{sec:case-2} }
 
+For Gillespie, the initial condition contains hundreds of particles
+such that the average over many realizations converges against the
+reaction-rate equation.
 \begin{itemize}
-		\item Same setup as in Section \ref{sec:case-1}
-		\item def. of failure rate
-		\item $10$-fold cross validation discarding fold with test set at the (temporal) beginning (reason: more robust hyperparameter estimation as statisticial dependence of our data is higher at the initial equilibration phase, shuffle didn't really work well)
+\item Same setup as in Section~\ref{sec:case-1} 
+\item definition of failure rate 
+\item $10$-fold cross validation discarding fold with test set at the (temporal)
+beginning (reason: more robust hyperparameter estimation as statistical
+dependence of our data is higher at the initial equilibration phase,
+shuffle didn't really work well) 
 \item CV was independently carried out 10 times to obtain avg/std 
-		\item comparison to LSQ: LSQ unreliable \textbf{[todo]}
+\item LSQ unreliable in comparison
 \item in the limit of small noise one gets a good fit 
-		\item obtain cutoff in noise free case -> compute failure rate (number of false positives and negatives)
+\item obtain cutoff in noise free case -> compute failure rate (number of
+false positives and negatives) 
 \end{itemize}
-
 \begin{figure}
-		\centering
-		\includegraphics[width=.7\columnwidth]{./figures_tex/case2}
-		\caption{Convergence of the estimation error when estimating the system described in Sec.~\ref{sec:case-1} with decreasing levels of noise by application of the minimization problem (\ref{method:minimizationproblem}) for $\alpha=0$ (LSQ) and a $(\alpha, \lambda)$ combination found through $10$-fold cross-validation (regularized).
-The level of noise is regulated by the number of measurements being averaged.
-The procedure was repeated 10 times with different realizations giving rise to the mean and standard deviation depicted by dark blue graph and shaded blue area, respectively.
-The failure rate is the number of incorrectly identified processes up to the cutoff value introduced in Sec.~\ref{sec:case-1}.
-The estimation error is given by the $L_1$ difference of the generating reaction rates $\xi$ and the estimated reaction rates $\hat{\xi}$.}
-		\label{fig:case2-convergence}
+\centering \includegraphics[width=1\columnwidth]{figures_tex/case2}
+\caption{\label{fig:case2-convergence}Convergence of the estimation error
+when estimating the system described in Sec.~\ref{sec:case-1} with
+decreasing levels of noise by application of the minimization problem
+(\ref{method:minimizationproblem}) for $\alpha=0$ (LSQ) and a $(\alpha,\lambda)$
+combination found through $10$-fold cross-validation (regularized).
+The level of noise is regulated by the number of measurements being
+averaged. The procedure was repeated 10 times with different realizations
+giving rise to the mean and standard deviation depicted by dark blue
+graph and shaded blue area, respectively. The failure rate is the
+number of incorrectly identified processes up to the cutoff value
+introduced in Sec.~\ref{sec:case-1}. The estimation error is given
+by the $L_{1}$ difference of the generating reaction rates $\xi$
+and the estimated reaction rates $\hat{\xi}$.}
 \end{figure}
 
-	\subsection{Multiple initial conditions}\label{sec:case-3}
-
+\subsection{Multiple initial conditions\label{sec:case-3}}
 \begin{itemize}
-		\item extend situation of Section \ref{sec:case-3} to multiple initial conditions (zipped / reorder)
-		\item $20$-fold CV discarding 1st fold (see sec 2)
+\item extend situation of Section~\ref{sec:case-3} to multiple initial
+conditions (zipped / reorder) 
+\item $20$-fold CV discarding 1st fold (see sec 4.2) 
 \item compute failure rate 
@@
 \end{itemize}
-
 \begin{figure}
-		\centering
-		\includegraphics[width=\columnwidth]{./figures_tex/case3}
-		\caption{Convergence of the estimation error when using two concatenated concentration timeseries with different initial conditions.
-\textbf{(a)}: A realization of the concatenated timeseries.
-The first initial condition is identical to what was used in Sec.~\ref{sec:case-1} and Sec.~\ref{sec:case-2}.
-The second initial condition prescribes positive initial concentrations for $\mathrm{mRNA}_\mathrm{A}$, $\mathrm{B}$, and $\mathrm{C}$ species.
-\textbf{(b)}: Analogously to Fig.~\ref{fig:case2-convergence} with the difference that 20-fold cross validation was used for hyperparameter estimation.}
+\centering \includegraphics[width=1\columnwidth]{figures_tex/case3}
+\caption{Convergence of the estimation error when using two concatenated concentration
+time series with different initial conditions. \textbf{(a)}: A realization
+of the concatenated timeseries. The first initial condition is identical
+to what was used in Sec.~\ref{sec:case-1} and Sec.~\ref{sec:case-2}.
+The second initial condition prescribes positive initial concentrations
+for $\mathrm{mRNA}_{\mathrm{A}}$, $\mathrm{B}$, and $\mathrm{C}$
+species. \textbf{(b)}: Analogously to Fig.~\ref{fig:case2-convergence}
+with the difference that 20-fold cross validation was used for hyperparameter
+estimation.}
 \end{figure}
@@
 \section{Conclusion}
-	\begin{colbox}{F8E0E0}
-		todo
-	\end{colbox}
-	In this work we have successfully applied and extended the SINDy method to not only parsimoniously detect potentially nonlinear terms in a dynamical system from noisy data, but also yield, in this case, a sparse set of rates with respect to generating reactions (\ref{method:the-reactions}).
+
+In this work we have successfully applied and extended the SINDy method
+to not only parsimoniously detect potentially nonlinear terms in a
+dynamical system from noisy data, but also yield, in this case, a
+sparse set of rates with respect to generating reactions (\ref{method:the-reactions}).
 
@@
 
-	\newpage
-% 	\bibliographystyle{alpha}
-	\bibliographystyle{abbrv}
+\newpage{} \bibliographystyle{abbrv}
 \bibliography{bibliography.bib}
-	\newpage	
+ \newpage{}
+
 \section{Appendix}\label{sec:appendix}
-	\begin{table}[h]
+	\begin{table}[h!]
 		\centering
-		\scalebox{.75}{
-			\begin{tabular}{rclcc}
-				\multicolumn{3}{c}{Reaction} & rate & description \\ \hline\noalign{\vskip .1cm}
-				$\mathrm{DNA}_\mathrm{A}$ &$\rightharpoonup$& $\mathrm{DNA}_\mathrm{A} + \mathrm{mRNA}_\mathrm{A}$ & $k_1 = 1.8$ & transcription of $\mathrm{mRNA}_\mathrm{A}$\\
-				$\mathrm{mRNA}_\mathrm{A}$ &$\rightharpoonup$& $\mathrm{mRNA}_\mathrm{A} + \mathrm{A}$ & $k_2 = 2.1$ & translation of $\mathrm{A}$ proteins\\
-				$\mathrm{mRNA}_\mathrm{A}$ &$\rightharpoonup$& $\emptyset$ & $k_3 = 1.3$ & $\mathrm{mRNA}_\mathrm{A}$ decay\\
-				$\mathrm{A}$ &$\rightharpoonup$& $\emptyset$ & $k_4 = 1.5$ & decay of $\mathrm{A}$ proteins\\
-				$\mathrm{DNA}_\mathrm{B}$ &$\rightharpoonup$& $\mathrm{DNA}_\mathrm{B} + \mathrm{mRNA}_\mathrm{B}$ & $k_5 = 2.2$ & transcription of $\mathrm{mRNA}_\mathrm{B}$\\
-				$\mathrm{mRNA}_\mathrm{B}$ &$\rightharpoonup$& $\mathrm{mRNA}_\mathrm{B} + \mathrm{B}$ & $k_6 = 2.0$ & translation of $\mathrm{B}$ proteins\\
-				$\mathrm{mRNA}_\mathrm{B}$ &$\rightharpoonup$& $\emptyset$ & $k_7 = 2.0$ & $\mathrm{mRNA}_\mathrm{B}$ decay\\
-				$\mathrm{B}$ &$\rightharpoonup$& $\emptyset$ & $k_8 = 2.5$ & decay of $\mathrm{B}$ proteins\\
-				$\mathrm{DNA}_\mathrm{C}$ &$\rightharpoonup$& $\mathrm{DNA}_\mathrm{C} + \mathrm{mRNA}_\mathrm{C}$ & $k_9 = 3.2$ & transcription of $\mathrm{mRNA}_\mathrm{C}$\\
-				$\mathrm{mRNA}_\mathrm{C}$ &$\rightharpoonup$& $\mathrm{mRNA}_\mathrm{C} + \mathrm{C}$ & $k_{10} = 3.0$ & translation of $\mathrm{C}$ proteins\\
-				$\mathrm{mRNA}_\mathrm{C}$ &$\rightharpoonup$& $\emptyset$ & $k_{11} = 2.3$ & $\mathrm{mRNA}_\mathrm{C}$ decay\\
-				$\mathrm{C}$ &$\rightharpoonup$& $\emptyset$ & $k_{12} = 2.5$ & decay of $\mathrm{C}$ proteins\\
-				$\mathrm{mRNA}_\mathrm{A} + \mathrm{A}$ &$\rightharpoonup$& $\mathrm{A}$ & $k_{13} = 0$ & self regulation of $\mathrm{A}$ proteins\\
-				$\mathrm{mRNA}_\mathrm{B} + \mathrm{B}$ &$\rightharpoonup$& $\mathrm{B}$ & $k_{14} = 0$ & self regulation of $\mathrm{B}$ proteins\\
-				$\mathrm{mRNA}_\mathrm{C} + \mathrm{C}$ &$\rightharpoonup$& $\mathrm{C}$ & $k_{15} = 0$ & self regulation of $\mathrm{C}$ proteins\\
-				$\mathrm{mRNA}_\mathrm{B} + \mathrm{A}$ &$\rightharpoonup$& $\mathrm{A}$ & $k_{16} = 0$ & cyclic regulation of $\mathrm{A}$ proteins\\
-				$\mathrm{mRNA}_\mathrm{C} + \mathrm{B}$ &$\rightharpoonup$& $\mathrm{B}$ & $k_{17} = 0$ & cyclic regulation of $\mathrm{B}$ proteins\\
-				$\mathrm{mRNA}_\mathrm{A} + \mathrm{C}$ &$\rightharpoonup$& $\mathrm{C}$ & $k_{18} = 0$ & cyclic regulation of $\mathrm{C}$ proteins\\
-				$\mathrm{mRNA}_\mathrm{C} + \mathrm{A}$ &$\rightharpoonup$& $\mathrm{A}$ & $k_{16} = 6.0$ & cyclic regulation of $\mathrm{A}$ proteins\\
-				$\mathrm{mRNA}_\mathrm{B} + \mathrm{C}$ &$\rightharpoonup$& $\mathrm{C}$ & $k_{17} = 4.0$ & cyclic regulation of $\mathrm{C}$ proteins\\
-				$\mathrm{mRNA}_\mathrm{A} + \mathrm{B}$ &$\rightharpoonup$& $\mathrm{B}$ & $k_{18} = 3.0$ & cyclic regulation of $\mathrm{B}$ proteins\\
-				$\mathrm{mRNA}_\mathrm{A} + \mathrm{A}$ &$\rightharpoonup$& $\mathrm{mRNA}_\mathrm{A}$ & $k_{19} = 0$ & artificial fusion\\
-				$\mathrm{mRNA}_\mathrm{B} + \mathrm{B}$ &$\rightharpoonup$& $\mathrm{mRNA}_\mathrm{B}$ & $k_{20} = 0$ & artificial fusion\\
-				$\mathrm{mRNA}_\mathrm{A} + \mathrm{B}$ &$\rightharpoonup$& $\mathrm{mRNA}_\mathrm{A}$ & $k_{21} = 0$ & artificial fusion\\
-				$\mathrm{mRNA}_\mathrm{B} + \mathrm{C}$ &$\rightharpoonup$& $\mathrm{mRNA}_\mathrm{B}$ & $k_{22} = 0$ & artificial fusion\\
-				$\mathrm{mRNA}_\mathrm{C} + \mathrm{A}$ &$\rightharpoonup$& $\mathrm{mRNA}_\mathrm{C}$ & $k_{23} = 0$ & artificial fusion\\
-				$\mathrm{mRNA}_\mathrm{A} + \mathrm{C}$ &$\rightharpoonup$& $\mathrm{mRNA}_\mathrm{A}$ & $k_{24} = 0$ & artificial fusion\\
-				$\mathrm{mRNA}_\mathrm{B} + \mathrm{A}$ &$\rightharpoonup$& $\mathrm{mRNA}_\mathrm{B}$ & $k_{25} = 0$ & artificial fusion\\
-				$\mathrm{A} + \mathrm{A}$ &$\rightharpoonup$& $\mathrm{A}$ & $k_{26} = 0$ & $\mathrm{A}$ regulates $\mathrm{A}$\\
-				$\mathrm{B} + \mathrm{B}$ &$\rightharpoonup$& $\mathrm{B}$ & $k_{27} = 0$ & $\mathrm{B}$ regulates $\mathrm{B}$\\
-				$\mathrm{C} + \mathrm{C}$ &$\rightharpoonup$& $\mathrm{C}$ & $k_{28} = 0$ & $\mathrm{C}$ regulates $\mathrm{C}$\\
-				$\mathrm{B} + \mathrm{A}$ &$\rightharpoonup$& $\mathrm{A}$ & $k_{29} = 0$ & artificial fusion between proteins\\
-				$\mathrm{C} + \mathrm{B}$ &$\rightharpoonup$& $\mathrm{B}$ & $k_{30} = 0$ & artificial fusion between proteins\\
-				$\mathrm{A} + \mathrm{C}$ &$\rightharpoonup$& $\mathrm{C}$ & $k_{31} = 0$ & artificial fusion between proteins\\
-				$\mathrm{C} + \mathrm{A}$ &$\rightharpoonup$& $\mathrm{A}$ & $k_{32} = 0$ & artificial fusion between proteins\\
-				$\mathrm{B} + \mathrm{C}$ &$\rightharpoonup$& $\mathrm{C}$ & $k_{33} = 0$ & artificial fusion between proteins\\
-				$\mathrm{A} + \mathrm{B}$ &$\rightharpoonup$& $\mathrm{B}$ & $k_{34} = 0$ & artificial fusion between proteins\\
-				$\mathrm{A}$ &$\rightharpoonup$& $\mathrm{B}$ & $k_{35} = 0$ & artificial conversion between proteins\\
-				$\mathrm{B}$ &$\rightharpoonup$& $\mathrm{C}$ & $k_{36} = 0$ & artificial conversion between proteins\\
-				$\mathrm{C}$ &$\rightharpoonup$& $\mathrm{A}$ & $k_{37} = 0$ & artificial conversion between proteins\\
-				$\mathrm{A}$ &$\rightharpoonup$& $\mathrm{C}$ & $k_{38} = 0$ & artificial conversion between proteins\\
-				$\mathrm{C}$ &$\rightharpoonup$& $\mathrm{B}$ & $k_{39} = 0$ & artificial conversion between proteins\\
-				$\mathrm{B}$ &$\rightharpoonup$& $\mathrm{A}$ & $k_{40} = 0$ & artificial conversion between proteins\\
-				$\mathrm{mRNA}_\mathrm{B} + \mathrm{mRNA}_\mathrm{C}$ &$\rightharpoonup$& $\mathrm{mRNA}_\mathrm{A}$ & $k_{41} = 0$ & artificial fusion\\
-				$\mathrm{mRNA}_\mathrm{C} + \mathrm{mRNA}_\mathrm{B}$ &$\rightharpoonup$& $\mathrm{mRNA}_\mathrm{C}$ & $k_{42} = 0$ & artificial fusion\\
-				$\mathrm{mRNA}_\mathrm{C} + \mathrm{A}$ &$\rightharpoonup$& $\mathrm{C}$ & $k_{43} = 0$ & artificial fusion\\
-				$\mathrm{mRNA}_\mathrm{C} + \mathrm{C}$ &$\rightharpoonup$& $\mathrm{C}$ & $k_{43} = 0$ & artificial fusion
-			\end{tabular}
-		}
-		\caption{Full set of ansatz reactions $\Theta$ used in Sec.~\ref{sec:results}.}
-		\label{tab:reaction-library}
+		\begin{tabular}{rclcc}\toprule
+			\multicolumn{3}{c}{Reaction} & rate  & description \\\midrule 
+			$\mathrm{DNA}_{\mathrm{A}}$  & $\rightharpoonup$ & $\mathrm{DNA}_{\mathrm{A}}+\mathrm{mRNA}_{\mathrm{A}}$  & $k_{1}=1.8$  & transcription of $\mathrm{mRNA}_{\mathrm{A}}$\\
+			$\mathrm{mRNA}_{\mathrm{A}}$  & $\rightharpoonup$ & $\mathrm{mRNA}_{\mathrm{A}}+\mathrm{A}$  & $k_{2}=2.1$  & translation of $\mathrm{A}$ proteins\\
+			$\mathrm{mRNA}_{\mathrm{A}}$  & $\rightharpoonup$ & $\emptyset$  & $k_{3}=1.3$  & $\mathrm{mRNA}_{\mathrm{A}}$ decay\\
+			$\mathrm{A}$  & $\rightharpoonup$ & $\emptyset$  & $k_{4}=1.5$  & decay of $\mathrm{A}$ proteins\\
+			$\mathrm{DNA}_{\mathrm{B}}$  & $\rightharpoonup$ & $\mathrm{DNA}_{\mathrm{B}}+\mathrm{mRNA}_{\mathrm{B}}$  & $k_{5}=2.2$  & transcription of $\mathrm{mRNA}_{\mathrm{B}}$\\
+			$\mathrm{mRNA}_{\mathrm{B}}$  & $\rightharpoonup$ & $\mathrm{mRNA}_{\mathrm{B}}+\mathrm{B}$  & $k_{6}=2.0$  & translation of $\mathrm{B}$ proteins\\
+			$\mathrm{mRNA}_{\mathrm{B}}$  & $\rightharpoonup$ & $\emptyset$  & $k_{7}=2.0$  & $\mathrm{mRNA}_{\mathrm{B}}$ decay\\
+			$\mathrm{B}$  & $\rightharpoonup$ & $\emptyset$  & $k_{8}=2.5$  & decay of $\mathrm{B}$ proteins\\
+			$\mathrm{DNA}_{\mathrm{C}}$  & $\rightharpoonup$ & $\mathrm{DNA}_{\mathrm{C}}+\mathrm{mRNA}_{\mathrm{C}}$  & $k_{9}=3.2$  & transcription of $\mathrm{mRNA}_{\mathrm{C}}$\\
+			$\mathrm{mRNA}_{\mathrm{C}}$  & $\rightharpoonup$ & $\mathrm{mRNA}_{\mathrm{C}}+\mathrm{C}$  & $k_{10}=3.0$  & translation of $\mathrm{C}$ proteins\\
+			$\mathrm{mRNA}_{\mathrm{C}}$  & $\rightharpoonup$ & $\emptyset$  & $k_{11}=2.3$  & $\mathrm{mRNA}_{\mathrm{C}}$ decay\\
+			$\mathrm{C}$  & $\rightharpoonup$ & $\emptyset$  & $k_{12}=2.5$  & decay of $\mathrm{C}$ proteins\\
+			$\mathrm{mRNA}_{\mathrm{A}}+\mathrm{A}$  & $\rightharpoonup$ & $\mathrm{A}$  & $k_{13}=0$  & self regulation of $\mathrm{A}$ proteins\\
+			$\mathrm{mRNA}_{\mathrm{B}}+\mathrm{B}$  & $\rightharpoonup$ & $\mathrm{B}$  & $k_{14}=0$  & self regulation of $\mathrm{B}$ proteins\\
+			$\mathrm{mRNA}_{\mathrm{C}}+\mathrm{C}$  & $\rightharpoonup$ & $\mathrm{C}$  & $k_{15}=0$  & self regulation of $\mathrm{C}$ proteins\\
+			$\mathrm{mRNA}_{\mathrm{B}}+\mathrm{A}$  & $\rightharpoonup$ & $\mathrm{A}$  & $k_{16}=0$  & cyclic regulation of $\mathrm{A}$ proteins\\
+			$\mathrm{mRNA}_{\mathrm{C}}+\mathrm{B}$  & $\rightharpoonup$ & $\mathrm{B}$  & $k_{17}=0$  & cyclic regulation of $\mathrm{B}$ proteins\\
+			$\mathrm{mRNA}_{\mathrm{A}}+\mathrm{C}$  & $\rightharpoonup$ & $\mathrm{C}$  & $k_{18}=0$  & cyclic regulation of $\mathrm{C}$ proteins\\
+			$\mathrm{mRNA}_{\mathrm{C}}+\mathrm{A}$  & $\rightharpoonup$ & $\mathrm{A}$  & $k_{16}=6.0$  & cyclic regulation of $\mathrm{A}$ proteins\\
+			$\mathrm{mRNA}_{\mathrm{B}}+\mathrm{C}$  & $\rightharpoonup$ & $\mathrm{C}$  & $k_{17}=4.0$  & cyclic regulation of $\mathrm{C}$ proteins\\
+			$\mathrm{mRNA}_{\mathrm{A}}+\mathrm{B}$  & $\rightharpoonup$ & $\mathrm{B}$  & $k_{18}=3.0$  & cyclic regulation of $\mathrm{B}$ proteins\\
+			$\mathrm{mRNA}_{\mathrm{A}}+\mathrm{A}$  & $\rightharpoonup$ & $\mathrm{mRNA}_{\mathrm{A}}$  & $k_{19}=0$  & artificial fusion\\
+			$\mathrm{mRNA}_{\mathrm{B}}+\mathrm{B}$  & $\rightharpoonup$ & $\mathrm{mRNA}_{\mathrm{B}}$  & $k_{20}=0$  & artificial fusion\\
+			$\mathrm{mRNA}_{\mathrm{A}}+\mathrm{B}$  & $\rightharpoonup$ & $\mathrm{mRNA}_{\mathrm{A}}$  & $k_{21}=0$  & artificial fusion\\
+			$\mathrm{mRNA}_{\mathrm{B}}+\mathrm{C}$  & $\rightharpoonup$ & $\mathrm{mRNA}_{\mathrm{B}}$  & $k_{22}=0$  & artificial fusion\\
+			$\mathrm{mRNA}_{\mathrm{C}}+\mathrm{A}$  & $\rightharpoonup$ & $\mathrm{mRNA}_{\mathrm{C}}$  & $k_{23}=0$  & artificial fusion\\
+			$\mathrm{mRNA}_{\mathrm{A}}+\mathrm{C}$  & $\rightharpoonup$ & $\mathrm{mRNA}_{\mathrm{A}}$  & $k_{24}=0$  & artificial fusion\\
+			$\mathrm{mRNA}_{\mathrm{B}}+\mathrm{A}$  & $\rightharpoonup$ & $\mathrm{mRNA}_{\mathrm{B}}$  & $k_{25}=0$  & artificial fusion\\
+			$\mathrm{A}+\mathrm{A}$  & $\rightharpoonup$ & $\mathrm{A}$  & $k_{26}=0$  & $\mathrm{A}$ regulates $\mathrm{A}$\\
+			$\mathrm{B}+\mathrm{B}$  & $\rightharpoonup$ & $\mathrm{B}$  & $k_{27}=0$  & $\mathrm{B}$ regulates $\mathrm{B}$\\
+			$\mathrm{C}+\mathrm{C}$  & $\rightharpoonup$ & $\mathrm{C}$  & $k_{28}=0$  & $\mathrm{C}$ regulates $\mathrm{C}$\\
+			$\mathrm{B}+\mathrm{A}$  & $\rightharpoonup$ & $\mathrm{A}$  & $k_{29}=0$  & artificial fusion between proteins\\
+			$\mathrm{C}+\mathrm{B}$  & $\rightharpoonup$ & $\mathrm{B}$  & $k_{30}=0$  & artificial fusion between proteins\\
+			$\mathrm{A}+\mathrm{C}$  & $\rightharpoonup$ & $\mathrm{C}$  & $k_{31}=0$  & artificial fusion between proteins\\
+			$\mathrm{C}+\mathrm{A}$  & $\rightharpoonup$ & $\mathrm{A}$  & $k_{32}=0$  & artificial fusion between proteins\\
+			$\mathrm{B}+\mathrm{C}$  & $\rightharpoonup$ & $\mathrm{C}$  & $k_{33}=0$  & artificial fusion between proteins\\
+			$\mathrm{A}+\mathrm{B}$  & $\rightharpoonup$ & $\mathrm{B}$  & $k_{34}=0$  & artificial fusion between proteins\\
+			$\mathrm{A}$  & $\rightharpoonup$ & $\mathrm{B}$  & $k_{35}=0$  & artificial conversion between proteins\\
+			$\mathrm{B}$  & $\rightharpoonup$ & $\mathrm{C}$  & $k_{36}=0$  & artificial conversion between proteins\\
+			$\mathrm{C}$  & $\rightharpoonup$ & $\mathrm{A}$  & $k_{37}=0$  & artificial conversion between proteins\\
+			$\mathrm{A}$  & $\rightharpoonup$ & $\mathrm{C}$  & $k_{38}=0$  & artificial conversion between proteins\\
+			$\mathrm{C}$  & $\rightharpoonup$ & $\mathrm{B}$  & $k_{39}=0$  & artificial conversion between proteins\\
+			$\mathrm{B}$  & $\rightharpoonup$ & $\mathrm{A}$  & $k_{40}=0$  & artificial conversion between proteins\\
+			$\mathrm{mRNA}_{\mathrm{B}}+\mathrm{mRNA}_{\mathrm{C}}$  & $\rightharpoonup$ & $\mathrm{mRNA}_{\mathrm{A}}$  & $k_{41}=0$  & artificial fusion\\
+			$\mathrm{mRNA}_{\mathrm{C}}+\mathrm{mRNA}_{\mathrm{B}}$  & $\rightharpoonup$ & $\mathrm{mRNA}_{\mathrm{C}}$  & $k_{42}=0$  & artificial fusion\\
+			$\mathrm{mRNA}_{\mathrm{C}}+\mathrm{A}$  & $\rightharpoonup$ & $\mathrm{C}$  & $k_{43}=0$  & artificial fusion\\\bottomrule
+		\end{tabular}\caption{\label{tab:reaction-library}Full set of ansatz reactions $\Theta$ used in Sec.~\ref{sec:results}.}
 	\end{table}

### 2018-09-03 docs/makefigures/case_config.py
transform number of measurements to noise level

@@ -239,7 +239,7 @@ def print_config(print_reactions=True):
         print(tabulate(data, headers=["Type", "Involved species", "Rate"]))
 
 
-def get_regulation_network_lma():
+def get_regulation_network_lma(init=1):
     regulation_network = RegulationNetwork()
     regulation_network.timestep = TIMESTEP
     regulation_network.realisations = 1.
@@ -249,7 +249,7 @@ def get_regulation_network_lma():
     regulation_network.target_time = TARGET_TIME
     print(len(regulation_network.desired_rates))
     print(regulation_network.get_bfc().n_basis_functions)
-    regulation_network.initial_states = [regulation_network.initial_states[1]]
+    regulation_network.initial_states = [regulation_network.initial_states[init]]
     analysis = regulation_network.generate_analysis_object(fname_prefix='case_1', fname_postfix='0')
     for i in range(len(regulation_network.initial_states)):
         analysis.generate_or_load_traj_lma(i, regulation_network.target_time,

### 2018-10-11 docs/manuscript_jcp.tex
jcp preparations

%% LyX 2.3.1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english,osajnl,jcp,preprint,groupedaddress,noshowpacs,showkeys,longbibliography,noeprint,nofootinbib,bibtex,ruled,vlined,english,natural,table,usenames,dvipsnames,a4paper]{revtex4-1}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\setcounter{secnumdepth}{3}
\usepackage{color}
\usepackage{babel}
\usepackage{array}
\usepackage{booktabs}
\usepackage{mathtools}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{microtype}
\usepackage[unicode=true,
bookmarks=false,
breaklinks=false,pdfborder={0 0 1},backref=section,colorlinks=true]
{hyperref}
%\hypersetup{%
%	linkcolor=BrickRed,urlcolor=blue!50!black,citecolor=blue!50!black}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{siunitx}

\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}

%\author[1]{Moritz Hoffmann\thanks{Equal contributions}}
%\author[1]{Christoph Fröhner\samethanks}
%\author[1]{Frank Noé\thanks{Corresponding author}}
%\affil[1]{Freie Universität Berlin, Fachbereich Mathematik und Informatik,\newline Arnimallee 6, 14195 Berlin, Germany}
%\affil[ ]{\textit{\{ moritz.hoffmann,christoph.froehner,frank.noe \}@fu-berlin.de}}

\newlength{\singlecolumnwidth}
\newlength{\doublecolumnwidth}
\singlecolumnwidth=3.37in
\doublecolumnwidth=6.69in

\makeatother

\linespread{0.956}

\begin{document}

	\author{Moritz Hoffmann}
	\email{moritz.hoffmann@fu-berlin.de}
	\thanks{Equal contributions}
	\affiliation{
		Freie Universität Berlin, Fachbereich Mathematik und Informatik,
		Arnimallee 6, 14195 Berlin, Germany
	}

	\author{Christoph Fröhner}
	\email{christoph.froehner@fu-berlin.de}
	\thanks{Equal contributions}
	\affiliation{
		Freie Universität Berlin, Fachbereich Mathematik und Informatik,
		Arnimallee 6, 14195 Berlin, Germany
	}

	\author{Frank Noé}
	\email{frank.noe@fu-berlin.de}
	\thanks{Corresponding author}
	\affiliation{
		Freie Universität Berlin, Fachbereich Mathematik und Informatik,
		Arnimallee 6, 14195 Berlin, Germany
	}

\title{Reactive SINDy: Discovering governing reactions from concentration
data}

\begin{abstract}
The inner workings of a biological cell or a chemical reaction can
be rationalized by the network of reactions, whose structure reveals
the most important functional mechanisms. For complex systems, these
reaction networks are not known a priori and cannot be efficiently
computed with \emph{ab initio} methods, therefore an important approach
goal is to estimate effective reaction networks from observations,
such as time series of the main species. Reaction networks estimated
with standard machine learning techniques such as least-squares regression
may fit the observations, but will typically contain spurious reactions.
Here we extend the sparse identification of nonlinear dynamics (SINDy)
method to vector-valued ansatz functions, each describing a particular
reaction process. The resulting sparse tensor regression method ``reactive
SINDy'' is able to estimate a parsimonious reaction network. We illustrate
that a gene regulation network can be correctly estimated from observed
time series.
\end{abstract}

\maketitle

\section{Introduction}

Mapping out the reaction networks behind biological processes, such
as gene regulation in cancer~\cite{Abramovitch2004}, is paramount
to understanding the mechanisms of life and disease. A well-known
example of gene regulation is the lactose operon whose crystal structure
was resolved in~\cite{Lewis1996} and dynamics were modeled in~\cite{Yildirim2003}.
The system's ``combinatorial control'' in E. coli cells was quantitatively
investigated in~\cite{Kuhlman2007}, in particular studying repression
and activation effects. These gene regulatory effects often appear
in complex networks~\cite{Shen-Orr2002} and there exist databases
resolving these for certain types of cells, e.g., E. coli cells~\cite{Gama-Castro2016}
and yeast cells~\cite{Lee2002}. Another example where mapping the
active reactions is important is that of chemical reactors~\cite{Roa2017},
where understanding which reactions are accessible for a given set
of educts and reaction conditions is important to design synthesis
pathways~\cite{cong1999high,kiwi2005microstructured}.

The traditional approach to determine a reaction network is to propose
the structure of the network based on chemical insight and subsequently
fit the parameters given available data~\cite{schoneberg2014explicit}.
To decipher complex reaction environments such as biological cells,
it would be desirable to have a data-driven approach that can answer
the question which reactions are underlying a given observation, e.g.,
the time series of a set of reactants. However, in sufficiently complex
reaction environments the number of reactive species and possible
reactions is practically unlimited -- as an illustration, consider
vast amount of possible isomerizations and post-translational modifications
for a single protein molecule. Therefore, the more specific formulation
is ``given observations of a set of chemical species, what is the
\emph{minimal set} of reactions necessary to explain their time evolution?''.
This formulation calls for a machine learning method that can infer
the reaction network underlying the observation data.

Knowledge about the reaction network can be applied to parameterize
other numerical methods to further investigate the processes at hand.
Such methods include particle-based approaches derived from the chemical
master equation~\cite{gillespie1977exact,winkelmann2016spatiotemporal,winkelmann2017hybrid,isaacson2009reaction},
as well as highly detailed but parameter-rich methods such as particle-based
or interacting-particle reaction~dynamics~\cite{schoneberg2013readdy,hoffmann2018readdy,frohner2018reversible,donev2018efficient,Andrews2017,van2005green,van2005simulating}
capable of fully resolving molecule positions in space and time --
see~\cite{schoneberg2014simulation,Andrews2018} for recent reviews.

Existing methods to infer regulatory networks include ARCANE~\cite{Margolin2006}
that uses experimental essay data and information theory, as well
as the likelihood approach presented in~\cite{Tian2007a} that takes
the stochasticity of observed reactant time series into account.

The method presented in this work can identify underlying complex
reaction networks from concentration time series by following the
law of parsimony, i.e., by inducing sparsity in the resulting reaction
network. This promotes the interpretability of the model and avoids
overfitting. We formulate the problem as data-driven identification
of a dynamical system, which renders the method consistent with and
an extension of the framework of sparse identification of nonlinear
dynamics (SINDy)~\cite{Brunton2015}. Specifically, the problem of
identifying a reaction network from time traces of reactant concentrations
can be solved by finding a linear combination from a library of nonlinear
ansatz functions that each corresponds to a reaction acting on a set
of reactants. With this formulation, the reaction rates can be determined
\emph{via} regression. Sparsity is induced by equipping the regression
algorithms with a sparsity inducing regularization. SINDy was investigated,
generalized, and applied in many different ways, e.g., including control~\cite{brunton2016sparse}
(SINDYc), in the context of partial differential equations~\cite{rudy2017data},
updating already existing models~\cite{quade2018sparse} (abrupt-SINDy),
and looking into convergence properties~\cite{zhang2018convergence}.

We extend and apply SINDy to the case of learning reaction networks
from non-equilibrium concentration data. Similar approaches make use
of SINDy but do not resolve specific reactions~\cite{mangan2016inferring},
use weak formulations to avoid numerical temporal derivatives~\cite{pantazis2017unified},
or use compressive sensing and sparse Bayesian learning~\cite{Pan2012}.

Our extension of the original SINDy method mostly involves estimating
parameters which are coupled across the equations of the arising dynamical
system. In the context of learning reaction networks this means that
we look for specific reactions and their rate constants that might
have lead to the observations instead of net flux across species.
We demonstrate the algorithm on a gene regulatory network in three
different scenarios of measurement: When there is no noise in the
data we can find, given sufficient amounts of data, all relevant processes
of the ground truth. If there is noise in the data we converge to
the correct reaction network and rates with decreasing levels of noise.
The final scenario generalizes the method to two measurements with
different initial conditions, also converging to the correct model
with decreasing levels of noise.

\section{Reactive SINDy: Sparse learning of reaction kinetics}

We are observing the concentrations of $S$ chemical species in time
$t$: 
\begin{align}
\mathbf{x}(t)=\begin{pmatrix}x_{1}(t)\\
\vdots\\
x_{S}(t)
\end{pmatrix}\in\mathbb{R}^{S}.\label{eq:concentration_vector}
\end{align}
We assume that their dynamics are governed by classical reaction-rate
equations subject to the law of mass action. A general expression
for the change of concentration of reactant $s$ as a result of order-0
reactions (creation or annihilation), order-1 reactions (transitions
of other species into $s$ or transitions of $s$ into other species),
order-2 reactions (production or consumption of $s$ by the encounter
of two species), etc, is given by:

\begin{equation}
\dot{x}_{s}=\sum_{i}\beta_{s,0}^{(i)}+\sum_{i}\beta_{s,1}^{(i)}x_{i}+\sum_{i,j}\beta_{s,2}^{(i,j)}x_{i}x_{j}+\ldots\label{eq:lma}
\end{equation}
where the $\beta_{s,k}^{(\ldots)}$-values are constants belonging
to the reactions of order $k$. These rate constants however can incorporate
several underlying reactions at once. For example, the two reactions
\begin{align}
s_{1}\xrightharpoonup{\xi_{1}}s_{2}\label{eq:example_reaction_1}\\
s_{1}\xrightharpoonup{\xi_{2}}s_{3}\label{eq:example_reaction_2}
\end{align}
both contribute to $\dot{x}_{1}=\beta_{1,1}^{(1)}x_{1}=-(\xi_{1}+\xi_{2})x_{1}$.
To disentangle (\ref{eq:lma}) into single reactions, we choose a
library of $R$ possible ansatz reactions that each represent a single
reaction: 
\begin{align}
\textbf{y}_{r}(\textbf{x}(t))=\begin{pmatrix}y_{r,1}(\textbf{x}(t))\\
\vdots\\
y_{r,S}(\textbf{x}(t))
\end{pmatrix},\quad r=1,\ldots,R.
\end{align}
With this ansatz, the reaction dynamics (\ref{eq:lma}) becomes a
set of linear equations with unknown parameters $\xi_{r}$ that represent
the sought macroscopic rate constants: 
\begin{align}
\dot{\textbf{x}}_{i}(t)=\sum_{r=1}^{R}y_{r,i}(\textbf{x}(t))\xi_{r},\quad i=1,\ldots,S,\label{method:the-system}
\end{align}
where $\xi_{r}$ are the to-be estimated macroscopic rate constants.
The two reactions in the previous example (\ref{eq:example_reaction_1}-\ref{eq:example_reaction_2})
would be modeled by the ansatz functions
\begin{align*}
y_{1}(\mathbf{x}) & =(-x_{1},x_{1},0)^{\top}\\
y_{2}(\mathbf{x}) & =(-x_{1},0,x_{1})^{\top},
\end{align*}
illustrating that the values of the coefficients $\xi_{1}$ and $\xi_{2}$
can be used to decide whether a single reaction is present and to
what degree.

Now suppose we have measured the concentration vector (\ref{eq:concentration_vector})
at $T$ time points $t_{1}<\cdots<t_{T}$. We represent these data
as a matrix 
\begin{align}
\textbf{X}=\left(\begin{array}{cccc}
\mathbf{x}(t_{1}) & \mathbf{x}(t_{2}) & \cdots & \mathbf{x}(t_{T})\end{array}\right)^{\top}\in\mathbb{R}^{T\times S}.
\end{align}
Given this matrix, a library $\Theta:\mathbb{R}^{T\times S}\to\mathbb{R}^{T\times S\times R},\;\mathbf{X}\mapsto\begin{pmatrix}\theta_{1}(\mathbf{X}) & \theta_{2}(\mathbf{X}) & \cdots & \theta_{R}(\mathbf{X})\end{pmatrix}$
of $R$ ansatz reactions can be proposed with corresponding reaction
functions 
\begin{align}
\theta_{r}(\mathbf{X})=\begin{pmatrix}\textbf{y}_{r}(\mathbf{X}_{1*})^{\top}\\
\vdots\\
\textbf{y}_{r}(\mathbf{X}_{T*})^{\top}
\end{pmatrix}\in\mathbb{R}^{T\times S},\quad r=1,\ldots,R,\label{method:the-reactions}
\end{align}
where $\textbf{X}_{i*}$ denotes the $i$-th row in $\textbf{X}$.
Applying the concentration trajectory to the library yields $\Theta(\textbf{X})\in\mathbb{R}^{T\times S\times R}$.

The goal is to find coefficients $\Xi=\begin{pmatrix}\xi_{1} & \xi_{2} & \cdots & \xi_{R}\end{pmatrix}^{\top}$,
so that 
\begin{align}
\dot{\textbf{X}}=\Theta(\textbf{X})\Xi=\sum_{r=1}^{R}\theta_{r}(\textbf{X})\xi_{r}.
\end{align}
In particular, the system is linear in the coefficients $\Xi$, which
makes regression tools such as elastic net regularization~\cite{Zou2005}
applicable. To this end, one can consider the regularized minimization
problem (reactive SINDy): 
\begin{align}
\hat{\Xi}=\underset{\Xi}{\arg\min}\left(\frac{1}{2T}\left\Vert \dot{\textbf{X}}-\Theta(\textbf{X})\Xi\right\Vert _{F}^{2}+\alpha\lambda\|\Xi\|_{1}+\alpha(1-\lambda)\|\Xi\|_{2}^{2}\right)\quad\text{subject to }\Xi\geq0.\label{method:minimizationproblem}
\end{align}
Here, $\|\cdot\|_{F}$ denotes the Frobenius norm, $\lambda\in[0,1]$
is a hyperparameter that interpolates linearly between LASSO~\cite{Tibshirani1996,Hastie2009}
and Ridge~\cite{Hoerl1} methods, and $\alpha\geq0$ is a hyperparameter
that, depending on $\lambda$, can induce sparsity and give preference
to smaller solutions in the $L_{1}$ or $L_{2}$ sense. For $\alpha=0$
the minimization problem reduces to standard least-squares (LSQ) with
the constraint $\Xi\geq0$. Reactive SINDy (\ref{method:minimizationproblem})
is therefore a generalization of the SINDy method to to the vector-valued
ansatz functions.

Since only the concentration data $\mathbf{X}$ is available but not
its temporal derivative, $\dot{\mathbf{X}}$ is approximated numerically
by second order finite differences with the exception of boundary
data. Once the pair $(\mathbf{X},\dot{\mathbf{X}})$ is obtained,
the problem becomes invariant under temporal reordering. Hence, when
presented with multiple trajectories the data matrices $\mathbf{X}_{i}$
and $\dot{\mathbf{X}}_{i}$ can simply be concatenated.

In order to solve (\ref{method:minimizationproblem}) the numerical
sequential least-squares minimizer SLSQP~\cite{Kraft1988} is applied
via the software package SciPy~\cite{SciPy}. Code related to this
paper can be found under \url{https://github.com/readdy/readdy_learn}.

\section{Results\label{sec:results}}

We demonstrate the method by estimating the reactions of a gene-regulatory
network from time series of concentrations of the involved molecules.
Let $S:=\{A,B,C\}$ be a set of three species of proteins which are
being translated each from their respective $\mathrm{mRNA}$ molecule.
Each $\mathrm{mRNA}$ in turn has a corresponding $\mathrm{DNA}$
which it is transcribed from. The proteins and $\mathrm{mRNA}$ molecules
decay over time whereas the $\mathrm{DNA}$ concentration remains
constant. The network contains reactions of the following form~\cite{Thattai2001a}
\begin{align*}
\mathrm{DNA}_{i} & \rightharpoonup\mathrm{DNA}_{i}+\mathrm{mRNA}_{i} &  & \text{(transcription)},\\
\mathrm{mRNA}_{i} & \rightharpoonup\mathrm{mRNA}_{i}+i &  & \text{(translation)},\\
\mathrm{mRNA}_{i} & \rightharpoonup\emptyset &  & \text{(decay of mRNA)},\\
i & \rightharpoonup\emptyset &  & \text{(decay of protein)},\\
i+\mathrm{mRNA}_{j} & \rightharpoonup i &  & \text{(regulation of protein }j\in S\text{),}
\end{align*}
for each of the species $i\in S$. These reactions model a regulation
of species $j$ by virtue of the fact that the transcription product
inhibits the transcription processes. In our example proteins of type
$\mathrm{A}$ regulate the $\mathrm{mRNA}_{\mathrm{B}}$ molecules,
proteins of type $\mathrm{B}$ regulate the $\mathrm{mRNA}_{\mathrm{C}}$
molecules and proteins of type $\mathrm{C}$ regulate the $\mathrm{mRNA}_{\mathrm{A}}$
molecules (Fig.~\ref{fig:network}). Using this reaction model, time
series of concentrations are generated using the rates given in Tab~\ref{tab:reaction-library}
under the initial condition described in Tab~\ref{tab:Initial-conditions}a,
which were chosen so that all the reactions in the reaction model
significantly contribute to the temporal evolution of the system's
concentrations. The generation samples the integrated equations equidistantly
with a discrete time step of $\tau=3\cdot10^{-3}$ yielding $667$
frames which amounts to a cumulative time of roughly $T=2$.

\begin{figure}
\centering \includegraphics[width=\singlecolumnwidth]{figures_tex/scheme}
\caption{\label{fig:network}The regulation network example described in Sec.~\ref{sec:results}.
Each circle depicts a species, each arrow corresponds to one reaction.
Blue arrows denote transcription from $\mathrm{DNA}$ to $\mathrm{mRNA}$,
green arrows denote translation from $\mathrm{mRNA}$ to protein,
and red arrows denote the regulatory network.}
\end{figure}
\begin{table}
\centering%
\begin{tabular}{cccccccccc}
 & $\mathrm{DNA}_{\mathrm{A}}$ & $\mathrm{mRNA}_{\mathrm{A}}$ & $\mathrm{A}$ & $\mathrm{DNA}_{\mathrm{B}}$ & $\mathrm{mRNA}_{\mathrm{B}}$ & $\mathrm{B}$ & $\mathrm{DNA}_{\mathrm{C}}$ & $\mathrm{mRNA}_{\mathrm{C}}$ & $\mathrm{C}$\tabularnewline
\midrule
\textbf{(a)} & $1$ & $2$ & $0$ & $1$ & $0$ & $3$ & $1$ & $0$ & $0$\tabularnewline
\textbf{(b)} & $1$ & $1.5$ & $0$ & $1$ & $0$ & $2$ & $1$ & $0$ & $1$\tabularnewline
\end{tabular}\caption{\label{tab:Initial-conditions}Initial conditions used to generate
concentration time series. Reaction rates can be found in Tab.~\ref{tab:reaction-library}.}
\end{table}
The proposed estimation method is applied to analyze these time series
of concentrations in order to recover the underlying reaction network
from data. To this end we use the library of ansatz functions given
in Tab.~\ref{tab:reaction-library}, which contains a large number
of possible reactions, only few of which are actually part of the
model.

\subsection{Learning the reaction network in the low-noise limit\label{sec:case-1}}

We first demonstrate that the true reaction network can be reconstructed
when using a finite amount of observation data without additional
measurement noise, i.e., the observations are reflecting the true
molecule concentrations at any given time point. The minimization
problem (\ref{method:minimizationproblem}) is solved using the concentration
time series shown in Fig.~\ref{fig:network}b.

We first set the hyperparameter $\alpha=0$ in the minimization problem
(\ref{method:minimizationproblem}), which results in constrained
least-squares regression without any of the regularization terms.
In this case we estimate a reaction network that can reproduce the
observations almost exactly (Fig.~\ref{fig:case-1-lsq}). However,
the result is mechanistically wrong as the sparsity pattern does not
match the reaction network used to generate the data. On the one hand
many spurious reactions are estimated that were not in the true reaction
scheme and would lead to wrong conclusions about the mechanism, such
as $\mathrm{A}+\mathrm{A}\rightharpoonup\mathrm{A}$ and $\mathrm{A}+\mathrm{C}\rightharpoonup\mathrm{C}$.
More dramatically, the reaction responsible for the decay of $\mathrm{A}$
particles is completely ignored (Fig.~\ref{fig:case-1-sparsity-pattern}).

\begin{figure}
\centering\includegraphics[width=\singlecolumnwidth]{figures_tex/trajectories_lsq}\caption{\label{fig:case-1-lsq}Concentration time series generated from integrating
the reaction network shown in Fig.~\ref{fig:network}a. The initial
condition prescribes positive concentration values only for $\mathrm{B}$
protein and $\mathrm{mRNA}_{\mathrm{A}}$ species (Tab.~\ref{tab:Initial-conditions}a).
This initial condition is used in the subsequent sections for further
analysis. Gray dots depict concentration time series yielded from
the LSQ rates estimated in Sec~\ref{sec:case-1}.}
\end{figure}
\begin{figure}
\centering \includegraphics[width=0.82\doublecolumnwidth]{figures_tex/sparsity_pattern}
\caption{\label{fig:case-1-sparsity-pattern}Estimated reaction rates in the
system described in Sec.~\ref{sec:case-1}. The y and x axes contain
reaction educts and products, respectively. A circle at position $(i,j)$
represents a reaction $i\rightharpoonup j$ whose rate has a linear
relation with the area of the circle. The black outlines denote the
reactions with which the system was generated and contain the respective
rate value. Red crosses denote reactions that were used as additional
ansatz reactions. Blue circles are estimated by LSQ and orange circles
depict rates which were obtained by solving the minimization problem
(\ref{method:minimizationproblem}). The latter rates are subject
to a cutoff $\kappa=0.22$ corresponding to the green circle's area
under which a sparse solution with the correct processes can be recovered.
If a certain rate was estimated in both cases, two wedges instead
of one circle are displayed.}
\end{figure}
Next, we sought sparse solutions by using $\alpha>0$ and additionally
eliminating reactions with rate constants smaller than a cutoff value
$\kappa$. For a suitable choice of hyperparameters $\alpha\approx1.91\cdot10^{-7}$,
$\lambda=1$, and $\kappa=0.22$, a sparse solution is obtained that
finds the correct reaction scheme and also recovers the decay reaction
(Fig.~\ref{fig:case-1-sparsity-pattern}).

The value of the cutoff was determined by comparing the magnitude
of estimated rates and finding a gap, see Fig.~\ref{fig:reaction-rates-sorted}.
The hyperparameter pair $(\alpha,\lambda)$ was obtained by a grid
search and evaluating the difference $\|\hat{\Xi}_{\alpha,\lambda}-\Xi\|_{1}$,
where $\hat{\Xi}_{\alpha,\lambda}$ is the estimated model under a
particular hyperparameter choice and $\Xi$ is the ground truth. If
the ground truth is unknown, a hyperparameter pair can be estimated
by utilizing cross-validation as in the following sections.

\subsection{Learning the reaction network from data with stochastic noise\label{sec:case-2} }

In contrast to Sec.~\ref{sec:case-1}, we now employ data that includes
measurement noise. Such noise can originate from uncertainties in
the experimental setup or from shot noise in single- or few-molecule
measurements. In gene regulatory networks such noise is commonly observed
when only few copy numbers of $\mathrm{mRNA}$ are present~\cite{Golding2005,Berg1978,Elowitz2002}.
In order to simulate noise from few copies of molecules, the system
of Sec.~\ref{sec:results} with initial conditions as given in Tab.~\ref{tab:Initial-conditions}a
is integrated using the Gillespie stochastic simulation algorithm
(SSA)~\cite{gillespie1976general,gillespie1977exact}. In the limit
of many particles and realizations, the Gillespie SSA converges to
the integrated reaction-rate equations subject to the law of mass
action. As our model is based on exactly these dynamics, the initial
condition's concentrations are interpreted in terms of hundreds of
particles. Each realization is then transformed back to a time series
of concentrations. We define the noise level as the mean-squared deviation
of the concentration time series from the integrated reaction-rate
equations. Data with different noise levels are prepared by averaging
multiple realizations of the time series obtained by the Gillespie
SSA.

It can be observed that decreasing levels of noise lead to fewer spurious
reactions when applying reactive SINDy (\ref{method:minimizationproblem}),
see Fig.~\ref{fig:case2-convergence}a. Also the estimation error
$\|\xi-\hat{\xi}\|_{1}$ with respect to the ground truth $\xi$ decreases
with decreasing levels of noise (Fig.~\ref{fig:case2-convergence}b).
In both cases, the regularized method with a suitable hyperparameter
pair $(\alpha,\lambda)$ performs better than LSQ.

\begin{figure}
\centering \includegraphics[width=\singlecolumnwidth]{figures_tex/case2}
\caption{\label{fig:case2-convergence}Convergence of the estimation error
when estimating the system described in Sec.~\ref{sec:case-1} with
varying levels of noise by application of reactive SINDy (\ref{method:minimizationproblem})
with and without regularization in blue and orange, respectively.
The procedure was independently repeated $10$ times with different
realizations giving rise to the mean and standard deviation depicted
by solid lines and shaded areas, respectively. \textbf{(a)}: The number
of detected spurious reactions up to the cutoff value introduced in
Sec.~\ref{sec:case-1} over different levels of noise. \textbf{(b)}:
The estimation error given by the mean absolute error between the
generating reaction rates $\xi$ and the estimated reaction rates
$\hat{\xi}$ over different levels of noise.}
\end{figure}
The hyperparameters $(\alpha,\lambda)$ are obtained by shuffling
the data and performing a $10$-fold cross validation.

\subsection{Learning the reaction network from multiple initial conditions\label{sec:case-3}}

Preparing the experiment that generates the data in different initial
conditions can help identifying the true reaction mechanisms as a
more diverse dataset makes it easier to confirm or exclude the participation
of specific reactions. This section extends the analysis of Sec.~\ref{sec:case-2}
to two initial conditions, where the first initial condition is identical
to the one used previously and the second initial condition is given
in Tab.~\ref{tab:Initial-conditions}b.

The corresponding time series are depicted in Fig.~\ref{fig:case3-convergence}a.
The gray graph corresponds to a sample trajectory generated by the
Gillespie SSA. For both initial conditions the same time step of $\tau=3\cdot10^{-3}$
has been applied, amounting to $2\cdot667=1334$ frames. Once the
data matrices 
\[
\mathbf{X}_{1}=\begin{pmatrix}\boldsymbol{x}_{1}(t_{1}) & \cdots & \boldsymbol{x}_{1}(t_{667})\end{pmatrix},\;\mathbf{X}_{2}=\begin{pmatrix}\boldsymbol{x}_{2}(t_{1}) & \cdots & \boldsymbol{x}_{2}(t_{667})\end{pmatrix}
\]
 and the corresponding derivatives $\dot{\mathbf{X}}_{1}$, $\dot{\mathbf{X}}_{2}$
have been obtained, the frames are concatenated so that 
\[
\mathbf{X}=\begin{pmatrix}\boldsymbol{x}_{1}(t_{1}) & \boldsymbol{x}_{2}(t_{1}) & \cdots & \boldsymbol{x}_{1}(t_{667}) & \boldsymbol{x}_{2}(t_{667})\end{pmatrix},
\]
analogously for $\dot{\mathbf{X}}$.

Similarly to Sec.~\ref{sec:case-2}, decreasing levels of noise lead
to fewer spurious reactions (Fig.~\ref{fig:case3-convergence}b)
and a smaller $L_{1}$ distance to the ground truth (Fig.~\ref{fig:case3-convergence}c).
Again applying the optimization problem with a suitable set of parameters
$(\alpha,\lambda,\kappa)$ performs better than LSQ. Compared to the
previous section the convergence is better due to twice as much available
data. At noise levels of smaller than roughly $10^{-6}$ the model
can reliably be recovered when using the regularized method.

\begin{figure}
\centering \includegraphics[width=\singlecolumnwidth]{figures_tex/case3}
\caption{\label{fig:case3-convergence}Convergence of estimation error of reaction
schemes from noisy gene-regulation data starting from two different
initial conditions under decreasing levels of noise. The minimization
problem (\ref{method:minimizationproblem}) was solved for $\alpha=0$
(LSQ) and with regularization. This was repeated $10$ times on different
sets of observation data generated by Gillespie SSA, giving rise to
mean and standard deviation (solid lines and shaded areas, respectively).
\textbf{(a)}: Concentration time series corresponding to the initial
conditions, generated by integrating the reaction-rate equations.
The first initial condition is identical to the one used in Sec.~\ref{sec:case-1}
and Sec.~\ref{sec:case-2}. The second initial condition (Tab.~\ref{tab:Initial-conditions}b)
prescribes positive initial concentrations for $\mathrm{mRNA}_{\mathrm{A}}$,
$\mathrm{B}$, and $\mathrm{C}$ species. The gray graphs are sample
realizations of integration using the Gillespie SSA. \textbf{(b)},\textbf{(c)}:
Analogously to Fig.~\ref{fig:case2-convergence} with the difference
that 20-fold cross validation was used for hyperparameter estimation.}
\end{figure}
The hyperparameters $(\alpha,\lambda)$ are obtained by shuffling
the data and performing a $20$-fold cross validation.

\section{Conclusion}

In this work we have extended the SINDy method to reactive SINDy,
not only parsimoniously detecting potentially nonlinear terms in a
dynamical system from noisy data, but also yielding, in this case,
a sparse set of rates with respect to generating reactions (\ref{method:the-reactions}).
Mathematically this has been achieved by permitting vector-valued
basis functions and obtaining a tensor linear regression problem.
We have applied this method on data generated from a gene regulation
network and could successfully recover the underlying model.

The studies of Sec.~\ref{sec:case-2} and Sec.~\ref{sec:case-3}
have shown that the applied regularization terms can mitigate noise
up to a certain degree compared to the unregularized method, so that
identification of the reaction network is more robust and closer to
the ground truth.

Potentially, this method could be used to identify reaction networks
from time series measurements even if the initial conditions are not
always exactly identical, as was demonstrated in Sec.~\ref{sec:case-3}.
One apparent limitation is that the method can only be applied if
the data stems from the equilibration phase, as the concentration-based
approach has derivatives equal zero in the equilibrium, which precludes
the reaction dynamics to be recovered. In future work, we will consider
the identification of reaction schemes from instantaneous fluctuations
of particle numbers in equilibrium.

\section*{Acknowledgments}

The authors are grateful to the Center for Theoretical Biological
Physics (CTBP, supported by NSF PHY-1427654) at Rice University for
hosting their sabbatical visit, during which part of this work was
performed.

We gratefully acknowledge funding from Deutsche Forschungsgemeinschaft
(SFB 958 / Project A04, TRR 186 / Project A12, SFB 1114 / Project
C03), Einstein Foundation Berlin (ECMath Project CH17) and European
Research Council (ERC CoG 772230 ``ScaleCell''). We are grateful
for inspiring discussions with Simon Olsson, Mohsen
Sadeghi, Felix Höfling and Christof Schütte.

\section{Appendix\label{sec:appendix}}
\begin{figure}[h!]

\centering\includegraphics[width=\singlecolumnwidth]{figures_tex/case1_cutoff}\caption{\label{fig:reaction-rates-sorted}Reaction rates sorted by their magnitude
to determine the cutoff $\kappa=0.22$ of Sec.~\ref{sec:case-1}.
The rates were estimated using the regularized minimization problem.}

\end{figure}
\begin{table}[h!]
\centering %
\resizebox{.8\textwidth}{!}{%
\begin{tabular}{rclcc}
\toprule 
\multicolumn{3}{c}{Reaction} & rate  & description \tabularnewline
\midrule 
$\mathrm{DNA}_{\mathrm{A}}$  & $\rightharpoonup$  & $\mathrm{DNA}_{\mathrm{A}}+\mathrm{mRNA}_{\mathrm{A}}$  & $k_{1}=1.8$  & transcription of $\mathrm{mRNA}_{\mathrm{A}}$\tabularnewline
$\mathrm{mRNA}_{\mathrm{A}}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{A}}+\mathrm{A}$  & $k_{2}=2.1$  & translation of $\mathrm{A}$ proteins\tabularnewline
$\mathrm{mRNA}_{\mathrm{A}}$  & $\rightharpoonup$  & $\emptyset$  & $k_{3}=1.3$  & $\mathrm{mRNA}_{\mathrm{A}}$ decay\tabularnewline
$\mathrm{A}$  & $\rightharpoonup$  & $\emptyset$  & $k_{4}=1.5$  & decay of $\mathrm{A}$ proteins\tabularnewline
$\mathrm{DNA}_{\mathrm{B}}$  & $\rightharpoonup$  & $\mathrm{DNA}_{\mathrm{B}}+\mathrm{mRNA}_{\mathrm{B}}$  & $k_{5}=2.2$  & transcription of $\mathrm{mRNA}_{\mathrm{B}}$\tabularnewline
$\mathrm{mRNA}_{\mathrm{B}}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{B}}+\mathrm{B}$  & $k_{6}=2.0$  & translation of $\mathrm{B}$ proteins\tabularnewline
$\mathrm{mRNA}_{\mathrm{B}}$  & $\rightharpoonup$  & $\emptyset$  & $k_{7}=2.0$  & $\mathrm{mRNA}_{\mathrm{B}}$ decay\tabularnewline
$\mathrm{B}$  & $\rightharpoonup$  & $\emptyset$  & $k_{8}=2.5$  & decay of $\mathrm{B}$ proteins\tabularnewline
$\mathrm{DNA}_{\mathrm{C}}$  & $\rightharpoonup$  & $\mathrm{DNA}_{\mathrm{C}}+\mathrm{mRNA}_{\mathrm{C}}$  & $k_{9}=3.2$  & transcription of $\mathrm{mRNA}_{\mathrm{C}}$\tabularnewline
$\mathrm{mRNA}_{\mathrm{C}}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{C}}+\mathrm{C}$  & $k_{10}=3.0$  & translation of $\mathrm{C}$ proteins\tabularnewline
$\mathrm{mRNA}_{\mathrm{C}}$  & $\rightharpoonup$  & $\emptyset$  & $k_{11}=2.3$  & $\mathrm{mRNA}_{\mathrm{C}}$ decay\tabularnewline
$\mathrm{C}$  & $\rightharpoonup$  & $\emptyset$  & $k_{12}=2.5$  & decay of $\mathrm{C}$ proteins\tabularnewline
$\mathrm{mRNA}_{\mathrm{A}}+\mathrm{A}$  & $\rightharpoonup$  & $\mathrm{A}$  & $k_{13}=0$  & self regulation of $\mathrm{A}$ proteins\tabularnewline
$\mathrm{mRNA}_{\mathrm{B}}+\mathrm{B}$  & $\rightharpoonup$  & $\mathrm{B}$  & $k_{14}=0$  & self regulation of $\mathrm{B}$ proteins\tabularnewline
$\mathrm{mRNA}_{\mathrm{C}}+\mathrm{C}$  & $\rightharpoonup$  & $\mathrm{C}$  & $k_{15}=0$  & self regulation of $\mathrm{C}$ proteins\tabularnewline
$\mathrm{mRNA}_{\mathrm{B}}+\mathrm{A}$  & $\rightharpoonup$  & $\mathrm{A}$  & $k_{16}=0$  & regulation of $\mathrm{mRNA}_{\mathrm{B}}$\tabularnewline
$\mathrm{mRNA}_{\mathrm{C}}+\mathrm{B}$  & $\rightharpoonup$  & $\mathrm{B}$  & $k_{17}=0$  & regulation of $\mathrm{mRNA}_{\mathrm{C}}$\tabularnewline
$\mathrm{mRNA}_{\mathrm{A}}+\mathrm{C}$  & $\rightharpoonup$  & $\mathrm{C}$  & $k_{18}=0$  & regulation of $\mathrm{mRNA}_{\mathrm{A}}$\tabularnewline
$\mathrm{mRNA}_{\mathrm{C}}+\mathrm{A}$  & $\rightharpoonup$  & $\mathrm{A}$  & $k_{16}=6.0$  & regulation of $\mathrm{mRNA}_{\mathrm{C}}$\tabularnewline
$\mathrm{mRNA}_{\mathrm{B}}+\mathrm{C}$  & $\rightharpoonup$  & $\mathrm{C}$  & $k_{17}=4.0$  & regulation of $\mathrm{mRNA}_{\mathrm{B}}$\tabularnewline
$\mathrm{mRNA}_{\mathrm{A}}+\mathrm{B}$  & $\rightharpoonup$  & $\mathrm{B}$  & $k_{18}=3.0$  & regulation of $\mathrm{mRNA}_{\mathrm{A}}$\tabularnewline
$\mathrm{mRNA}_{\mathrm{A}}+\mathrm{A}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{A}}$  & $k_{19}=0$  & artificial fusion\tabularnewline
$\mathrm{mRNA}_{\mathrm{B}}+\mathrm{B}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{B}}$  & $k_{20}=0$  & artificial fusion\tabularnewline
$\mathrm{mRNA}_{\mathrm{A}}+\mathrm{B}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{A}}$  & $k_{21}=0$  & artificial fusion\tabularnewline
$\mathrm{mRNA}_{\mathrm{B}}+\mathrm{C}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{B}}$  & $k_{22}=0$  & artificial fusion\tabularnewline
$\mathrm{mRNA}_{\mathrm{C}}+\mathrm{A}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{C}}$  & $k_{23}=0$  & artificial fusion\tabularnewline
$\mathrm{mRNA}_{\mathrm{A}}+\mathrm{C}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{A}}$  & $k_{24}=0$  & artificial fusion\tabularnewline
$\mathrm{mRNA}_{\mathrm{B}}+\mathrm{A}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{B}}$  & $k_{25}=0$  & artificial fusion\tabularnewline
$\mathrm{A}+\mathrm{A}$  & $\rightharpoonup$  & $\mathrm{A}$  & $k_{26}=0$  & $\mathrm{A}$ regulates $\mathrm{A}$\tabularnewline
$\mathrm{B}+\mathrm{B}$  & $\rightharpoonup$  & $\mathrm{B}$  & $k_{27}=0$  & $\mathrm{B}$ regulates $\mathrm{B}$\tabularnewline
$\mathrm{C}+\mathrm{C}$  & $\rightharpoonup$  & $\mathrm{C}$  & $k_{28}=0$  & $\mathrm{C}$ regulates $\mathrm{C}$\tabularnewline
$\mathrm{B}+\mathrm{A}$  & $\rightharpoonup$  & $\mathrm{A}$  & $k_{29}=0$  & artificial fusion\tabularnewline
$\mathrm{C}+\mathrm{B}$  & $\rightharpoonup$  & $\mathrm{B}$  & $k_{30}=0$  & artificial fusion\tabularnewline
$\mathrm{A}+\mathrm{C}$  & $\rightharpoonup$  & $\mathrm{C}$  & $k_{31}=0$  & artificial fusion\tabularnewline
$\mathrm{C}+\mathrm{A}$  & $\rightharpoonup$  & $\mathrm{A}$  & $k_{32}=0$  & artificial fusion\tabularnewline
$\mathrm{B}+\mathrm{C}$  & $\rightharpoonup$  & $\mathrm{C}$  & $k_{33}=0$  & artificial fusion\tabularnewline
$\mathrm{A}+\mathrm{B}$  & $\rightharpoonup$  & $\mathrm{B}$  & $k_{34}=0$  & artificial fusion\tabularnewline
$\mathrm{A}$  & $\rightharpoonup$  & $\mathrm{B}$  & $k_{35}=0$  & artificial conversion\tabularnewline
$\mathrm{B}$  & $\rightharpoonup$  & $\mathrm{C}$  & $k_{36}=0$  & artificial conversion\tabularnewline
$\mathrm{C}$  & $\rightharpoonup$  & $\mathrm{A}$  & $k_{37}=0$  & artificial conversion\tabularnewline
$\mathrm{A}$  & $\rightharpoonup$  & $\mathrm{C}$  & $k_{38}=0$  & artificial conversion\tabularnewline
$\mathrm{C}$  & $\rightharpoonup$  & $\mathrm{B}$  & $k_{39}=0$  & artificial conversion\tabularnewline
$\mathrm{B}$  & $\rightharpoonup$  & $\mathrm{A}$  & $k_{40}=0$  & artificial conversion\tabularnewline
$\mathrm{mRNA}_{\mathrm{B}}+\mathrm{mRNA}_{\mathrm{C}}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{A}}$  & $k_{41}=0$  & artificial fusion\tabularnewline
$\mathrm{mRNA}_{\mathrm{C}}+\mathrm{mRNA}_{\mathrm{B}}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{C}}$  & $k_{42}=0$  & artificial fusion\tabularnewline
$\mathrm{mRNA}_{\mathrm{C}}+\mathrm{A}$  & $\rightharpoonup$  & $\mathrm{C}$  & $k_{43}=0$  & artificial fusion\tabularnewline
\bottomrule
\end{tabular}}\caption{\label{tab:reaction-library}Full set of ansatz reactions $\Theta$
used in Sec.~\ref{sec:results}. The given rate constants define
the ground truth reaction model.}
\end{table}

\linespread{1.5}

\clearpage
\newpage{}  \bibliographystyle{abbrv}
\bibliography{bibliography}
\newpage{}

\end{document}

### 2018-10-11 docs/makefigures/convert.sh

@@ -1,0 +1,5 @@
+convert -flatten -density 300 -units PixelsPerInch sparsity_pattern.png -resize 1011 sparsity_pattern.tiff
+convert -flatten -density 300 -units PixelsPerInch trajectories_lsq.png -resize 1011 trajectories_lsq.tiff
+convert -flatten -density 300 -units PixelsPerInch case1_cutoff.png -resize 1645.74 case1_cutoff.tiff
+convert -flatten -density 300 -units PixelsPerInch case2.png -resize 1011 case2.tiff
+convert -flatten -density 300 -units PixelsPerInch case3.png -resize 1011 case3.tiff

### 2018-11-28 docs/manuscript_jcp.tex

@@
 %% Do not edit unless you really know what you are doing.
-\documentclass[english,osajnl,jcp,preprint,groupedaddress,noshowpacs,showkeys,longbibliography,noeprint,nofootinbib,bibtex,ruled,vlined,english,natural,table,usenames,dvipsnames,a4paper]{revtex4-1}
+\documentclass[english,osajnl,jcp,reprint,groupedaddress,noshowpacs,showkeys,longbibliography,noeprint,nofootinbib,bibtex,ruled,vlined,english,natural,table,usenames,dvipsnames,a4paper]{revtex4-1}
 \usepackage[T1]{fontenc}
@@
 \linespread{0.956}
+
+\newif\ifthisispreprint
+% \thisispreprinttrue
+\thisispreprintfalse
 
@@
 \begin{align}
-\hat{\Xi}=\underset{\Xi}{\arg\min}\left(\frac{1}{2T}\left\Vert \dot{\textbf{X}}-\Theta(\textbf{X})\Xi\right\Vert _{F}^{2}+\alpha\lambda\|\Xi\|_{1}+\alpha(1-\lambda)\|\Xi\|_{2}^{2}\right)\quad\text{subject to }\Xi\geq0.\label{method:minimizationproblem}
+\hat{\Xi}&=\underset{\Xi}{\arg\min}\Bigg(\frac{1}{2T}\left\Vert \dot{\textbf{X}}-\Theta(\textbf{X})\Xi\right\Vert _{F}^{2} \label{method:minimizationproblem} \\
+&\quad+\alpha\lambda\|\Xi\|_{1}+\alpha(1-\lambda)\|\Xi\|_{2}^{2}\Bigg)\quad\text{subject to }\Xi\geq0.\nonumber
 \end{align}
@@
 i & \rightharpoonup\emptyset &  & \text{(decay of protein)},\\
-i+\mathrm{mRNA}_{j} & \rightharpoonup i &  & \text{(regulation of protein }j\in S\text{),}
+i+\mathrm{mRNA}_{j} & \rightharpoonup i &  & \text{(regulation of }j\in S\text{),}
 \end{align*}
@@
 
-\begin{figure}
-\centering \includegraphics[width=\singlecolumnwidth]{figures_tex/scheme}
+\begin{figure}[h!]
+\centering \includegraphics[width=\singlecolumnwidth]{figures_jcp/1.pdf}
 \caption{\label{fig:network}The regulation network example described in Sec.~\ref{sec:results}.
@@
 \centering%
+\ifthisispreprint
+% ignore
+\else
+\resizebox{.95\columnwidth}{!}{%
+\fi
 \begin{tabular}{cccccccccc}
@@
 \textbf{(b)} & $1$ & $1.5$ & $0$ & $1$ & $0$ & $2$ & $1$ & $0$ & $1$\tabularnewline
-\end{tabular}\caption{\label{tab:Initial-conditions}Initial conditions used to generate
+\end{tabular}
+\ifthisispreprint
+% ignore
+\else
+}
+\fi
+\caption{\label{tab:Initial-conditions}Initial conditions \textbf{(a)} and \textbf{(b)} used to generate
 concentration time series. Reaction rates can be found in Tab.~\ref{tab:reaction-library}.}
@@
 particles is completely ignored (Fig.~\ref{fig:case-1-sparsity-pattern}).
-
 \begin{figure}
-\centering\includegraphics[width=\singlecolumnwidth]{figures_tex/trajectories_lsq}\caption{\label{fig:case-1-lsq}Concentration time series generated from integrating
+\centering\includegraphics[width=\singlecolumnwidth]{figures_jcp/2.pdf}\caption{\label{fig:case-1-lsq}Concentration time series generated from integrating
 the reaction network shown in Fig.~\ref{fig:network}a. The initial
@@
 \end{figure}
+
+\ifthisispreprint
+% ignore
+\else
+\onecolumngrid
+\fi
+
 \begin{figure}
-\centering \includegraphics[width=0.82\doublecolumnwidth]{figures_tex/sparsity_pattern}
+\centering \includegraphics[width=0.82\doublecolumnwidth]{figures_jcp/3.pdf}
 \caption{\label{fig:case-1-sparsity-pattern}Estimated reaction rates in the
@@
 \end{figure}
+\ifthisispreprint
+% ignore
+\else
+\twocolumngrid
+\fi
+
 Next, we sought sparse solutions by using $\alpha>0$ and additionally
@@
 \begin{figure}
-\centering \includegraphics[width=\singlecolumnwidth]{figures_tex/case2}
+\centering \includegraphics[width=\singlecolumnwidth]{figures_jcp/4.pdf}
 \caption{\label{fig:case2-convergence}Convergence of the estimation error
@@
 \begin{figure}
-\centering \includegraphics[width=\singlecolumnwidth]{figures_tex/case3}
+\centering \includegraphics[width=\singlecolumnwidth]{figures_jcp/5.pdf}
 \caption{\label{fig:case3-convergence}Convergence of estimation error of reaction
@@
 
-\centering\includegraphics[width=\singlecolumnwidth]{figures_tex/case1_cutoff}\caption{\label{fig:reaction-rates-sorted}Reaction rates sorted by their magnitude
+\centering\includegraphics[width=\singlecolumnwidth]{figures_jcp/6.pdf}\caption{\label{fig:reaction-rates-sorted}Reaction rates sorted by their magnitude
 to determine the cutoff $\kappa=0.22$ of Sec.~\ref{sec:case-1}.
@@
 \end{figure}
+
+\ifthisispreprint
+% ignore
+\else
+\onecolumngrid
+\fi
+
 \begin{table}[h!]
 \centering %
-\resizebox{.8\textwidth}{!}{%
+%\resizebox{.9\textwidth}{!}{%
 \begin{tabular}{rclcc}
@@
 \bottomrule
-\end{tabular}}\caption{\label{tab:reaction-library}Full set of ansatz reactions $\Theta$
+\end{tabular}\caption{\label{tab:reaction-library}Full set of ansatz reactions $\Theta$
 used in Sec.~\ref{sec:results}. The given rate constants define
@@
 
-\linespread{1.5}
+\ifthisispreprint
+% ignore
+\else
+\twocolumngrid
+\fi
+
+\linespread{1.45}
 

### 2018-11-28 readdy_learn/analyze/cross_validation.py
update

@@ -56,10 +56,10 @@ class CrossValidation(object):
     def _solve(self, traj, alpha, l1_ratio, tol=1e-16):
         initial_value = _np.zeros((self._bfc.n_basis_functions,), dtype=_np.float64)
         estimator = _estimator.ReaDDyElasticNetEstimator(traj, self._bfc, alpha=alpha, l1_ratio=l1_ratio,
-                                                         maxiter=30000, method='SLSQP', verbose=False, approx_jac=False,
+                                                         maxiter=5000, method='SLSQP', verbose=False, approx_jac=False,
                                                          options={'ftol': tol}, rescale=False,
                                                          init_xi=initial_value,
-                                                         constrained=True)
+                                                         constrained=False)
 
         estimator.fit(None)
         if estimator.success_:
@@ -71,8 +71,8 @@ class CrossValidation(object):
     def _score(self, traj, rates):
         initial_value = _np.zeros((self._bfc.n_basis_functions,), dtype=_np.float64)
         estimator = _estimator.ReaDDyElasticNetEstimator(traj, self._bfc, alpha=0, l1_ratio=0,
-                                                         maxiter=30000, method='SLSQP', verbose=False, approx_jac=False,
-                                                         options={'ftol': 1e-12}, rescale=False,
+                                                         maxiter=5000, method='SLSQP', verbose=False, approx_jac=False,
+                                                         options={'ftol': 1e-13}, rescale=False,
                                                          init_xi=initial_value,
                                                          constrained=True)
         estimator.coefficients_ = rates
@@ -98,10 +98,11 @@ class CrossValidation(object):
             assert test_traj.n_time_steps == len(test)
             assert train_traj.dcounts_dt.shape[0] == len(train)
             assert test_traj.dcounts_dt.shape[0] == len(test)
-            tolerances_to_try = [1e-16, 1e-15, 1e-14, 1e-13, 1e-12, 1e-11, 1e-10, 1e-9, 1e-8]
+            tolerances_to_try = [1e-15, 1e-14, 1e-13, 1e-12, 1e-11, 1e-10, 1e-9, 1e-8]
 
             rates = None
             for tol in tolerances_to_try:
+                print(f"Solving for tolerance {tol}")
                 try:
                     rates = self._solve(train_traj, alpha, l1_ratio, tol=tol)
                     break

### 2018-11-30 readdy_learn/analyze/estimator.py
mapk

@@ -160,7 +160,7 @@ class ReaDDyElasticNetEstimator(BaseEstimator):
         jac = False if self.approx_jac else \
             lambda x: opt.elastic_net_objective_fun_jac(x, self.alpha, self.l1_ratio, large_theta, expected)
 
-        options = {'disp': False, 'maxiter': self.maxiter}
+        options = {'disp': False, 'maxiter': self.maxiter, 'iprint': 2}
         if self.method == 'L-BFGS-B':
             options['maxfun'] = self.maxiter
         options.update(self.options)

### 2018-11-30 readdy_learn/example/mapk.py

@@ -4,6 +4,17 @@ import itertools
 
 import readdy_learn.analyze.basis as _basis
 
+
+N_STIMULUS = 1
+SPECIES_NAMES = ["S", "MAPKKK", "MAPKKK*", "MAPKK", "MAPKK*", "MAPK", "MAPK*", "TF", "TF*"]
+N_SPECIES = len(SPECIES_NAMES)
+TIMESTEP = 1e-3
+
+INITIAL_STATES = [
+    [N_STIMULUS, 1, 0, 1, 0, 1, 0, 1, 0]
+]
+
+
 ALL_BOGUS_CONVERSION_OPS = [
     lambda result: result.add_double_conversion((1, 3), (1, 4)),
     lambda result: result.add_double_conversion((1, 5), (1, 6)),
@@ -39,20 +50,20 @@ def conversion_ops_range(start, end):
 
 class MAPKConfiguration(object):
 
-    def __init__(self, conversion_ops_selection):
+    def __init__(self, conversion_ops_selection=_np.arange(len(ALL_BOGUS_CONVERSION_OPS))):
         self.conversion_ops = [ALL_BOGUS_CONVERSION_OPS[i] for i in conversion_ops_selection]
         self.n_bogus = len(self.conversion_ops)
         self.n_real = 8
         self.n_total = self.n_real + self.n_bogus
         self.rates = _np.array([
             1.,
-            1000.,
             1.,
-            1000.,
             1.,
-            1000.,
             1.,
-            1000.,
+            1.,
+            1.,
+            1.,
+            1.,
         ])
         self.rates = _np.concatenate((self.rates, _np.zeros((self.n_bogus,))))
 
@@ -72,16 +83,6 @@ class MAPKConfiguration(object):
             op(result)
 
         return result
-
-
-N_STIMULUS = 100
-SPECIES_NAMES = ["S", "MAPKKK", "MAPKKK*", "MAPKK", "MAPKK*", "MAPK", "MAPK*", "TF", "TF*"]
-N_SPECIES = len(SPECIES_NAMES)
-TIMESTEP = 1e-3
-
-INITIAL_STATES = [
-    [N_STIMULUS, 1000, 0, 1000, 0, 1000, 0, 1000, 0]
-]
 
 
 def generate_kmc(target_time: float, n_realizations: int, config: MAPKConfiguration, njobs: int):
[…]

### 2018-11-30 readdy_learn/example/mapk_enqueue_command.sh

@@ -1,0 +1,29 @@
+#!/usr/bin/env bash
+#SBATCH --job-name=rlearn
+#SBATCH --workdir=/home/clonker/workspace/readdy_learn/allegro
+#SBATCH --output=/home/clonker/NO_BACKUP/rlearn/rlearn.%j.%a.out
+#SBATCH --partition=small
+#SBATCH --exclude=cmp245
+#SBATCH --nodes=1
+#SBATCH --cpus-per-task=1
+#SBATCH --mem=2990M
+#SBATCH --time=40:00:00
+#SBATCH --mail-type=end
+#SBATCH --mail-user=clonker@gmail.com
+
+echo "Running enqueue_command .."
+
+if [[ $# -eq 0 ]]; then
+  echo "Specify the command to run"
+  exit 1
+fi
+
+hostname
+date
+echo "Calling command:"
+echo $1
+
+eval $1
+
+echo "Done .."
+

### 2018-11-30 readdy_learn/analyze/cross_validation.py
push

@@ -59,7 +59,7 @@ class CrossValidation(object):
                                                          maxiter=5000, method='SLSQP', verbose=False, approx_jac=False,
                                                          options={'ftol': tol}, rescale=False,
                                                          init_xi=initial_value,
-                                                         constrained=False)
+                                                         constrained=True)
 
         estimator.fit(None)
         if estimator.success_:
@@ -72,7 +72,7 @@ class CrossValidation(object):
         initial_value = _np.zeros((self._bfc.n_basis_functions,), dtype=_np.float64)
         estimator = _estimator.ReaDDyElasticNetEstimator(traj, self._bfc, alpha=0, l1_ratio=0,
                                                          maxiter=5000, method='SLSQP', verbose=False, approx_jac=False,
-                                                         options={'ftol': 1e-13}, rescale=False,
+                                                         options={'ftol': 1e-16}, rescale=False,
                                                          init_xi=initial_value,
                                                          constrained=True)
         estimator.coefficients_ = rates
@@ -98,7 +98,7 @@ class CrossValidation(object):
             assert test_traj.n_time_steps == len(test)
             assert train_traj.dcounts_dt.shape[0] == len(train)
             assert test_traj.dcounts_dt.shape[0] == len(test)
-            tolerances_to_try = [1e-15, 1e-14, 1e-13, 1e-12, 1e-11, 1e-10, 1e-9, 1e-8]
+            tolerances_to_try = [1e-16, 1e-15, 1e-14, 1e-13, 1e-12, 1e-11, 1e-10, 1e-9, 1e-8]
 
             rates = None
             for tol in tolerances_to_try:

### 2018-11-30 readdy_learn/analyze/cross_validation.py
push

@@ -89,9 +89,11 @@ class CrossValidation(object):
         if self._splitter == 'shuffle':
             splitter = _ShuffleSplit(n_splits=self.n_splits, test_size=.5)
         else:
+            print("Running kfold with n_splits={}".format(self.n_splits))
             splitter = _KFold(n_splits=self.n_splits)
         scores = []
         for train, test in splitter.split(_np.arange(n_steps_total)):
+            print("Test:", test)
             train_traj = self._obtain_trajs_subset(train)
             test_traj = self._obtain_trajs_subset(test)
             assert train_traj.n_time_steps == len(train)

### 2018-11-30 readdy_learn/example/predator_prey.py
push

@@ -6,18 +6,18 @@ import readdy_learn.analyze.basis as _basis
 FRICTION_PREY = .1
 FRICTION_PREDATOR = .1
 
-BETA = 200.
+BETA = 1
 
 INITIAL_STATES = _np.array([
-    [1500, 1500],
+    [1, 1],
 ])
 
 RATES = _np.array([
-    10.,  # X + X -> 0
-    10.,  # Y + Y -> 0
-    1000 * BETA,  # X -> X + X
+    .1,  # X + X -> 0
+    .1,  # Y + Y -> 0
+    1 * BETA,  # X -> X + X
     BETA,  # X + Y -> Y + Y
-    1000 * BETA,  # Y -> 0
+    1 * BETA,  # Y -> 0
     0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0  # bogus
 ])
 
@@ -135,7 +135,7 @@ def solve(counts, dcounts_dt, alpha, l1_ratio):
                                                    maxiter=30000, method='SLSQP', verbose=True, approx_jac=False,
                                                    options={'ftol': tol}, rescale=False,
                                                    init_xi=_np.zeros_like(RATES),
-                                                   constrained=False)
+                                                   constrained=True)
 
         estimator.fit(None)
         if estimator.success_:

### 2018-11-30 docs/manuscript_jcp.tex
update

@@
 identifying a reaction network from time traces of reactant concentrations
-can be solved by finding a linear combination from a library of nonlinear
-ansatz functions that each corresponds to a reaction acting on a set
+can be solved by finding a linear combination from a library of candidate
+nonlinear functions (ansatz functions) that each corresponds to a reaction acting on a set
 of reactants. With this formulation, the reaction rates can be determined
@@
 The two reactions in the previous example (\ref{eq:example_reaction_1}-\ref{eq:example_reaction_2})
-would be modeled by the ansatz functions
+would be modeled by the functions
 \begin{align*}
-y_{1}(\mathbf{x}) & =(-x_{1},x_{1},0)^{\top}\\
+y_{1}(\mathbf{x}) & =(-x_{1},x_{1},0)^{\top},\\
 y_{2}(\mathbf{x}) & =(-x_{1},0,x_{1})^{\top},
@@
 Given this matrix, a library $\Theta:\mathbb{R}^{T\times S}\to\mathbb{R}^{T\times S\times R},\;\mathbf{X}\mapsto\begin{pmatrix}\theta_{1}(\mathbf{X}) & \theta_{2}(\mathbf{X}) & \cdots & \theta_{R}(\mathbf{X})\end{pmatrix}$
-of $R$ ansatz reactions can be proposed with corresponding reaction
+of $R$ candidate (ansatz) reactions can be proposed with corresponding reaction
 functions 
@@
 
-\ifthisispreprint
+%\ifthisispreprint
 % ignore
-\else
-\onecolumngrid
-\fi
+%\else
+%\onecolumngrid
+%\fi
 
-\begin{figure}
+\begin{figure*}
 \centering \includegraphics[width=0.82\doublecolumnwidth]{figures_jcp/3.pdf}
@@
 of one circle are displayed.}
-\end{figure}
-\ifthisispreprint
+\end{figure*}
+%\ifthisispreprint
 % ignore
-\else
-\twocolumngrid
-\fi
+%\else
+%\twocolumngrid
+%\fi
 
@@
 
-The value of the cutoff was determined by comparing the magnitude
+The value of the cutoff $\kappa$ was determined by comparing the magnitude
 of estimated rates and finding a gap, see Fig.~\ref{fig:reaction-rates-sorted}.
@@
 \[
-\mathbf{X}=\begin{pmatrix}\boldsymbol{x}_{1}(t_{1}) & \boldsymbol{x}_{2}(t_{1}) & \cdots & \boldsymbol{x}_{1}(t_{667}) & \boldsymbol{x}_{2}(t_{667})\end{pmatrix},
+\mathbf{X}=\begin{pmatrix}\boldsymbol{x}_{1}(t_{1}) & \cdots & \boldsymbol{x}_{1}(t_{667}) & \boldsymbol{x}_{2}(t_{1}) & \cdots & \boldsymbol{x}_{2}(t_{667})\end{pmatrix},
 \]
@@
 
-\ifthisispreprint
+%\ifthisispreprint
 % ignore
-\else
-\onecolumngrid
-\fi
+%\else
+%\onecolumngrid
+%\fi
 
-\begin{table}[h!]
+\begin{table*}[h!]
 \centering %
@@
 the ground truth reaction model.}
-\end{table}
+\end{table*}
 
-\ifthisispreprint
+%\ifthisispreprint
 % ignore
-\else
-\twocolumngrid
-\fi
+%\else
+%\twocolumngrid
+%\fi
 

### 2018-12-04 docs/manuscript_jcp.tex
run mapk cv again with more points, plot results for mapk and predator-prey, wip manuscript

@@
 the constraint $\Xi\geq0$. Reactive SINDy (\ref{method:minimizationproblem})
-is therefore a generalization of the SINDy method to to the vector-valued
+is therefore a generalization of the SINDy method to vector-valued
 ansatz functions.
@@
 
+\subsection{Application to MAPK cascade\label{sec:mapk}}
+
+\subsection{Application to Lotka-Volterra system\label{sec:lv}}
+
 \section{Conclusion}
@@
 \bottomrule
-\end{tabular}\caption{\label{tab:reaction-library}Full set of ansatz reactions $\Theta$
-used in Sec.~\ref{sec:results}. The given rate constants define
-the ground truth reaction model.}
+\end{tabular}\caption{\label{tab:reaction-library}Full set of ansatz reactions $\Theta$ used in Sec.~\ref{sec:results} for the gene-regulatory network.
+The given rate constants define the ground truth reaction model.}
+\end{table*}
+
+\begin{table*}[h!]
+    \centering %
+    %\resizebox{.9\textwidth}{!}{%
+    \begin{tabular}{rclcc}
+        \toprule 
+        \multicolumn{3}{c}{Reaction} & rate  & description \tabularnewline
+        \midrule 
+        $\mathrm{DNA}_{\mathrm{A}}$  & $\rightharpoonup$  & $\mathrm{DNA}_{\mathrm{A}}+\mathrm{mRNA}_{\mathrm{A}}$  & $k_{1}=1.8$  & transcription of $\mathrm{mRNA}_{\mathrm{A}}$\tabularnewline
+        $\mathrm{mRNA}_{\mathrm{A}}+\mathrm{A}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{A}}$  & $k_{19}=0$  & artificial fusion\tabularnewline
+        $\mathrm{mRNA}_{\mathrm{B}}+\mathrm{B}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{B}}$  & $k_{20}=0$  & artificial fusion\tabularnewline
+        \bottomrule
+    \end{tabular}\caption{\label{tab:reaction-library-mapk}Full set of ansatz reactions $\Theta$ used in Sec.~\ref{sec:mapk} for the MAPK system.
+The given rate constants define the ground truth reaction model.}
+\end{table*}
+
+\begin{table*}[h!]
+    \centering %
+    %\resizebox{.9\textwidth}{!}{%
+    \begin{tabular}{rclcc}
+        \toprule 
+        \multicolumn{3}{c}{Reaction} & rate  & description \tabularnewline
+        \midrule 
+        $\mathrm{DNA}_{\mathrm{A}}$  & $\rightharpoonup$  & $\mathrm{DNA}_{\mathrm{A}}+\mathrm{mRNA}_{\mathrm{A}}$  & $k_{1}=1.8$  & transcription of $\mathrm{mRNA}_{\mathrm{A}}$\tabularnewline
+        $\mathrm{mRNA}_{\mathrm{A}}+\mathrm{A}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{A}}$  & $k_{19}=0$  & artificial fusion\tabularnewline
+        $\mathrm{mRNA}_{\mathrm{B}}+\mathrm{B}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{B}}$  & $k_{20}=0$  & artificial fusion\tabularnewline
+        \bottomrule
+    \end{tabular}\caption{\label{tab:reaction-library-lv}Full set of ansatz reactions $\Theta$ used in Sec.~\ref{sec:lv} for the Lotka-Volterra system.
+The given rate constants define the ground truth reaction model.}
 \end{table*}

### 2018-12-04 docs/makefigures/mapk_and_pprey.py

@@ -1,0 +1,133 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+if __name__ == '__main__':
+    plt.style.use("./rlearn.mplstyle")
+
+    # MAPK
+    print(5 * "-----" + "MAPK" + 5 * "-----")
+    mapk_file = "../../mapk_pathway/mapk_data.npz"
+    with np.load(mapk_file) as f:
+        time = f["time"]
+        counts = f["counts"]
+        dcounts_dt = f["dcounts_dt"]
+        alphas_gs = f["alphas_gs"]
+        l1_errs_gs = f["l1_errs_gs"]
+        estimated_rates_gs = f["estimated_rates_gs"]
+        best_estimated_rates_gs = f["best_estimated_rates_gs"]
+        alphas_cv = f["alphas_cv"]
+        scores_cv = f["scores_cv"]
+        rates_cv = f["rates_cv"]
+        rates_lsq = f["rates_lsq"]
+        rates_desired = f["rates_desired"]
+        activities = f["activities"]
+        stimulus = f["stimulus"]
+        init_state_activities = f["init_state_activities"]
+
+    # plot MAPK activation curve
+    plt.semilogx(stimulus, activities)
+    plt.xlabel('stimulus')
+    plt.ylabel('activity')
+    plt.vlines(init_state_activities, *plt.ylim(), color="xkcd:red", label="initial conditions", lw=0.5)
+    plt.legend()
+    plt.gcf().tight_layout()
+    plt.savefig("mapk_activation.pdf")
+    #plt.show()
+    plt.clf()
+
+    # @todo plot data ?
+
+    # plot cross validation scores
+    sel = np.argmin(-scores_cv)
+    plt.semilogx(alphas_cv, -scores_cv)
+    #plt.plot([alphas_cv[sel]], [-scores_cv[sel]], 'x', markersize=20)
+    plt.ylim(0., 0.5e-9)
+    xmin, xmax = plt.xlim()
+    plt.xlim(xmin, 2e-8)
+    plt.xlabel(r"Hyperparameter $\alpha$")
+    plt.ylabel(r"Mean loss in val. set")
+    plt.gcf().tight_layout()
+    plt.savefig("mapk_cross_validation.pdf", bbox_inches="tight")
+    #plt.show()
+    plt.clf()
+
+    # plot resulting models
+    plt.plot(rates_cv, 'o', label="regularized")
+    plt.plot(rates_lsq, 'd', label="LSQ")
[…]

### 2018-12-05 docs/manuscript_jcp.tex
rename fig '6.pdf' to '8.pdf' due to two new figures beforehand, finalize plotting of mapk and pprey, specify ansatz reactions

@@
 
-\centering\includegraphics[width=\singlecolumnwidth]{figures_jcp/6.pdf}\caption{\label{fig:reaction-rates-sorted}Reaction rates sorted by their magnitude
+\centering\includegraphics[width=\singlecolumnwidth]{figures_jcp/8.pdf}\caption{\label{fig:reaction-rates-sorted}Reaction rates sorted by their magnitude
 to determine the cutoff $\kappa=0.22$ of Sec.~\ref{sec:case-1}.
@@
     %\resizebox{.9\textwidth}{!}{%
-    \begin{tabular}{rclcc}
+    \begin{tabular}{rclc}
         \toprule 
-        \multicolumn{3}{c}{Reaction} & rate  & description \tabularnewline
+        \multicolumn{3}{c}{Reaction} & rate \tabularnewline
         \midrule 
-        $\mathrm{DNA}_{\mathrm{A}}$  & $\rightharpoonup$  & $\mathrm{DNA}_{\mathrm{A}}+\mathrm{mRNA}_{\mathrm{A}}$  & $k_{1}=1.8$  & transcription of $\mathrm{mRNA}_{\mathrm{A}}$\tabularnewline
-        $\mathrm{mRNA}_{\mathrm{A}}+\mathrm{A}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{A}}$  & $k_{19}=0$  & artificial fusion\tabularnewline
-        $\mathrm{mRNA}_{\mathrm{B}}+\mathrm{B}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{B}}$  & $k_{20}=0$  & artificial fusion\tabularnewline
+        $\mathrm{S}        + \mathrm{MAPKKK}$  & $\rightharpoonup$  &  $\mathrm{S      } + \mathrm{MAPKKK*}$ & $k_{1} =1$ \tabularnewline
+        $\mathrm{MAPKKK*}    \mathrm{      }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK }                   $ & $k_{2} =1$ \tabularnewline
+        $\mathrm{MAPKKK*}  + \mathrm{MAPKK }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK*} + \mathrm{MAPKK* }$ & $k_{3} =1$ \tabularnewline
+        $\mathrm{MAPKK*}     \mathrm{      }$  & $\rightharpoonup$  &  $\mathrm{MAPKK  }                   $ & $k_{4} =1$ \tabularnewline
+        $\mathrm{MAPKK*}   + \mathrm{MAPK  }$  & $\rightharpoonup$  &  $\mathrm{MAPKK* } + \mathrm{MAPK*  }$ & $k_{5} =1$ \tabularnewline
+        $\mathrm{MAPK*}      \mathrm{      }$  & $\rightharpoonup$  &  $\mathrm{MAPK   }                   $ & $k_{6} =1$ \tabularnewline
+        $\mathrm{MAPK*}    + \mathrm{TF    }$  & $\rightharpoonup$  &  $\mathrm{MAPK*  } + \mathrm{TF*    }$ & $k_{7} =1$ \tabularnewline
+        $\mathrm{TF*}        \mathrm{      }$  & $\rightharpoonup$  &  $\mathrm{TF     }                   $ & $k_{8} =1$ \tabularnewline
+        $\mathrm{MAPKKK}   + \mathrm{MAPKK }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK } + \mathrm{MAPKK* }$ & $k_{9} =0$ \tabularnewline
+        $\mathrm{MAPKKK}   + \mathrm{MAPK  }$  & $\rightharpoonup$  &  $\mathrm{MAPK*  }                   $ & $k_{10}=0$ \tabularnewline
+        $\mathrm{MAPKKK}   + \mathrm{TF    }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK } + \mathrm{TF*    }$ & $k_{11}=0$ \tabularnewline
+        $\mathrm{MAPKKK*}  + \mathrm{MAPK  }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK*} + \mathrm{MAPK*  }$ & $k_{12}=0$ \tabularnewline
+        $\mathrm{MAPKKK*}  + \mathrm{TF    }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK*} + \mathrm{TF*    }$ & $k_{13}=0$ \tabularnewline
+        $\mathrm{MAPKK}    + \mathrm{TF    }$  & $\rightharpoonup$  &  $\mathrm{MAPKK  } + \mathrm{TF*    }$ & $k_{14}=0$ \tabularnewline
+        $\mathrm{MAPKK*}   + \mathrm{TF    }$  & $\rightharpoonup$  &  $\mathrm{MAPKK* } + \mathrm{TF*    }$ & $k_{15}=0$ \tabularnewline
+        $\mathrm{MAPK}     + \mathrm{TF    }$  & $\rightharpoonup$  &  $\mathrm{MAPK   } + \mathrm{TF*    }$ & $k_{16}=0$ \tabularnewline
+        $\mathrm{MAPKK}    + \mathrm{MAPK  }$  & $\rightharpoonup$  &  $\mathrm{MAPKK  } + \mathrm{MAPK*  }$ & $k_{17}=0$ \tabularnewline
+        $\mathrm{MAPKKK}   + \mathrm{MAPKK*}$  & $\rightharpoonup$  &  $\mathrm{MAPKKK } + \mathrm{MAPKK  }$ & $k_{18}=0$ \tabularnewline
+        $\mathrm{MAPKKK}   + \mathrm{MAPK* }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK } + \mathrm{MAPK   }$ & $k_{19}=0$ \tabularnewline
+        $\mathrm{MAPKKK}   + \mathrm{TF*   }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK } + \mathrm{TF     }$ & $k_{20}=0$ \tabularnewline
+        $\mathrm{MAPKKK*}  + \mathrm{MAPKK*}$  & $\rightharpoonup$  &  $\mathrm{MAPKKK*} + \mathrm{MAPKK  }$ & $k_{21}=0$ \tabularnewline
+        $\mathrm{MAPKKK*}  + \mathrm{MAPK* }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK*} + \mathrm{MAPK   }$ & $k_{22}=0$ \tabularnewline
+        $\mathrm{MAPKKK*}  + \mathrm{TF*   }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK*} + \mathrm{TF     }$ & $k_{23}=0$ \tabularnewline
+        $\mathrm{MAPKK}    + \mathrm{MAPK* }$  & $\rightharpoonup$  &  $\mathrm{MAPKK  } + \mathrm{MAPK   }$ & $k_{24}=0$ \tabularnewline
+        $\mathrm{MAPKK}    + \mathrm{TF*   }$  & $\rightharpoonup$  &  $\mathrm{MAPKK  } + \mathrm{TF     }$ & $k_{25}=0$ \tabularnewline
+        $\mathrm{MAPKK*}   + \mathrm{MAPK* }$  & $\rightharpoonup$  &  $\mathrm{MAPKK* } + \mathrm{MAPK   }$ & $k_{26}=0$ \tabularnewline
+        $\mathrm{MAPKK*}   + \mathrm{TF*   }$  & $\rightharpoonup$  &  $\mathrm{MAPKK* } + \mathrm{TF     }$ & $k_{27}=0$ \tabularnewline
+        $\mathrm{MAPK}     + \mathrm{TF*   }$  & $\rightharpoonup$  &  $\mathrm{MAPK   } + \mathrm{TF     }$ & $k_{28}=0$ \tabularnewline
+        $\mathrm{MAPK*}    + \mathrm{TF*   }$  & $\rightharpoonup$  &  $\mathrm{MAPK*  } + \mathrm{TF     }$ & $k_{29}=0$ \tabularnewline
         \bottomrule
@@
     %\resizebox{.9\textwidth}{!}{%
-    \begin{tabular}{rclcc}
+    \begin{tabular}{rclc}
         \toprule 
-        \multicolumn{3}{c}{Reaction} & rate  & description \tabularnewline
+        \multicolumn{3}{c}{Reaction} & rate \tabularnewline
         \midrule 
-        $\mathrm{DNA}_{\mathrm{A}}$  & $\rightharpoonup$  & $\mathrm{DNA}_{\mathrm{A}}+\mathrm{mRNA}_{\mathrm{A}}$  & $k_{1}=1.8$  & transcription of $\mathrm{mRNA}_{\mathrm{A}}$\tabularnewline
-        $\mathrm{mRNA}_{\mathrm{A}}+\mathrm{A}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{A}}$  & $k_{19}=0$  & artificial fusion\tabularnewline
-        $\mathrm{mRNA}_{\mathrm{B}}+\mathrm{B}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{B}}$  & $k_{20}=0$  & artificial fusion\tabularnewline
+        $\mathrm{X} + \mathrm{X}$ & $\rightharpoonup$ & $\emptyset             $ & $k_{1} =0.1$ \tabularnewline
+        $\mathrm{Y} + \mathrm{Y}$ & $\rightharpoonup$ & $\emptyset             $ & $k_{2} =0.1$ \tabularnewline
+        $\mathrm{X}             $ & $\rightharpoonup$ & $\mathrm{X} + \mathrm{X}$ & $k_{3} =1  $ \tabularnewline
+        $\mathrm{X} + \mathrm{Y}$ & $\rightharpoonup$ & $\mathrm{Y} + \mathrm{Y}$ & $k_{4} =1  $ \tabularnewline
+        $\mathrm{Y}             $ & $\rightharpoonup$ & $\emptyset             $ & $k_{5} =1  $ \tabularnewline
+        $\mathrm{X} + \mathrm{Y}$ & $\rightharpoonup$ & $\mathrm{X} + \mathrm{X}$ & $k_{6} =0  $ \tabularnewline
+        $\mathrm{X}             $ & $\rightharpoonup$ & $\emptyset             $ & $k_{7} =0  $ \tabularnewline
+        $\mathrm{Y} + \mathrm{Y}$ & $\rightharpoonup$ & $\mathrm{Y}             $ & $k_{8} =0  $ \tabularnewline
+        $\mathrm{Y}             $ & $\rightharpoonup$ & $\mathrm{Y} + \mathrm{Y}$ & $k_{9} =0  $ \tabularnewline
+        $\mathrm{X} + \mathrm{X}$ & $\rightharpoonup$ & $\mathrm{X}             $ & $k_{10}=0  $ \tabularnewline
+        $\mathrm{X} + \mathrm{Y}$ & $\rightharpoonup$ & $\mathrm{X}             $ & $k_{11}=0  $ \tabularnewline
+        $\mathrm{X} + \mathrm{Y}$ & $\rightharpoonup$ & $\mathrm{Y}             $ & $k_{12}=0  $ \tabularnewline
+        $\mathrm{X} + \mathrm{X}$ & $\rightharpoonup$ & $\mathrm{Y}             $ & $k_{13}=0  $ \tabularnewline
+        $\mathrm{X}             $ & $\rightharpoonup$ & $\mathrm{Y}             $ & $k_{14}=0  $ \tabularnewline
+        $\mathrm{Y}             $ & $\rightharpoonup$ & $\mathrm{X}             $ & $k_{15}=0  $ \tabularnewline
+        $\mathrm{X}             $ & $\rightharpoonup$ & $\mathrm{Y} + \mathrm{Y}$ & $k_{16}=0  $ \tabularnewline
         \bottomrule

### 2018-12-05 docs/makefigures/mapk_and_pprey.py

@@ -1,4 +1,6 @@
 import matplotlib.pyplot as plt
+from matplotlib.gridspec import GridSpec
+from matplotlib.ticker import LogLocator
 import numpy as np
 
 if __name__ == '__main__':
@@ -24,18 +26,9 @@ if __name__ == '__main__':
         stimulus = f["stimulus"]
         init_state_activities = f["init_state_activities"]
 
-    # plot MAPK activation curve
-    plt.semilogx(stimulus, activities)
-    plt.xlabel('stimulus')
-    plt.ylabel('activity')
-    plt.vlines(init_state_activities, *plt.ylim(), color="xkcd:red", label="initial conditions", lw=0.5)
-    plt.legend()
-    plt.gcf().tight_layout()
-    plt.savefig("mapk_activation.pdf")
-    #plt.show()
-    plt.clf()
+    # plot data ?
+    print(f"Number of frames: {len(time)}")
 
-    # @todo plot data ?
 
     # plot cross validation scores
     sel = np.argmin(-scores_cv)
@@ -51,29 +44,61 @@ if __name__ == '__main__':
     #plt.show()
     plt.clf()
 
-    # plot resulting models
-    plt.plot(rates_cv, 'o', label="regularized")
-    plt.plot(rates_lsq, 'd', label="LSQ")
-    #plt.vlines([7.5], 0, 1, 'grey', 'dashed')
-    plt.plot(rates_desired, 'x', label="ground truth")
-    labels = list(map(str, range(1, len(rates_desired) + 1)))
-    for i in range(len(labels)):
-        if i % 3 == 0:
-            pass
-        else:
-            labels[i] = ""
-    plt.xticks(ticks=range(len(rates_desired)), labels=labels, fontsize=11)
-    plt.legend()
-    plt.xlabel(r"Ansatz reaction \#")
-    plt.ylabel("Rate constant in a.u.")
-    plt.gcf().tight_layout()
-    plt.savefig("mapk_result_models.pdf", bbox_inches="tight")
+    def plot_mapk_activation_curve():
+        plt.semilogx(stimulus, activities, label="MAPK cascade response")
+        plt.xlabel('Stimulus in a.u.')
+        plt.ylabel('Activity in a.u.')
+        ymin, ymax = plt.ylim()
+        plt.ylim(ymin, 0.35)
[…]

### 2018-12-05 readdy_learn/example/mapk.py

@@ -6,6 +6,7 @@ import readdy_learn.analyze.basis as _basis
 
 
 N_STIMULUS = 1
+# ids             0    1         2          3        4         5       6        7     8
 SPECIES_NAMES = ["S", "MAPKKK", "MAPKKK*", "MAPKK", "MAPKK*", "MAPK", "MAPK*", "TF", "TF*"]
 N_SPECIES = len(SPECIES_NAMES)
 TIMESTEP = 1e-3
@@ -16,27 +17,27 @@ INITIAL_STATES = [
 
 
 ALL_BOGUS_CONVERSION_OPS = [
-    lambda result: result.add_double_conversion((1, 3), (1, 4)),
-    lambda result: result.add_double_conversion((1, 5), (1, 6)),
-    lambda result: result.add_double_conversion((1, 7), (1, 8)),
-    lambda result: result.add_double_conversion((2, 5), (2, 6)),
-    lambda result: result.add_double_conversion((2, 7), (2, 8)),
-    lambda result: result.add_double_conversion((3, 7), (3, 8)),
-    lambda result: result.add_double_conversion((4, 7), (4, 8)),
-    lambda result: result.add_double_conversion((5, 7), (5, 8)),  # wrong kinese partners
-    lambda result: result.add_double_conversion((3, 5), (3, 6)),  # wrong kinese partners
-    lambda result: result.add_double_conversion((1, 4), (1, 3)),
-    lambda result: result.add_double_conversion((1, 6), (1, 5)),
-    lambda result: result.add_double_conversion((1, 8), (1, 7)),
-    lambda result: result.add_double_conversion((2, 4), (2, 3)),  # backward deactivation
-    lambda result: result.add_double_conversion((2, 6), (2, 5)),  # backward deactivation
-    lambda result: result.add_double_conversion((2, 8), (2, 7)),  # backward deactivation
-    lambda result: result.add_double_conversion((3, 6), (3, 5)),
-    lambda result: result.add_double_conversion((3, 8), (3, 7)),
-    lambda result: result.add_double_conversion((4, 6), (4, 5)),  # backward deactivation
-    lambda result: result.add_double_conversion((4, 8), (4, 7)),  # backward deactivation
-    lambda result: result.add_double_conversion((5, 8), (5, 7)),
[…]

### 2018-12-05 readdy_learn/example/predator_prey.py

@@ -28,23 +28,23 @@ TIMESTEP = 2e-7
 def bfc():
     result = _basis.BasisFunctionConfiguration(len(SPECIES_NAMES))
 
-    result.add_fusion(0, 0, None)
-    result.add_fusion(1, 1, None)
-    result.add_fission(0, 0, 0)
-    result.add_double_conversion([0, 1], [1, 1])
-    result.add_decay(1)
+    result.add_fusion(0, 0, None)                 # 1:  X + X -> 0
+    result.add_fusion(1, 1, None)                 # 2:  Y + Y -> 0
+    result.add_fission(0, 0, 0)                   # 3:  X     -> X + X
+    result.add_double_conversion([0, 1], [1, 1])  # 4:  X + Y -> Y + Y
+    result.add_decay(1)                           # 5:  Y     -> 0
 
-    result.add_double_conversion([0, 1], [0, 0])
-    result.add_decay(0)
-    result.add_fusion(1, 1, 1)
-    result.add_fission(1, 1, 1)
-    result.add_fusion(0, 0, 0)
-    result.add_fusion(0, 1, 0)
-    result.add_fusion(0, 1, 1)
-    result.add_fission(0, 0, 1)
-    result.add_conversion(0, 1)
-    result.add_conversion(1, 0)
-    result.add_fission(0, 1, 1)
+    result.add_double_conversion([0, 1], [0, 0])  # 6:  X + Y -> X + X
+    result.add_decay(0)                           # 7:  X     -> 0
+    result.add_fusion(1, 1, 1)                    # 8:  Y + Y -> Y
+    result.add_fission(1, 1, 1)                   # 9:  Y     -> Y + Y
+    result.add_fusion(0, 0, 0)                    # 10: X + X -> X
+    result.add_fusion(0, 1, 0)                    # 11: X + Y -> X
+    result.add_fusion(0, 1, 1)                    # 12: X + Y -> Y
+    result.add_fission(0, 0, 1)                   # 13: X + X -> Y
+    result.add_conversion(0, 1)                   # 14: X     -> Y
+    result.add_conversion(1, 0)                   # 15: Y     -> X
+    result.add_fission(0, 1, 1)                   # 16: X     -> Y + Y
 
     return result
 

### 2018-12-05 docs/manuscript_jcp.tex
add figures with captions, add description to reaction tables

--- 2018-12-04 docs/manuscript_jcp.tex
@@
 \subsection{Application to MAPK cascade\label{sec:mapk}}
+\begin{figure}
+    \centering
+    \includegraphics[width=\singlecolumnwidth]{figures_jcp/6.pdf}
+    \caption{\label{fig:mapk}Application of reactive SINDy to the MAPK pathway system.
+\textbf{(a)}~The response curve of the MAPK cascade as a function of external stimulus given as a constant concentration $[\mathrm{S}]$.
+The activity is the long-time concentration of activated transcription factors $[\mathrm{TF*}]$.
+Dashed lines show the values of $[\mathrm{S}]$ at which concentration time series data was generated.
+\textbf{(b)}~Estimated rate coefficients of candidate reactions after application of reactive SINDy (regularized) to the time series data.
+LSQ estimation and the ground truth model for comparison.}
+\end{figure}
 
 \subsection{Application to Lotka-Volterra system\label{sec:lv}}
+\begin{figure}
+    \centering
+    \includegraphics[width=\singlecolumnwidth]{figures_jcp/7.pdf}
+    \caption{\label{fig:lv}Application of reactive SINDy to the Lotka-Volterra system with social friction.
+\textbf{(a)}~Concentration data as a function of time for predator and prey.
+\textbf{(b)}~Estimated rate coefficients of candidate reactions after application of reactive SINDy (regularized) to the time series data.
+LSQ estimation and the ground truth model for comparison.}
+\end{figure}
 
@@
         \midrule 
-        $\mathrm{DNA}_{\mathrm{A}}$  & $\rightharpoonup$  & $\mathrm{DNA}_{\mathrm{A}}+\mathrm{mRNA}_{\mathrm{A}}$  & $k_{1}=1.8$  & transcription of $\mathrm{mRNA}_{\mathrm{A}}$\tabularnewline
-        $\mathrm{mRNA}_{\mathrm{A}}+\mathrm{A}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{A}}$  & $k_{19}=0$  & artificial fusion\tabularnewline
-        $\mathrm{mRNA}_{\mathrm{B}}+\mathrm{B}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{B}}$  & $k_{20}=0$  & artificial fusion\tabularnewline
+        $\mathrm{S}        + \mathrm{MAPKKK}$  & $\rightharpoonup$  &  $\mathrm{S      } + \mathrm{MAPKKK*}$ & $k_{1} =1$ & external stimulus activates MAPKKK\tabularnewline
+        $\mathrm{MAPKKK*}    \mathrm{      }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK }                   $ & $k_{2} =1$ & dephosphorylation\tabularnewline
+        $\mathrm{MAPKKK*}  + \mathrm{MAPKK }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK*} + \mathrm{MAPKK* }$ & $k_{3} =1$ & phosphorylation of MAPKK\tabularnewline
+        $\mathrm{MAPKK*}     \mathrm{      }$  & $\rightharpoonup$  &  $\mathrm{MAPKK  }                   $ & $k_{4} =1$ & dephosphorylation\tabularnewline
+        $\mathrm{MAPKK*}   + \mathrm{MAPK  }$  & $\rightharpoonup$  &  $\mathrm{MAPKK* } + \mathrm{MAPK*  }$ & $k_{5} =1$ & phosphorylation of MAPK\tabularnewline
+        $\mathrm{MAPK*}      \mathrm{      }$  & $\rightharpoonup$  &  $\mathrm{MAPK   }                   $ & $k_{6} =1$ & dephosphorylation\tabularnewline
+        $\mathrm{MAPK*}    + \mathrm{TF    }$  & $\rightharpoonup$  &  $\mathrm{MAPK*  } + \mathrm{TF*    }$ & $k_{7} =1$ & phosphorylation of transcription factor\tabularnewline
+        $\mathrm{TF*}        \mathrm{      }$  & $\rightharpoonup$  &  $\mathrm{TF     }                   $ & $k_{8} =1$ & dephosphorylation\tabularnewline
+        $\mathrm{MAPKKK}   + \mathrm{MAPKK }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK } + \mathrm{MAPKK* }$ & $k_{9} =0$ & artificial reaction\tabularnewline
+        $\mathrm{MAPKKK}   + \mathrm{MAPK  }$  & $\rightharpoonup$  &  $\mathrm{MAPK*  }                   $ & $k_{10}=0$ & artificial reaction\tabularnewline
+        $\mathrm{MAPKKK}   + \mathrm{TF    }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK } + \mathrm{TF*    }$ & $k_{11}=0$ & artificial reaction\tabularnewline
+        $\mathrm{MAPKKK*}  + \mathrm{MAPK  }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK*} + \mathrm{MAPK*  }$ & $k_{12}=0$ & artificial reaction\tabularnewline
+        $\mathrm{MAPKKK*}  + \mathrm{TF    }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK*} + \mathrm{TF*    }$ & $k_{13}=0$ & artificial reaction\tabularnewline
+        $\mathrm{MAPKK}    + \mathrm{TF    }$  & $\rightharpoonup$  &  $\mathrm{MAPKK  } + \mathrm{TF*    }$ & $k_{14}=0$ & artificial reaction\tabularnewline
+        $\mathrm{MAPKK*}   + \mathrm{TF    }$  & $\rightharpoonup$  &  $\mathrm{MAPKK* } + \mathrm{TF*    }$ & $k_{15}=0$ & artificial reaction\tabularnewline
+        $\mathrm{MAPK}     + \mathrm{TF    }$  & $\rightharpoonup$  &  $\mathrm{MAPK   } + \mathrm{TF*    }$ & $k_{16}=0$ & artificial reaction\tabularnewline
+        $\mathrm{MAPKK}    + \mathrm{MAPK  }$  & $\rightharpoonup$  &  $\mathrm{MAPKK  } + \mathrm{MAPK*  }$ & $k_{17}=0$ & artificial reaction\tabularnewline
+        $\mathrm{MAPKKK}   + \mathrm{MAPKK*}$  & $\rightharpoonup$  &  $\mathrm{MAPKKK } + \mathrm{MAPKK  }$ & $k_{18}=0$ & artificial reaction\tabularnewline
+        $\mathrm{MAPKKK}   + \mathrm{MAPK* }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK } + \mathrm{MAPK   }$ & $k_{19}=0$ & artificial reaction\tabularnewline
+        $\mathrm{MAPKKK}   + \mathrm{TF*   }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK } + \mathrm{TF     }$ & $k_{20}=0$ & artificial reaction\tabularnewline
+        $\mathrm{MAPKKK*}  + \mathrm{MAPKK*}$  & $\rightharpoonup$  &  $\mathrm{MAPKKK*} + \mathrm{MAPKK  }$ & $k_{21}=0$ & artificial reaction\tabularnewline
+        $\mathrm{MAPKKK*}  + \mathrm{MAPK* }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK*} + \mathrm{MAPK   }$ & $k_{22}=0$ & artificial reaction\tabularnewline
+        $\mathrm{MAPKKK*}  + \mathrm{TF*   }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK*} + \mathrm{TF     }$ & $k_{23}=0$ & artificial reaction\tabularnewline
+        $\mathrm{MAPKK}    + \mathrm{MAPK* }$  & $\rightharpoonup$  &  $\mathrm{MAPKK  } + \mathrm{MAPK   }$ & $k_{24}=0$ & artificial reaction\tabularnewline
+        $\mathrm{MAPKK}    + \mathrm{TF*   }$  & $\rightharpoonup$  &  $\mathrm{MAPKK  } + \mathrm{TF     }$ & $k_{25}=0$ & artificial reaction\tabularnewline
+        $\mathrm{MAPKK*}   + \mathrm{MAPK* }$  & $\rightharpoonup$  &  $\mathrm{MAPKK* } + \mathrm{MAPK   }$ & $k_{26}=0$ & artificial reaction\tabularnewline
+        $\mathrm{MAPKK*}   + \mathrm{TF*   }$  & $\rightharpoonup$  &  $\mathrm{MAPKK* } + \mathrm{TF     }$ & $k_{27}=0$ & artificial reaction\tabularnewline
+        $\mathrm{MAPK}     + \mathrm{TF*   }$  & $\rightharpoonup$  &  $\mathrm{MAPK   } + \mathrm{TF     }$ & $k_{28}=0$ & artificial reaction\tabularnewline
+        $\mathrm{MAPK*}    + \mathrm{TF*   }$  & $\rightharpoonup$  &  $\mathrm{MAPK*  } + \mathrm{TF     }$ & $k_{29}=0$ & artificial reaction\tabularnewline
         \bottomrule
@@
         \midrule 
-        $\mathrm{DNA}_{\mathrm{A}}$  & $\rightharpoonup$  & $\mathrm{DNA}_{\mathrm{A}}+\mathrm{mRNA}_{\mathrm{A}}$  & $k_{1}=1.8$  & transcription of $\mathrm{mRNA}_{\mathrm{A}}$\tabularnewline
-        $\mathrm{mRNA}_{\mathrm{A}}+\mathrm{A}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{A}}$  & $k_{19}=0$  & artificial fusion\tabularnewline
-        $\mathrm{mRNA}_{\mathrm{B}}+\mathrm{B}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{B}}$  & $k_{20}=0$  & artificial fusion\tabularnewline
+        $\mathrm{X} + \mathrm{X}$ & $\rightharpoonup$ & $\emptyset             $  & $k_{1} =0.1$ & social friction of prey\tabularnewline
+        $\mathrm{Y} + \mathrm{Y}$ & $\rightharpoonup$ & $\emptyset             $  & $k_{2} =0.1$ & social friction of predator\tabularnewline
+        $\mathrm{X}             $ & $\rightharpoonup$ & $\mathrm{X} + \mathrm{X}$ & $k_{3} =1  $ & prey growth\tabularnewline
+        $\mathrm{X} + \mathrm{Y}$ & $\rightharpoonup$ & $\mathrm{Y} + \mathrm{Y}$ & $k_{4} =1  $ & predator eats prey\tabularnewline
+        $\mathrm{Y}             $ & $\rightharpoonup$ & $\emptyset             $  & $k_{5} =1  $ & predator decays\tabularnewline
+        $\mathrm{X} + \mathrm{Y}$ & $\rightharpoonup$ & $\mathrm{X} + \mathrm{X}$ & $k_{6} =0  $ & artificial reaction\tabularnewline
+        $\mathrm{X}             $ & $\rightharpoonup$ & $\emptyset             $  & $k_{7} =0  $ & artificial reaction\tabularnewline
+        $\mathrm{Y} + \mathrm{Y}$ & $\rightharpoonup$ & $\mathrm{Y}             $ & $k_{8} =0  $ & artificial reaction\tabularnewline
+        $\mathrm{Y}             $ & $\rightharpoonup$ & $\mathrm{Y} + \mathrm{Y}$ & $k_{9} =0  $ & artificial reaction\tabularnewline
+        $\mathrm{X} + \mathrm{X}$ & $\rightharpoonup$ & $\mathrm{X}             $ & $k_{10}=0  $ & artificial reaction\tabularnewline
+        $\mathrm{X} + \mathrm{Y}$ & $\rightharpoonup$ & $\mathrm{X}             $ & $k_{11}=0  $ & artificial reaction\tabularnewline
+        $\mathrm{X} + \mathrm{Y}$ & $\rightharpoonup$ & $\mathrm{Y}             $ & $k_{12}=0  $ & artificial reaction\tabularnewline
+        $\mathrm{X} + \mathrm{X}$ & $\rightharpoonup$ & $\mathrm{Y}             $ & $k_{13}=0  $ & artificial reaction\tabularnewline
+        $\mathrm{X}             $ & $\rightharpoonup$ & $\mathrm{Y}             $ & $k_{14}=0  $ & artificial reaction\tabularnewline
+        $\mathrm{Y}             $ & $\rightharpoonup$ & $\mathrm{X}             $ & $k_{15}=0  $ & artificial reaction\tabularnewline
+        $\mathrm{X}             $ & $\rightharpoonup$ & $\mathrm{Y} + \mathrm{Y}$ & $k_{16}=0  $ & artificial reaction\tabularnewline
         \bottomrule

### 2018-12-05 docs/makefigures/mapk_and_pprey.py

@@ -46,8 +46,8 @@ if __name__ == '__main__':
 
     def plot_mapk_activation_curve():
         plt.semilogx(stimulus, activities, label="MAPK cascade response")
-        plt.xlabel('Stimulus in a.u.')
-        plt.ylabel('Activity in a.u.')
+        plt.xlabel(r'Stimulus $[\mathrm{S}]$ in a.u.')
+        plt.ylabel(r'Activity $[\mathrm{TF*}]$ in a.u.')
         ymin, ymax = plt.ylim()
         plt.ylim(ymin, 0.35)
         plt.vlines(init_state_activities, *plt.ylim(), color="xkcd:red", linestyle="--",label="initial conditions", lw=1)

### 2018-12-05 docs/manuscript_jcp.tex
text for mapk and pprey wip

@@
 the correct reaction network and rates with decreasing levels of noise.
-The final scenario generalizes the method to two measurements with
+The third scenario generalizes the method to two measurements with
 different initial conditions, also converging to the correct model
 with decreasing levels of noise.
+
+We additionally demonstrate the algorithm
+on time series data of the mitogen activated protein kinases (MAPK) pathway as 
+an example for a bimodal system and on time series data of the
+Lotka-Volterra system which describes predator-prey dynamics subject to
+social friction. In both systems reactive SINDy recovers the generating 
+reaction network whereas non-sparse estimation detects many false processes.
 
@@
 \subsection{Application to MAPK cascade\label{sec:mapk}}
+The reactive SINDy method is applied to the mitogen activated protein 
+kinases (MAPK) pathway~\cite{Xing1996} which is an important regulatory mechanism of 
+biological cells to respond to stimuli and is involved in proliferation, differentiation, 
+inflammation, and apoptosis~\cite{Zhang2002mapk}. 
+Single-cell MAPK kinetics can be observed experimentally~\cite{Ryu2018integrated}.
+Mathematically MAPK kinetics are often modelled using reaction 
+rate equations~\cite{Kolch2005kinases,Orton2005computational} which 
+enables analysis using reactive SINDy.
+
+Generally a MAPK pathway consists of multiple stages of kinases that 
+are either inactive or active, denoted by ``*''. Their activation occurs 
+due to phosphorylation catalyzed by the upstream kinase of the previous 
+stage, dephosphorylation is catalyzed by phosphatases. When the kinase is 
+active it can activate other downstream kinases of the next stage. 
+The initial activation is often due to an external stimulus. 
+The response of the whole cascade is the amount of activated substrate 
+after the final stage, typically measured as a function of the initial stimulus.
+
+Here we model the MAPK pathway with three stages of kinases $\mathrm{MAPK}$, $\mathrm{MAPKK}$, and $\mathrm{MAPKKK}$.
+The initial stimulus is called $\mathrm{S}$ and the final substrate to be activated is a transcription factor $\mathrm{TF}$.
+The ground truth reaction network reads
+\begin{align*}
+\text{Activations:} &\\
+\mathrm{S}        + \mathrm{MAPKKK}  & \rightharpoonup \mathrm{S      } + \mathrm{MAPKKK*} \\
+\mathrm{MAPKKK*}  + \mathrm{MAPKK }  & \rightharpoonup \mathrm{MAPKKK*} + \mathrm{MAPKK* } \\
+\mathrm{MAPKK*}   + \mathrm{MAPK  }  & \rightharpoonup \mathrm{MAPKK* } + \mathrm{MAPK*  } \\
+\mathrm{MAPK*}    + \mathrm{TF    }  & \rightharpoonup \mathrm{MAPK*  } + \mathrm{TF*    }  \\
+\text{Deactivations:}&\\
+\mathrm{MAPKKK*}    \mathrm{      }  & \rightharpoonup \mathrm{MAPKKK }  \\
+\mathrm{MAPKK*}     \mathrm{      }  & \rightharpoonup \mathrm{MAPKK  }  \\
+\mathrm{MAPK*}      \mathrm{      }  & \rightharpoonup \mathrm{MAPK   }  \\
+\mathrm{TF*}        \mathrm{      }  & \rightharpoonup \mathrm{TF     }.
+\end{align*}
+For simplicity we assume phosphatase to be abundant such that deactivations 
+effectively become first order reactions.
+The external stimulus $\mathrm{S}$ is not consumed such that time integration of these 
+reactions yields a steady state in which the concentration of response $[\mathrm{TF}]$ 
+can be measured as a function of the stimulus concentration $[\mathrm{S}]$.
+Using the rate constants given in Tab.~\ref{tab:reaction-library-mapk} we obtain the 
+response curve given in Fig.~\ref{fig:mapk}a.
+
+We generate concentration time series data of the MAPK reactions above at 
+three different initial conditions, each differing in the amount of stimulus.
+They are marked in Fig.~\ref{fig:mapk}a by vertical dashed lines.
+These time series are concatenated to yield a dataset of 300 frames in total.
+We use the library $\Theta$ of candidate reactions Tab.~\ref{tab:reaction-library-mapk}.
+To determine the hyperparameter $\alpha=6.6\times 10^{-9}$ we shuffle the data and perform
+15-fold cross validation. We obtain the estimated 
+rate constants by solving the minimization problem (\ref{method:minimizationproblem}) 
+with $\lambda=1$. The results are given in Fig.~\ref{fig:mapk}b.
+{\color{red}TODO describe}
+
 \begin{figure}
@@
 \textbf{(a)}~The response curve of the MAPK cascade as a function of external stimulus given as a constant concentration $[\mathrm{S}]$.
-The activity is the long-time concentration of activated transcription factors $[\mathrm{TF*}]$.
+The activity is the steady state concentration of activated transcription factors $[\mathrm{TF*}]$.
 Dashed lines show the values of $[\mathrm{S}]$ at which concentration time series data was generated.
 \textbf{(b)}~Estimated rate coefficients of candidate reactions after application of reactive SINDy (regularized) to the time series data.
-LSQ estimation and the ground truth model for comparison.}
+Least-squares estimation (LSQ) and the ground truth model for comparison.}
 \end{figure}
@@
 \subsection{Application to Lotka-Volterra system\label{sec:lv}}
+As biological pathways often exhibit oscillatory behavior~\cite{Iqbal2008tnf} which can stem 
+from positive or negative feedback loops~\cite{Shin2009positive} we apply reactive SINDy to
+an idealized oscillatory system, namely the Lotka-Volterra system. 
+The predator-prey dynamics of two species $\mathrm{A}$ (prey) and $\mathrm{B}$ 
+(predator) is defined by the reactions
+\begin{align*}
+\mathrm{A}              & \rightharpoonup \mathrm{A} + \mathrm{A} &  & \text{(prey growth)},\\
+\mathrm{A} + \mathrm{B} & \rightharpoonup \mathrm{B} + \mathrm{B} &  & \text{(predator eats prey)},\\
+\mathrm{B}              & \rightharpoonup \emptyset &  & \text{(predator decay)},\\
+\mathrm{A} + \mathrm{A} & \rightharpoonup \emptyset &  & \text{(prey friction)},\\
+\mathrm{B} + \mathrm{B} & \rightharpoonup \emptyset &  & \text{(predator friction)}.
+\end{align*}
+From this model we generate concentration time series data with $200$ frames 
+which is displayed in Fig.~\ref{fig:lv}a. We construct the library $\Theta$ 
+of candidate reactions given in Tab.~\ref{tab:reaction-library-lv}. 
+We determine the hyperparameter $\alpha=2.7\times10^{-7}$ by shuffling 
+the data and performing 5-fold cross validation. We obtain the estimated 
+rate constants by solving the minimization problem (\ref{method:minimizationproblem}) 
+with $\lambda=1$. The results are given in Fig.~\ref{fig:lv}b.
+{\color{red}TODO describe}
+
 \begin{figure}
@@
     \caption{\label{fig:lv}Application of reactive SINDy to the Lotka-Volterra system with social friction.
-\textbf{(a)}~Concentration data as a function of time for predator and prey.
+\textbf{(a)}~Concentration data as a function of time for predator and prey species.
 \textbf{(b)}~Estimated rate coefficients of candidate reactions after application of reactive SINDy (regularized) to the time series data.
-LSQ estimation and the ground truth model for comparison.}
+Least-squares estimation (LSQ) estimation and the ground truth model for comparison.}
 \end{figure}
@@
         \midrule 
-        $\mathrm{X} + \mathrm{X}$ & $\rightharpoonup$ & $\emptyset             $  & $k_{1} =0.1$ & social friction of prey\tabularnewline
-        $\mathrm{Y} + \mathrm{Y}$ & $\rightharpoonup$ & $\emptyset             $  & $k_{2} =0.1$ & social friction of predator\tabularnewline
-        $\mathrm{X}             $ & $\rightharpoonup$ & $\mathrm{X} + \mathrm{X}$ & $k_{3} =1  $ & prey growth\tabularnewline
-        $\mathrm{X} + \mathrm{Y}$ & $\rightharpoonup$ & $\mathrm{Y} + \mathrm{Y}$ & $k_{4} =1  $ & predator eats prey\tabularnewline
-        $\mathrm{Y}             $ & $\rightharpoonup$ & $\emptyset             $  & $k_{5} =1  $ & predator decays\tabularnewline
-        $\mathrm{X} + \mathrm{Y}$ & $\rightharpoonup$ & $\mathrm{X} + \mathrm{X}$ & $k_{6} =0  $ & artificial reaction\tabularnewline
-        $\mathrm{X}             $ & $\rightharpoonup$ & $\emptyset             $  & $k_{7} =0  $ & artificial reaction\tabularnewline
-        $\mathrm{Y} + \mathrm{Y}$ & $\rightharpoonup$ & $\mathrm{Y}             $ & $k_{8} =0  $ & artificial reaction\tabularnewline
-        $\mathrm{Y}             $ & $\rightharpoonup$ & $\mathrm{Y} + \mathrm{Y}$ & $k_{9} =0  $ & artificial reaction\tabularnewline
-        $\mathrm{X} + \mathrm{X}$ & $\rightharpoonup$ & $\mathrm{X}             $ & $k_{10}=0  $ & artificial reaction\tabularnewline
-        $\mathrm{X} + \mathrm{Y}$ & $\rightharpoonup$ & $\mathrm{X}             $ & $k_{11}=0  $ & artificial reaction\tabularnewline
-        $\mathrm{X} + \mathrm{Y}$ & $\rightharpoonup$ & $\mathrm{Y}             $ & $k_{12}=0  $ & artificial reaction\tabularnewline
-        $\mathrm{X} + \mathrm{X}$ & $\rightharpoonup$ & $\mathrm{Y}             $ & $k_{13}=0  $ & artificial reaction\tabularnewline
-        $\mathrm{X}             $ & $\rightharpoonup$ & $\mathrm{Y}             $ & $k_{14}=0  $ & artificial reaction\tabularnewline
-        $\mathrm{Y}             $ & $\rightharpoonup$ & $\mathrm{X}             $ & $k_{15}=0  $ & artificial reaction\tabularnewline
-        $\mathrm{X}             $ & $\rightharpoonup$ & $\mathrm{Y} + \mathrm{Y}$ & $k_{16}=0  $ & artificial reaction\tabularnewline
+        $\mathrm{A} + \mathrm{A}$ & $\rightharpoonup$ & $\emptyset             $  & $k_{1} =0.1$ & social friction of prey\tabularnewline
+        $\mathrm{B} + \mathrm{B}$ & $\rightharpoonup$ & $\emptyset             $  & $k_{2} =0.1$ & social friction of predator\tabularnewline
+        $\mathrm{A}             $ & $\rightharpoonup$ & $\mathrm{A} + \mathrm{A}$ & $k_{3} =1  $ & prey growth\tabularnewline
+        $\mathrm{A} + \mathrm{B}$ & $\rightharpoonup$ & $\mathrm{B} + \mathrm{B}$ & $k_{4} =1  $ & predator eats prey\tabularnewline
+        $\mathrm{B}             $ & $\rightharpoonup$ & $\emptyset             $  & $k_{5} =1  $ & predator decays\tabularnewline
+        $\mathrm{A} + \mathrm{B}$ & $\rightharpoonup$ & $\mathrm{A} + \mathrm{A}$ & $k_{6} =0  $ & artificial reaction\tabularnewline
+        $\mathrm{A}             $ & $\rightharpoonup$ & $\emptyset             $  & $k_{7} =0  $ & artificial reaction\tabularnewline
+        $\mathrm{B} + \mathrm{B}$ & $\rightharpoonup$ & $\mathrm{B}             $ & $k_{8} =0  $ & artificial reaction\tabularnewline
+        $\mathrm{B}             $ & $\rightharpoonup$ & $\mathrm{B} + \mathrm{B}$ & $k_{9} =0  $ & artificial reaction\tabularnewline
+        $\mathrm{A} + \mathrm{A}$ & $\rightharpoonup$ & $\mathrm{A}             $ & $k_{10}=0  $ & artificial reaction\tabularnewline
+        $\mathrm{A} + \mathrm{B}$ & $\rightharpoonup$ & $\mathrm{A}             $ & $k_{11}=0  $ & artificial reaction\tabularnewline
+        $\mathrm{A} + \mathrm{B}$ & $\rightharpoonup$ & $\mathrm{B}             $ & $k_{12}=0  $ & artificial reaction\tabularnewline
+        $\mathrm{A} + \mathrm{A}$ & $\rightharpoonup$ & $\mathrm{B}             $ & $k_{13}=0  $ & artificial reaction\tabularnewline
+        $\mathrm{A}             $ & $\rightharpoonup$ & $\mathrm{B}             $ & $k_{14}=0  $ & artificial reaction\tabularnewline
+        $\mathrm{B}             $ & $\rightharpoonup$ & $\mathrm{A}             $ & $k_{15}=0  $ & artificial reaction\tabularnewline
+        $\mathrm{A}             $ & $\rightharpoonup$ & $\mathrm{B} + \mathrm{B}$ & $k_{16}=0  $ & artificial reaction\tabularnewline
         \bottomrule

### 2018-12-05 docs/makefigures/mapk_and_pprey.py

@@ -34,6 +34,7 @@ if __name__ == '__main__':
     sel = np.argmin(-scores_cv)
     plt.semilogx(alphas_cv, -scores_cv)
     #plt.plot([alphas_cv[sel]], [-scores_cv[sel]], 'x', markersize=20)
+    print(f"Hyperparameter alpha {alphas_cv[sel]}")
     plt.ylim(0., 0.5e-9)
     xmin, xmax = plt.xlim()
     plt.xlim(xmin, 2e-8)
@@ -122,8 +123,8 @@ if __name__ == '__main__':
     def plot_pprey_traj():
         plt.plot(time, counts[:, 0], label='prey')
         plt.plot(time, counts[:, 1], label='predator')
-        plt.xlabel("Time")
-        plt.ylabel("Concentration")
+        plt.xlabel("Time in a.u.")
+        plt.ylabel("Concentration in a.u.")
         plt.legend()
 
     def plot_pprey_models():
@@ -166,6 +167,7 @@ if __name__ == '__main__':
     sel = np.argmin(-scores_cv)
     plt.semilogx(alphas_cv, -scores_cv)
     #plt.plot([alphas_cv[sel]], [-scores_cv[sel]], 'x', markersize=20)
+    print(f"Hyperparameter alpha {alphas_cv[sel]}")
     plt.ylim(6.45e-7, 6.7e-7)
     xmin, xmax = plt.xlim()
     plt.xlim(1e-9, 1e-5)

### 2018-12-05 docs/makefigures/mapk_and_pprey.py
wip

@@ -57,19 +57,28 @@ if __name__ == '__main__':
 
 
     def plot_mapk_models():
+        from matplotlib.ticker import MultipleLocator, FormatStrFormatter
+
+        majorLocator = MultipleLocator(3)
+        minorLocator = MultipleLocator(1)
+        x_values = [r"$\theta_{%s}$" % (i+1) for i in range(len(rates_lsq))]
         plt.plot(rates_cv, 'o', label="regularized")
         plt.plot(rates_lsq, 'd', label="LSQ")
         # plt.vlines([7.5], 0, 1, 'grey', 'dashed')
         plt.plot(rates_desired, 'x', label="ground truth")
-        labels = list(map(str, range(1, len(rates_desired) + 1)))
-        for i in range(len(labels)):
-            if i % 3 == 0:
-                pass
-            else:
-                labels[i] = ""
-        plt.xticks(ticks=range(len(rates_desired)), labels=labels, fontsize=11)
+        #plt.xticks([0, 5, 10], [r'$a^2$', r'$b^2$', r'$c^2$'])
+        labels = [x for (i,x) in enumerate(x_values) if i % 3 == 0]  # list(map(str, range(1, len(rates_desired) + 1)))
+        #for i in range(len(labels)):
+        #    if i % 3 == 0:
+        #        labels[i] = x_values[i]
+        #    else:
+        #        labels[i] = ""
+        plt.gca().xaxis.set_major_locator(majorLocator)
+        plt.gca().xaxis.set_minor_locator(minorLocator)
+        plt.xticks(ticks=[i*3 for i in range(len(labels))], labels=labels, fontsize=11)
+        plt.xlim([-.5, None])
         plt.legend()
-        plt.xlabel(r"Ansatz reaction \#")
+        plt.xlabel(r"Ansatz reaction $\theta_r$")
         plt.ylabel("Rate constant in a.u.")
 
     # Combined figure of activation and results

### 2018-12-06 docs/manuscript_jcp.tex
wip
describe results of mapk and pprey, update figure 7

@@
 an example for a bimodal system and on time series data of the
-Lotka-Volterra system which describes predator-prey dynamics subject to
+Lotka-Volterra system which describes oscillatory predator-prey dynamics subject to
 social friction. In both systems reactive SINDy recovers the generating 
@@
 \begin{align*}
-\text{Activations:} &\\
+\intertext{Activations:}
 \mathrm{S}        + \mathrm{MAPKKK}  & \rightharpoonup \mathrm{S      } + \mathrm{MAPKKK*} \\
@@
 \mathrm{MAPKK*}   + \mathrm{MAPK  }  & \rightharpoonup \mathrm{MAPKK* } + \mathrm{MAPK*  } \\
-\mathrm{MAPK*}    + \mathrm{TF    }  & \rightharpoonup \mathrm{MAPK*  } + \mathrm{TF*    }  \\
-\text{Deactivations:}&\\
+\mathrm{MAPK*}    + \mathrm{TF    }  & \rightharpoonup \mathrm{MAPK*  } + \mathrm{TF*    }
+\intertext{Deactivations:}
 \mathrm{MAPKKK*}    \mathrm{      }  & \rightharpoonup \mathrm{MAPKKK }  \\
@@
 The external stimulus $\mathrm{S}$ is not consumed such that time integration of these 
-reactions yields a steady state in which the concentration of response $[\mathrm{TF}]$ 
+reactions yields a steady state in which the response, i.e. the concentration $[\mathrm{TF*}]$ 
 can be measured as a function of the stimulus concentration $[\mathrm{S}]$.
@@
 We generate concentration time series data of the MAPK reactions above at 
-three different initial conditions, each differing in the amount of stimulus.
+three different initial conditions, each differing in the amount of stimulus $[\mathrm{S}]$.
 They are marked in Fig.~\ref{fig:mapk}a by vertical dashed lines.
@@
 with $\lambda=1$. The results are given in Fig.~\ref{fig:mapk}b.
-{\color{red}TODO describe}
+Least-squares estimation detects 5 of the 8 reaction processes that belong to the 
+ground truth model. However it also detects 12 other reaction 
+processes ($\theta_{18}$ - $\theta_{29}$) which are not in the ground truth.
+Reactive SINDy estimation detects all reactions of the ground truth, 
+two processes ($\theta_{4}$ and $\theta_{8}$) show deviations in 
+rate constants. Generally reactive SINDy yields a sparse model which allows further
+simplification of the reaction network by dropping out reaction processes that lie 
+beneath a certain cutoff. In this case for example a cutoff of $0.25$ would directly recover
+the ground truth reaction network.
+Quantitatively, one may consider the $L_1$ norm of the relative distance 
+of estimated rate constants $\hat{\xi}_r$ to the non-zero rate 
+constants of the ground truth $\xi_r$
+\begin{equation*}
+\sum_{r=1}^{8} \left|(\hat{\xi}_{r} - \xi_r)/\xi_r\right|
+\end{equation*}
+which yields $167\%$ error for least-squares and $21\%$ error for reactive SINDy.
 
@@
 Dashed lines show the values of $[\mathrm{S}]$ at which concentration time series data was generated.
-\textbf{(b)}~Estimated rate coefficients of candidate reactions after application of reactive SINDy (regularized) to the time series data.
+\textbf{(b)}~Estimated rate coefficients of candidate reactions (see Tab.~\ref{tab:reaction-library-mapk}) after application of reactive SINDy (regularized) to the time series data.
 Least-squares estimation (LSQ) and the ground truth model for comparison.}
@@
 with $\lambda=1$. The results are given in Fig.~\ref{fig:lv}b.
-{\color{red}TODO describe}
+Least-squares estimation detects all reactions of the ground truth model but also
+two false processes ($\theta_{6}$ and $\theta_{7}$) with a higher rate than
+the two first true processes ($\theta_{1}$ and $\theta_{2}$). Reactive SINDy detects
+the true reaction network with minor deviations in rate constants. As in 
+Sec.~\ref{sec:mapk} we consider the $L_1$ norm of the relative distance to the ground truth
+for non-zero rate constants
+\begin{equation*}
+\sum_{r=1}^{5} \left|(\hat{\xi}_{r} - \xi_r)/\xi_r\right|
+\end{equation*}
+which yields $75\%$ error for least-squares and $7\%$ error for reactive SINDy.
 
@@
 \textbf{(a)}~Concentration data as a function of time for predator and prey species.
-\textbf{(b)}~Estimated rate coefficients of candidate reactions after application of reactive SINDy (regularized) to the time series data.
+\textbf{(b)}~Estimated rate coefficients of candidate reactions (see Tab.~\ref{tab:reaction-library-lv}) after application of reactive SINDy (regularized) to the time series data.
 Least-squares estimation (LSQ) estimation and the ground truth model for comparison.}
@@
 We have applied this method on data generated from a gene regulation
-network and could successfully recover the underlying model.
+network, a MAPK pathway, and a Lotka-Volterra system and 
+could successfully recover the underlying reaction networks.
 
@@
 the ground truth.
-
 Potentially, this method could be used to identify reaction networks
@@
 always exactly identical, as was demonstrated in Sec.~\ref{sec:case-3}.
+
 One apparent limitation is that the method can only be applied if
@@
 approach has derivatives equal zero in the equilibrium, which precludes
-the reaction dynamics to be recovered. In future work, we will consider
+the reaction dynamics to be recovered. Thus, in the case of oscillatory 
+systems the reaction network can be recovered robustly.
+
+In future work, we will consider
 the identification of reaction schemes from instantaneous fluctuations

### 2018-12-06 docs/makefigures/mapk_and_pprey.py

@@ -137,19 +137,27 @@ if __name__ == '__main__':
         plt.legend()
 
     def plot_pprey_models():
+        from matplotlib.ticker import MultipleLocator, FormatStrFormatter
+
+        majorLocator = MultipleLocator(3)
+        minorLocator = MultipleLocator(1)
+        x_values = [r"$\theta_{%s}$" % (i + 1) for i in range(len(rates_lsq))]
         plt.plot(rates_cv, 'o', label="regularized")
         plt.plot(rates_lsq, 'd', label="LSQ")
         # plt.vlines([7.5], 0, 1, 'grey', 'dashed')
         plt.plot(rates_desired, 'x', label="ground truth")
-        labels = list(map(str, range(1, len(rates_desired) + 1)))
-        for i in range(len(labels)):
-            if i % 2 == 0:
-                pass
-            else:
-                labels[i] = ""
-        plt.xticks(ticks=range(len(rates_desired)), labels=labels, fontsize=11)
+        # plt.xticks([0, 5, 10], [r'$a^2$', r'$b^2$', r'$c^2$'])
+        labels = [x for (i, x) in enumerate(x_values) if i % 3 == 0]  # list(map(str, range(1, len(rates_desired) + 1)))
+        # for i in range(len(labels)):
+        #    if i % 3 == 0:
+        #        labels[i] = x_values[i]
+        #    else:
+        #        labels[i] = ""
+        plt.gca().xaxis.set_major_locator(majorLocator)
+        plt.gca().xaxis.set_minor_locator(minorLocator)
+        plt.xticks(ticks=[i * 3 for i in range(len(labels))], labels=labels, fontsize=11)
         plt.legend()
-        plt.xlabel(r"Ansatz reaction \#")
+        plt.xlabel(r"Ansatz reaction $\theta_r$")
         plt.ylabel("Rate constant in a.u.")
 
     # Combined figure of traj and result

### 2018-12-07 docs/manuscript_jcp.tex
comment out intertext (check if it breaks latexdiff)
break align in two separate ones, such that latexdiff works
wip

@@
 \begin{abstract}
-The inner workings of a biological cell or a chemical reaction can
+The inner workings of a biological cell or a chemical reactor can
 be rationalized by the network of reactions, whose structure reveals
@@
 for the change of concentration of reactant $s$ as a result of order-0
-reactions (creation or annihilation), order-1 reactions (transitions
-of other species into $s$ or transitions of $s$ into other species),
+reactions (creation), order-1 reactions (transitions
+of other species into $s$, transitions of $s$ into other species, or annihilation),
 order-2 reactions (production or consumption of $s$ by the encounter
@@
 The initial stimulus is called $\mathrm{S}$ and the final substrate to be activated is a transcription factor $\mathrm{TF}$.
-The ground truth reaction network reads
+The ground truth reaction network consists of activation/phosphorylation reactions
 \begin{align*}
-\intertext{Activations:}
 \mathrm{S}        + \mathrm{MAPKKK}  & \rightharpoonup \mathrm{S      } + \mathrm{MAPKKK*} \\
@@
 \mathrm{MAPKK*}   + \mathrm{MAPK  }  & \rightharpoonup \mathrm{MAPKK* } + \mathrm{MAPK*  } \\
-\mathrm{MAPK*}    + \mathrm{TF    }  & \rightharpoonup \mathrm{MAPK*  } + \mathrm{TF*    }
-\intertext{Deactivations:}
-\mathrm{MAPKKK*}    \mathrm{      }  & \rightharpoonup \mathrm{MAPKKK }  \\
-\mathrm{MAPKK*}     \mathrm{      }  & \rightharpoonup \mathrm{MAPKK  }  \\
-\mathrm{MAPK*}      \mathrm{      }  & \rightharpoonup \mathrm{MAPK   }  \\
-\mathrm{TF*}        \mathrm{      }  & \rightharpoonup \mathrm{TF     }.
+\mathrm{MAPK*}    + \mathrm{TF    }  & \rightharpoonup \mathrm{MAPK*  } + \mathrm{TF*    } \\
+\end{align*}
+and deactivation/dephosphorylation reactions
+\begin{align*}
+\mathrm{MAPKKK*}                     & \rightharpoonup \mathrm{MAPKKK }  \\
+\mathrm{MAPKK*}                      & \rightharpoonup \mathrm{MAPKK  }  \\
+\mathrm{MAPK*}                       & \rightharpoonup \mathrm{MAPK   }  \\
+\mathrm{TF*}                         & \rightharpoonup \mathrm{TF     }.
 \end{align*}

### 2018-12-08 docs/manuscript_jcp.tex
wip

@@
 
-Here we model the MAPK pathway with three stages of kinases $\mathrm{MAPK}$, $\mathrm{MAPKK}$, and $\mathrm{MAPKKK}$.
+Here the MAPK pathway is modeled with three stages of kinases $\mathrm{MAPK}$, $\mathrm{MAPKK}$, and $\mathrm{MAPKKK}$.
 The initial stimulus is called $\mathrm{S}$ and the final substrate to be activated is a transcription factor $\mathrm{TF}$.
@@
 The external stimulus $\mathrm{S}$ is not consumed such that time integration of these 
-reactions yields a steady state in which the response, i.e. the concentration $[\mathrm{TF*}]$ 
+reactions yields a steady state in which the response, i.e., the concentration $[\mathrm{TF*}]$ 
 can be measured as a function of the stimulus concentration $[\mathrm{S}]$.
@@
 three different initial conditions, each differing in the amount of stimulus $[\mathrm{S}]$.
-They are marked in Fig.~\ref{fig:mapk}a by vertical dashed lines.
-These time series are concatenated to yield a dataset of 300 frames in total.
-We use the library $\Theta$ of candidate reactions Tab.~\ref{tab:reaction-library-mapk}.
-To determine the hyperparameter $\alpha=6.6\times 10^{-9}$ we shuffle the data and perform
-15-fold cross validation. We obtain the estimated 
-rate constants by solving the minimization problem (\ref{method:minimizationproblem}) 
+The response yielded by the chosen initial conditions is marked in Fig.~\ref{fig:mapk}a by vertical dashed lines.
+The concatenated time series is a dataset of 300 frames in total.
+We use the library $\Theta$ of ansatz reactions Tab.~\ref{tab:reaction-library-mapk}.
+The hyperparameter $\alpha=6.6\times 10^{-9}$ was determined by shuffling the data and performing
+15-fold cross validation. The estimated 
+rate constants were obtained by solving the minimization problem (\ref{method:minimizationproblem}) 
 with $\lambda=1$. The results are given in Fig.~\ref{fig:mapk}b.
 Least-squares estimation detects 5 of the 8 reaction processes that belong to the 
-ground truth model. However it also detects 12 other reaction 
-processes ($\theta_{18}$ - $\theta_{29}$) which are not in the ground truth.
+ground truth model. However it also detects 12 spurious reaction 
+processes ($\theta_{18}$ - $\theta_{29}$).
 Reactive SINDy estimation detects all reactions of the ground truth, 
@@
 simplification of the reaction network by dropping out reaction processes that lie 
-beneath a certain cutoff. In this case for example a cutoff of $0.25$ would directly recover
+beneath a certain cutoff. In this case for example a cutoff of $\kappa=0.25$ would directly recover
 the ground truth reaction network.
@@
 
-\subsection{Application to Lotka-Volterra system\label{sec:lv}}
+\subsection{Application to Lotka--Volterra system\label{sec:lv}}
 As biological pathways often exhibit oscillatory behavior~\cite{Iqbal2008tnf} which can stem 
 from positive or negative feedback loops~\cite{Shin2009positive} we apply reactive SINDy to
-an idealized oscillatory system, namely the Lotka-Volterra system. 
+an idealized oscillatory system, namely the Lotka--Volterra system. 
 The predator-prey dynamics of two species $\mathrm{A}$ (prey) and $\mathrm{B}$ 
-(predator) is defined by the reactions
+(predator) is defined by the reaction network
 \begin{align*}

### 2018-12-08 docs/manuscript_jcp.tex
wip

@@
 an example for a bimodal system and on time series data of the
-Lotka-Volterra system which describes oscillatory predator-prey dynamics subject to
+Lotka--Volterra system which describes oscillatory predator-prey dynamics subject to
 social friction. In both systems reactive SINDy recovers the generating 
@@
 \end{align*}
-From this model we generate concentration time series data with $200$ frames 
-which is displayed in Fig.~\ref{fig:lv}a. We construct the library $\Theta$ 
-of candidate reactions given in Tab.~\ref{tab:reaction-library-lv}. 
-We determine the hyperparameter $\alpha=2.7\times10^{-7}$ by shuffling 
-the data and performing 5-fold cross validation. We obtain the estimated 
-rate constants by solving the minimization problem (\ref{method:minimizationproblem}) 
-with $\lambda=1$. The results are given in Fig.~\ref{fig:lv}b.
+From this model we generated concentration time series data with $200$ frames 
+which is displayed in Fig.~\ref{fig:lv}a. The library 
+of ansatz reactions $\Theta$ is given in Tab.~\ref{tab:reaction-library-lv}. 
+The hyperparameter $\alpha=2.7\times10^{-7}$ was determined by shuffling 
+the data and performing 5-fold cross validation. The estimated 
+rate constants were obtained by solving the minimization problem (\ref{method:minimizationproblem}) 
+with $\lambda=1$. The results are depicted in Fig.~\ref{fig:lv}b.
 Least-squares estimation detects all reactions of the ground truth model but also
-two false processes ($\theta_{6}$ and $\theta_{7}$) with a higher rate than
-the two first true processes ($\theta_{1}$ and $\theta_{2}$). Reactive SINDy detects
+two spurious processes ($\theta_{6}$ and $\theta_{7}$) with a higher rate than
+the first two underlying processes ($\theta_{1}$ and $\theta_{2}$). Reactive SINDy recovers
 the true reaction network with minor deviations in rate constants. As in 
-Sec.~\ref{sec:mapk} we consider the $L_1$ norm of the relative distance to the ground truth
+Sec.~\ref{sec:mapk}, considering the $L_1$ norm of the relative distance to the ground truth
 for non-zero rate constants
@@
 \end{equation*}
-which yields $75\%$ error for least-squares and $7\%$ error for reactive SINDy.
+yields $75\%$ error for least-squares and $7\%$ error for reactive SINDy.
 
@@
     \includegraphics[width=\singlecolumnwidth]{figures_jcp/7.pdf}
-    \caption{\label{fig:lv}Application of reactive SINDy to the Lotka-Volterra system with social friction.
+    \caption{\label{fig:lv}Application of reactive SINDy to the Lotka--Volterra system with social friction.
 \textbf{(a)}~Concentration data as a function of time for predator and prey species.
@@
 We have applied this method on data generated from a gene regulation
-network, a MAPK pathway, and a Lotka-Volterra system and 
+network, a MAPK pathway, and a Lotka--Volterra system and 
 could successfully recover the underlying reaction networks.
@@
         \bottomrule
-    \end{tabular}\caption{\label{tab:reaction-library-lv}Full set of ansatz reactions $\Theta$ used in Sec.~\ref{sec:lv} for the Lotka-Volterra system.
+    \end{tabular}\caption{\label{tab:reaction-library-lv}Full set of ansatz reactions $\Theta$ used in Sec.~\ref{sec:lv} for the Lotka--Volterra system.
 The given rate constants define the ground truth reaction model.}

### 2018-12-10 docs/revision_reply/revision_reply.tex
wip

\documentclass[11pt,english]{article}
\usepackage[T1]{fontenc}
\usepackage{microtype}
\usepackage[utf8]{inputenc}
\usepackage{babel}
\usepackage{hyperref}
\setlength\parindent{0pt}

\usepackage{url}
\def\UrlFont{\fontfamily{lmtt}\selectfont \footnotesize}

\title{}

\begin{document}
\pagenumbering{gobble}

\textbf{{\Large Response to the reviewers}}\newline

\textbf{{\large Reviewer \# 2}}\newline

Thank you for the detailed review. We have addressed the comments in the
following way\newline

\textbf{Original comment:}
{This paper describes the modification of an approach called SINDY that first appeared in PNAS in 2015. It is used to estimate parameters in a system of chemical reactions in the presence or absence of noise. The method is applied to a rather artificial gene regulatory models in both a deterministic and stochastic setting. I have the following minor comments and one major comment.\\
1. The authors refer to ``ansatz'' functions. I know what ansatz means in German but what is meant here? - please clarify.}\newline

\textbf{Answer:}
The meaning of the word ``ansatz'' is now clarified when it first occurs (page 1).\newline

\textbf{Original comment:}
{2. Bottom of page 3, column 1. The authors mention a suitable choice of hyper parameters. But why is alpha =$1.91\times 10^{-7}$ suitable? And at this stage I still do not know what Kappa represents.\\
3. Page 3, line -5, Col. 2. There is a line that makes no sense.}\newline

\textbf{Answer:}
There was a formatting error which caused certain pieces of
text to disappear and randomly re-appear at a different place in the
document. We hope that now the choice of alpha (cross validation) and
the meaning of kappa (cutoff) are clarified and the line that made no
sense is put back into context.\newline

\textbf{Original comment:}
{My major comment is that the test problem is artificial and boring, with very little interesting dynamics. I would like to see the approach work on a problem that has some bimodality, for example. Furthermore, I would also like to see the method applied to a realistic biological test problem such as the MAPK cascade. Otherwise the impact of the work is quite low. This is not negotiable in terms of acceptance.}\newline

\textbf{Answer:}
We now applied the method to a MAPK cascade as well as a predator-prey model that contains
temporal bimodality and show that we can recover the correct reaction
network with the correct regularization hyperparameters.

\end{document}

### 2018-12-11 docs/resubmission_jcp/cover_letter/resubmission_cover_letter.tex
resubmission preparations

--- 2018-12-10 docs/revision_reply/revision_reply.tex
@@
-\documentclass[11pt,english]{article}
+\documentclass[11pt,english]{scrartcl}
 \usepackage[T1]{fontenc}
@@
 
-\title{}
-
 \begin{document}
@@
 
-\textbf{{\Large Response to the reviewers}}\newline
+\begin{flushleft}
+    \large
+    \textbf{Re-submission of manuscript:} Reactive SINDy: Discovering governing reactions from concentration data
+\end{flushleft}
+\begin{flushright}
+    \today
+\end{flushright}
+~\\
 
-\textbf{{\large Reviewer \# 2}}\newline
+Dear editorial board members,\\
+
+please find attached our revised manuscript in tex format and another compiled pdf document highlighting the changes. Further we give the response to the reviewers below.\\
+
+\textbf{{\large Response to Reviewer \#1}}\newline
+
+Thank you for the review.\newline
+
+\textbf{{\large Response to Reviewer \#2}}\newline
 
@@
 temporal bimodality and show that we can recover the correct reaction
-network with the correct regularization hyperparameters.
+network with the correct regularization hyperparameters.\newline
+
+Best regards,\\
+
+Moritz Hoffmann, Christoph Fröhner, and Frank Noé
 

### 2018-12-11 docs/resubmission_jcp/manuscript.tex
resubmission preparations

--- 2018-12-08 docs/manuscript_jcp.tex
@@
 social friction. In both systems reactive SINDy recovers the generating 
-reaction network whereas non-sparse estimation detects many false processes.
+reaction network whereas non-sparse estimation detects many spurious processes.
 
@@
 \begin{figure}[h!]
-\centering \includegraphics[width=\singlecolumnwidth]{figures_jcp/1.pdf}
+\centering \includegraphics[width=\singlecolumnwidth]{1.pdf}
 \caption{\label{fig:network}The regulation network example described in Sec.~\ref{sec:results}.
@@
 \begin{figure}
-\centering\includegraphics[width=\singlecolumnwidth]{figures_jcp/2.pdf}\caption{\label{fig:case-1-lsq}Concentration time series generated from integrating
+\centering\includegraphics[width=\singlecolumnwidth]{2.pdf}\caption{\label{fig:case-1-lsq}Concentration time series generated from integrating
 the reaction network shown in Fig.~\ref{fig:network}a. The initial
@@
 \begin{figure*}
-\centering \includegraphics[width=0.82\doublecolumnwidth]{figures_jcp/3.pdf}
+\centering \includegraphics[width=0.82\doublecolumnwidth]{3.pdf}
 \caption{\label{fig:case-1-sparsity-pattern}Estimated reaction rates in the
@@
 \begin{figure}
-\centering \includegraphics[width=\singlecolumnwidth]{figures_jcp/4.pdf}
+\centering \includegraphics[width=\singlecolumnwidth]{4.pdf}
 \caption{\label{fig:case2-convergence}Convergence of the estimation error
@@
 \begin{figure}
-\centering \includegraphics[width=\singlecolumnwidth]{figures_jcp/5.pdf}
+\centering \includegraphics[width=\singlecolumnwidth]{5.pdf}
 \caption{\label{fig:case3-convergence}Convergence of estimation error of reaction
@@
     \centering
-    \includegraphics[width=\singlecolumnwidth]{figures_jcp/6.pdf}
+    \includegraphics[width=\singlecolumnwidth]{6.pdf}
     \caption{\label{fig:mapk}Application of reactive SINDy to the MAPK pathway system.
@@
     \centering
-    \includegraphics[width=\singlecolumnwidth]{figures_jcp/7.pdf}
+    \includegraphics[width=\singlecolumnwidth]{7.pdf}
     \caption{\label{fig:lv}Application of reactive SINDy to the Lotka--Volterra system with social friction.
@@
 
-\centering\includegraphics[width=\singlecolumnwidth]{figures_jcp/8.pdf}\caption{\label{fig:reaction-rates-sorted}Reaction rates sorted by their magnitude
+\centering\includegraphics[width=\singlecolumnwidth]{8.pdf}\caption{\label{fig:reaction-rates-sorted}Reaction rates sorted by their magnitude
 to determine the cutoff $\kappa=0.22$ of Sec.~\ref{sec:case-1}.
